### 2017-09-06 Tex/paper.tex
Tex file

\documentclass[headsepline=true, abstracton]{scrartcl}
%\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{verbatim}
\usepackage[a4paper,text={140mm,215mm},centering,headsep=10mm,footskip=10mm]{geometry}
\usepackage[urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black,bookmarks]{hyperref}
\usepackage{aliascnt}
\usepackage{lmodern}
\usepackage{mdwlist}
\usepackage{listings}
\usepackage[activate]{pdfcprot}
\usepackage{graphicx}
\usepackage{slashed}
\usepackage{mathrsfs}
\usepackage{geometry}
\usepackage{float}
\usepackage{oldgerm}
\usepackage{setspace}
\usepackage{dsfont}
%\usepackage{mathtools}
\usepackage[all]{xy}
\usepackage{cite}
\usepackage{url}
\pagestyle{headings}
\newcommand{\myclearpage}{\clearpage}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\newenvironment{gelaber}{}{}
\newenvironment{preamble}{}{}
\newcommand{\tostar}{\overset{*}{\lower0.5em\hbox{$\smash{\scriptscriptstyle\rightharpoonup}$}}}
\newtheorem{mydef}{Definition}
\newtheorem{bem}{Bemerkung}
\usepackage{tikz}
\newcommand\circlearound[1]{%
  \tikz[baseline]\node[draw,shape=circle,anchor=base] {#1} ;}

\begin{document}

\setcounter{section}{-1}

\renewcommand{\refname}{Bibliography}

\onehalfspacing
\setlength{\headsep}{15mm}

\thispagestyle{plain}

\title{\Large Analyzing the Supreme Court Citation Network}
\maketitle

\begin{abstract}
\noindent 
\end{abstract}

 \section*{Introduction}

 Coming...

 \section*{The Exponential Random Configuration Model}
Let $c(t)\in \{0,1\}^N$ be a vector indicating which Supreme Court case has been cited at time $t$, where $c_i(t)=1, i \in \{1, \dots , N\}$ indicates that the $i$th case has been cited at time $t$ and $c_i(t)=0$ indicates that the $i$th case has not been cited at time $t$. Furthermore, let
$$\mathcal{C}_t(N)=\{c(t)\in \{0,1\}^N: c_i(t)\in \{0,1\} \}$$ 
be the set of all possible citation combinations at time $t$. Note that the cardinality of $\mathcal{C}_t(N)$ increases exponentially for every newly added case, which results in $2^N$ elements.\\
The probability function of the ERCM is defined as
\begin{equation}
P_{\theta}(c(t)~|~c(t-))=\cfrac{\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}{\sum_{c(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}
\label{ercm}
\end{equation}
where $c_{t-}\in \{0,1\}^{N \times (t-1)}$ is a matrix that indicates which cases have been citing in each other before time $t$, $\theta \in \mathbb{R}^q$ is a q-dimensional vector of parameters,  $h: \mathcal{C}_t(N) \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{c(t)^*\in \mathcal{C}}\exp(\theta^T\cdot h(c(t)|c(t-)))$ is a normalization constant that ensures that (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.\\[0.3cm]

% Network Statistics
The generative process of a model are informed by
the decision regarding which network statistics $h(\cdot)$ are incorporated. We include the following statistics for the Supreme Court citation network:
$$h_{edges}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{i=1}^Nc_i(t)$$
the number of citations made at time $t$. 
$$h_{outstar}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
the number of weighted outstars occuring at time $t$. We argue that it should me more likely to cite more recent cases than cases that have been decided further in the past. For the weight 
$$w(a,b):= \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
we define $a$ and $b$ as the elapsed time since case $i$ and $j$ have been introduced to the network.
$$h_{triangle}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot
c_j(t_{-i}) \cdot w(a,b)$$
where $c_j(t_{-i})$ indicates whether case $j$ was cited at the time case $i$ was introduced into the network. Just as for the outstar statistic, we include a weighting factor to favor more recent cases. \\[0.3cm]

%Change Statistic
The individual entries $c_i(t)$ can be taken as a manifestation of single Bernoulli variables $C_i(t)$. This interpretation allows the following calculation regarding the conditional distribution of $C_i(t)$:
%
\begin{eqnarray*}
\dfrac{P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~|~ C_i(t)^c=c_i(t)^c)} &=&
\dfrac{P_{\theta}(C_i(t)=1 ~,~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~,~ C_i(t)^c=c_i(t)^c)} \\
                           &=&\dfrac{P_{\theta}(C(t)= c_i^+(t))}{P_{\theta}(C(t)=c_i^-(t))}\\
                           &=&\dfrac{\exp(\theta^T \cdot h(c_i^+(t)~|~c(t-)))}{\exp(\theta^T \cdot h(c_i^-(t)~|~c(t-)))}\\
                           &=& \exp(\theta^T \cdot (h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-)))
\end{eqnarray*}
%
This implies the following equation:
%
\begin{equation}
\text{logit}(P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c))= \theta^T \cdot (h(c_i^+(t)~~|c(t-)) - h(c_i^-(t)~|~c(t-)))
\label{Logit}
\end{equation}
In the equation above the following notations were used:
%
\begin{itemize}
\item $c_i^+(t)$ emerges from $c(t)$, while assuming $c_i(t)=1$
\item $c_i^-(t)$ emerges from $c(t)$, while assuming $c_i(t)=0$
\item The condition $C_i(t)^c=c_i(t)^c$ is short for: $C_j(t)=c_j(t)$ for all $j\in \{1,\dots,N\}$ with $i \neq j$
\item The expression $(\Delta c_i)(t):=h(c_i^+(t)~|~c(t-)) - h(c_i^-(t))$ is called the \textit{change statistic}. The $k$th component of $(\Delta c_i)(t)$ captures the difference between citation networks $c_i^+(t)$ and $c_i^-(t)$ on the $k$th integrated statistic in the model
\end{itemize}

\section*{Estimation}

% Pseudo Likelihood
\subsection*{Maximum Pseudo-Likelihood Estimator}
One can assume that the dyads are independent of each other, which means that
the random variables $C_i(t)$ inside the random vector $C(t)$ are independent of each other.
In this case, the equation (\ref{Logit}) reduces to
$$logit(P_{\theta}(C_i(t) = 1)) = \theta^T \cdot (\Delta c_i)(t)$$
This corresponds with the logistic regression approach, where the observations of
the dependent variables are simply edge values of the observed citation vector,
and the observations of the covariate values are given as the scores of every single
change statistic. Therefore, the resulting likelihood function is of the following form:
\begin{equation}
\text{lik}(\theta)= P_{\theta}(C(t)=c(t))= \prod_{i} \dfrac{ \exp \left(\theta^T \Delta(c_i)(t) \right)}{1+\exp \left(\theta^T \Delta(c_i)(t) \right)}
\label{PseudoLik}
\end{equation}

\subsection*{Maximum Likelihood Estimator}
The more rigorous technique is to estimate the parameters directly with the log-likelihood function derived from (\ref{ercm}), which has the following form:
%
\begin{equation}
\text{loglik}(\theta)=\theta^T \cdot h(c(t)| c(t-))-\log(\kappa(\theta))
\label{loglik}
\end{equation}
%
The problem resulting from estimating the parameters with (\ref{loglik}) is that the term
%
$$\kappa(\theta):= \sum_{c(t)^* \in \mathcal{C}(N)} \exp(\theta^T \cdot h(c(t)^*|c(t-)))$$ 
%
which sums up the weighted statistics of all possible binar vectors of length $N$, has to be evaluated.

\subsection*{MCMC}

 \section*{Results/Discussion}

\newpage

\bibliography{bib} 
\bibliographystyle{plain}

\end{document}

### 2017-09-13 Tex/paper.tex
MCMLE estimation and MH algorithm

@@
 \end{equation}
-where $c_{t-}\in \{0,1\}^{N \times (t-1)}$ is a matrix that indicates which cases have been citing in each other before time $t$, $\theta \in \mathbb{R}^q$ is a q-dimensional vector of parameters,  $h: \mathcal{C}_t(N) \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{c(t)^*\in \mathcal{C}}\exp(\theta^T\cdot h(c(t)|c(t-)))$ is a normalization constant that ensures that (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.\\[0.3cm]
+where $c(t-)\in \{0,1\}^{N \times (t-1)}$ is a matrix that indicates which cases have been citing in each other before time $t$, $\theta \in \mathbb{R}^q$ is a q-dimensional vector of parameters,  $h: \mathcal{C}_t(N) \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{c(t)^*\in \mathcal{C}}\exp(\theta^T\cdot h(c(t)|c(t-)))$ is a normalization constant that ensures that (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.\\[0.3cm]
 
@@
 $$h_{outstar}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
-the number of weighted outstars occuring at time $t$. We argue that it should me more likely to cite more recent cases than cases that have been decided further in the past. For the weight 
+the number of weighted outstars occuring at time $t$. We argue that it should be more likely to cite more recent cases than cases that have been decided further in the past. For the weight 
 $$w(a,b):= \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
@@
 which sums up the weighted statistics of all possible binar vectors of length $N$, has to be evaluated.
-
-\subsection*{MCMC}
+However, the cardinality of $\mathcal{C}(N)$ (\#$(\mathcal{C})=2^N$) is incredibly large and a direkt calculation of this sum is for already small $N$ not feasible.
+\\[0.3cm]
+An solution for this limitation is based on the following consideration: Fix a vector of parameters $\theta_0 \in \Theta$ from the underlying parameter range $\Theta$ and compute for $\theta \in \Theta$ the expected value
+%
+\begin{eqnarray*}
+\mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right]&=& 
+\sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot \mathbb{P}_{\theta_0}(C(t)=c(t))\\
+&=& \sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot 
+\frac{\exp(\theta_0^T \cdot \Gamma(c(t)))}{\kappa(\theta_0)}\\
+&=& \frac{1}{\kappa(\theta_0)} \sum_{c(t) \in \mathcal{C}(N)}\exp\left(\theta^T \cdot \Gamma(c(t))\right)\\
+&=&\frac{\kappa(\theta)}{\kappa(\theta_0)}
+\end{eqnarray*}
+%
+This equation offers the following possibility: If one draws $L$ random vectors $c^{(1)}(t), \dots ,c^{(L)}(t)$ out of a distribution $\mathbb{P}_{\theta_0}$ appropriately, one gets with the law of big numbers and a big enough sample $L$ the following relation:
+%
+\begin{equation}
+\frac{1}{L}\cdot \sum_{i=1}^{L}  \exp\left((\theta - \theta_0)^T \cdot \Gamma(c^{(i)}(t))\right)
+~~\approx~~ \mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right] = \frac{\kappa(\theta)}{\kappa(\theta_0)}
+\label{konver}
+\end{equation}
+%
+This approximate can then be used to approximate the log likelihood function.\\[0.4cm]
+Next, we will discuss how a sufficient number of suitable drawings $c^{(1)}(t), \dots ,c^{(L)}(t)$ can be sampled from the distribution $\mathbb{P}_{\theta_0}$. \\
+For this purpose, the Markov Chain Monte Carlo (MCMC) methods can be used.
+\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
+To be able to compute the approximate likelihood function one needs a sufficiently large number of random vectors from the distribution $\mathbb{P}_{\theta_0}$.
+Snijders \cite{Snijders.2002b} introduces an approach to sample random networks for the ERGM framework by using \textit{MCMC methods}.
+We adapt this approach for sampling appropriate binary vectors for the ERCM.
+\\[0.4cm]
+\textit{Metropolis-Hastings Algorithm}\\\\
+Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ to start with (e.g., the observed vector). For $k \in \{0,...,L-1\}$ recursively proceed as follows:\\
+\begin{enumerate}
+\item Randomly choose a number $i\in \{1,\dots, N\}$
+\item Compute, using the equation (\ref{Logit}) the value
+$$\pi := \dfrac{\mathbb{P}_{\theta}(C_{i}(t) \neq c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}{\mathbb{P}_{\theta}(C_{i}(t) = c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}$$
+\item Define $\delta:= \min\{1, \pi\}$ and draw a random number $Z$ from Bin$(1, \delta)$. If
+\begin{itemize}
+\item $Z=0$, let $c^{(k+1)}(t) := c^{(k)}(t)$ 
+\item $Z=1$, define $c^{(k+1)}(t)$ as
+$$c_{p}^{(k+1)}(t)=\begin{cases}
+1-c_{p}^{(k)}(t)& \text{if}~ p=i \\
+c_{p}^{(k)}(t) &\text{if}~ p \neq i 
+\end{cases}$$
+\end{itemize}
+\item Start at step 1 with $c^{(k+1)}(t)$.
+\end{enumerate}
+The first $B \ll L$ vectors are discarded as Burn-In.
 

### 2017-09-13 R-Code/ERCM.R
MPLE estimation

@@ -36,10 +36,10 @@ for(i in 1:209946){
 
 ## down-weighted triangle statistic 1787
 
-triangles<- function(c_t){
+triangles<- function(c_t, time_t){
   # get right colum of time t
-  col.is.a.match <- apply(citation.matrix, 2, identical, c_t)
-  time_t<- which(col.is.a.match)
+  #col.is.a.match <- apply(citation.matrix, 2, identical, c_t)
+  #time_t<- which(col.is.a.match)
   
   # which rows have an entry 1
   citation.row<- which(c_t == 1)
@@ -73,9 +73,9 @@ edges<- function(c_t){
 }
 
 # down-weighted outstar statistic
-outstar<- function(c_t){
-  col.is.a.match <- apply(citation.matrix, 2, identical, c_t)
-  time_t<- which(col.is.a.match)
+outstar<- function(c_t, time_t){
+  #col.is.a.match <- apply(citation.matrix, 2, identical, c_t)
+  #time_t<- which(col.is.a.match)
   
   # which rows have an entry 1
   citation.row<- which(c_t == 1)
@@ -93,10 +93,64 @@ outstar<- function(c_t){
       year.c_t_a <- scc1[citation.row[j],4]
       year.c_t_a_b <- scc1[citation.row[i],4]
       
-      ost.count <- ost.count + sqrt(((year.c_t-year.c_t_a+1)*(year.c_t-year.c_t_a_b+1)/year.c_t^2))
+      ost.count <- ost.count + ((year.c_t-year.c_t_a+1)*sqrt((year.c_t-year.c_t_a_b+1)/year.c_t^2))
       
       }
   }
   return(ost.count)
 }
 
+############### Estimation
+
+# Pseudolikelihood
+
+# estimate coefficients for time t=1431 | condition on years before 1869
+
+# get vector c1431
+time.t<- 5000
+
+cit.mat.row<- which(scc1[,71]==time.t) # could be multiple rows
+#c1431<- citation.matrix[1:max(cit.mat.row), 1431] 
+ct<- citation.matrix[,time.t]
+
+# create matrix to store results, such that we can fit a logistic regression model; edges is intercept -> no column for edges
+
+change.mat<- matrix(0, max(cit.mat.row), 3)
+colnames(change.mat)<- c("citation", "outstar", "triangle")
+
+# fill in right values
+
+change.mat[,1]<- ct[1:max(cit.mat.row)]
+
[…]

### 2017-09-20 Tex/paper.tex
Gibbs sampling

@@
 \item The condition $C_i(t)^c=c_i(t)^c$ is short for: $C_j(t)=c_j(t)$ for all $j\in \{1,\dots,N\}$ with $i \neq j$
-\item The expression $(\Delta c_i)(t):=h(c_i^+(t)~|~c(t-)) - h(c_i^-(t))$ is called the \textit{change statistic}.
+\item The expression $(\Delta c_i)(t):=h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-))$ is called the \textit{change statistic}.
 The $k$th component of $(\Delta c_i)(t)$ captures the difference between citation networks $c_i^+(t)$ and $c_i^-(t)$ on the $k$th integrated statistic in the model
@@
 For this purpose, the Markov Chain Monte Carlo (MCMC) methods can be used.
-\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
+\subsection*{Gibbs sampling for the ERCM}
 To be able to compute the approximate likelihood function one needs a sufficiently large number of random vectors from the distribution $\mathbb{P}_{\theta_0}$.
@@
 \\[0.4cm]
-\textit{Metropolis-Hastings Algorithm}\\\\
+\textit{Gibbs sampling}\\\\
+Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g.
+observed vector).
+Afterwards, the length $L$ of the respective sub-sequence is determined.
+For $k \in \{0,...,L-1\}$ execute the following steps recursively (here the vector in its $k$th iteration is denoted as $c^{(k)}(t)$):\\
+\begin{enumerate}
+\item Randomly choose a number $i \in \{1,\dots, N\}$
+\item Compute using the likelihood the value 
+$$\pi:= \mathbb{P}_{\theta}(C_{i}(t)=1 | C_{i}^c(t)=(c_{i}^{(k)}(t))^c)=\cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$$
+\item Draw a random number $Z$ from Bin$(1, \pi)$. If
+\begin{itemize}
+\item $Z=0$, define $c^{(k+1)}(t)$ via
+$$c_{p}^{(k+1)}(t)=\begin{cases}
+0& \text{if}~ p=i \\
+c_{p}^{(k)}(t) &\text{if}~ p \neq i 
+\end{cases}$$
+\item $Z=1$, define $c^{(k+1)}(t)$ via
+$$c_p^{(k+1)}(t)=\begin{cases}
+1& \text{if}~ p=i \\
+c_p^{(k)}(t) &\text{if}~ p \neq i 
+\end{cases}$$
+\end{itemize}
+\item Start at step 1 with $c^{(k+1)}(t)$.
+\end{enumerate}
+\noindent The depicted algorithm provides a sequence of random vectors $c^{(0)}(t),...,c^{(L)}(t)$.
+Since the orignial vector was chosen randomly and the first simulated vectors are very dependent on the chosen mvector (only one entry is changed per iteration!), usually the first $B$ networks, where $N \ll B \ll L$, are discarded as the so called \textit{Burn-In}.
+\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
 Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ to start with (e.g., the observed vector). For $k \in \{0,...,L-1\}$ recursively proceed as follows:\\

### 2017-10-12 Tex/paper.tex
algorithmics

--- 2017-09-13 Tex/paper.tex
@@
 \usepackage{cite}
+%\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{algorithm2e}
 \usepackage{url}
@@
 \\[0.4cm]
-\textit{Metropolis-Hastings Algorithm}\\\\
+\textit{Gibbs sampling}\\
+%Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g.
+observed vector).
+Afterwards, the length $L$ of the respective sub-sequence is determined.
+For $k \in \{0,...,L-1\}$ execute the following steps recursively (here the vector in its $k$th iteration is denoted as $c^{(k)}(t)$):\\
+
+\begin{algorithm}[H]
+ Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g. observed vector)\\
+ \For{i in 1:N}{
+  Compute $\pi:= \cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$\\
+  Draw a random number Z from Bin(1,$\pi$)\\
+  \eIf{Z=1}{
+   set $c^{(k+1)}_i = 1$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
+   }{
+   set $c^{(k+1)}_i = 0$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
+  }
+ }
+ Start all over using $c^{(k+1)}$\\[0.3cm]
+ \caption{Simulation of vectors of $\mathbb{P}_\theta$ using Gibbs sampling}
+\end{algorithm}
+\vspace{0.5cm}
+%\begin{enumerate}
+%\item Randomly choose a number $i \in \{1,\dots, N\}$
+%\item Compute using the likelihood the value 
+%$$\pi:= \mathbb{P}_{\theta}(C_{i}(t)=1 | C_{i}^c(t)=(c_{i}^{(k)}(t))^c)=\cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$$
+%\item Draw a random number $Z$ from Bin$(1, \pi)$. %If
+%\begin{itemize}
+%\item $Z=0$, define $c^{(k+1)}(t)$ via
+%$$c_{p}^{(k+1)}(t)=\begin{cases}
+%0& \text{if}~ p=i \\
+%c_{p}^{(k)}(t) &\text{if}~ p \neq i 
+%\end{cases}$$
+%\item $Z=1$, define $c^{(k+1)}(t)$ via
+%$$c_p^{(k+1)}(t)=\begin{cases}
+%1& \text{if}~ p=i \\
+%c_p^{(k)}(t) &\text{if}~ p \neq i 
+%\end{cases}$$
+%\end{itemize}
+%\item Start at step 1 with $c^{(k+1)}(t)$.
+%\end{enumerate}
+\noindent Using the depicted algorithm, a sequence of random vectors $c^{(0)}(t),...,c^{(L)}(t)$ can be simulated.
+Since the orignial vector was chosen randomly and the first simulated vectors are very dependent on the chosen mvector (only one entry is changed per iteration!), usually the first $B$ vectors, where $N \ll B \ll L$, are discarded as the so called \textit{Burn-In}.
+\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
 Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ to start with (e.g., the observed vector). For $k \in \{0,...,L-1\}$ recursively proceed as follows:\\

### 2017-10-12 R-Code/ERCM3.R
3 new predictors (MQ score, Issue Area, and Year)

@@ -2,6 +2,7 @@
 
 scc1<- scc
 
+
 # create empty adjacency matrix
 memory.limit(30000)
 adjacency.matrix<- matrix(0, 26803, 26803)
@@ -25,16 +26,16 @@ adjacency.matrix<- adjacency.matrix[17987:26803, 17987:26803]
 
 # write function that cuts out the necessary adjacency matrix for a time t
 
-AM.t<- function(time.t){ # first id with MQ scores is 3136, last 5251
+#AM.t<- function(time.t){ # first id with MQ scores is 3136, last 5251
   # what is the id of the case/cases that entered the network at time t
-  latest.case<- max(which(time.t==scc1[, 71]))
-  AM<- adjacency.matrix[1:(latest.case-17986), 1:(latest.case-17986)]
-  return(AM)
-}
+#  latest.case<- max(which(time.t==scc1[, 71]))
+#  AM<- adjacency.matrix[1:(latest.case-17986), 1:(latest.case-17986)]
+#  return(AM)
+#}
 
 
 #### statistics
-time.t<- 5251
+time.t<- 3500
 latest.case<- max(which(time.t==scc1[, 71]))
 year.latest.case <- scc1[latest.case, 4]
 
@@ -48,8 +49,30 @@ lines(age,decayFun(age,0.05),col="blue",lwd=2)
 lines(age,decayFun(age,0.025),col="grey60",lwd=2)
 legend("topright",legend=c("1","0.25","0.10","0.05","0.025"),lwd=2,col=c("black","red","green","blue","grey60"),title="Beta")
 
+#######################################
+# adjust Am.t such that it can also cut mq.matrix into the right shape
 
-A<- AM.t(time.t)
+# mod=1 adjacency matrix, mod=2, Marting Quinn score matrix, mod=3 Issue are, mod=4, Year difference matrix
+AM.t<- function(time.t, mod){ # first id with MQ scores is 3136, last 5251
+  # what is the id of the case/cases that entered the network at time t
+  latest.case<- max(which(time.t==scc1[, 71]))
+  AM<- adjacency.matrix[1:(latest.case-17986), 1:(latest.case-17986)]
+  MQ<- mq.matrix[1:(latest.case-17986), 1:(latest.case-17986)]
+  IA <- same.issue.area[1:(latest.case-17986), 1:(latest.case-17986)]
[…]

### 2017-11-22 Tex/paper.tex
first results

@@
 \usepackage{listings}
+\usepackage[table,xcdraw]{xcolor}
 \usepackage[activate]{pdfcprot}
@@
 
- \section*{Results/Discussion}
+ \section*{Results}
+  \subsection*{Descriptive Results}
+ The supreme court citation network from $1937 - 2005$ consists of $8817$ cases which got voted at 2116 different time points.
+The network has a total of $93,263$ ties, of which 452 are mutual.
+The number of triangles in the network is $211,855$.
+The in- and outdegree distribution is visualized in figures \ref{indegree_dist} and \ref{outdegree_dist}.
+The maximum indegree is $190$ and the maximum outdegree is $159$.
+
+\begin{figure}
+\includegraphics[scale=0.5]{degree_distribution}
+\caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2005. There are cases with an indegree >50, but they are not captured in this figure.}
+ \label{degree_dist}
+\vspace{-.25cm}
+\end{figure}
+
+\subsection*{Inferential Results}
+
+\begin{table}[]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
+Edges                   & -5.364                                                 & -5.411                                                     & -4.155                                                     \\ \hline
+Instar(2)               & 0.031                                                   & 0.025                                                      & 0.037                                                      \\ \hline
+Outstar(2)              & 0.016                                                   & -0.004                                                      & 0.017                                                      \\ \hline
+Mutual                  & 4.472                                                   & 3.821                                                 & 5.069                                                      \\ \hline
+Triangle                & 1.275                                                   & 1.194                                                      & 1.401                                                      \\ \hline
+MQ Score                & 0.174                                                   & 0.082                                                      & 0.244                                                      \\ \hline
+Same Issue Area         & 1.113                                                   & 0.972                                                      & 1.152                                                      \\ \hline
+Year Difference         & -0.092                                                   & -0.114                                                     & -0.085                                                          \\ \hline
+\end{tabular}
+\caption{Bootstrappped MPLE Results for the timep period when Fred M. Vinson was chief justice.}
+\label{bootMPLE_vinson}
+\end{table}
+
+\begin{table}[]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
+Edges                   & -3.571                                                 & -3.597                                                     & -3.460                                                     \\ \hline
+Instar(2)               & 0.014                                                   & 0.013                                                      & 0.016                                                      \\ \hline
+Outstar(2)              & 0.001                                                   & -0.001                                                      & 0.004                                                      \\ \hline
+Mutual                  & 9.678                                                   & 8.308                                                 & 11.261                                                      \\ \hline
+Triangle                & 1.526                                                   & 1.417                                                      & 1.527                                                      \\ \hline
+MQ Score                & -0.0001                                                   & -0.043                                                      & 0.012                                                      \\ \hline
+Same Issue Area         & 0.826                                                   & 0.718                                                      & 0.822                                                      \\ \hline
+Year Difference         & -0.017                                                   & -0.018                                                     & -0.015                                                          \\ \hline
+\end{tabular}
+\caption{Bootstrappped MPLE Results for the timep period when William Rehnquist was chief justice.}
+\label{bootMPLE_vinson}
+\end{table}
 

### 2017-11-22 Tex/paper.tex
Inferential Results

@@
 {\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -5.364                                                 & -5.411                                                     & -4.155                                                     \\ \hline
-Instar(2)               & 0.031                                                   & 0.025                                                      & 0.037                                                      \\ \hline
-Outstar(2)              & 0.016                                                   & -0.004                                                      & 0.017                                                      \\ \hline
-Mutual                  & 4.472                                                   & 3.821                                                 & 5.069                                                      \\ \hline
-Triangle                & 1.275                                                   & 1.194                                                      & 1.401                                                      \\ \hline
-MQ Score                & 0.174                                                   & 0.082                                                      & 0.244                                                      \\ \hline
-Same Issue Area         & 1.113                                                   & 0.972                                                      & 1.152                                                      \\ \hline
-Year Difference         & -0.092                                                   & -0.114                                                     & -0.085                                                          \\ \hline
+Edges                   & -5.445                                                 & -5.514                                                     & -5.296                                                     \\ \hline
+Instar(2)               & 0.033                                                   & 0.028                                                      & 0.040                                                      \\ \hline
+Outstar(2)              & 0.015                                                   & -0.001                                                      & 0.016                                                      \\ \hline
+Mutual                  & 4.717                                                   & 4.093                                                 & 5.422                                                      \\ \hline
+Triangle                & 1.337                                                   & 1.251                                                      & 1.475                                                      \\ \hline
+MQ Score                & -0.009                                                   & -0.035                                                      & 0.087                                                      \\ \hline
+Same Issue Area         & 1.205                                                   & 1.064                                                      & 1.237                                                      \\ \hline
+Year Difference         & -0.076                                                   & -0.101                                                     & -0.072                                                          \\ \hline
 \end{tabular}
@@
 {\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -3.571                                                 & -3.597                                                     & -3.460                                                     \\ \hline
-Instar(2)               & 0.014                                                   & 0.013                                                      & 0.016                                                      \\ \hline
-Outstar(2)              & 0.001                                                   & -0.001                                                      & 0.004                                                      \\ \hline
-Mutual                  & 9.678                                                   & 8.308                                                 & 11.261                                                      \\ \hline
-Triangle                & 1.526                                                   & 1.417                                                      & 1.527                                                      \\ \hline
-MQ Score                & -0.0001                                                   & -0.043                                                      & 0.012                                                      \\ \hline
-Same Issue Area         & 0.826                                                   & 0.718                                                      & 0.822                                                      \\ \hline
-Year Difference         & -0.017                                                   & -0.018                                                     & -0.015                                                          \\ \hline
+Edges                   & -5.042                                                 & -5.107                                                     & -4.952                                                     \\ \hline
+Instar(2)               & 0.028                                                   & 0.026                                                      & 0.033                                                      \\ \hline
+Outstar(2)              & 0.018                                                   & 0.017                                                      & 0.022                                                      \\ \hline
+Mutual                  & 3.158                                                   & 1.469                                                 & 3.508                                                      \\ \hline
+Triangle                & 1.500                                                   & 1.391                                                      & 1.506                                                      \\ \hline
+MQ Score                & -0.106                                                   & -0.142                                                      & -0.067                                                      \\ \hline
+Same Issue Area         & 1.334                                                   & 1.274                                                      & 1.378                                                      \\ \hline
+Year Difference         & -0.031                                                   & -0.040                                                     & -0.029                                                          \\ \hline
+\end{tabular}
+\caption{Bootstrappped MPLE Results for the timep period when Earl Warren was chief justice.}
+\label{bootMPLE_vinson}
+\end{table}
+
+\begin{table}[]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
+Edges                   & -3.961                                                 & -4.036                                                     & -3.911                                                     \\ \hline
+Instar(2)               & 0.016                                                   & 0.014                                                      & 0.037                                                      \\ \hline
+Outstar(2)              & 0.004                                                   & 0.003                                                      & 0.006                                                      \\ \hline
+Mutual                  & 2.528                                                   & 1.193                                                 & 3.007                                                      \\ \hline
+Triangle                & 1.372                                                   & 1.248                                                      & 1.341                                                      \\ \hline
+MQ Score                & -0.084                                                   & -0.096                                                      & -0.054                                                      \\ \hline
+Same Issue Area         & 0.957                                                   & 0.917                                                      & 0.995                                                      \\ \hline
+Year Difference         & -0.017                                                   & -0.019                                                     & -0.015                                                          \\ \hline
+\end{tabular}
+\caption{Bootstrappped MPLE Results for the timep period when Warren E. Burger was chief justice.}
+\label{bootMPLE_vinson}
+\end{table}
+
+\begin{table}[]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
+Edges                   & -3.618                                                 & -3.666                                                     & -3.539                                                     \\ \hline
+Instar(2)               & 0.015                                                   & 0.014                                                      & 0.016                                                      \\ \hline
+Outstar(2)              & 0.001                                                   & -0.002                                                      & 0.004                                                      \\ \hline
+Mutual                  & 9.667                                                   & 8.218                                                 & 11.253                                                      \\ \hline
+Triangle                & 1.546                                                   & 1.433                                                      & 1.551                                                      \\ \hline
+MQ Score                & -0.007                                                   & -0.023                                                      & 0.032                                                      \\ \hline
+Same Issue Area         & 0.852                                                   & 0.751                                                      & 0.846                                                      \\ \hline
+Year Difference         & -0.016                                                   & -0.018                                                     & -0.014                                                          \\ \hline
 \end{tabular}

### 2017-11-29 Tex/paper.tex
descriptive and inferential results

--- 2017-10-12 Tex/paper.tex
@@
 
- \section*{Results/Discussion}
+ \section*{Results}
+  \subsection*{Descriptive Results}
+The supreme court citation network from $1937 - 2005$ consists of $8817$ cases which got voted at $2116$ different time points.
+The network has a total of $93,263$ ties, of which 452 are mutual.
+The number of triangles in the network is $211,855$.
+The in- and outdegree distribution is visualized in figures \ref{indegree_dist} and \ref{outdegree_dist}.
+The maximum indegree is $190$ and the maximum outdegree is $159$.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Terms} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Total Number Cases} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Cases/Term} \\ \hline
+CE Hughes*              & 1937 - 1941                                          & 629                                                               & 125.5                                                     \\ \hline
+HF Stone                & 1942 - 1946                                          & 766                                                               & 153.2                                                     \\ \hline
+FM Vinson               & 1946 - 1953                                          & 788                                                               & 98.5                                                      \\ \hline
+E Warren                & 1954 - 1969                                          & 2159                                                              & 127.0                                                     \\ \hline
+WE Burger               & 1970 - 1986                                          & 2805                                                              & 155.8                                                     \\ \hline
+W Rehnquist**            & 1987 - 2001                                          & 1670                                                              & 83.5                                                      \\ \hline
+\end{tabular}
+\caption[caption]{For the time range of interest (1937 - 2001) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
+\\\hspace{\textwidth} ** W Rehnquist served as chief justice from 1987 - 2005.}
+\label{my-label}
+\end{table}
+
+\begin{figure}[H]
+\includegraphics[scale=0.5]{degree_distribution}
+\caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2001. There are cases with an indegree >50, but they are not captured in this figure.}
+ \label{degree_dist}
+\vspace{-.25cm}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[scale=0.5]{number_cases}
+\caption{Number of cases in each term. Different colors indicate different chief justices.}
+ \label{number_cases}
+\vspace{-.25cm}
+\end{figure}
+
+ \begin{figure}[H]
+\includegraphics[scale=0.5]{number_citations}
+\caption{Number of citations for the 1937-2001 time period. Citations for cases prior 1937 are not considered in this figure. Different colors indicate different chief justices.}
+ \label{number_citations}
+\vspace{-.25cm}
+\end{figure}
+
+  \begin{figure}[H]
+\includegraphics[scale=0.5]{number_votes_supporting}
+\caption{Number of Votes that were supporting cases between 1937-2001. Different colors indicate terms with different chief justices. }
+ \label{number_supporting}
+\vspace{-.25cm}
+\end{figure}
+
+\newpage
+
+\subsection*{Inferential Results}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
+Edges                   & -5.319                                                 & -5.392                                                     & -5.243                                                     \\ \hline
+Instar(2)               & 0.018                                                   & 0.017                                                      & 0.019                                                      \\ \hline
+Outstar(2)              & 0.018                                                  & 0.016                                                      & 0.022                                                      \\ \hline
+Mutual                  & 3.179                                                   & 2.694                                                 & 3.565                                                      \\ \hline
+Triangle                & 1.531                                                   & 1.493                                                      & 1.563                                                      \\ \hline
+MQ Score                & 0.041                                                   & 0.018                                                      & 0.063                                                      \\ \hline
+Same Issue Area         & 1.061                                                   & 1.028                                                      & 1.099                                                      \\ \hline
+Year Difference         & 0.010                                                   & 0.008                                                     & 0.013                                                          \\ \hline
+\end{tabular}
+\caption{Bootstrappped MPLE Results for the time period 1937-2001}
+\label{bootMPLE_total}
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
+Edges                   & -5.391                                                 & -5.506                                                     & -5.244                                                     \\ \hline
+Instar(2)               & 0.035                                                   & 0.028                                                      & 0.042                                                      \\ \hline
+Outstar(2)              & 0.015                                                   & 0.0003                                                      & 0.017                                                      \\ \hline
+Mutual                  & 4.785                                                   & 3.616                                                 & 5.749                                                      \\ \hline
+Triangle                & 1.377                                                   & 1.271                                                      & 1.513                                                      \\ \hline
+MQ Score                & -0.018                                                   & -0.095                                                      & 0.061                                                      \\ \hline
+Same Issue Area         & 1.229                                                   & 1.126                                                      & 1.327                                                      \\ \hline
+Year Difference         & -0.077                                                   & -0.096                                                     & -0.061                                                          \\ \hline
+\end{tabular}
+\caption{Bootstrappped MPLE Results for the time period when Fred M. Vinson was chief justice.}
+\label{bootMPLE_vinson}
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
+Edges                   & -5.032                                                 & -5.140                                                     & -4.931                                                     \\ \hline
+Instar(2)               & 0.028                                                   & 0.025                                                      & 0.032                                                      \\ \hline
+Outstar(2)              & 0.018                                                   & 0.015                                                      & 0.022                                                      \\ \hline
+Mutual                  & 3.166                                                   & 1.826                                                 & 4.565                                                      \\ \hline
+Triangle                & 1.520                                                   & 1.449                                                      & 1.599                                                      \\ \hline
+MQ Score                & -0.101                                                   & -0.149                                                      & -0.052                                                      \\ \hline
+Same Issue Area         & 1.349                                                   & 1.289                                                      & 1.408                                                      \\ \hline
+Year Difference         & -0.031                                                   & -0.038                                                     & -0.026                                                          \\ \hline
+\end{tabular}
+\caption{Bootstrappped MPLE Results for the time period when Earl Warren was chief justice.}
+\label{bootMPLE_warren}
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
+Edges                   & -3.952                                                 & -4.017                                                     & -3.882                                                     \\ \hline
+Instar(2)               & 0.015                                                   & 0.013                                                      & 0.016                                                      \\ \hline
+Outstar(2)              & 0.004                                                   & 0.002                                                      & 0.006                                                      \\ \hline
+Mutual                  & 2.558                                                   & 1.425                                                 & 3.719                                                      \\ \hline
+Triangle                & 1.383                                                   & 1.333                                                      & 1.440                                                      \\ \hline
+MQ Score                & -0.084                                                   & -0.101                                                      & -0.059                                                      \\ \hline
+Same Issue Area         & 0.959                                                   & 0.911                                                      & 1.008                                                      \\ \hline
+Year Difference         & -0.017                                                   & -0.019                                                     & -0.015                                                          \\ \hline
+\end{tabular}
+\caption{Bootstrappped MPLE Results for the time period when Warren E. Burger was chief justice.}
+\label{bootMPLE_burger}
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
+Edges                   & -3.614                                                 & -3.701                                                     & -3.525                                                     \\ \hline
+Instar(2)               & 0.015                                                   & 0.013                                                      & 0.016                                                      \\ \hline
+Outstar(2)              & 0.002                                                   & -0.002                                                      & 0.004                                                      \\ \hline
+Mutual                  & 9.635                                                  & 6.426                                                 & 10.834                                                      \\ \hline
+Triangle                & 1.551                                                   & 1.489                                                      & 1.620                                                      \\ \hline
+MQ Score                & -0.007                                                   & -0.039                                                      & 0.028                                                      \\ \hline
+Same Issue Area         & 0.855                                                   & 0.794                                                      & 0.916                                                      \\ \hline
+Year Difference         & -0.016                                                   & -0.018                                                     & -0.014                                                          \\ \hline
+\end{tabular}
+\caption{Bootstrappped MPLE Results for the time period when William Rehnquist was chief justice.}
+\label{bootMPLE_rehnquist}
+\end{table}
 

### 2017-12-01 Tex/paper.tex
added introduction

@@
 
- Coming...
+Majority opinions written by the US Supreme Court exercise their authority and influence, in part, through their roles as precedents in future Supreme Court decisions and opinions [CITES].
+Research on the nature of {\em Stare Decisis}---the Court's reliance on precedent---has been mixed, with some finding that past opinions influence and constrain future opinions [CITES], and others finding that, while referenced in opinions, past cases do not effect the Court's current behavior [CITES].
 

### 2017-12-01 Tex/paper.tex
updating revisions to sections

@@
 
- \section*{Introduction}
+ \section{Introduction}
 
@@
 
- \section*{The Exponential Random Configuration Model}
+\section{Network Approaches to Studying Citations}
+
+\textbf{TED}, would you write a section that covers the ways in which network scholars have studied citations in the past, focusing on scientific papers, patents, and court cases?
+
+ \section{The Exponential Random Configuration Model}
 Let $c(t)\in \{0,1\}^N$ be a vector indicating which Supreme Court case has been cited at time $t$, where $c_i(t)=1, i \in \{1, \dots , N\}$ indicates that the $i$th case has been cited at time $t$ and $c_i(t)=0$ indicates that the $i$th case has not been cited at time $t$.
@@
 
-\section*{Estimation}
+\section{Estimation}
 
@@
 
- \section*{Results}
+ \section{Results}
   \subsection*{Descriptive Results}
@@
 
-\subsection*{Inferential Results}
+\subsection{Inferential Results}
+
+\textbf{Model updates}
+\begin{itemize}
+\item Add sender issue area and receiver issue area node covariates (i.e., nodeifactor and nodeofactor applied to issue area). 
+\item In the GLM, multiply each variable by the sender time covariate.
+\item Write the bootstrap coefficients to a file, send to Bruce so he can write up a code for summarizing the over-time trends in effects.
+\item Add a receiver variable (i.e., nodeicov) that equals the absolute difference between the maximum MQ score of a justice in the majority and a minimum MQ score of a justice in the majority.
+\item Add a receiver variable (nodeicov) equivalent to the number of justices in the majority
+
+\end{itemize}
 

### 2017-12-02 Tex/paper.tex
writing intro

@@
 \usepackage{mdwlist}
+\usepackage{natbib}
 \usepackage{listings}
@@
 
-\setcounter{section}{-1}
-
 \renewcommand{\refname}{Bibliography}
@@
 
-Majority opinions written by the US Supreme Court exercise their authority and influence, in part, through their roles as precedents in future Supreme Court decisions and opinions [CITES].
-Research on the nature of {\em Stare Decisis}---the Court's reliance on precedent---has been mixed, with some finding that past opinions influence and constrain future opinions [CITES], and others finding that, while referenced in opinions, past cases do not effect the Court's current behavior [CITES].
+Majority opinions written by the United States Supreme Court exercise their authority and influence, in part, through their roles as precedents in future Supreme Court decisions and opinions.
+The findings regarding the extent and exact nature of the influences of precedent have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,gillman2001s,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained}.
+Despite a considerable body of research that focuses on the way in which relevant precedents shape decision making on the Court, relatively little work has focused on understanding which past opinions are cited by an opinion.
+Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited by a case.
+We follow \citet{fowler2007network} and \citet{fowler2008authority}, and treat the citations between cases as forming a network.
 
@@
 To be able to compute the approximate likelihood function one needs a sufficiently large number of random vectors from the distribution $\mathbb{P}_{\theta_0}$.
-Snijders \cite{Snijders.2002b} introduces an approach to sample random networks for the ERGM framework by using \textit{MCMC methods}.
+Snijders \citet{Snijders.2002b} introduces an approach to sample random networks for the ERGM framework by using \textit{MCMC methods}.
 We adapt this approach for sampling appropriate binary vectors for the ERCM.
@@
 \bibliography{bib} 
-\bibliographystyle{plain}
+\bibliographystyle{apsr}
 

### 2017-12-02 Tex/paper.tex
adding some theory

@@
 Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited by a case.
-We follow \citet{fowler2007network} and \citet{fowler2008authority}, and treat the citations between cases as forming a network.
+We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics}.
+
+\section{Prior work on court citations}
+
+Measures of citations have been used in several recent studies of courts...\citet{hinkle2016transmission} study the citations between US state supreme courts.
 
@@
 \textbf{TED}, would you write a section that covers the ways in which network scholars have studied citations in the past, focusing on scientific papers, patents, and court cases?
+
+\section{Complex Dependence in Court Citations}
+
+\textbf{Transitivity:} Discussing past edges in the citation network leads to transitivity. If i cites j, it is also likely to discuss the cases on which j is based.
+
+The statute thus addresses the risk of a morally unjustifiable death sentence, not by minimizing it as precedent unmistakably requires, but by guaranteeing that in equipoise cases the risk will be realized, by ?placing a ?thumb [on] death?s side of the scale,?
+?
+Sochor v.
+Florida, 504 U.
+S.
+527, 532 (1992) (quoting Stringer v.
+Black, 503 U.
+S.
+222, 232 (1992); alteration in original).
+
+https://supreme.justia.com/cases/federal/us/548/163/dissent2.html
+
+``Never before the decision in Union Gas had we suggested that the bounds of Article III could be expanded by Congress operating pursuant to any constitutional provision other than the Fourteenth Amendment.
+Indeed, it had seemed fundamental that Congress could not expand the jurisdiction of the federal courts beyond the bounds of Article III.
+Marbury v.
+Madison, 1 Cranch 137 (1803).
+The plurality's citation of prior decisions for support was based upon what we believe to be a misreading of precedent.
+See Union Gas, 491 U.
+S., at 40-41 (SCALIA, J., dissenting).
+The plurality claimed support for its decision from a case holding the unremarkable, and completely unrelated, proposition that the States may waive their sovereign immunity, see id., at 14-15 (citing Parden v.
+Terminal Railway of Ala.
+Docks Dept., 377 U.
+S.
+184 (1964)), and cited as precedent propositions that had been merely assumed for the sake of argument in earlier cases, see 491 U.
+S., at 15 (citing Welch v.
+Texas Dept.
+of Highways and Public Transp., 483 U.
+S., at 475-476, and n.
+5, and County of Oneida v.
+Oneida Indian Nation of N.
+Y., 470 U.
+S., at 252).''
+
+https://supreme.justia.com/cases/federal/us/517/44/case.html
+
+Mutuality:
+
+\textbf{Also popularity} If j served as precedent for i, j is authoritative regarding the rule of i, and should also be cited.
 

### 2017-12-02 Tex/paper.tex
updating theory section

@@
 
-Mutuality:
+\textbf{Mutuality:} When deciding a set of related cases, the courts opinions regularly reference each other in order to reinforce their arguments...
+
+On a more specific level, Western argues that flight engineers must meet the same stringent qualifications as pilots, and that it was therefore quite logical to extend to flight engineers the FAA's age 60 retirement rule for pilots.
+Although the FAA's rule for pilots, adopted for safety reasons, is relevant evidence in the airline's BFOQ defense, it is not to be accorded conclusive weight.
+Johnson v.
+Mayor and City Council of Baltimore, ante at 472 U.
+S.
+370-371.
+The extent to which the rule is probative varies with the weight of the evidence supporting its safety rationale and "the congruity between the .
+.
+.
+occupations at issue." Ante at 472 U.
+S.
+371.
+In this case, the evidence clearly established that the FAA, Western, and other airlines all recognized that the qualifications for a flight engineer were less rigorous than those required for a pilot.
+[Footnote 28]
+
+https://supreme.justia.com/cases/federal/us/472/400/case.html
+
+The city, supported by several amici, argues for affirmance nonetheless.
+It asserts first that the federal civil service statute is not just a federal retirement provision unrelated to the ADEA, but in fact establishes age as a BFOQ for federal firefighters based on factors that properly go into that determination under the ADEA, see Western Air Lines, Inc.
+v.
+Criswell, post p.
+472 U.
+S.
+400.
+Second, the city asserts, a congressional finding that age is a BFOQ for a certain occupation is dispositive of that determination with respect to nonfederal employees in that occupation.
+We consider each of these contentions in turn.
+
+\url{https://supreme.justia.com/cases/federal/us/472/353/case.html#370}
 
 \textbf{Also popularity} If j served as precedent for i, j is authoritative regarding the rule of i, and should also be cited.
+Citations beget citations.
+Citation confers authority.
+Cases are even discussed in opinions in terms of the extent of influence they have had on other cases,
+
+"The first case in which this Court struck down a statute under the Equal Protection Clause of the Fourteenth Amendment was Strauder v.
+West Virginia, 100 U.
+S.
+303, decided in the 1879 Term.
+[Footnote 2/1] In the 1961 Term, we squarely held that the manner of apportionment of members of a state legislature raised a justiciable question under the Equal Protection Clause, Baker v.
+Carr, 369 U.
+S.
+186.
+That case was followed by numerous others, e.g.: that one person could not be given twice or 10 times the voting power of another person in a state-wide election merely because he lived in a rural area..."
+
+https://supreme.justia.com/cases/federal/us/400/112/case.html
 

### 2017-12-03 Tex/paper.tex
added theory regarding sociality

@@
 https://supreme.justia.com/cases/federal/us/400/112/case.html
+
+\textbf{Sociality} Citations sent beget citations sent.
+For each case that applies, there is likely a case that needs to be discussed in terms of why it does not apply.
+Justices often clarify not only those legal principles that apply, but often those that do not.
+When multiple lines of legal reasoning are drawn upon to justify a decision, often more are discussed in terms of their lack of applicability.
+
+"1.
+The "overbreadth" doctrine is not applicable here.
+There is nothing in the record to indicate that § 28.04 will have any different impact on any third parties' interests in free speech than it has on appellees' interests, and appellees have failed to identify any significant difference between their claim that § 28.04 is invalid on overbreadth grounds and their claim that it is unconstitutional when applied to their signs during a political campaign.
+Thus, it is inappropriate to entertain an overbreadth challenge to § 28.04.
+Pp.
+466 U.
+S.
+796-803."
+
+https://supreme.justia.com/cases/federal/us/466/789/
+
+``No allegations of racial or class-based invidiously discriminatory animus are required to establish a cause of action under the first part of  1985(2).
+The statutory provisions now codified at § 1985 were originally enacted as § 2 of the Civil Rights Act of 1871, and the substantive meaning of the 1871 Act has not been changed.
+The provisions relating to institutions and processes of the Federal Government (including the first part of § 1985(2)) -- unlike those encompassing activity that is usually of primary state concern (including the second part of § 1985(2) and the part of § 1985(3) involved in Griffin, supra -- contain no language requiring that the conspirators act with intent to deprive their victims of the equal protection of the laws.
+Thus, the reasoning of Griffin is not applicable here, and, given the structure of § 2 of the 1871 Act, it is clear''
+
+https://supreme.justia.com/cases/federal/us/460/719/
 

### 2017-12-05 Tex/paper.tex
new inferential results

--- 2017-12-02 Tex/paper.tex
@@
 https://supreme.justia.com/cases/federal/us/400/112/case.html
+
+\textbf{Sociality} Citations sent beget citations sent.
+For each case discussed, often related cases to discuss, even those that do not cite each other.
+For each case that applies, there is likely a case that needs to be discussed in terms of why it does not apply.
+Justices often clarify not only those legal principles that apply, but often those that do not.
+When multiple lines of legal reasoning are drawn upon to justify a decision, often more are discussed in terms of their lack of applicability.
+
+"1.
+The "overbreadth" doctrine is not applicable here.
+There is nothing in the record to indicate that \S 28.04 will have any different impact on any third parties' interests in free speech than it has on appellees' interests, and appellees have failed to identify any significant difference between their claim that \S 28.04 is invalid on overbreadth grounds and their claim that it is unconstitutional when applied to their signs during a political campaign.
+Thus, it is inappropriate to entertain an overbreadth challenge to \S 28.04.
+Pp.
+466 U.
+S.
+796-803."
+
+https://supreme.justia.com/cases/federal/us/466/789/
+
+''No allegations of racial or class-based invidiously discriminatory animus are required to establish a cause of action under the first part of  1985(2).
+The statutory provisions now codified at \S 1985 were originally enacted as \S 2 of the Civil Rights Act of 1871, and the substantive meaning of the 1871 Act has not been changed.
+The provisions relating to institutions and processes of the Federal Government (including the first part of \S 1985(2)) -- unlike those encompassing activity that is usually of primary state concern (including the second part of \S 1985(2) and the part of \P 1985(3) involved in Griffin, supra -- contain no language requiring that the conspirators act with intent to deprive their victims of the equal protection of the laws.
+Thus, the reasoning of Griffin is not applicable here, and, given the structure of \S 2 of the 1871 Act, it is clear''
+
+https://supreme.justia.com/cases/federal/us/460/719/
 
@@
 \begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+>{\columncolor[HTML]{EFEFEF}}l |l|l|l|l|}
 \hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -5.319                                                 & -5.392                                                     & -5.243                                                     \\ \hline
-Instar(2)               & 0.018                                                   & 0.017                                                      & 0.019                                                      \\ \hline
-Outstar(2)              & 0.018                                                  & 0.016                                                      & 0.022                                                      \\ \hline
-Mutual                  & 3.179                                                   & 2.694                                                 & 3.565                                                      \\ \hline
-Triangle                & 1.531                                                   & 1.493                                                      & 1.563                                                      \\ \hline
-MQ Score                & 0.041                                                   & 0.018                                                      & 0.063                                                      \\ \hline
-Same Issue Area         & 1.061                                                   & 1.028                                                      & 1.099                                                      \\ \hline
-Year Difference         & 0.010                                                   & 0.008                                                     & 0.013                                                          \\ \hline
+                                       & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}Lower Bound & \cellcolor[HTML]{EFEFEF}Upper Bound & \cellcolor[HTML]{EFEFEF}Significance \\ \hline
+Edges                                  & -5.151                           & -5.269                              & -5.038                              & *                                    \\ \hline
+Instar(2)                              & 0.00026                          & 0.00023                             & 0.00030                             & *                                    \\ \hline
+Outstar(2)                             & 0.00010                          & 0.00004                             & 0.00018                             & *                                    \\ \hline
+Mutual                                 & 0.190                            & 0.139                               & 0.260                               & *                                    \\ \hline
+Triangle                               & 0.039                            & 0.038                               & 0.041                               & *                                    \\ \hline
+Martin Quinn Score                     & -0.0017                          & -0.0021                             & -0.0010                             & *                                    \\ \hline
+Same Issue Area Homophily              & 0.027                            & 0.026                               & 0.028                               & *                                    \\ \hline
+Year Difference                        & -0.0005                          & -0.0006                             & -0.0004                             & *                                    \\ \hline
+$(\text{Year Differnce})^2$            & 7.498e-08                        & 4.629e-08                           & 1.057e-07                           & *                                    \\ \hline
+{\small Receiver Abs Diff MQ Score in Majority} & 0.002                            & -0.005                              & 0.009                               &                                      \\ \hline
+{\small Receiver Number Justices in Majority}   & -0.012                           & -0.022                              & -0.0002                             & *                                    \\ \hline
+Receiver  Sender Year                  & 0.0005                           & 0.0004                              & 0.0006                              & *                                    \\ \hline
+Sender Same Issue Area 2               & 0.155                            & 0.091                               & 0.214                               & *                                    \\ \hline
+Sender Same Issue Area 3               & -0.095                           & -0.189                              & -0.008                              & *                                    \\ \hline
+Sender Same Issue Area 4               & 0.401                            & 0.324                               & 0.481                               & *                                    \\ \hline
+Sender Same Issue Area 5               & 0.262                            & 0.089                               & 0.450                               & *                                    \\ \hline
+Sender Same Issue Area 6               & 0.350                            & 0.222                               & 0.472                               & *                                    \\ \hline
+Sender Same Issue Area 7               & 0.438                            & 0.361                               & 0.520                               & *                                    \\ \hline
+Sender Same Issue Area 8               & 0.064                            & 0.002                               & 0.128                               & *                                    \\ \hline
+Sender Same Issue Area 9               & 0.181                            & 0.123                               & 0.244                               & *                                    \\ \hline
+Sender Same Issue Area 10              & 0.305                            & 0.224                               & 0.380                               & *                                    \\ \hline
+Sender Same Issue Area 11              & -0.187                           & -0.371                              & -0.003                              & *                                    \\ \hline
+Sender Same Issue Area 12              & -0.244                           & -0.348                              & -0.149                              & *                                    \\ \hline
+Sender Same Issue Area 13              & -0.022                           & -0.270                              & 0.202                               &                                      \\ \hline
+Sender Same Issue Area 14              & -0.418                           & -0.613                              & -0.269                              & *                                    \\ \hline
+Receiver Same Issue Area 2             & 0.204                            & 0.126                               & 0.278                               & *                                    \\ \hline
+Receiver Same Issue Area 3             & 0.077                            & -0.030                              & 0.183                               &                                      \\ \hline
+Receiver Same Issue Area 4             & 0.459                            & 0.352                               & 0.573                               & *                                    \\ \hline
+Receiver Same Issue Area 5             & 0.320                            & 0.094                               & 0.529                               & *                                    \\ \hline
+Receiver Same Issue Area 6             & 0.411                            & 0.194                               & 0.615                               & *                                    \\ \hline
+Receiver Same Issue Area 7             & 0.391                            & 0.312                               & 0.480                               & *                                    \\ \hline
+Receiver Same Issue Area 8             & 0.210                            & 0.146                               & 0.274                               & *                                    \\ \hline
+Receiver Same Issue Area 9             & 0.230                            & 0.161                               & 0.315                               & *                                    \\ \hline
+Receiver Same Issue Area 10            & 0.495                            & 0.393                               & 0.569                               & *                                    \\ \hline
+Receiver Same Issue Area 11            & 0.196                            & -0.033                              & 0.389                               &                                      \\ \hline
+Receiver Same Issue Area 12            & 0.177                            & 0.026                               & 0.348                               & *                                    \\ \hline
+Receiver Same Issue Area 13            & 0.896                            & 0.555                               & 1.194                               & *                                    \\ \hline
+Receiver Same Issue Area 14            & -0.137                           & -0.645                              & 0.233                               &                                      \\ \hline
 \end{tabular}
-\caption{Bootstrappped MPLE Results for the time period 1937-2001}
+\caption{Bootstrapped MPLE Results for the time period 1937-2001}
 \label{bootMPLE_total}
@@
 \begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+>{\columncolor[HTML]{EFEFEF}}l |l|l|l|l|}
 \hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -5.391                                                 & -5.506                                                     & -5.244                                                     \\ \hline
-Instar(2)               & 0.035                                                   & 0.028                                                      & 0.042                                                      \\ \hline
-Outstar(2)              & 0.015                                                   & 0.0003                                                      & 0.017                                                      \\ \hline
-Mutual                  & 4.785                                                   & 3.616                                                 & 5.749                                                      \\ \hline
-Triangle                & 1.377                                                   & 1.271                                                      & 1.513                                                      \\ \hline
-MQ Score                & -0.018                                                   & -0.095                                                      & 0.061                                                      \\ \hline
-Same Issue Area         & 1.229                                                   & 1.126                                                      & 1.327                                                      \\ \hline
-Year Difference         & -0.077                                                   & -0.096                                                     & -0.061                                                          \\ \hline
+                                            & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}Lower Bound & \cellcolor[HTML]{EFEFEF}Upper Bound & \cellcolor[HTML]{EFEFEF}Significance \\ \hline
+Edges                                       & -4.524                           & -4.736                              & -4.331                              & *                                    \\ \hline
+Instar(2)                                   & 0.00041                          & 0.00038                             & 0.00045                             & *                                    \\ \hline
+Outstar(2)                                  & 0.00012                          & 0.00007                             & 0.00017                             & *                                    \\ \hline
+Mutual                                      & 0.060                            & 0.034                               & 0.089                               & *                                    \\ \hline
+Triangle                                    & 0.033                            & 0.032                               & 0.035                               & *                                    \\ \hline
+Martin Quinn Score                          & -0.0001                          & -0.0007                             & 0.0006                              &                                      \\ \hline
+Same Issue Area Homophily                   & 0.025                            & 0.024                               & 0.026                               & *                                    \\ \hline
+Year Difference                             & -0.001                           & -0.0012                             & -0.0009                             & *                                    \\ \hline
+$(\text{Year Differnce})^2$                 & 4.108e-07                        & 3.413e-07                           & 4.909e-07                           & *                                    \\ \hline
+{\small Receiver Abs Diff MQ Score in Majority} & 0.036                            & 0.027                               & 0.047                               & *                                    \\ \hline
+{\small Receiver Number Justices in Majority}        & -0.004                           & -0.018                              & 0.011                               &                                      \\ \hline
+Receiver  Sender Year                       & -0.040                           & -0.0002                             & 0.0001                              &                                      \\ \hline
+Sender Same Issue Area 2                    & 0.172                            & 0.098                               & 0.249                               & *                                    \\ \hline
+Sender Same Issue Area 3                    & -0.187                           & -0.301                              & -0.067                              & *                                    \\ \hline
+Sender Same Issue Area 4                    & 0.535                            & 0.424                               & 0.625                               & *                                    \\ \hline
+Sender Same Issue Area 5                    & 0.662                            & 0.439                               & 0.860                               & *                                    \\ \hline
+Sender Same Issue Area 6                    & 0.656                            & 0.468                               & 0.831                               & *                                    \\ \hline
+Sender Same Issue Area 7                    & 0.478                            & 0.353                               & 0.598                               & *                                    \\ \hline
+Sender Same Issue Area 8                    & 0.255                            & 0.176                               & 0.336                               & *                                    \\ \hline
+Sender Same Issue Area 9                    & 0.385                            & 0.301                               & 0.469                               & *                                    \\ \hline
+Sender Same Issue Area 10                   & 0.431                            & 0.295                               & 0.548                               & *                                    \\ \hline
+Sender Same Issue Area 11                   & 0.165                            & -0.204                              & 0.444                               &                                      \\ \hline
+Sender Same Issue Area 12                   & 0.168                            & 0.016                               & 0.302                               & *                                    \\ \hline
+Sender Same Issue Area 13                   & 0.694                            & 0.162                               & 1.102                               & *                                    \\ \hline
+Sender Same Issue Area 14                   & -0.054                           & -0.525                              & 0.325                               &                                      \\ \hline
+Receiver Same Issue Area 2                  & 0.212                            & 0.122                               & 0.311                               & *                                    \\ \hline
+Receiver Same Issue Area 3                  & -0.140                           & -0.291                              & 0.017                               &                                      \\ \hline
+Receiver Same Issue Area 4                  & 0.538                            & 0.375                               & 0.699                               & *                                    \\ \hline
+Receiver Same Issue Area 5                  & 0.426                            & 0.189                               & 0.653                               & *                                    \\ \hline
+Receiver Same Issue Area 6                  & 0.512                            & 0.197                               & 0.777                               & *                                    \\ \hline
+Receiver Same Issue Area 7                  & 0.510                            & 0.385                               & 0.649                               & *                                    \\ \hline
+Receiver Same Issue Area 8                  & 0.372                            & 0.283                               & 0.468                               & *                                    \\ \hline
+Receiver Same Issue Area 9                  & 0.243                            & 0.134                               & 0.347                               & *                                    \\ \hline
+Receiver Same Issue Area 10                 & 0.555                            & 0.388                               & 0.709                               & *                                    \\ \hline
+Receiver Same Issue Area 11                 & -0.318                           & -0.847                              & 0.094                               &                                      \\ \hline
+Receiver Same Issue Area 12                 & 0.626                            & 0.478                               & 0.798                               & *                                    \\ \hline
+Receiver Same Issue Area 13                 & 1.342                            & 0.561                               & 1.696                               & *                                    \\ \hline
+Receiver Same Issue Area 14                 & NA                               & NA                                  & NA                                  &                                      \\ \hline
 \end{tabular}
-\caption{Bootstrappped MPLE Results for the time period when Fred M. Vinson was chief justice.}
-\label{bootMPLE_vinson}
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
-\hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -5.032                                                 & -5.140                                                     & -4.931                                                     \\ \hline
-Instar(2)               & 0.028                                                   & 0.025                                                      & 0.032                                                      \\ \hline
-Outstar(2)              & 0.018                                                   & 0.015                                                      & 0.022                                                      \\ \hline
-Mutual                  & 3.166                                                   & 1.826                                                 & 4.565                                                      \\ \hline
-Triangle                & 1.520                                                   & 1.449                                                      & 1.599                                                      \\ \hline
-MQ Score                & -0.101                                                   & -0.149                                                      & -0.052                                                      \\ \hline
-Same Issue Area         & 1.349                                                   & 1.289                                                      & 1.408                                                      \\ \hline
-Year Difference         & -0.031                                                   & -0.038                                                     & -0.026                                                          \\ \hline
-\end{tabular}
-\caption{Bootstrappped MPLE Results for the time period when Earl Warren was chief justice.}
-\label{bootMPLE_warren}
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
-\hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -3.952                                                 & -4.017                                                     & -3.882                                                     \\ \hline
-Instar(2)               & 0.015                                                   & 0.013                                                      & 0.016                                                      \\ \hline
-Outstar(2)              & 0.004                                                   & 0.002                                                      & 0.006                                                      \\ \hline
-Mutual                  & 2.558                                                   & 1.425                                                 & 3.719                                                      \\ \hline
-Triangle                & 1.383                                                   & 1.333                                                      & 1.440                                                      \\ \hline
-MQ Score                & -0.084                                                   & -0.101                                                      & -0.059                                                      \\ \hline
-Same Issue Area         & 0.959                                                   & 0.911                                                      & 1.008                                                      \\ \hline
-Year Difference         & -0.017                                                   & -0.019                                                     & -0.015                                                          \\ \hline
-\end{tabular}
-\caption{Bootstrappped MPLE Results for the time period when Warren E. Burger was chief justice.}
+\caption{Bootstrapped MPLE Results for the time period when Warren E. Burger was chief justice.}
 \label{bootMPLE_burger}
@@
 \begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+>{\columncolor[HTML]{EFEFEF}}l |l|l|l|l|}
 \hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -3.614                                                 & -3.701                                                     & -3.525                                                     \\ \hline
-Instar(2)               & 0.015                                                   & 0.013                                                      & 0.016                                                      \\ \hline
-Outstar(2)              & 0.002                                                   & -0.002                                                      & 0.004                                                      \\ \hline
-Mutual                  & 9.635                                                  & 6.426                                                 & 10.834                                                      \\ \hline
-Triangle                & 1.551                                                   & 1.489                                                      & 1.620                                                      \\ \hline
-MQ Score                & -0.007                                                   & -0.039                                                      & 0.028                                                      \\ \hline
-Same Issue Area         & 0.855                                                   & 0.794                                                      & 0.916                                                      \\ \hline
-Year Difference         & -0.016                                                   & -0.018                                                     & -0.014                                                          \\ \hline
+                                       & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}Lower Bound & \cellcolor[HTML]{EFEFEF}Upper Bound & \cellcolor[HTML]{EFEFEF}Significance \\ \hline
+Edges                                  & -4.271                           & -4.735                              & -3.779                              & *                                    \\ \hline
+Instar(2)                              & 0.00034                          & 0.00030                             & 0.00037                             & *                                    \\ \hline
+Outstar(2)                             & 0.0001                           & 0.00004                             & 0.00015                             & *                                    \\ \hline
+Mutual                                 & 0.200                            & 0.123                               & 0.231                               & *                                    \\ \hline
+Triangle                               & 0.028                            & 0.027                               & 0.029                               & *                                    \\ \hline
+Martin Quinn Score                     & -0.0002                          & -0.0008                             & 0.0006                              &                                      \\ \hline
+Same Issue Area Homophily              & 0.0018                           & 0.016                               & 0.019                               & *                                    \\ \hline
+Year Difference                        & -0.0007                          & -0.0008                             & -0.0006                             & *                                    \\ \hline
+$(\text{Year Differnce})^2$            & 1.326e-07                        & 1.003e-07                           & 1.677e-07                           & *                                    \\ \hline
+{\small Receiver Abs Diff MQ Score in Majority} & -0.0003                          & -0.012                              & 0.013                               &                                     \\ \hline
+{\small Receiver Number Justices in Majority}   & 0.025                            & 0.006                               & 0.041                               & *                                     \\ \hline
+Receiver  Sender Year                  & 0.00001                          & -0.0002                             & 0.0002                              &                                      \\ \hline
+Sender Same Issue Area 2               & 0.202                            & 0.101                               & 0.302                               & *                                    \\ \hline
+Sender Same Issue Area 3               & -0.304                           & -0.449                              & -0.159                              & *                                    \\ \hline
+Sender Same Issue Area 4               & 0.443                            & 0.310                               & 0.586                               & *                                    \\ \hline
+Sender Same Issue Area 5               & 0.141                            & -0.104                              & 0.378                               &                                     \\ \hline
+Sender Same Issue Area 6               & 0.774                            & 0.568                               & 0.962                               & *                                    \\ \hline
+Sender Same Issue Area 7               & 0.604                            & 0.477                               & 0.732                               & *                                    \\ \hline
+Sender Same Issue Area 8               & 0.467                            & 0.384                               & 0.564                               & *                                    \\ \hline
+Sender Same Issue Area 9               & 0.390                            & 0.292                               & 0.495                               & *                                    \\ \hline
+Sender Same Issue Area 10              & 0.450                            & 0.316                               & 0.591                               & *                                    \\ \hline
+Sender Same Issue Area 11              & 0.062                            & -0.341                              & 0.376                               &                                      \\ \hline
+Sender Same Issue Area 12              & 0.512                            & 0.353                               & 0.674                               & *                                    \\ \hline
+Sender Same Issue Area 13              & 0.602                            & 0.238                               & 0.915                               & *                                    \\ \hline
+Sender Same Issue Area 14              & 1.442                            & 0.761                               & 1.908                               &  *                                    \\ \hline
+Receiver Same Issue Area 2             & 0.171                            & 0.040                               & 0.317                               & *                                    \\ \hline
+Receiver Same Issue Area 3             & -0.127                           & -0.310                              & 0.049                               &                                      \\ \hline
+Receiver Same Issue Area 4             & 0.326                            & 0.143                               & 0.480                               & *                                    \\ \hline
+Receiver Same Issue Area 5             & 0.338                            & 0.037                               & 0.613                               & *                                    \\ \hline
+Receiver Same Issue Area 6             & 0.371                            & 0.117                               & 0.596                               & *                                    \\ \hline
+Receiver Same Issue Area 7             & 0.305                            & 0.137                               & 0.453                               & *                                    \\ \hline
+Receiver Same Issue Area 8             & 0.284                            & 0.166                               & 0.405                               & *                                    \\ \hline
+Receiver Same Issue Area 9             & 0.398                            & 0.271                               & 0.523                               & *                                    \\ \hline
+Receiver Same Issue Area 10            & 0.419                            & 0.264                               & 0.591                               & *                                    \\ \hline
+Receiver Same Issue Area 11            & 0.670                            & 0.395                               & 0.914                               & *                                    \\ \hline
+Receiver Same Issue Area 12            & 0.490                            & 0.287                               & 0.687                               & *                                    \\ \hline
+Receiver Same Issue Area 13            & 1.045                            & 0.816                               & 1.247                               & *                                    \\ \hline
+Receiver Same Issue Area 14            & NA                               & NA                                  & NA                                  &                                      \\ \hline
 \end{tabular}
-\caption{Bootstrappped MPLE Results for the time period when William Rehnquist was chief justice.}
+\caption{Bootstrapped MPLE Results for the time period when William Rehnquist was chief justice.}
 \label{bootMPLE_rehnquist}

### 2017-12-12 Tex/paper.tex
updating theory
New inferential results

--- 2017-12-02 Tex/paper.tex
@@
 \begin{table}[H]
+\footnotesize
 \centering
 \begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
+>{\columncolor[HTML]{EFEFEF}}l |l|l|l|l|} 
 \hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -5.319                                                 & -5.392                                                     & -5.243                                                     \\ \hline
-Instar(2)               & 0.018                                                   & 0.017                                                      & 0.019                                                      \\ \hline
-Outstar(2)              & 0.018                                                  & 0.016                                                      & 0.022                                                      \\ \hline
-Mutual                  & 3.179                                                   & 2.694                                                 & 3.565                                                      \\ \hline
-Triangle                & 1.531                                                   & 1.493                                                      & 1.563                                                      \\ \hline
-MQ Score                & 0.041                                                   & 0.018                                                      & 0.063                                                      \\ \hline
-Same Issue Area         & 1.061                                                   & 1.028                                                      & 1.099                                                      \\ \hline
-Year Difference         & 0.010                                                   & 0.008                                                     & 0.013                                                          \\ \hline
+                                                   & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}Lower Bound & \cellcolor[HTML]{EFEFEF}Upper Bound & \cellcolor[HTML]{EFEFEF}Significance \\
+                                                    \hline
+Edges                                              & -5.533                           & -5.733                              & -5.388                              & *                                    \\ \hline
+Instar(2)                                          & 0.031                            & 0.027                               & 0.036                               & *                                    \\ \hline
+Outstar(2)                                         & 0.022                            & 0.020                               & 0.032                               & *                                    \\ \hline
+Mutual                                             & 3.316                            & 2.622                               & 3.983                               & *                                    \\ \hline
+Triangle                                           & 1.490                            & 1.410                               & 1.560                               & *                                    \\ \hline
+Martin Quinn Score                                 & 0.080                            & 0.022                               & 0.126                               & *                                    \\ \hline
+Same Issue Area Homophily                          & 1.378                            & 1.313                               & 1.451                               & *                                    \\ \hline
+Year Difference                                    & -0.077                           & -0.090                              & -0.064                              & *                                    \\ \hline
+$(\text{Year Differnce})^2$                        & 0.0029                           & 0.0025                              & 0.0032                              & *                                    \\ \hline
+Receiver Abs Diff MQ Score in Majority             & -0.030                           & -0.048                              & -0.014                              & *                                    \\ \hline
+Receiver Number Justices in Majority               & -0.115                           & -0.135                              & -0.089                              & *                                    \\ \hline
+Receiver  Sender Year                              & 0.0007                           & 0.0006                              & 0.0009                              & *                                    \\ \hline
+Sender Same Issue Area 2                           & 0.162                            & 0.105                               & 0.217                               & *                                    \\ \hline
+Sender Same Issue Area 3                           & -0.317                           & -0.403                              & -0.236                              & *                                    \\ \hline
+Sender Same Issue Area 4                           & 0.497                            & 0.426                               & 0.570                               & *                                    \\ \hline
+Sender Same Issue Area 5                           & 0.332                            & 0.174                               & 0.495                               & *                                    \\ \hline
+Sender Same Issue Area 6                           & 0.527                            & 0.402                               & 0.635                               & *                                    \\ \hline
+Sender Same Issue Area 7                           & 0.327                            & 0.251                               & 0.394                               & *                                    \\ \hline
+Sender Same Issue Area 8                           & 0.076                            & 0.021                               & 0.129                               & *                                    \\ \hline
+Sender Same Issue Area 9                           & 0.261                            & 0.209                               & 0.315                               & *                                    \\ \hline
+Sender Same Issue Area 10                          & 0.369                            & 0.301                               & 0.432                               & *                                    \\ \hline
+Sender Same Issue Area 11                          & -0.046                           & -0.233                              & 0.151                               &                                      \\ \hline
+Sender Same Issue Area 12                          & 0.0073                           & -0.075                              & 0.089                               &                                      \\ \hline
+Sender Same Issue Area 13                          & 0.381                            & 0.162                               & 0.564                               & *                                    \\ \hline
+Sender Same Issue Area 14                          & 0.151                            & -0.019                              & 0.300                               &                                      \\ \hline
+Receiver Same Issue Area 2                         & 0.231                            & 0.166                               & 0.300                               & *                                    \\ \hline
+Receiver Same Issue Area 3                         & -0.101                           & -0.209                              & -0.006                              & *                                    \\ \hline
+Receiver Same Issue Area 4                         & 0.516                            & 0.425                               & 0.612                               & *                                    \\ \hline
+Receiver Same Issue Area 5                         & 0.351                            & 0.183                               & 0.534                               & *                                    \\ \hline
+Receiver Same Issue Area 6                         & 0.461                            & 0.284                               & 0.628                               & *                                    \\ \hline
+Receiver Same Issue Area 7                         & 0.361                            & 0.281                               & 0.437                               & *                                    \\ \hline
+Receiver Same Issue Area 8                         & 0.153                            & 0.097                               & 0.221                               & *                                    \\ \hline
+Receiver Same Issue Area 9                         & 0.279                            & 0.221                               & 0.346                               & *                                    \\ \hline
+Receiver Same Issue Area 10                        & 0.492                            & 0.413                               & 0.570                               & *                                    \\ \hline
+Receiver Same Issue Area 11                        & 0.247                            & 0.056                               & 0.411                               & *                                    \\ \hline
+Receiver Same Issue Area 12                        & 0.339                            & 0.246                               & 0.430                               & *                                    \\ \hline
+Receiver Same Issue Area 13                        & 1.060                            & 0.845                               & 1.239                               & *                                    \\ \hline
+Receiver Same Issue Area 14                        & 0.557                            & 0.255                               & 0.793                               & *                                    \\ \hline
+Instar(2) $\times$ Sender Year                     & -0.00025                         & -0.00034                            & -0.00017                            & *                                    \\ \hline
+Outstar(2) $\times$ Sender Year                    & -0.00036                         & -0.00056                            & -0.00028                            & *                                    \\ \hline
+Mutual $\times$ Sender Year                        & -0.010                           & -0.043                              & 0.029                               &                                      \\ \hline
+Triangle $\times$ Sender Year                      & -0.0004                          & -0.0021                             & 0.0016                              &                                      \\ \hline
+Martin Quinn Score $\times$ Sender Year            & -0.0025                          & -0.0035                             & -0.0013                             & *                                    \\ \hline
+Same Issue Area $\times$ Sender Year               & -0.0058                          & -0.0076                             & -0.0043                             & *                                    \\ \hline
+Year Difference $\times$ Sender Year               & 0.0004                           & 0.0002                              & 0.0007                              & *                                    \\ \hline
+$\text{Year Difference}^2 \times$ Sender Year      & -0.000037                        & -0.000044                           & -0.000030                           & *                                    \\ \hline
+MQ Score in Majority $\times$ Sender Year & 0.0013                           & 0.0009                              & 0.0017                              & *                                    \\ \hline
+Justices in Majority $\times$ Sender Year   & 0.0029                           & 0.0023                              & 0.0035                              & *                                    \\ \hline
 \end{tabular}
-\caption{Bootstrappped MPLE Results for the time period 1937-2001}
-\label{bootMPLE_total}
+\caption{Bootstrapped MPLE Results for the time period $1937-2001$.
+A '*' indicates that the 2.5th and 97.5th quantile of the variable does not include '0' and as a result is statistically significant.}
+\label{my-label}
 \end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
-\hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -5.391                                                 & -5.506                                                     & -5.244                                                     \\ \hline
-Instar(2)               & 0.035                                                   & 0.028                                                      & 0.042                                                      \\ \hline
-Outstar(2)              & 0.015                                                   & 0.0003                                                      & 0.017                                                      \\ \hline
-Mutual                  & 4.785                                                   & 3.616                                                 & 5.749                                                      \\ \hline
-Triangle                & 1.377                                                   & 1.271                                                      & 1.513                                                      \\ \hline
-MQ Score                & -0.018                                                   & -0.095                                                      & 0.061                                                      \\ \hline
-Same Issue Area         & 1.229                                                   & 1.126                                                      & 1.327                                                      \\ \hline
-Year Difference         & -0.077                                                   & -0.096                                                     & -0.061                                                          \\ \hline
-\end{tabular}
-\caption{Bootstrappped MPLE Results for the time period when Fred M. Vinson was chief justice.}
-\label{bootMPLE_vinson}
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
-\hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -5.032                                                 & -5.140                                                     & -4.931                                                     \\ \hline
-Instar(2)               & 0.028                                                   & 0.025                                                      & 0.032                                                      \\ \hline
-Outstar(2)              & 0.018                                                   & 0.015                                                      & 0.022                                                      \\ \hline
-Mutual                  & 3.166                                                   & 1.826                                                 & 4.565                                                      \\ \hline
-Triangle                & 1.520                                                   & 1.449                                                      & 1.599                                                      \\ \hline
-MQ Score                & -0.101                                                   & -0.149                                                      & -0.052                                                      \\ \hline
-Same Issue Area         & 1.349                                                   & 1.289                                                      & 1.408                                                      \\ \hline
-Year Difference         & -0.031                                                   & -0.038                                                     & -0.026                                                          \\ \hline
-\end{tabular}
-\caption{Bootstrappped MPLE Results for the time period when Earl Warren was chief justice.}
-\label{bootMPLE_warren}
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
-\hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -3.952                                                 & -4.017                                                     & -3.882                                                     \\ \hline
-Instar(2)               & 0.015                                                   & 0.013                                                      & 0.016                                                      \\ \hline
-Outstar(2)              & 0.004                                                   & 0.002                                                      & 0.006                                                      \\ \hline
-Mutual                  & 2.558                                                   & 1.425                                                 & 3.719                                                      \\ \hline
-Triangle                & 1.383                                                   & 1.333                                                      & 1.440                                                      \\ \hline
-MQ Score                & -0.084                                                   & -0.101                                                      & -0.059                                                      \\ \hline
-Same Issue Area         & 0.959                                                   & 0.911                                                      & 1.008                                                      \\ \hline
-Year Difference         & -0.017                                                   & -0.019                                                     & -0.015                                                          \\ \hline
-\end{tabular}
-\caption{Bootstrappped MPLE Results for the time period when Warren E. Burger was chief justice.}
-\label{bootMPLE_burger}
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
-\hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Estimate} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Lower Bound} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Upper Bound} \\ \hline
-Edges                   & -3.614                                                 & -3.701                                                     & -3.525                                                     \\ \hline
-Instar(2)               & 0.015                                                   & 0.013                                                      & 0.016                                                      \\ \hline
-Outstar(2)              & 0.002                                                   & -0.002                                                      & 0.004                                                      \\ \hline
-Mutual                  & 9.635                                                  & 6.426                                                 & 10.834                                                      \\ \hline
-Triangle                & 1.551                                                   & 1.489                                                      & 1.620                                                      \\ \hline
-MQ Score                & -0.007                                                   & -0.039                                                      & 0.028                                                      \\ \hline
-Same Issue Area         & 0.855                                                   & 0.794                                                      & 0.916                                                      \\ \hline
-Year Difference         & -0.016                                                   & -0.018                                                     & -0.014                                                          \\ \hline
-\end{tabular}
-\caption{Bootstrappped MPLE Results for the time period when William Rehnquist was chief justice.}
-\label{bootMPLE_rehnquist}
-\end{table}
-
 \newpage

### 2017-12-14 R-Code/bootstrap_coefficient_summaries.R
example plot code

@@ -1,0 +1,38 @@
+# code assumes that working directory is set
+# to top-level project
+bootstrapResults <- read.csv("./R-Code/bootresults.csv",stringsAsFactors=F,row.names=1)
+
+# Christian, can you correct this? The sender time values should be set to map onto the range
+# of sender times in our data
+sender_times <- 1:60
+sender_years <- 1941:2000
+
+density_coefficients <- c("edges","nodeocov.sender.time")
+i2star_coefficients <- c("istar2","instar2_sendertime")
+o2star_coefficients <- c("ostar2","outstar2_sendertime")
+mutual_coefficients <- c("mutual","mutual_sendertime")
+triangle_coefficients <- c("triangle","triangle_sendertime")
+mq_coefficients <- c("edgecov.mq.t","mq_sendertime")
+sameissue_coefficients <- c("edgecov.same.issue.area.t","sameissuearea_sendertime")
+absdiffmq_coefficients <- c("nodeicov.AbsDiffMQscores","AbsDiffMQscores_sendertime")
+numberjusticespro_coefficients <- c("nodeicov.NumberJusticesPro","NumberJusticesPro_sendertime")
+
+# example trend CI plot for MQ coef
+mq_coef_trend <- matrix(NA,length(sender_times),3)
+colnames(mq_coef_trend) <- c("mean","lower","upper")
+for(t in sender_times){
+  effect.t <- bootstrapResults[,mq_coefficients[1]]+t*bootstrapResults[,mq_coefficients[2]]
+  mq_coef_trend[t,] <- c(mean(effect.t),quantile(effect.t,prob=c(0.025,.975)))
+}
+
+pdf("./Tex/images/mq_coef_trend.pdf",width=6,height=3.5)
+par(las=1,mar=c(4,4,1,1))
+plot(sender_years,mq_coef_trend[,1],type="n",ylab="effect",xlab="year",ylim=c(min(c(mq_coef_trend)),max(c(mq_coef_trend))))
+polygon(c(sender_years,rev(sender_years)),c(mq_coef_trend[,2],rev(mq_coef_trend[,3])),col="lightblue",border=NA)
+lines(sender_years,mq_coef_trend[,1],lwd=2,col="grey50")
+abline(h=0,lty=2,lwd=2)
+dev.off()
+
+
+
+

### 2017-12-15 R-Code/bootstrap_coefficient_summaries.R
bootstrap_coefficient_summaries.R updated

@@ -1,11 +1,11 @@
 # code assumes that working directory is set
 # to top-level project
-bootstrapResults <- read.csv("./R-Code/bootresults.csv",stringsAsFactors=F,row.names=1)
+bootstrapResults <- read.csv("bootresults.csv",stringsAsFactors=F,row.names=1)
 
 # Christian, can you correct this? The sender time values should be set to map onto the range
 # of sender times in our data
-sender_times <- 1:60
-sender_years <- 1941:2000
+sender_times <- 1:65
+sender_years <- 1937:2001
 
 density_coefficients <- c("edges","nodeocov.sender.time")
 i2star_coefficients <- c("istar2","instar2_sendertime")
@@ -25,14 +25,146 @@ for(t in sender_times){
   mq_coef_trend[t,] <- c(mean(effect.t),quantile(effect.t,prob=c(0.025,.975)))
 }
 
-pdf("./Tex/images/mq_coef_trend.pdf",width=6,height=3.5)
+pdf("./images/mq_coef_trend.pdf",width=6,height=3.5)
 par(las=1,mar=c(4,4,1,1))
-plot(sender_years,mq_coef_trend[,1],type="n",ylab="effect",xlab="year",ylim=c(min(c(mq_coef_trend)),max(c(mq_coef_trend))))
+plot(sender_years,mq_coef_trend[,1],type="n",ylab="Effect",xlab="Year", main="Martin-Quinn Score" ,ylim=c(min(c(mq_coef_trend)),max(c(mq_coef_trend))))
 polygon(c(sender_years,rev(sender_years)),c(mq_coef_trend[,2],rev(mq_coef_trend[,3])),col="lightblue",border=NA)
 lines(sender_years,mq_coef_trend[,1],lwd=2,col="grey50")
 abline(h=0,lty=2,lwd=2)
 dev.off()
 
 
+# trend CI plot for edges coef
+edges_coef_trend <- matrix(NA,length(sender_times),3)
+colnames(edges_coef_trend) <- c("mean","lower","upper")
+for(t in sender_times){
+  effect.t <- bootstrapResults[,density_coefficients[1]]+t*bootstrapResults[,density_coefficients[2]]
+  edges_coef_trend[t,] <- c(mean(effect.t),quantile(effect.t,prob=c(0.025,.975)))
+}
+
[…]

### 2017-12-22 Tex/paper.tex
reviewing past work on citaitons

@@
 
+We are not the first to study what predicts citations on the Supreme Court...
+
+We are the first to examine at the level of the citation, and, as a result, be able to incorporate complex dependence into our explanation of citations.
+
 \section{Prior work on court citations}
@@
 Measures of citations have been used in several recent studies of courts...\citet{hinkle2016transmission} study the citations between US state supreme courts.
+
+\begin{itemize}
+
+\item \citet{clark2010locating} Estimates a latent coordinate model of supreme court opinions based on the network of supreme court citations.
+Find that the majority opinion falls at the ideal point of the median member of the majority coalition.
+Account for negative citations.
+WE COULD USE THIS DATASET AS A SECOND APPLICATION
+
+\item \citet{cross2010determinants} Look at the total number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
+(case level).
+Found case issue area, opinion ideology, opinion author, vote margin was not significant.
+Negative citations decreased future positive citations.
+
+\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
+Show that the rate of citations to a case goes down very quickly.
+THE EFFECT OF AN INDEPENDENT VARIABLE ON A PRECEDENT IS DEPENDENT UPON THE AGE OF THE CASE.
+
+\item \citet{benjamin2012standing} Study the propensity for cases to be overruled and cited in other negative ways.
+They find that cases with large and broad majorities rare less likely to be cited negatively.
+
+\item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
+
+\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
+Lower if stautory cases.
+Increases with ideological variance of the majority coalition.
+Increases with the number of concurring opinions.
+
+\end{itemize}
 

### 2017-12-22 Tex/paper.tex
added coefficient plots to ms

@@
 \usepackage{verbatim}
-\usepackage[a4paper,text={140mm,215mm},centering,headsep=10mm,footskip=10mm]{geometry}
+\usepackage[a4paper,text={160mm,215mm},centering,headsep=10mm,footskip=10mm]{geometry}
 \usepackage[urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black,bookmarks]{hyperref}
@@
 
-We are not the first to study what predicts citations on the Supreme Court...
+We are not the first to ask what predicts the citations in US Supreme Court Opinions.
+Indeed, a voluminous body of work has sought to explain how many times an opinion is cited \citep[e.g.,][]{cross2010determinants,benjamin2012standing}, when in a lifecycle an opinion is cited \citep[e.g.,][]{black2013citation,spriggs2001explaining}, and how many cases are cited by an opinion \citep[e.g.,][]{lupu2013strategic}---all focused on the US Supreme Court.
+One common feature of the research design in all of these studies is that the observations are at the case or case-time level.
+The outcome variables in these analyses are defined as measures of the number of citations to a case over a period of time, the number of citations to a case at a particular time, or a measurement on the cases cited by a case.
+None of these studies treats citation in its micro-level form---as a relationship between two opinions, the citing opinion and the cited opinion.
+We are aware of one prior study, \citet{clark2010locating}, in which a statistical model is used to analyze dyadic citations between cases.
+However, citet{clark2010locating} uses a dyadic latent variable model in order to estimate ideal points for Supreme Court Opinions, but does not use any explanatory variables to predict the formation of citation ties between opinions.
+We build upon this literature both methodologically and substantively.
+Methodologically, we develop a novel statistical framework for modeling directed dyadic citations.
+Second, we apply this methodology to a half-century of directed dyadic citations between US Supreme Court opinions.
 
-We are the first to examine at the level of the citation, and, as a result, be able to incorporate complex dependence into our explanation of citations.
+There exist two broad reasons why empirical analyses of citations are best defined on the directed dyad level, not the case level.
+The first is that directed dyadic analyses can test both dyadic and case-level hypotheses.
+For example, case-level analyses can model whether opinions supported by a liberal majority coalition are more likely than those supported by a conservative majority coalition to be cited heavily in the future, but they cannot precisely model whether liberal cases will be cited more by liberal cases than by conservative cases.
+Thus, the first reason for analyzing citations at the dyadic level it to expand the set of hypotheses that can be represented in the model.
+The second reason for studying citations at the directed dyadic level is that, as articulated in the growing literature on legal citation networks, citations form complex networks in which a citation at one point in time may influence future citations.
+This phenomenon of complex dependence is very common in networks of many types, but processes specific to Supreme Court citations create interdependence in citations.
+For example, if opinion $i$ relies heavily on opinion $j$ as precedent, opinion $i$ is likely to discuss the legal basis for opinion $j$, and as a consequence cite some of the opinions cited by opinion $j$.
+Suppose opinion $k$ is cited by opinion $j$.
+Opinion $k$ is more likely to be cited by opinion $i$ because opinion $i$ relies heavily on $j$, and opinion $j$ cites $k$.
+This is a special case of a very common process on networks referred to as ``triad closure''.
+Complex dependence is theoretically interesting on its own merits, but the effects of covariates cannot be reliably identified---either in terms of coefficient values or standard errors---without accounting for the interdependence inherent in networks \citep{desmaraisstatistical}.
 
-\section{Prior work on court citations}
+In what follows we develop a theoretical case that citations on the US Supreme Court are characterized by forms of complex dependence that are common in networks.
+We then develop an extension of a model---the exponential random graph model---that can incorporate both exogenous covariates and complex forms of interdependence into a directed dyadic analysis of citations.
+Finally, we develop and estimate a specification of this model in an analysis of US Supreme Court citations between 1937 and 2001.
+We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's predictive performance.
 
-Measures of citations have been used in several recent studies of courts...\citet{hinkle2016transmission} study the citations between US state supreme courts.
+\section{Complex Interdependence in Supreme Court Citations}
 
-\begin{itemize}
-
-\item \citet{clark2010locating} Estimates a latent coordinate model of supreme court opinions based on the network of supreme court citations.
-Find that the majority opinion falls at the ideal point of the median member of the majority coalition.
-Account for negative citations.
-WE COULD USE THIS DATASET AS A SECOND APPLICATION
-
-\item \citet{cross2010determinants} Look at the total number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
-(case level).
-Found case issue area, opinion ideology, opinion author, vote margin was not significant.
-Negative citations decreased future positive citations.
-
-\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
-Show that the rate of citations to a case goes down very quickly.
-THE EFFECT OF AN INDEPENDENT VARIABLE ON A PRECEDENT IS DEPENDENT UPON THE AGE OF THE CASE.
-
-\item \citet{benjamin2012standing} Study the propensity for cases to be overruled and cited in other negative ways.
-They find that cases with large and broad majorities rare less likely to be cited negatively.
-
-\item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
-
-\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
-Lower if stautory cases.
-Increases with ideological variance of the majority coalition.
-Increases with the number of concurring opinions.
-
-\end{itemize}
-
-\section{Network Approaches to Studying Citations}
-
-\textbf{TED}, would you write a section that covers the ways in which network scholars have studied citations in the past, focusing on scientific papers, patents, and court cases?
-
-\section{Complex Dependence in Court Citations}
+When it comes to the development and testing of theory, the defining feature of networks is that the fundamental element under study---the relationship between two units (i.e., the citation from one opinion to another) is a piece of a complex web of relations that cannot be fully understood without considering how the singular relationship fits into that complex web.
 
@@
 https://supreme.justia.com/cases/federal/us/460/719/
+
+\section{Network Approaches to Studying Citations}
+
+\textbf{TED}, would you write a section that covers the ways in which network scholars have studied citations in the past, focusing on scientific papers, patents, and court cases?
 
@@
 
+\section{Prior work on court citations}
+
+Measures of citations have been used in several recent studies of courts...\citet{hinkle2016transmission} study the citations between US state supreme courts.
+
+\begin{itemize}
+
+\item \citet{clark2010locating} Estimates a latent coordinate model of supreme court opinions based on the network of supreme court citations.
+Find that the majority opinion falls at the ideal point of the median member of the majority coalition.
+Account for negative citations.
+WE COULD USE THIS DATASET AS A SECOND APPLICATION
+
+\item \citet{cross2010determinants} Look at the total number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
+(case level).
+Found case issue area, opinion ideology, opinion author, vote margin was not significant.
+Negative citations decreased future positive citations.
+
+\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
+Show that the rate of citations to a case goes down very quickly.
+THE EFFECT OF AN INDEPENDENT VARIABLE ON A PRECEDENT IS DEPENDENT UPON THE AGE OF THE CASE.
+
+\item \citet{benjamin2012standing} Study the propensity for cases to be overruled and cited in other negative ways.
+They find that cases with large and broad majorities rare less likely to be cited negatively.
+
+\item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
+
+\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
+Lower if stautory cases.
+Increases with ideological variance of the majority coalition.
+Increases with the number of concurring opinions.
+
+\end{itemize}
+
  \section{Results}
@@
 
+  \begin{figure}[H]
+  \begin{tabular}{cc}
+ (a) MQ Score Difference & (b) Ideological Breadth \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend.pdf} \\
+
+  (c) Same Issue & (d) Majority Size \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend.pdf} \\
+
+   (e) Reciprocity & (f) Transitivity \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend.pdf} \\
+
+    (g) Sociality & (h) Popularity \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend.pdf} \\
+
+\end{tabular}
+\caption{Effects}
+ \label{fig:coeftrends}
+\vspace{-.25cm}
+\end{figure}
+
 \bibliography{bib} 

### 2017-12-23 Tex/paper.tex
added prediction experiment

@@
 
-When it comes to the development and testing of theory, the defining feature of networks is that the fundamental element under study---the relationship between two units (i.e., the citation from one opinion to another) is a piece of a complex web of relations that cannot be fully understood without considering how the singular relationship fits into that complex web.
+When it comes to the development and testing of theory, the defining feature of networks is that the fundamental element under study---the relationship between two units (i.e., the citation from one opinion to another) is a piece of a complex web of relations.
+The formation (or lack thereof) of that relationship cannot be fully understood without considering how the relationship fits into the complex web.
+Analytical designs that account only for covariates in explaining tie formation
 
@@
 
+\subsection{Predictive Performance}
+
+Our case for studying legal citaions at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence.
+To quantify this contribution, we use out-of-sample prediction [CITES].
+Predicting out-of-sample offers an unbiased and general purpose way to evaluate the contribution, in terms of model fit, of one or more terms/parameters in a model.
+Unlike in-sample measures of model fit, out-of-sample methods are highly robust in avoiding overfitting, and work when we cannot accurately calculate the value of the likelihood function, as in the current case.
+
+In our prediction experiment we randomly split the directed dyads into an 80\% training set and a 20\% test set.
+The parameters of the model are estimated using the directed dyads in the training set, and the parameters are used to form the conditional probability of a tie for all of the directed dyads in the test set.
+Directed dyads for which the conditional probability of a tie exceeds 0.5 are predicted to be citations.
+The experiment is run with the full model, and with a model that excludes all of the dependence terms (i.e., all terms involving reciprocity, in and out stars, and triangles)---the independent dyads model.
+We run this experiment for 10 iterations.
+Predictive performance is evaluated with three common and related measures---precision (i.e., the proportion of predicted citations that are actually citations), recall (the proportion of actual citations that are predicted to be citations), and the F1 score (the harmonic mean of precision and recall).
+All three measures are bounded between 0 and 1, with higher scores indicating better performance.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllll}
+\hline \hline
+& \multicolumn{2}{c}{Independent Model} & \multicolumn{2}{c}{Full Model} \\
+  \hline
+ & mean & range & mean & range \\ 
+  \hline
+precision & 0.5499 & (0.5384, 0.5619) & 0.8605 & (0.8526, 0.8666) \\ 
+  recall & 0.0827 & (0.0811, 0.0843) & 0.5858 & (0.5817, 0.5891) \\ 
+  F1 score & 0.1438 & (0.141, 0.1463) & 0.6971 & (0.6941, 0.7008) \\ 
+   \hline \hline
+\end{tabular}
+\caption{The predictive performance of the directed dyadic models, over ten 80/20 train/test splits.}
+\label{tab:prediction}
+\end{table}
+
+We see from Table \ref{tab:prediction} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms.
+The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set.
+This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.
+
 \bibliography{bib} 

### 2017-12-23 R-Code/holdOutExperiment.R

@@ -1,0 +1,106 @@
+scdata <- read.csv("mpledata_all.csv",stringsAsFactors=F)
+
+set.seed(123)
+
+niter = 10
+predict.results <- matrix(0,niter,6)
+
+for(i in 1:niter){
+
+train.obs <- sample(1:nrow(scdata),round(0.8*nrow(scdata)))
+
+train.data <- scdata[train.obs,]
+
+library(biglm)
+
+# MPLE
+mple.est <- bigglm(edgeij ~ istar2+ostar2+mutual+triangle+ edgecov.mq.t+ edgecov.same.issue.area.t+ edgecov.year.diff.t +edgecov.year.diff.square.t +  
+                nodeicov.AbsDiffMQscores+ nodeicov.NumberJusticesPro+ nodeocov.sender.time + nodeifactor.SameIssueArea.2+ nodeifactor.SameIssueArea.3+ 
+                nodeifactor.SameIssueArea.4+nodeifactor.SameIssueArea.5+ nodeifactor.SameIssueArea.6+ nodeifactor.SameIssueArea.7+ 
+                nodeifactor.SameIssueArea.8+nodeifactor.SameIssueArea.9+  nodeifactor.SameIssueArea.10+
+                nodeifactor.SameIssueArea.11+ nodeifactor.SameIssueArea.12+ nodeifactor.SameIssueArea.13+nodeifactor.SameIssueArea.14+ 
+                nodeofactor.SameIssueArea.2+ nodeofactor.SameIssueArea.3+ nodeofactor.SameIssueArea.4+nodeofactor.SameIssueArea.5+ 
+                nodeofactor.SameIssueArea.6+nodeofactor.SameIssueArea.7+
+                nodeofactor.SameIssueArea.8+nodeofactor.SameIssueArea.9+nodeofactor.SameIssueArea.10+ nodeofactor.SameIssueArea.11+ 
+                nodeofactor.SameIssueArea.12+ nodeofactor.SameIssueArea.13+nodeofactor.SameIssueArea.14+
+                instar2_sendertime+outstar2_sendertime+ mutual_sendertime+ triangle_sendertime+mq_sendertime +sameissuearea_sendertime+
+                yeardiff_sendertime + yeardiffsquare_sendertime+ AbsDiffMQscores_sendertime + NumberJusticesPro_sendertime,family=binomial(logit),data=train.data)
+summary(mple.est)
+
+save(list="mple.est",file="training_full.RData")
+
+
+# MPLE
[…]

### 2017-12-24 Tex/paper.tex
added in Ted's section

@@
 
-\textbf{TED}, would you write a section that covers the ways in which network scholars have studied citations in the past, focusing on scientific papers, patents, and court cases?
+A citation is a directed link between two documents that indicates the citing document attributes the cited to be in some manner relevant to it.
+Collective patterns of citations within a certain domain make up citation networks, which are most commonly found in three social systems, scientific publishing, patents, and the judicial system.
+When raw citation data is formally represented as a network, the nodes are usually the documents themselves, and each directed arc, usually unweighted, is the existence of one or more cites from the sending to the receiving document.
+Given the permanent nature of documents in citations networks, these networks are in large acyclic \citep{leicht2007large,karrer2009random}.
+Relatedly, citation networks grow as new documents and citations enter the network, but established arcs persist in all but exceptional cases.
+
+	In its raw form a citation is a link between two documents, but it is the result of human judgment about the utility of the cite \citep{fowler2008authority}.
+Studies of citation networks can be broadly categorized according to whether they treat this set of observed behavioural patterns as information that can be used to understand different aspects of the system, or treat them as outcomes to be modeled.
+We discuss these in turn.
+
+	First, as these arcs are observed patterns of behaviour, they can be used to infer latent characteristics of nodes or the relations between nodes that produced them \citep{batagelj2003efficient}.
+Moreover, since each judgment is embedded in the broader network of citations, which influences the judgment beyond the characteristics of the two directly linked documents, inferences about the neighborhood of these two documents are possible as well.
+A common application of this general logic is to obtain local or global network indices, such as Kleinberg centrality at the node level \citep{kleinberg1999authoritative,fowler2007network} or global hierarchy \citep{mones2014universal}, then use them to draw conclusions about networks or compare networks across contexts such as time periods \citep[e.g.][]{vazquez2001statistics,fowler2008authority,greenberg2009citation,lupu2012precedent,lupu2013strategic,dawson2014current,jaffe2017patent}.
+
+	Still within the category of using observed network constellation to make inferences about components of the system, identification of communities using methods for clustering, such as modularity maximization \citep[e.g.][]{kajikawa2007creating,shibata2011detecting,chen2010community} and stochastic block modeling \citep{jo2009citing}; and subnetwork identification \citep[e.g.][]{batagelj2017emergence} have been used to analyze relevance of topics within an academic field or to determine trends in technological advancement \citep{verspagen2007mapping,erdi2013prediction}.
+Main path analysis \citep{hummon1989connectivity}, which determines the main path through an acyclic network, is a particularly useful way to examine trends as it allows researchers identify structures of knowledge flow.
+Other methods include classifying documents into groups based on similarities in their citation profiles of cites over time, which can be examined as a function of time to see if there are temporal patterns \citep{leicht2007large}.
+The identification of communities is often coupled with classification of these communities by the researcher using historical knowledge.
+
+	Researchers are also interested in understanding the mechanisms that drive citations between documents.
+Work in this area generally proceed by specifying a set of citation behaviour, then proposing a model to capture the combination of these behavioural rules \citep{simkin2007mathematical}.
+Assessment of the model is based on how well it fit citation distributions.
+Researchers working in this area tend to focus on modeling the growth of the citation network as governed by a degree-based mechanism such as preferential attachment \citep{barabasi1999emergence} or the recursive search \citep{vazquez2001statistics}, and the age of the paper \citep{jeong2003measuring,eom2011characterizing,wang2013quantifying}.
+Regarding age, the previous standard approach was to treat the probability of citation as a separable function of degree and age \citep{hajra2005aging,hajra2006modelling}, then examine the distribution of citations, scaled to the rate of new publications, by age.
+More recently, work has been done in relaxing the assumption that the effect of degree is static, instead allowing it to vary with time \citep{wang2008measuring}.
+
+	Another approach to understanding the generative process of citation networks is to examine the existence of network motifs, or subnetwork structures, that can be interpreted as measuring different generative mechanisms.
+One way of doing so is to compare the observed network statistics to a null distribution.
+This null model must also be characterized by features of citation networks discussed earlier, meaning that it must be a directed acyclic graph with unweighted arcs \citep{carstens2016topology}.
+Two sets of uniform random graph models have been proposed for directed acyclic graphs.
+In the former, \citet{karrer2009random} introduce two uniform random graph models based respectively on fixed-degree sequence and fixed-expected degree sequence.
+Both models preserve the requisite characteristics of citation networks with the exception that they can introduce weighted edges with low probability, which leads to networks from the null distribution to be sampled with nonequal probability.
+\citet{carstens2014uniform,carstens2016topology} introduces a model that preserves the requisite characteristics and also do not introduce weighted arcs.
+Loosely related is an egocentric framework, introduced by \citet{vu2011dynamic}, that models continuous-time network data through a multivariate counting process, which can be used to count network motifs.
+
+	\section{Extra Notes}
+	This section contains some of the extra notes I cut out from the draft that might be useful.
+	\subsection{Broader definitions of citation networks}
+	More broadly, citation networks can have as nodes the document producers, including scientists \citep{ji2016coauthorship}, judges \citep{landes1998judicial}, journals \citep{rice1988citation}, or aggregated units \citep{gelter2012networks,pan2012world}.
+In these cases, which are essentially exercises in aggregation, the resulting network can take on different characteristics such as weighted or cyclic edges.
+Other networks derived from the raw citation network is the co-citation or co-cited-by networks whereby two documents are connected if they co-cite or are co-cited by another document \citep{van2005reference}.
+
+	\subsection{Weighted edges and multiplexity}
+	Recognition, and subsequent quantification, of different levels of complexity in citation arcs allows for a host of advancements in citation network analysis.
+Weights can indicate relevance \citep{liu2014citations} or be used to incorporate the temporal dimension into citation networks \citep{fujita2014detecting}.
+Moving to multiplexity leads to further advancements.
+For example, \citet{greenberg2009citation} accounted for level of support and through simulations based on arc typed switching demonstrated how citation bias against critical articles yielded differences in beliefs within a medical discipline.
+\citet{bommarito2010distance} argued that articles contain different types of information and accounting for the specific piece of information that resulted in the cite can yield clustering that are more interpretable.
+In all cases however, substantial work has to be done into reclassifying the weightedness or multiplexity of citation arcs \citep{zhang2007semantics}.
+
+	\subsection{Misc. notes}
+	\begin{itemize}
+		\item To control for age, \citet{clough2015transitive} propose transitive reduction of citation networks (i.e.
+removal of redundant information ties) as it will primarily remove citation arcs that are disparate in age.
+		\item Most citation networks are directed acyclic graphs, but not all.
+Different versions of the same document can also cause problems (such as strong network components).
+The ``preprint transformation'' is a solution to small strong components \citep{batagelj2017emergence}.
+		\item Date of publication can be assigned as level to nodes in acyclic networks \citep{batagelj2017emergence}.
+Depending on the citation network at hand, temporal ordering can be difficult, as documents can share publication dates.
+To overcome this, begin each iteration by sampling from one of the possible orderings.
+\citep{carstens2016topology}.
+	\end{itemize}
 

### 2017-12-27 Tex/paper.tex
updated transitivity description

@@
 The formation (or lack thereof) of that relationship cannot be fully understood without considering how the relationship fits into the complex web.
-Analytical designs that account only for covariates in explaining tie formation
+Analytical designs that account only for covariates in explaining tie formation are incomplete theoretically, and, as a consequence, are subject to a form of omitted variable bias [CITE ISQ].
+Citations in legal opinions are unique in terms of the windows into network dependencies offered by the texts of the opinions.
+A number of common structural dependencies that are found in networks are likely to apply to citations in Supreme Court opinions.
+In this section we present these dependence forms, and document the mechanisms by which they arise through quotations in archetypal example opinions.
 
-\textbf{Transitivity:} Discussing past edges in the citation network leads to transitivity. If i cites j, it is also likely to discuss the cases on which j is based.
+\textbf{Transitivity:} In a network of directed relations (e.g., A cites B, but B doesn't cite A) transitivity refers to the tendency for A to send a tie to C if A sends a tie to B and B sends a tie to C [CITE FROM SYLLABUS].
+In undirected networks, transitivity is simply the process by which friends of friends become friends (i.e., a friend of a friend is a friend).
+The term, ``transitive closure'' refers to a tie forming from A to C in response to extant ties from A to B and B to C.
+When writing opinions, Supreme Court justices present the legal bases for their rulings, which often involves discussing the most primary/relevant precedents underpinning these legal bases, but also the precedents and legal rules on which the primary precedents were based.
+This process of presenting several layers/levels of precedent in an opinion follows the structure of transitive closure exactly---opinion A cites opinion B as a primary precedent, and then cites opinion C because opinion B cites opinion C.
+The two examples presented below illustrate this process.
 

### 2017-12-28 Tex/paper.tex
finishing dependency effects

@@
 
-\textbf{Mutuality:} When deciding a set of related cases, the courts opinions regularly reference each other in order to reinforce their arguments...
+\textbf{Reciprocity:} Reciprocity (also referred to as mutuality) is the tendency for node B to send a tie to node A in response to or coordination with A sending a tie to B [CITE FROM SYLLABUS].
+It is typically not possible for reciprocated ties to form in legal opinions.
+Most citations reference past opinions that were issued before the citing opinion's case was even argued before the Court.
+However, opinions written within the same Supreme Court term are often drafted in tandem, and can cite each other reciprocally.
+Opinion A citing opinion B within the same term represents a signal that opinion B is relevant to the legal reasoning underpinning opinion A.
+Unlike the citations themselves, the applicability of legal rules or lines of reasoning across cases is not directed---if A is relevant to B, B is highly likely to be relevant to A.
+We expect opinions written within the same term to exhibit a high degree of reciprocity.
+Below we provide two example passages from opinions that illustrate the phenomenon of within-term reciprocity.
 
@@
 Second, the city asserts, a congressional finding that age is a BFOQ for a certain occupation is dispositive of that determination with respect to nonfederal employees in that occupation.
-We consider each of these contentions in turn.
 
@@
 
-\textbf{Also popularity} If j served as precedent for i, j is authoritative regarding the rule of i, and should also be cited.
-Citations beget citations.
-Citation confers authority.
-Cases are even discussed in opinions in terms of the extent of influence they have had on other cases,
+\textbf{Popularity} Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent [CITE FROM SYLLABUS].
+Citations to an opinion signal both the Court's awareness of the legal reasoning of the case and the Court's evaluation that the opinion is an authoritative precedent.
+The more citations, the stronger this signal.
+Landmark cases, or those that establish new legal rules, are particularly authoritative and accrue citations from most future opinions that follow the respective line of reasoning.
+The passage below, from XXXX, in which an authoritative opinion is referenced, and even discussed in terms of the number of other cases by which it was followed.
+We consider each of these contentions in turn.
+Based on this popularity phenomenon, we expect that new citations will be disproportionately directed at cases to which many past citations have accrued.
 
@@
 
-\textbf{Sociality} Citations sent beget citations sent.
-For each case discussed, often related cases to discuss, even those that do not cite each other.
-For each case that applies, there is likely a case that needs to be discussed in terms of why it does not apply.
+\textbf{Sociality} Sociality is the tendency for ties sent to beget more ties sent [CITE].
+Structurally, this means that there are few nodes that send a moderate number of ties---the number of ties sent is either small or fairly large.
+In terms of Supreme Court opinions,  for each opinion discussed, that discussion is likely to raise other tangential issues on which the justices will want to draw upon past opinions.
+Furthermore, for each opinion that applies to and is cited by the current opinion, there is often a case/opinion that needs to be discussed in terms of why it does not apply.
 Justices often clarify not only those legal principles that apply, but often those that do not.

### 2017-12-28 Tex/paper.tex
revising model definition

@@
 Justices often clarify not only those legal principles that apply, but often those that do not.
-When multiple lines of legal reasoning are drawn upon to justify a decision, often more are discussed in terms of their lack of applicability.
+The passages below represent examples of this process of citation proliferation, or sociality.
 
@@
 
+Researchers from several fields have used network analysis to analyze citations---legal, patent, and scientific.
+Before describing the network model for Supreme Court citations that we develop, we review the methods researchers have previously used to study citations..
 A citation is a directed link between two documents that indicates the citing document attributes the cited to be in some manner relevant to it.
@@
 
- \section{The Exponential Random Configuration Model}
-Let $c(t)\in \{0,1\}^N$ be a vector indicating which Supreme Court case has been cited at time $t$, where $c_i(t)=1, i \in \{1, \dots , N\}$ indicates that the $i$th case has been cited at time $t$ and $c_i(t)=0$ indicates that the $i$th case has not been cited at time $t$.
-Furthermore, let
+ \section{The Citation Temporal Exponential Random Graph Model}
+We build upon existing methods for network analysis to define a model that fits the constraints to which Supreme Court citation networks are subject.
+The Supreme Court citation network is nearly a directed acyclic graph---a graph in which there can be no loops/cycles (e.g., ties cannot be reciprocated, which would constitute a two-node loop).
+If two cases are decided in different terms, the later case can cite the earlier case, but not the reverse.
+However, as noted above, if two opinions are written in the same term, they can cite each other reciprocally.
+We implement a version of the temporal exponential random graph model (TERGM)---a model that is commonly used for longitudinal network data [CITES]---that is subject to the constraints specific to the Supreme Court citation network.
+The TERGM is a model that can simultaneously incorporate structural dependencies such as transitivity and reciprocity and exogenous covariates  (e.g.
+the issue area of the citing case, the ideological distance between the opinion authors of the two cases in the dyad) to explain tie formation.
+If the ties do not depend upon each other, and the coefficients associated with the structural dependencies are either assumed or estimated to be zero, the TERGM reduces to a logistic regression in which the observation is a directed dyad [CITE OLD PA].
+In this section we describe the citation TERGM (c-TERGM), and explain estimation methods.
+
+Let $c_ij \in \{0,1\}$ be a binary indicator of whether case $i$ cites $j$. Furthermore, let
 $$\mathcal{C}_t(N)=\{c(t)\in \{0,1\}^N: c_i(t)\in \{0,1\} \}$$ 
@@
 Unlike in-sample measures of model fit, out-of-sample methods are highly robust in avoiding overfitting, and work when we cannot accurately calculate the value of the likelihood function, as in the current case.
+Out-of-sample prediction is a common way to evaluate methods for modeling ties in networks, and has recently been applied to TERGMs in particular [CITE PA AND IEEE].
 

### 2017-12-29 Tex/paper.tex
revising notation

@@
 
-Let $c_ij \in \{0,1\}$ be a binary indicator of whether case $i$ cites $j$. Furthermore, let
-$$\mathcal{C}_t(N)=\{c(t)\in \{0,1\}^N: c_i(t)\in \{0,1\} \}$$ 
-be the set of all possible citation combinations at time $t$. Note that the cardinality of $\mathcal{C}_t(N)$ increases exponentially for every newly added case, which results in $2^N$ elements.\\
-The probability function of the ERCM is defined as
+Let $c_{ij} \in \{0,1\}$ be a binary indicator of whether case $i$ cites $j$. Furthermore, let
+$$\mathcal{C}_t=\{c(t)\in \{0,1\}^{N_t}: c_{ij}(t)\in \{0,1\} \}$$ 
+be the set of all possible citation combinations for opinions issued in term $t$, where $c_{ij}(t)$ is a binary indicator of whether case $i$ cites case $j$, and case $i$ is decided at time $t$.
+Note that the cardinality of $\mathcal{C}_t$ increases exponentially for every newly added case.\\ The probability function of the c-TERGM  is defined as
 \begin{equation}
-P_{\theta}(c(t)~|~c(t-))=\cfrac{\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}{\sum_{c(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}
+P_{\theta}(c(t)~|~c(t-))=\cfrac{\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}{\sum_{z(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(z(t)~|~c(t-)\big)\big)}
 \label{ercm}
 \end{equation}
-where $c(t-)\in \{0,1\}^{N \times (t-1)}$ is a matrix that indicates which cases have been citing in each other before time $t$, $\theta \in \mathbb{R}^q$ is a q-dimensional vector of parameters,  $h: \mathcal{C}_t(N) \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{c(t)^*\in \mathcal{C}}\exp(\theta^T\cdot h(c(t)|c(t-)))$ is a normalization constant that ensures that (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.\\[0.3cm]
+where $c(t-)$ is a matrix that indicates the citations among cases decided before time $t$, $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C}_t \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{z(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(z(t)~|~c(t-)\big))$ is a normalization constant that ensures that (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.\\[0.3cm]
 

### 2017-12-29 Tex/paper.tex
revising notation

@@
 \end{equation}
-where $c(t-)$ is a matrix that indicates the citations among cases decided before time $t$, $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C}_t \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{z(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(z(t)~|~c(t-)\big))$ is a normalization constant that ensures that (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.\\[0.3cm]
+where $c(t-)$ is a matrix that indicates the citations among cases decided before time $t$, $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C}_t \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{z(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(z(t)~|~c(t-)\big))$ is a normalization constant that ensures that Equation (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.
 
 % Network Statistics
-The generative process of a model are informed by
-the decision regarding which network statistics $h(\cdot)$ are incorporated. We include the following statistics for the Supreme Court citation network:
-$$h_{edges}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{i=1}^Nc_i(t)$$
-the number of citations made at time $t$. 
-$$h_{outstar}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
-the number of weighted outstars occuring at time $t$. We argue that it should be more likely to cite more recent cases than cases that have been decided further in the past. For the weight 
-$$w(a,b):= \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
-we define $a$ and $b$ as the elapsed time since case $i$ and $j$ have been introduced to the network.
+The c-TERGM is specified through 
+the decision regarding which network statistics $h(\cdot)$ to incorporate.
+We include the following statistics for the Supreme Court citation network, which are derived from the interdependence hypotheses described in the previous section:
+$$h_{edges}:\mathcal{C}_t\to \mathbb{R}~~~, ~~~c(t) \to \sum_{i(t)} c_{ij}, $$
+the number of citations made at time $t$, where $i(t)$ is the set of opinions issued at time $t$.  $h_{edges}$ performs the function of an intercept, and models the expected value of any given edge.
+$$h_{outstar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~c(t) \to \sum_{i(t)} {\sum_{j \neq i} c_{ij} \choose 2} $$
+the number of out-two-stars occuring at time $t$.
+An out-two-star is a configuration in which one node sends to two other nodes.
+The number of these configurations grows quadratically as the origins of ties concentrate on a few highly active/social senders.
+The out-two-stars configuration is commonly used to model sociality in ERGMs.
+$$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~c(t) \to \sum_{j(t)} {\sum_{i \neq j} c_{ij} \choose 2} $$
+the number of in-two-stars occuring at time $t$.
+An in-two-star is a configuration in which one node receives ties from two other nodes.
+The number of these configurations grows quadratically as the destinations of ties concentrate on a few highly popular recipients.
+The out-two-stars configuration is commonly used to model popularity in ERGMs.
+[HERE]
+
 $$h_{triangle}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot
@@
 \\[0.3cm]
-
-%Change Statistic
-The individual entries $c_i(t)$ can be taken as a manifestation of single Bernoulli variables $C_i(t)$.
-This interpretation allows the following calculation regarding the conditional distribution of $C_i(t)$:
-%
-\begin{eqnarray*}
-\dfrac{P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~|~ C_i(t)^c=c_i(t)^c)} &=&
-\dfrac{P_{\theta}(C_i(t)=1 ~,~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~,~ C_i(t)^c=c_i(t)^c)} \\
-                           &=&\dfrac{P_{\theta}(C(t)= c_i^+(t))}{P_{\theta}(C(t)=c_i^-(t))}\\
-                           &=&\dfrac{\exp(\theta^T \cdot h(c_i^+(t)~|~c(t-)))}{\exp(\theta^T \cdot h(c_i^-(t)~|~c(t-)))}\\
-                           &=& \exp(\theta^T \cdot (h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-)))
-\end{eqnarray*}
-%
-This implies the following equation:
-%
-\begin{equation}
-\text{logit}(P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c))= \theta^T \cdot (h(c_i^+(t)~~|c(t-)) - h(c_i^-(t)~|~c(t-)))
-\label{Logit}
-\end{equation}
-In the equation above the following notations were used:
-%
-\begin{itemize}
-\item $c_i^+(t)$ emerges from $c(t)$, while assuming $c_i(t)=1$
-\item $c_i^-(t)$ emerges from $c(t)$, while assuming $c_i(t)=0$
-\item The condition $C_i(t)^c=c_i(t)^c$ is short for: $C_j(t)=c_j(t)$ for all $j\in \{1,\dots,N\}$ with $i \neq j$
-\item The expression $(\Delta c_i)(t):=h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-))$ is called the \textit{change statistic}.
-The $k$th component of $(\Delta c_i)(t)$ captures the difference between citation networks $c_i^+(t)$ and $c_i^-(t)$ on the $k$th integrated statistic in the model
-\end{itemize}
-
-\section{Estimation}
-
-% Pseudo Likelihood
-\subsection*{Maximum Pseudo-Likelihood Estimator}
-One can assume that the dyads are independent of each other, which means that
-the random variables $C_i(t)$ inside the random vector $C(t)$ are independent of each other.
-In this case, the equation (\ref{Logit}) reduces to
-$$logit(P_{\theta}(C_i(t) = 1)) = \theta^T \cdot (\Delta c_i)(t)$$
-This corresponds with the logistic regression approach, where the observations of
-the dependent variables are simply edge values of the observed citation vector,
-and the observations of the covariate values are given as the scores of every single
-change statistic. Therefore, the resulting likelihood function is of the following form:
-\begin{equation}
-\text{lik}(\theta)= P_{\theta}(C(t)=c(t))= \prod_{i} \dfrac{ \exp \left(\theta^T \Delta(c_i)(t) \right)}{1+\exp \left(\theta^T \Delta(c_i)(t) \right)}
-\label{PseudoLik}
-\end{equation}
-
-\subsection*{Maximum Likelihood Estimator}
-The more rigorous technique is to estimate the parameters directly with the log-likelihood function derived from (\ref{ercm}), which has the following form:
-%
-\begin{equation}
-\text{loglik}(\theta)=\theta^T \cdot h(c(t)| c(t-))-\log(\kappa(\theta))
-\label{loglik}
-\end{equation}
-%
-The problem resulting from estimating the parameters with (\ref{loglik}) is that the term
-%
-$$\kappa(\theta):= \sum_{c(t)^* \in \mathcal{C}(N)} \exp(\theta^T \cdot h(c(t)^*|c(t-)))$$ 
-%
-which sums up the weighted statistics of all possible binar vectors of length $N$, has to be evaluated.
-However, the cardinality of $\mathcal{C}(N)$ (\#$(\mathcal{C})=2^N$) is incredibly large and a direkt calculation of this sum is for already small $N$ not feasible.
-\\[0.3cm]
-An solution for this limitation is based on the following consideration: Fix a vector of parameters $\theta_0 \in \Theta$ from the underlying parameter range $\Theta$ and compute for $\theta \in \Theta$ the expected value
-%
-\begin{eqnarray*}
-\mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right]&=& 
-\sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot \mathbb{P}_{\theta_0}(C(t)=c(t))\\
-&=& \sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot 
-\frac{\exp(\theta_0^T \cdot \Gamma(c(t)))}{\kappa(\theta_0)}\\
-&=& \frac{1}{\kappa(\theta_0)} \sum_{c(t) \in \mathcal{C}(N)}\exp\left(\theta^T \cdot \Gamma(c(t))\right)\\
-&=&\frac{\kappa(\theta)}{\kappa(\theta_0)}
-\end{eqnarray*}
-%
-This equation offers the following possibility: If one draws $L$ random vectors $c^{(1)}(t), \dots ,c^{(L)}(t)$ out of a distribution $\mathbb{P}_{\theta_0}$ appropriately, one gets with the law of big numbers and a big enough sample $L$ the following relation:
-%
-\begin{equation}
-\frac{1}{L}\cdot \sum_{i=1}^{L}  \exp\left((\theta - \theta_0)^T \cdot \Gamma(c^{(i)}(t))\right)
-~~\approx~~ \mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right] = \frac{\kappa(\theta)}{\kappa(\theta_0)}
-\label{konver}
-\end{equation}
-%
-This approximate can then be used to approximate the log likelihood function.\\[0.4cm]
-Next, we will discuss how a sufficient number of suitable drawings $c^{(1)}(t), \dots ,c^{(L)}(t)$ can be sampled from the distribution $\mathbb{P}_{\theta_0}$. \\
-For this purpose, the Markov Chain Monte Carlo (MCMC) methods can be used.
-\subsection*{Gibbs sampling for the ERCM}
-To be able to compute the approximate likelihood function one needs a sufficiently large number of random vectors from the distribution $\mathbb{P}_{\theta_0}$.
-Snijders \citet{Snijders.2002b} introduces an approach to sample random networks for the ERGM framework by using \textit{MCMC methods}.
-We adapt this approach for sampling appropriate binary vectors for the ERCM.
-\\[0.4cm]
-\textit{Gibbs sampling}\\
-%Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g.
-observed vector).
-Afterwards, the length $L$ of the respective sub-sequence is determined.
-For $k \in \{0,...,L-1\}$ execute the following steps recursively (here the vector in its $k$th iteration is denoted as $c^{(k)}(t)$):\\
-
-\begin{algorithm}[H]
- Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g. observed vector)\\
- \For{i in 1:N}{
-  Compute $\pi:= \cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$\\
-  Draw a random number Z from Bin(1,$\pi$)\\
-  \eIf{Z=1}{
-   set $c^{(k+1)}_i = 1$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
-   }{
-   set $c^{(k+1)}_i = 0$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
-  }
- }
- Start all over using $c^{(k+1)}$\\[0.3cm]
- \caption{Simulation of vectors of $\mathbb{P}_\theta$ using Gibbs sampling}
-\end{algorithm}
-\vspace{0.5cm}
-%\begin{enumerate}
-%\item Randomly choose a number $i \in \{1,\dots, N\}$
-%\item Compute using the likelihood the value 
-%$$\pi:= \mathbb{P}_{\theta}(C_{i}(t)=1 | C_{i}^c(t)=(c_{i}^{(k)}(t))^c)=\cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$$
-%\item Draw a random number $Z$ from Bin$(1, \pi)$. %If
-%\begin{itemize}
-%\item $Z=0$, define $c^{(k+1)}(t)$ via
-%$$c_{p}^{(k+1)}(t)=\begin{cases}
-%0& \text{if}~ p=i \\
-%c_{p}^{(k)}(t) &\text{if}~ p \neq i 
-%\end{cases}$$
-%\item $Z=1$, define $c^{(k+1)}(t)$ via
-%$$c_p^{(k+1)}(t)=\begin{cases}
-%1& \text{if}~ p=i \\
-%c_p^{(k)}(t) &\text{if}~ p \neq i 
-%\end{cases}$$
-%\end{itemize}
-%\item Start at step 1 with $c^{(k+1)}(t)$.
-%\end{enumerate}
-\noindent Using the depicted algorithm, a sequence of random vectors $c^{(0)}(t),...,c^{(L)}(t)$ can be simulated.
-Since the orignial vector was chosen randomly and the first simulated vectors are very dependent on the chosen mvector (only one entry is changed per iteration!), usually the first $B$ vectors, where $N \ll B \ll L$, are discarded as the so called \textit{Burn-In}.
-\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
-Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ to start with (e.g., the observed vector). For $k \in \{0,...,L-1\}$ recursively proceed as follows:\\
-\begin{enumerate}
-\item Randomly choose a number $i\in \{1,\dots, N\}$
-\item Compute, using the equation (\ref{Logit}) the value
-$$\pi := \dfrac{\mathbb{P}_{\theta}(C_{i}(t) \neq c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}{\mathbb{P}_{\theta}(C_{i}(t) = c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}$$
-\item Define $\delta:= \min\{1, \pi\}$ and draw a random number $Z$ from Bin$(1, \delta)$. If
-\begin{itemize}
-\item $Z=0$, let $c^{(k+1)}(t) := c^{(k)}(t)$ 
-\item $Z=1$, define $c^{(k+1)}(t)$ as
-$$c_{p}^{(k+1)}(t)=\begin{cases}
-1-c_{p}^{(k)}(t)& \text{if}~ p=i \\
-c_{p}^{(k)}(t) &\text{if}~ p \neq i 
-\end{cases}$$
-\end{itemize}
-\item Start at step 1 with $c^{(k+1)}(t)$.
-\end{enumerate}
-The first $B \ll L$ vectors are discarded as Burn-In.
 
@@
 \end{document}
+
+% Network Statistics
+The c-TERGM is specified through 
+the decision regarding which network statistics $h(\cdot)$ to incorporate. We include the following statistics for the Supreme Court citation network:
+$$h_{edges}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{i=1}^Nc_i(t)$$
+the number of citations made at time $t$.  $h_{edges}$ is like an intercept, and models the expected value of any given edge.
+$$h_{outstar}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
+the number of weighted outstars occuring at time $t$. We argue that it should be more likely to cite more recent cases than cases that have been decided further in the past. For the weight 
+$$w(a,b):= \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
+we define $a$ and $b$ as the elapsed time since case $i$ and $j$ have been introduced to the network.
+$$h_{triangle}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot
+c_j(t_{-i}) \cdot w(a,b)$$
+where $c_j(t_{-i})$ indicates whether case $j$ was cited at the time case $i$ was introduced into the network.
+Just as for the outstar statistic, we include a weighting factor to favor more recent cases.
+\\[0.3cm]
+
+%Change Statistic
+The individual entries $c_i(t)$ can be taken as a manifestation of single Bernoulli variables $C_i(t)$.
+This interpretation allows the following calculation regarding the conditional distribution of $C_i(t)$:
+%
+\begin{eqnarray*}
+\dfrac{P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~|~ C_i(t)^c=c_i(t)^c)} &=&
+\dfrac{P_{\theta}(C_i(t)=1 ~,~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~,~ C_i(t)^c=c_i(t)^c)} \\
+                           &=&\dfrac{P_{\theta}(C(t)= c_i^+(t))}{P_{\theta}(C(t)=c_i^-(t))}\\
+                           &=&\dfrac{\exp(\theta^T \cdot h(c_i^+(t)~|~c(t-)))}{\exp(\theta^T \cdot h(c_i^-(t)~|~c(t-)))}\\
+                           &=& \exp(\theta^T \cdot (h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-)))
+\end{eqnarray*}
+%
+This implies the following equation:
+%
+\begin{equation}
+\text{logit}(P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c))= \theta^T \cdot (h(c_i^+(t)~~|c(t-)) - h(c_i^-(t)~|~c(t-)))
+\label{Logit}
+\end{equation}
+In the equation above the following notations were used:
+%
+\begin{itemize}
+\item $c_i^+(t)$ emerges from $c(t)$, while assuming $c_i(t)=1$
+\item $c_i^-(t)$ emerges from $c(t)$, while assuming $c_i(t)=0$
+\item The condition $C_i(t)^c=c_i(t)^c$ is short for: $C_j(t)=c_j(t)$ for all $j\in \{1,\dots,N\}$ with $i \neq j$
+\item The expression $(\Delta c_i)(t):=h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-))$ is called the \textit{change statistic}.
+The $k$th component of $(\Delta c_i)(t)$ captures the difference between citation networks $c_i^+(t)$ and $c_i^-(t)$ on the $k$th integrated statistic in the model
+\end{itemize}
+
+\section{Estimation}
+
+% Pseudo Likelihood
+\subsection*{Maximum Pseudo-Likelihood Estimator}
+One can assume that the dyads are independent of each other, which means that
+the random variables $C_i(t)$ inside the random vector $C(t)$ are independent of each other.
+In this case, the equation (\ref{Logit}) reduces to
+$$logit(P_{\theta}(C_i(t) = 1)) = \theta^T \cdot (\Delta c_i)(t)$$
+This corresponds with the logistic regression approach, where the observations of
+the dependent variables are simply edge values of the observed citation vector,
+and the observations of the covariate values are given as the scores of every single
+change statistic. Therefore, the resulting likelihood function is of the following form:
+\begin{equation}
+\text{lik}(\theta)= P_{\theta}(C(t)=c(t))= \prod_{i} \dfrac{ \exp \left(\theta^T \Delta(c_i)(t) \right)}{1+\exp \left(\theta^T \Delta(c_i)(t) \right)}
+\label{PseudoLik}
+\end{equation}
+
+\subsection*{Maximum Likelihood Estimator}
+The more rigorous technique is to estimate the parameters directly with the log-likelihood function derived from (\ref{ercm}), which has the following form:
+%
+\begin{equation}
+\text{loglik}(\theta)=\theta^T \cdot h(c(t)| c(t-))-\log(\kappa(\theta))
+\label{loglik}
+\end{equation}
+%
+The problem resulting from estimating the parameters with (\ref{loglik}) is that the term
+%
+$$\kappa(\theta):= \sum_{c(t)^* \in \mathcal{C}(N)} \exp(\theta^T \cdot h(c(t)^*|c(t-)))$$ 
+%
+which sums up the weighted statistics of all possible binar vectors of length $N$, has to be evaluated.
+However, the cardinality of $\mathcal{C}(N)$ (\#$(\mathcal{C})=2^N$) is incredibly large and a direkt calculation of this sum is for already small $N$ not feasible.
+\\[0.3cm]
+An solution for this limitation is based on the following consideration: Fix a vector of parameters $\theta_0 \in \Theta$ from the underlying parameter range $\Theta$ and compute for $\theta \in \Theta$ the expected value
+%
+\begin{eqnarray*}
+\mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right]&=& 
+\sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot \mathbb{P}_{\theta_0}(C(t)=c(t))\\
+&=& \sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot 
+\frac{\exp(\theta_0^T \cdot \Gamma(c(t)))}{\kappa(\theta_0)}\\
+&=& \frac{1}{\kappa(\theta_0)} \sum_{c(t) \in \mathcal{C}(N)}\exp\left(\theta^T \cdot \Gamma(c(t))\right)\\
+&=&\frac{\kappa(\theta)}{\kappa(\theta_0)}
+\end{eqnarray*}
+%
+This equation offers the following possibility: If one draws $L$ random vectors $c^{(1)}(t), \dots ,c^{(L)}(t)$ out of a distribution $\mathbb{P}_{\theta_0}$ appropriately, one gets with the law of big numbers and a big enough sample $L$ the following relation:
+%
+\begin{equation}
+\frac{1}{L}\cdot \sum_{i=1}^{L}  \exp\left((\theta - \theta_0)^T \cdot \Gamma(c^{(i)}(t))\right)
+~~\approx~~ \mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right] = \frac{\kappa(\theta)}{\kappa(\theta_0)}
+\label{konver}
+\end{equation}
+%
+This approximate can then be used to approximate the log likelihood function.\\[0.4cm]
+Next, we will discuss how a sufficient number of suitable drawings $c^{(1)}(t), \dots ,c^{(L)}(t)$ can be sampled from the distribution $\mathbb{P}_{\theta_0}$. \\
+For this purpose, the Markov Chain Monte Carlo (MCMC) methods can be used.
+\subsection*{Gibbs sampling for the ERCM}
+To be able to compute the approximate likelihood function one needs a sufficiently large number of random vectors from the distribution $\mathbb{P}_{\theta_0}$.
+Snijders \citet{Snijders.2002b} introduces an approach to sample random networks for the ERGM framework by using \textit{MCMC methods}.
+We adapt this approach for sampling appropriate binary vectors for the ERCM.
+\\[0.4cm]
+\textit{Gibbs sampling}\\
+%Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g.
+observed vector).
+Afterwards, the length $L$ of the respective sub-sequence is determined.
+For $k \in \{0,...,L-1\}$ execute the following steps recursively (here the vector in its $k$th iteration is denoted as $c^{(k)}(t)$):\\
+
+\begin{algorithm}[H]
+ Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g. observed vector)\\
+ \For{i in 1:N}{
+  Compute $\pi:= \cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$\\
+  Draw a random number Z from Bin(1,$\pi$)\\
+  \eIf{Z=1}{
+   set $c^{(k+1)}_i = 1$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
+   }{
+   set $c^{(k+1)}_i = 0$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
+  }
+ }
+ Start all over using $c^{(k+1)}$\\[0.3cm]
+ \caption{Simulation of vectors of $\mathbb{P}_\theta$ using Gibbs sampling}
+\end{algorithm}
+\vspace{0.5cm}
+%\begin{enumerate}
+%\item Randomly choose a number $i \in \{1,\dots, N\}$
+%\item Compute using the likelihood the value 
+%$$\pi:= \mathbb{P}_{\theta}(C_{i}(t)=1 | C_{i}^c(t)=(c_{i}^{(k)}(t))^c)=\cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$$
+%\item Draw a random number $Z$ from Bin$(1, \pi)$. %If
+%\begin{itemize}
+%\item $Z=0$, define $c^{(k+1)}(t)$ via
+%$$c_{p}^{(k+1)}(t)=\begin{cases}
+%0& \text{if}~ p=i \\
+%c_{p}^{(k)}(t) &\text{if}~ p \neq i 
+%\end{cases}$$
+%\item $Z=1$, define $c^{(k+1)}(t)$ via
+%$$c_p^{(k+1)}(t)=\begin{cases}
+%1& \text{if}~ p=i \\
+%c_p^{(k)}(t) &\text{if}~ p \neq i 
+%\end{cases}$$
+%\end{itemize}
+%\item Start at step 1 with $c^{(k+1)}(t)$.
+%\end{enumerate}
+\noindent Using the depicted algorithm, a sequence of random vectors $c^{(0)}(t),...,c^{(L)}(t)$ can be simulated.
+Since the orignial vector was chosen randomly and the first simulated vectors are very dependent on the chosen mvector (only one entry is changed per iteration!), usually the first $B$ vectors, where $N \ll B \ll L$, are discarded as the so called \textit{Burn-In}.
+\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
+Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ to start with (e.g., the observed vector). For $k \in \{0,...,L-1\}$ recursively proceed as follows:\\
+\begin{enumerate}
+\item Randomly choose a number $i\in \{1,\dots, N\}$
+\item Compute, using the equation (\ref{Logit}) the value
+$$\pi := \dfrac{\mathbb{P}_{\theta}(C_{i}(t) \neq c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}{\mathbb{P}_{\theta}(C_{i}(t) = c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}$$
+\item Define $\delta:= \min\{1, \pi\}$ and draw a random number $Z$ from Bin$(1, \delta)$. If
+\begin{itemize}
+\item $Z=0$, let $c^{(k+1)}(t) := c^{(k)}(t)$ 
+\item $Z=1$, define $c^{(k+1)}(t)$ as
+$$c_{p}^{(k+1)}(t)=\begin{cases}
+1-c_{p}^{(k)}(t)& \text{if}~ p=i \\
+c_{p}^{(k)}(t) &\text{if}~ p \neq i 
+\end{cases}$$
+\end{itemize}
+\item Start at step 1 with $c^{(k+1)}(t)$.
+\end{enumerate}
+The first $B \ll L$ vectors are discarded as Burn-In.

### 2017-12-30 Tex/paper.tex
finished notation revisions

@@
 
-Let $c_{ij} \in \{0,1\}$ be a binary indicator of whether case $i$ cites $j$. Furthermore, let
-$$\mathcal{C}_t=\{c(t)\in \{0,1\}^{N_t}: c_{ij}(t)\in \{0,1\} \}$$ 
-be the set of all possible citation combinations for opinions issued in term $t$, where $c_{ij}(t)$ is a binary indicator of whether case $i$ cites case $j$, and case $i$ is decided at time $t$.
-Note that the cardinality of $\mathcal{C}_t$ increases exponentially for every newly added case.\\ The probability function of the c-TERGM  is defined as
+Let $C$ be the case-to-case adjacency matrix such that $C_{ij} \in \{0,1\}$ is a binary indicator of whether case $i$ cites $j$. Furthermore, let
+$$\mathcal{C} =\{C \in \{0,1\}^{(N \cdot (N-1))/2}: C_{ij} \in \{0,1\} \}$$ 
+be the set of all possible adjacency matrices among $N$ cases.
+Note that the cardinality of $\mathcal{C}$ increases exponentially for every newly added case.
+The probability function of the c-TERGM  is defined as
 \begin{equation}
-P_{\theta}(c(t)~|~c(t-))=\cfrac{\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}{\sum_{z(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(z(t)~|~c(t-)\big)\big)}
+P_{\theta}(C)=\cfrac{\exp\big(\theta^T\cdot h\big(C)\big)\big)}{\sum_{Z^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(Z^*)\big)\big)}
 \label{ercm}
 \end{equation}
-where $c(t-)$ is a matrix that indicates the citations among cases decided before time $t$, $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C}_t \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{z(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(z(t)~|~c(t-)\big))$ is a normalization constant that ensures that Equation (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.
+where $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C} \to \mathbb{R}^q$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{Z^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(Z^*)\big)$ is a normalization constant that ensures that Equation (\ref{ercm}) defines a probability function on $\mathcal{C}$.
 
@@
 We include the following statistics for the Supreme Court citation network, which are derived from the interdependence hypotheses described in the previous section:
-$$h_{edges}:\mathcal{C}_t\to \mathbb{R}~~~, ~~~c(t) \to \sum_{i(t)} c_{ij}, $$
-the number of citations made at time $t$, where $i(t)$ is the set of opinions issued at time $t$.  $h_{edges}$ performs the function of an intercept, and models the expected value of any given edge.
-$$h_{outstar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~c(t) \to \sum_{i(t)} {\sum_{j \neq i} c_{ij} \choose 2} $$
-the number of out-two-stars occuring at time $t$.
+$$h_{edges}:\mathcal{C} \to \mathbb{R}~~~, ~~~C \to \sum_{ij} C_{ij}, $$
+the number of citations.  $h_{edges}$ performs the function of an intercept, and models the expected value of any given edge.
+$$h_{outstar}:\mathcal{C} \to \mathbb{R}~~~, ~~~C \to \sum_{i=1}^N {\sum_{j \neq i} c_{ij} \choose 2} $$
+the number of out-two-stars.
 An out-two-star is a configuration in which one node sends to two other nodes.
@@
 The out-two-stars configuration is commonly used to model sociality in ERGMs.
-$$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~c(t) \to \sum_{j(t)} {\sum_{i \neq j} c_{ij} \choose 2} $$
-the number of in-two-stars occuring at time $t$.
+$$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C \to \sum_{j=1}^N {\sum_{i \neq j} c_{ij} \choose 2} $$
+the number of in-two-stars.
 An in-two-star is a configuration in which one node receives ties from two other nodes.
@@
 The out-two-stars configuration is commonly used to model popularity in ERGMs.
-[HERE]
+\begin{align*}
+h_{triangle}:\mathcal{C} \to \mathbb{R}~~~, ~~~C \to & \sum_{j<i<k} C_{ij}\cdot C_{jk} \cdot C_{ki} + C_{ji}\cdot C_{jk} \cdot C_{ki} + C_{ij}\cdot C_{kj} \cdot C_{ki} \\ & + C_{ji}\cdot C_{kj} \cdot C_{ki} + C_{ij}\cdot C_{jk} \cdot C_{ik} + C_{ji}\cdot C_{jk} \cdot C_{ik}  \\ & + C_{ij}\cdot C_{kj} \cdot C_{ik} + C_{ji}\cdot C_{kj} \cdot C_{ik},
+\end{align*}
+The number of triangles in the network.
+Triangles measure triad closure, as discussed above.
+$$h_{covariate}:\mathcal{C} \to \mathbb{R}~~~, ~~~C \to \sum_{ij} C_{ij}X_{ij}, $$ the effect of an exogenous covariate ($X$).
+We include several exogenous covariates based on this standard statistic formulation.
+These covariates are discussed below.
 
-$$h_{triangle}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot
-c_j(t_{-i}) \cdot w(a,b)$$
-where $c_j(t_{-i})$ indicates whether case $j$ was cited at the time case $i$ was introduced into the network.
-Just as for the outstar statistic, we include a weighting factor to favor more recent cases.
-\\[0.3cm]
+The structure of this model is very similar to a conventional (T)ERGM setup.
+The main difference is that  $\mathcal{C}$ excludes adjacency matrices in which there are loops that include edges sent at different time points.
+Loops can exist in the Supreme Court citation network, but only among cases decided in the same term.
+We estimate the c-TERGM via bootstrapped pseudolikelihood, as introduced by [CITE PHYSICA].
 

### 2017-12-30 Tex/paper.tex
revising literature setup

@@
 \section{Prior work on court citations}
+
+Theorizing that legal citations arise through network interdependence processes  represents a potential contribution to the literature.
+To demonstrate the value of our contribution we must also incorporate established non-network processes that have been found to affect legal citations.
+We incorporate these processes through the specification of $h_{covariate}$ to be included in the c-TERGM.
 

### 2017-12-31 Tex/paper.tex
revising lit`
finished revising literature section

@@
 
-\section{Prior work on court citations}
+\section{Correlates of Legal Citations}
 
@@
 To demonstrate the value of our contribution we must also incorporate established non-network processes that have been found to affect legal citations.
-We incorporate these processes through the specification of $h_{covariate}$ to be included in the c-TERGM.
+We incorporate these processes through the specification of $h_{covariate}$  terms to be included in the c-TERGM.
 
-Measures of citations have been used in several recent studies of courts...\citet{hinkle2016transmission} study the citations between US state supreme courts.
+We incorporate a statistic to model the degree to which cases cite those that are similar in terms of the ideological positions of the justices who supported the decision.
+We account for this effect following citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
+\citet{clark2010locating} estimates a latent coordinate model of Cupreme Court opinions based on the network of case-to-case citations.
+They find that the majority opinion falls at the ideal point of the median member of the majority coalition in the case.
+We include a covariate term in which $X_{ij}$ is the absolute difference between the Martin-Quinn [CITE MQ] scores of the median justices in the majority coalitions for cases $i$ and $j$.
+We expect this variable to have a negative effect, which would correspond to cases citing those to which they are ideologically similar.%WE COULD USE THIS DATASET AS A SECOND APPLICATION
 
-\begin{itemize}
+We include two sets of dummy variables that account for the issue areas of cases.
+In one set of dummy variable---sender intercepts---the variable $X_{ij}$ indicates the issue area of the sending case ($i$).
+In the set of receiver intercepts $X_{ij}$ indicates the issue are of the receiver case.
+Issue area data comes from the Supreme Court Database (SCDB) [CITE SCDB].
+We include these variables because \citet{cross2010determinants} find that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
 
-\item \citet{clark2010locating} Estimates a latent coordinate model of supreme court opinions based on the network of supreme court citations.
-Find that the majority opinion falls at the ideal point of the median member of the majority coalition.
-Account for negative citations.
-WE COULD USE THIS DATASET AS A SECOND APPLICATION
+We model the way in which citations to a case depend upon the age of a case.
+For this we use a second-order polynomial in which one covariate $X_{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X_{ij}$ is the squared age of case $j$ at the time that case $i$ is decided.
+We include these covariates because \citet{black2013citation} find that the number of citations to a Supreme Court case over time depends significantly on the age of the case, characterized by a sharp drop off and leveling out with age.
 
-\item \citet{cross2010determinants} Look at the total number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
+\citet{benjamin2012standing} Study the propensity for cases to be overruled and cited in other negative ways.
+They find that cases with majority coalitions that are large and ideologically broad are less likely to be cited negatively.
+In our data we do not differentiate between negative and positive citaitons, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis.
+We include one covariate in which $X_{ij}$ is the number of justices in the majority coalition for case $j$.
+We also include another covariate in which $X_{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$.
+We expect both of these covariates to have positive effects.
+
+%%%% DONT DELETE THESE COMMENTS, INCLUDE IDEAS FOR ADDITIONAL COVARIATES %%%%%
+
+%\citet{cross2010determinants} find that the number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
 (case level).
@@
 
-\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
+%\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
 Show that the rate of citations to a case goes down very quickly.
@@
 
-\item \citet{benjamin2012standing} Study the propensity for cases to be overruled and cited in other negative ways.
-They find that cases with large and broad majorities rare less likely to be cited negatively.
+% \item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
 
-\item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
-
-\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
+%\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
 Lower if stautory cases.
@@
 Increases with the number of concurring opinions.
-
-\end{itemize}
 

### 2017-12-31 Network Visualizations/visualizations.tex

%preamble
%basics
\documentclass[12pt]{article}
\usepackage{setspace}
\singlespacing
\usepackage{fancyhdr}
\usepackage[margin=1in]{geometry}
\frenchspacing
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Citation Networks}
\fancyhead[R]{\thepage}
\renewcommand{\headrulewidth}{0.5pt}
\usepackage{csquotes}

%figures graphs equations
\usepackage{graphics}
\usepackage{tikz}
\usepackage{amsmath,amsfonts,amsthm}
\setlength\parindent{20pt}
\usepackage{cancel}
\usepackage[normalem]{ulem}
\usepackage{siunitx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{colortbl}

%colours
\usepackage{color}
\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\definecolor{mylightgray}{gray}{.9}

%references
\usepackage[numbers]{natbib}
\bibliographystyle{plainnat}
\setcitestyle{aysep={}}
\newcommand\posscite[1]{\citeauthor{#1}\textcolor{cyan}{'s} (\citeyear{#1})}
\usepackage[hyperfootnotes=false]{hyperref}
\hypersetup{
    colorlinks=true,
	citecolor=mygray,
	linkcolor=black,
	urlcolor=mygray}

%code
\usepackage{lmodern}
\usepackage{listings}
\lstset{basicstyle=\selectfont\ttfamily\footnotesize,showstringspaces=false,language=R, frame=single, breaklines=true,commentstyle=\color{mygreen},keywordstyle=\color{blue}}

%title
\title{Citation Networks Visualizations}

\begin{document}

\section{Visualizations}
\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\textwidth]{{citations1}.pdf}
		\caption{Supreme Court Citation Network, 1937--2002. Nodes are supreme court cases, with size based on incoming citations. Salient  cases (\textit{Oxford Guide}) are in yellow. Each network layer contains cases decided under a different chief justice. Because our data is temporally bound, cases for the Hughes and Rehnquist courts are not complete. Edges between two nodes indicate the later document cites the earlier one.}
	\end{center}
\end{figure}

\newpage

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=0.6\textwidth]{{citations2}.pdf}
		\caption{Supreme Court Citation Network, 1937--2002. Nodes are supreme court cases, with size based on incoming citations. Salient  cases (\textit{Oxford Guide}) are in yellow. Each network layer contains cases decided in a given year; the top layer is 2002 and the bottom layer is 1937. Edges between two nodes indicate that the later document cites the earlier one.}
	\end{center}
\end{figure}

\newpage
\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{{citations3}.pdf}
		\includegraphics[width=0.49\textwidth]{{citations4}.pdf}
		\caption{Supreme Court Citation Network, 1985--1989. Nodes are supreme court cases, with size based on incoming citations. Important cases are in yellow. On the left, each network layer contains cases decided in a given year; the top layer is 1989 and the bottom layer is 1985. On the right, the five layers are collapsed. Edges between two nodes indicate that the later document cites the earlier one.}
	\end{center}
\end{figure}

\end{document}

### 2018-01-02 Tex/paper.tex
revising citation lit section

@@
 Researchers from several fields have used network analysis to analyze citations---legal, patent, and scientific.
-Before describing the network model for Supreme Court citations that we develop, we review the methods researchers have previously used to study citations..
-A citation is a directed link between two documents that indicates the citing document attributes the cited to be in some manner relevant to it.
-Collective patterns of citations within a certain domain make up citation networks, which are most commonly found in three social systems, scientific publishing, patents, and the judicial system.
-When raw citation data is formally represented as a network, the nodes are usually the documents themselves, and each directed arc, usually unweighted, is the existence of one or more cites from the sending to the receiving document.
-Given the permanent nature of documents in citations networks, these networks are in large acyclic \citep{leicht2007large,karrer2009random}.
+Before describing our approach to modeling the Supreme Court citations, we review the methods researchers have previously used to study citations.
+A citation is a directed link between two documents that indicates the citing document attributes the cited to be relevant to the evidentiary basis of the citing document.
+Collective patterns of citations within a certain domain make up citation networks.
+When raw citation data is formally represented as a network, the nodes are usually the documents themselves, and each directed arc is the existence of one or more cites from the sending to the receiving document.
+Given the permanent nature of documents in citations networks, these networks are acyclic \citep{leicht2007large,karrer2009random}.
 Relatedly, citation networks grow as new documents and citations enter the network, but established arcs persist in all but exceptional cases.
@@
 	In its raw form a citation is a link between two documents, but it is the result of human judgment about the utility of the cite \citep{fowler2008authority}.
-Studies of citation networks can be broadly categorized according to whether they treat this set of observed behavioural patterns as information that can be used to understand different aspects of the system, or treat them as outcomes to be modeled.
+Studies of citation networks can be broadly categorized according to whether they treat this set of observed behavioral patterns as information that can be used to understand different aspects of the system, or treat them as outcomes to be modeled.
 We discuss these in turn.

### 2018-01-02 Tex/paper.tex
revising theory section

@@
 
-\textbf{Transitivity:} In a network of directed relations (e.g., A cites B, but B doesn't cite A) transitivity refers to the tendency for A to send a tie to C if A sends a tie to B and B sends a tie to C [CITE FROM SYLLABUS].
+\textbf{Transitivity:} In a network of directed relations (e.g., A cites B, but B doesn't cite A) transitivity refers to the tendency for A to send a tie to C if A sends a tie to B and B sends a tie to C \citep{holland1971transitivity}.
 In undirected networks, transitivity is simply the process by which friends of friends become friends (i.e., a friend of a friend is a friend).
@@
 
-The statute thus addresses the risk of a morally unjustifiable death sentence, not by minimizing it as precedent unmistakably requires, but by guaranteeing that in equipoise cases the risk will be realized, by ?placing a ?thumb [on] death?s side of the scale,?
-?
-Sochor v.
+In the first example, a passage from Kansas v.
+Marsh (548 U.S.
+163, 2006)---a case considering the constitutionality of a death sentence statute in Kansas.
+In this example, the case Stringer v.
+Black is cited by Kansas v.
+Marsh as a case that is quoted by Sochor v.
+Florida.
+The primary precedent under discussion in this passage of the opinion is Sochor v.
+Florida, but Stringer v.
+Black is cited as a result of its role in the Sochor v.
+Florida opinion.
+\begin{quotation}
+The statute thus addresses the risk of a morally unjustifiable death sentence, not by minimizing it as precedent unmistakably requires, but by guaranteeing that in equipoise cases the risk will be realized, by ``placing a `thumb [on] death's side of the scale,' '' Sochor v.
 Florida, 504 U.
@@
 222, 232 (1992); alteration in original).
-
-https://supreme.justia.com/cases/federal/us/548/163/dissent2.html
-
-``Never before the decision in Union Gas had we suggested that the bounds of Article III could be expanded by Congress operating pursuant to any constitutional provision other than the Fourteenth Amendment.
+\end{quotation} %https://supreme.justia.com/cases/federal/us/548/163/dissent2.html
+The second example is a passage from Seminole Tribe of Fla. v. Florida, 
+(517 U.S. 44 1996)---a case addressing the rights of groups and citizens to sue states in federal court. In this example, Pennsylvania v. Union Gas Co
+(491 U.S. 1, 1989) is the primary precedent under discussion, and several cases are cited and discussed in terms of their roles as precedents in the Union Gas opinion. 
+\begin{quotation}
+Never before the decision in Union Gas had we suggested that the bounds of Article III could be expanded by Congress operating pursuant to any constitutional provision other than the Fourteenth Amendment.
 Indeed, it had seemed fundamental that Congress could not expand the jurisdiction of the federal courts beyond the bounds of Article III.
@@
 Y., 470 U.
-S., at 252).''
+S., at 252).'
+\end{quotation}% https://supreme.justia.com/cases/federal/us/517/44/case.html
 
-https://supreme.justia.com/cases/federal/us/517/44/case.html
-
-\textbf{Reciprocity:} Reciprocity (also referred to as mutuality) is the tendency for node B to send a tie to node A in response to or coordination with A sending a tie to B [CITE FROM SYLLABUS].
+\noindent \textbf{Reciprocity:} Reciprocity (also referred to as mutuality) is the tendency for node B to send a tie to node A in response to or coordination with A sending a tie to B \citep{garlaschelli2004patterns}.
 It is typically not possible for reciprocated ties to form in legal opinions.
@@
 
-On a more specific level, Western argues that flight engineers must meet the same stringent qualifications as pilots, and that it was therefore quite logical to extend to flight engineers the FAA's age 60 retirement rule for pilots.
+The first case in our example reciprocal dyad is a passage from Western Air Lines v.
+Criswell (472 U.S.
+400, 1985)---a case considering mandatory retirement in the context of age discrimination laws.
+The second case in the dyad, Johnson v.
+Mayor of Baltimore (472 U.S.
+353, 1985) is another case considering whether mandatory retirement violates the Age Discrimination in Employment Act.
+These cases addressed very similar legal questions, which increased the likelihood that they would inform each other, and the opinions were written within the same term, which made it possible for them to cite each other.
+\begin{quotation}
+{\em From Western Air Lines:} On a more specific level, Western argues that flight engineers must meet the same stringent qualifications as pilots, and that it was therefore quite logical to extend to flight engineers the FAA's age 60 retirement rule for pilots.
 Although the FAA's rule for pilots, adopted for safety reasons, is relevant evidence in the airline's BFOQ defense, it is not to be accorded conclusive weight.
@@
 In this case, the evidence clearly established that the FAA, Western, and other airlines all recognized that the qualifications for a flight engineer were less rigorous than those required for a pilot.
-[Footnote 28]
-
-https://supreme.justia.com/cases/federal/us/472/400/case.html
-
-The city, supported by several amici, argues for affirmance nonetheless.
+\\~\\
+{\em From Johnson:} The city, supported by several amici, argues for affirmance nonetheless.
 It asserts first that the federal civil service statute is not just a federal retirement provision unrelated to the ADEA, but in fact establishes age as a BFOQ for federal firefighters based on factors that properly go into that determination under the ADEA, see Western Air Lines, Inc.
@@
 Second, the city asserts, a congressional finding that age is a BFOQ for a certain occupation is dispositive of that determination with respect to nonfederal employees in that occupation.
+\end{quotation} % https://supreme.justia.com/cases/federal/us/472/400/case.html \url{https://supreme.justia.com/cases/federal/us/472/353/case.html#370}
 
-\url{https://supreme.justia.com/cases/federal/us/472/353/case.html#370}
-
-\textbf{Popularity} Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent [CITE FROM SYLLABUS].
+\noindent \textbf{Popularity} Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent \citep{barabasi1999emergence}.
 Citations to an opinion signal both the Court's awareness of the legal reasoning of the case and the Court's evaluation that the opinion is an authoritative precedent.
@@
 Landmark cases, or those that establish new legal rules, are particularly authoritative and accrue citations from most future opinions that follow the respective line of reasoning.
-The passage below, from XXXX, in which an authoritative opinion is referenced, and even discussed in terms of the number of other cases by which it was followed.
-We consider each of these contentions in turn.
-Based on this popularity phenomenon, we expect that new citations will be disproportionately directed at cases to which many past citations have accrued.
-
-"The first case in which this Court struck down a statute under the Equal Protection Clause of the Fourteenth Amendment was Strauder v.
+The passage below, from Oregon v.
+Mitchell (400 U.S.
+112, 1970)---a case on the legality of state age restrictions on voting in federal elections---illustrates this popularity dynamic.
+In this opinion passage Baker v.
+Carr is cited in reference to its role as a landmark precedent, and noted for the number of other cases by which it has been followed.
+which an authoritative opinion is referenced, and even discussed in terms of the number of other cases by which it was followed.
+\begin{quotation}
+The first case in which this Court struck down a statute under the Equal Protection Clause of the Fourteenth Amendment was Strauder v.
 West Virginia, 100 U.
@@
 That case was followed by numerous others, e.g.: that one person could not be given twice or 10 times the voting power of another person in a state-wide election merely because he lived in a rural area..."
+\end{quotation} %https://supreme.justia.com/cases/federal/us/400/112/case.html
 
-https://supreme.justia.com/cases/federal/us/400/112/case.html
-
-\textbf{Sociality} Sociality is the tendency for ties sent to beget more ties sent [CITE].
+\textbf{Activity} Activity, or sender activation, is the tendency for ties sent to beget more ties sent \citep{howard2016understanding}.
 Structurally, this means that there are few nodes that send a moderate number of ties---the number of ties sent is either small or fairly large.
@@
 Justices often clarify not only those legal principles that apply, but often those that do not.
-The passages below represent examples of this process of citation proliferation, or sociality.
+The passages below represent examples of this process of citation proliferation, or activation.
 
@@
 
-	In its raw form a citation is a link between two documents, but it is the result of human judgment about the utility of the cite \citep{fowler2008authority}.
+[HERE]	In its raw form a citation is a link between two documents, but it is the result of human judgment about the utility of the cite \citep{fowler2008authority}.
 Studies of citation networks can be broadly categorized according to whether they treat this set of observed behavioral patterns as information that can be used to understand different aspects of the system, or treat them as outcomes to be modeled.
@@
 The number of these configurations grows quadratically as the origins of ties concentrate on a few highly active/social senders.
-The out-two-stars configuration is commonly used to model sociality in ERGMs.
+The out-two-stars configuration is commonly used to model activity (sender activation) in ERGMs.
 $$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C \to \sum_{j=1}^N {\sum_{i \neq j} c_{ij} \choose 2} $$
@@
 
-    (g) Sociality & (h) Popularity \\
+    (g) Activity & (h) Popularity \\
 \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend.pdf} \\
@@
 This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.
+
+\section{conclusion}
+
+MAJOR LIMITATION---DO NOT CONSIDER SIGN OF CASES
 

### 2018-01-02 Tex/paper.tex
revising descriptive section

@@
 
-\textbf{Activity} Activity, or sender activation, is the tendency for ties sent to beget more ties sent \citep{howard2016understanding}.
+\noindent \textbf{Activity} Activity, or sender activation, is the tendency for ties sent to beget more ties sent \citep{howard2016understanding}.
 Structurally, this means that there are few nodes that send a moderate number of ties---the number of ties sent is either small or fairly large.
@@
 Justices often clarify not only those legal principles that apply, but often those that do not.
-The passages below represent examples of this process of citation proliferation, or activation.
-
-"1.
-The "overbreadth" doctrine is not applicable here.
-There is nothing in the record to indicate that \S 28.04 will have any different impact on any third parties' interests in free speech than it has on appellees' interests, and appellees have failed to identify any significant difference between their claim that \S 28.04 is invalid on overbreadth grounds and their claim that it is unconstitutional when applied to their signs during a political campaign.
-Thus, it is inappropriate to entertain an overbreadth challenge to \S 28.04.
-Pp.
-466 U.
-S.
-796-803."
-
-https://supreme.justia.com/cases/federal/us/466/789/
-
-''No allegations of racial or class-based invidiously discriminatory animus are required to establish a cause of action under the first part of  1985(2).
+The example passage that illustrates activity is from Kush v.
+Rutledge (460 U.S.
+719, 1983)---a case addressing witness intimidation and due process protections.
+In the Kush opinion, Griffin v.
+Beckenridge (i.e., Griffin) is discussed at length in terms of its lack of applicability.
+It is common that opinions distinguish among a variety of legal rules in terms of their applicability to a case, which means that for each rule that is established in an applicable precedent, one or more need to be discussed in terms of why they do not apply.
+\begin{quotation}
+No allegations of racial or class-based invidiously discriminatory animus are required to establish a cause of action under the first part of  1985(2).
 The statutory provisions now codified at \S 1985 were originally enacted as \S 2 of the Civil Rights Act of 1871, and the substantive meaning of the 1871 Act has not been changed.
@@
 Thus, the reasoning of Griffin is not applicable here, and, given the structure of \S 2 of the 1871 Act, it is clear''
-
-https://supreme.justia.com/cases/federal/us/460/719/
+\end{quotation} % https://supreme.justia.com/cases/federal/us/460/719/
 
@@
 
+The dependence processes discussed in the previous section represent new theoretical claims regarding the factors that account for citation formation among US Supreme Court opinions.
+Before we can test these claims, we consider methods for analyzing citation data as a network.
 Researchers from several fields have used network analysis to analyze citations---legal, patent, and scientific.
-Before describing our approach to modeling the Supreme Court citations, we review the methods researchers have previously used to study citations.
+Before describing our approach to modeling Supreme Court citations, we review the methods researchers have previously used to study citations.
 A citation is a directed link between two documents that indicates the citing document attributes the cited to be relevant to the evidentiary basis of the citing document.
@@
 
-[HERE]	In its raw form a citation is a link between two documents, but it is the result of human judgment about the utility of the cite \citep{fowler2008authority}.
-Studies of citation networks can be broadly categorized according to whether they treat this set of observed behavioral patterns as information that can be used to understand different aspects of the system, or treat them as outcomes to be modeled.
-We discuss these in turn.
-
-	First, as these arcs are observed patterns of behaviour, they can be used to infer latent characteristics of nodes or the relations between nodes that produced them \citep{batagelj2003efficient}.
-Moreover, since each judgment is embedded in the broader network of citations, which influences the judgment beyond the characteristics of the two directly linked documents, inferences about the neighborhood of these two documents are possible as well.
-A common application of this general logic is to obtain local or global network indices, such as Kleinberg centrality at the node level \citep{kleinberg1999authoritative,fowler2007network} or global hierarchy \citep{mones2014universal}, then use them to draw conclusions about networks or compare networks across contexts such as time periods \citep[e.g.][]{vazquez2001statistics,fowler2008authority,greenberg2009citation,lupu2012precedent,lupu2013strategic,dawson2014current,jaffe2017patent}.
-
-	Still within the category of using observed network constellation to make inferences about components of the system, identification of communities using methods for clustering, such as modularity maximization \citep[e.g.][]{kajikawa2007creating,shibata2011detecting,chen2010community} and stochastic block modeling \citep{jo2009citing}; and subnetwork identification \citep[e.g.][]{batagelj2017emergence} have been used to analyze relevance of topics within an academic field or to determine trends in technological advancement \citep{verspagen2007mapping,erdi2013prediction}.
-Main path analysis \citep{hummon1989connectivity}, which determines the main path through an acyclic network, is a particularly useful way to examine trends as it allows researchers identify structures of knowledge flow.
-Other methods include classifying documents into groups based on similarities in their citation profiles of cites over time, which can be examined as a function of time to see if there are temporal patterns \citep{leicht2007large}.
-The identification of communities is often coupled with classification of these communities by the researcher using historical knowledge.
-
-	Researchers are also interested in understanding the mechanisms that drive citations between documents.
-Work in this area generally proceed by specifying a set of citation behaviour, then proposing a model to capture the combination of these behavioural rules \citep{simkin2007mathematical}.
-Assessment of the model is based on how well it fit citation distributions.
-Researchers working in this area tend to focus on modeling the growth of the citation network as governed by a degree-based mechanism such as preferential attachment \citep{barabasi1999emergence} or the recursive search \citep{vazquez2001statistics}, and the age of the paper \citep{jeong2003measuring,eom2011characterizing,wang2013quantifying}.
-Regarding age, the previous standard approach was to treat the probability of citation as a separable function of degree and age \citep{hajra2005aging,hajra2006modelling}, then examine the distribution of citations, scaled to the rate of new publications, by age.
+Networks researchers have developed approaches to understanding the mechanisms that drive citations between documents.
+Work in this area generally proceeds by specifying a set of citation behaviors, then proposing a model to capture the combination of these behavioral rules \citep{simkin2007mathematical}.
+Assessment of the model is based on how well it fits citation distributions.
+Researchers working in this area tend to focus on modeling the growth of the citation network as governed by a degree-based mechanism such as preferential attachment \citep{barabasi1999emergence, vazquez2001statistics}, and the age of the paper \citep{jeong2003measuring,eom2011characterizing,wang2013quantifying}.
+Regarding age, the previous standard approach was to treat the probability of citation as function of degree (i.e., the number of citations previously accrued by the document) and age \citep{hajra2005aging,hajra2006modelling}, then examine the distribution of citations.
 More recently, work has been done in relaxing the assumption that the effect of degree is static, instead allowing it to vary with time \citep{wang2008measuring}.
@@
 Another approach to understanding the generative process of citation networks is to examine the existence of network motifs, or subnetwork structures, that can be interpreted as measuring different generative mechanisms.
-One way of doing so is to compare the observed network statistics to a null distribution.
+For example, a triangle in which case A cites both cases B and C, and case B cites case C, is a type of motif---a triadic motif.
+If researchers find that a triangle is a highly prevalent motif in a network, that finding would suggest that triad closure is a prevalent process by which the network is generated.
+The analysis of network motifs is done by comparing the prevalence of the motifs in the observed network to the prevalence of the motifs in networks drawn from a relevant null distribution.
 This null model must also be characterized by features of citation networks discussed earlier, meaning that it must be a directed acyclic graph with unweighted arcs \citep{carstens2016topology}.
-Two sets of uniform random graph models have been proposed for directed acyclic graphs.
-In the former, \citet{karrer2009random} introduce two uniform random graph models based respectively on fixed-degree sequence and fixed-expected degree sequence.
-Both models preserve the requisite characteristics of citation networks with the exception that they can introduce weighted edges with low probability, which leads to networks from the null distribution to be sampled with nonequal probability.
-\citet{carstens2014uniform,carstens2016topology} introduces a model that preserves the requisite characteristics and also do not introduce weighted arcs.
-Loosely related is an egocentric framework, introduced by \citet{vu2011dynamic}, that models continuous-time network data through a multivariate counting process, which can be used to count network motifs.
+\citet{karrer2009random} proposes the use of null models based on fixed-degree sequences.
+In these models, the nodes in the networks (i.e., opinions) have the same, or close to the same, numbers of ties.
+Conditional on the number of ties accrued by each node, ties are randomly re-wired such that there is no systematic pattern regarding the nodes to which any node connects.
 
-	\section{Extra Notes}
-	This section contains some of the extra notes I cut out from the draft that might be useful.
-	\subsection{Broader definitions of citation networks}
-	More broadly, citation networks can have as nodes the document producers, including scientists \citep{ji2016coauthorship}, judges \citep{landes1998judicial}, journals \citep{rice1988citation}, or aggregated units \citep{gelter2012networks,pan2012world}.
-In these cases, which are essentially exercises in aggregation, the resulting network can take on different characteristics such as weighted or cyclic edges.
-Other networks derived from the raw citation network is the co-citation or co-cited-by networks whereby two documents are connected if they co-cite or are co-cited by another document \citep{van2005reference}.
-
-	\subsection{Weighted edges and multiplexity}
-	Recognition, and subsequent quantification, of different levels of complexity in citation arcs allows for a host of advancements in citation network analysis.
-Weights can indicate relevance \citep{liu2014citations} or be used to incorporate the temporal dimension into citation networks \citep{fujita2014detecting}.
-Moving to multiplexity leads to further advancements.
-For example, \citet{greenberg2009citation} accounted for level of support and through simulations based on arc typed switching demonstrated how citation bias against critical articles yielded differences in beliefs within a medical discipline.
-\citet{bommarito2010distance} argued that articles contain different types of information and accounting for the specific piece of information that resulted in the cite can yield clustering that are more interpretable.
-In all cases however, substantial work has to be done into reclassifying the weightedness or multiplexity of citation arcs \citep{zhang2007semantics}.
-
-	\subsection{Misc. notes}
-	\begin{itemize}
-		\item To control for age, \citet{clough2015transitive} propose transitive reduction of citation networks (i.e.
-removal of redundant information ties) as it will primarily remove citation arcs that are disparate in age.
-		\item Most citation networks are directed acyclic graphs, but not all.
-Different versions of the same document can also cause problems (such as strong network components).
-The ``preprint transformation'' is a solution to small strong components \citep{batagelj2017emergence}.
-		\item Date of publication can be assigned as level to nodes in acyclic networks \citep{batagelj2017emergence}.
-Depending on the citation network at hand, temporal ordering can be difficult, as documents can share publication dates.
-To overcome this, begin each iteration by sampling from one of the possible orderings.
-\citep{carstens2016topology}.
-	\end{itemize}
+From this brief review we see two broad components of past methods for analyzing citation networks.
+First, node features, age in particular, are used to model the rate at which a document is cited.
+Second, theoretical models are used to account for prevalent graph structures.
+In the next section we describe a methodology that can be used to integrate node attributes and graph structures/motifs into a comprehensive model of US Supreme Court citations.
 
@@
 However, as noted above, if two opinions are written in the same term, they can cite each other reciprocally.
-We implement a version of the temporal exponential random graph model (TERGM)---a model that is commonly used for longitudinal network data [CITES]---that is subject to the constraints specific to the Supreme Court citation network.
+We implement a version of the temporal exponential random graph model (TERGM)---a model that is commonly used for longitudinal network data \citep[e.g.,][]{cranmer2012complex,desmarais2013forecasting,clark2013multimember,block2018change,graif2017neighborhood}---that is subject to the constraints specific to the Supreme Court citation network.
 The TERGM is a model that can simultaneously incorporate structural dependencies such as transitivity and reciprocity and exogenous covariates  (e.g.
 the issue area of the citing case, the ideological distance between the opinion authors of the two cases in the dyad) to explain tie formation.
-If the ties do not depend upon each other, and the coefficients associated with the structural dependencies are either assumed or estimated to be zero, the TERGM reduces to a logistic regression in which the observation is a directed dyad [CITE OLD PA].
+If the ties do not depend upon each other, and the coefficients associated with the structural dependencies are either assumed or estimated to be zero, the TERGM reduces to a logistic regression in which the observation is a directed dyad \citep{cranmer2010inferential}.
 In this section we describe the citation TERGM (c-TERGM), and explain estimation methods.
@@
 Loops can exist in the Supreme Court citation network, but only among cases decided in the same term.
-We estimate the c-TERGM via bootstrapped pseudolikelihood, as introduced by [CITE PHYSICA].
+The normalizing constant in the denominator of the probability function (i.e., Equation \ref{ercm}) is computationally intractable.
+This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with ERGM family models.
+Methods of Monte Carlo approximation are well established \citep{hunter2006inference,van2009framework,hummel2012improving}, but with over 8,000 vertices, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}.
+To estimate the parameters of the c-TERGM, we rely on two features of the model and data that facilitate the application of an approximate inference method that is faster than Monte Carlo approximation---maximum pseudolikelihood.
+First, our sample size is very large.
+Second, under the c-TERGM we assume that citations are formed in each time period conditional upon the citations that exist in the past.
+We estimate the c-TERGM via bootstrapped pseudolikelihood \citep{desmarais2012statistical,desmarais2010consistent}.
 
@@
 They find that the majority opinion falls at the ideal point of the median member of the majority coalition in the case.
-We include a covariate term in which $X_{ij}$ is the absolute difference between the Martin-Quinn [CITE MQ] scores of the median justices in the majority coalitions for cases $i$ and $j$.
+We include a covariate term in which $X_{ij}$ is the absolute difference between the Martin-Quinn scores \citep{martin2002dynamic} of the median justices in the majority coalitions for cases $i$ and $j$.
 We expect this variable to have a negative effect, which would correspond to cases citing those to which they are ideologically similar.%WE COULD USE THIS DATASET AS A SECOND APPLICATION
@@
 In the set of receiver intercepts $X_{ij}$ indicates the issue are of the receiver case.
-Issue area data comes from the Supreme Court Database (SCDB) [CITE SCDB].
+Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
 We include these variables because \citet{cross2010determinants} find that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
@@
 They find that cases with majority coalitions that are large and ideologically broad are less likely to be cited negatively.
-In our data we do not differentiate between negative and positive citaitons, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis.
+In our data we do not differentiate between negative and positive citations, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis.
 We include one covariate in which $X_{ij}$ is the number of justices in the majority coalition for case $j$.
@@
 
- \section{Results}
-  \subsection*{Descriptive Results}
-The supreme court citation network from $1937 - 2005$ consists of $8817$ cases which got voted at $2116$ different time points.
-The network has a total of $93,263$ ties, of which 452 are mutual.
+ \section{Empirical Analysis}
+Our three data sources for this study include the SCDB, Martin-Quinn scores, and Supreme Court citation data provided by James Fowler and colleagues \citep{fowler2007network,fowler2008authority}.
+We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2001).
+The Supreme Court citation network from $1937 - 2001$ consists of $8,817$.
+The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
+The network has a total of $93,263$ ties.
 The number of triangles in the network is $211,855$.
-The in- and outdegree distribution is visualized in figures \ref{indegree_dist} and \ref{outdegree_dist}.
+The in- and outdegree distributions (i.e., the distributions of the number of citations sent and received by cases, respectively), are visualized in figures \ref{indegree_dist} and \ref{outdegree_dist}.
 The maximum indegree is $190$ and the maximum outdegree is $159$.
 
-\begin{table}[H]
+\begin{table}[htp]
 \centering
@@
 \\\hspace{\textwidth} ** W Rehnquist served as chief justice from 1987 - 2005.}
-\label{my-label}
+\label{tab:chiefs}
 \end{table}
 
-\begin{figure}[H]
+The degree distributions indicate that there is a long tail to both the number of citations sent and received.
+These long-tailed distributions provide preliminary evidence of both popularity and activity dynamics.
+Figure \ref{fig:networkviz} displays the complete network.
+We see here that the densest rates of tie formation tend to be between consecutive courts (e.g., the Stone Court is much more tightly tied to the Hughes Court than the Rehnquist Court is to the Hughes Court).
+This pattern lends preliminary/descriptive support to the hypothesis that the rate of citations to a case decrease over time.
+
+\begin{figure}[htp]
 \includegraphics[scale=0.5]{degree_distribution}
@@
 
-\begin{figure}[H]
-\includegraphics[scale=0.5]{number_cases}
-\caption{Number of cases in each term. Different colors indicate different chief justices.}
- \label{number_cases}
+\begin{figure}[htp]
+\includegraphics[scale=0.35,clip=true,trim=.5cm 0cm 0cm 2cm]{images/citations1}
+\caption{ Supreme Court Citation Network, 1937?2001.
+Nodes are Supreme Court cases, with size based on incoming citations.
+Salient cases ({\em Oxford Guide}) are in yellow.
+Each network layer contains cases decided under a different chief justice.
+Because our data is temporally bound, cases for the Hughes and Rehnquist courts are not complete.
+}
+ \label{fig:networkviz}
 \vspace{-.25cm}
@@
 
- \begin{figure}[H]
-\includegraphics[scale=0.5]{number_citations}
-\caption{Number of citations for the 1937-2001 time period. Citations for cases prior 1937 are not considered in this figure. Different colors indicate different chief justices.}
- \label{number_citations}
-\vspace{-.25cm}
-\end{figure}
+\subsection{c-TERGM Results}
 
-  \begin{figure}[H]
-\includegraphics[scale=0.5]{number_votes_supporting}
-\caption{Number of Votes that were supporting cases between 1937-2001. Different colors indicate terms with different chief justices. }
- \label{number_supporting}
-\vspace{-.25cm}
-\end{figure}
+INTERPRETATION OF EFFECTS
 
-\newpage
-
-\subsection{Inferential Results}
-
-\textbf{Model updates}
-\begin{itemize}
-\item Add sender issue area and receiver issue area node covariates (i.e., nodeifactor and nodeofactor applied to issue area). 
-\item In the GLM, multiply each variable by the sender time covariate.
-\item Write the bootstrap coefficients to a file, send to Bruce so he can write up a code for summarizing the over-time trends in effects.
-\item Add a receiver variable (i.e., nodeicov) that equals the absolute difference between the maximum MQ score of a justice in the majority and a minimum MQ score of a justice in the majority.
-\item Add a receiver variable (nodeicov) equivalent to the number of justices in the majority
-
-\end{itemize}
-
-\begin{table}[H]
+\begin{table}[htp]
 \footnotesize
@@
 The first $B \ll L$ vectors are discarded as Burn-In.
+
+First, as these arcs are observed patterns of behaviour, they can be used to infer latent characteristics of nodes or the relations between nodes that produced them \citep{batagelj2003efficient}.
+Moreover, since each judgment is embedded in the broader network of citations, which influences the judgment beyond the characteristics of the two directly linked documents, inferences about the neighborhood of these two documents are possible as well.
+A common application of this general logic is to obtain local or global network indices, such as Kleinberg centrality at the node level \citep{kleinberg1999authoritative,fowler2007network} or global hierarchy \citep{mones2014universal}, then use them to draw conclusions about networks or compare networks across contexts such as time periods \citep[e.g.][]{vazquez2001statistics,fowler2008authority,greenberg2009citation,lupu2012precedent,lupu2013strategic,dawson2014current,jaffe2017patent}.
+
+	Still within the category of using observed network constellation to make inferences about components of the system, identification of communities using methods for clustering, such as modularity maximization \citep[e.g.][]{kajikawa2007creating,shibata2011detecting,chen2010community} and stochastic block modeling \citep{jo2009citing}; and subnetwork identification \citep[e.g.][]{batagelj2017emergence} have been used to analyze relevance of topics within an academic field or to determine trends in technological advancement \citep{verspagen2007mapping,erdi2013prediction}.
+Main path analysis \citep{hummon1989connectivity}, which determines the main path through an acyclic network, is a particularly useful way to examine trends as it allows researchers identify structures of knowledge flow.
+Other methods include classifying documents into groups based on similarities in their citation profiles of cites over time, which can be examined as a function of time to see if there are temporal patterns \citep{leicht2007large}.
+The identification of communities is often coupled with classification of these communities by the researcher using historical knowledge.
+
+	\section{Extra Notes}
+	This section contains some of the extra notes I cut out from the draft that might be useful.
+	\subsection{Broader definitions of citation networks}
+	More broadly, citation networks can have as nodes the document producers, including scientists \citep{ji2016coauthorship}, judges \citep{landes1998judicial}, journals \citep{rice1988citation}, or aggregated units \citep{gelter2012networks,pan2012world}.
+In these cases, which are essentially exercises in aggregation, the resulting network can take on different characteristics such as weighted or cyclic edges.
+Other networks derived from the raw citation network is the co-citation or co-cited-by networks whereby two documents are connected if they co-cite or are co-cited by another document \citep{van2005reference}.
+
+	\subsection{Weighted edges and multiplexity}
+	Recognition, and subsequent quantification, of different levels of complexity in citation arcs allows for a host of advancements in citation network analysis.
+Weights can indicate relevance \citep{liu2014citations} or be used to incorporate the temporal dimension into citation networks \citep{fujita2014detecting}.
+Moving to multiplexity leads to further advancements.
+For example, \citet{greenberg2009citation} accounted for level of support and through simulations based on arc typed switching demonstrated how citation bias against critical articles yielded differences in beliefs within a medical discipline.
+\citet{bommarito2010distance} argued that articles contain different types of information and accounting for the specific piece of information that resulted in the cite can yield clustering that are more interpretable.
+In all cases however, substantial work has to be done into reclassifying the weightedness or multiplexity of citation arcs \citep{zhang2007semantics}.
+
+	\subsection{Misc. notes}
+	\begin{itemize}
+		\item To control for age, \citet{clough2015transitive} propose transitive reduction of citation networks (i.e.
+removal of redundant information ties) as it will primarily remove citation arcs that are disparate in age.
+		\item Most citation networks are directed acyclic graphs, but not all.
+Different versions of the same document can also cause problems (such as strong network components).
+The ``preprint transformation'' is a solution to small strong components \citep{batagelj2017emergence}.
+		\item Date of publication can be assigned as level to nodes in acyclic networks \citep{batagelj2017emergence}.
+Depending on the citation network at hand, temporal ordering can be difficult, as documents can share publication dates.
+To overcome this, begin each iteration by sampling from one of the possible orderings.
+\citep{carstens2016topology}.
+	\end{itemize}
+
+	\begin{figure}[H]
+\includegraphics[scale=0.5]{number_cases}
+\caption{Number of cases in each term. Different colors indicate different chief justices.}
+ \label{number_cases}
+\vspace{-.25cm}
+\end{figure}
+
+ \begin{figure}[H]
+\includegraphics[scale=0.5]{number_citations}
+\caption{Number of citations for the 1937-2001 time period. Citations for cases prior 1937 are not considered in this figure. Different colors indicate different chief justices.}
+ \label{number_citations}
+\vspace{-.25cm}
+\end{figure}
+
+  \begin{figure}[H]
+\includegraphics[scale=0.5]{number_votes_supporting}
+\caption{Number of Votes that were supporting cases between 1937-2001. Different colors indicate terms with different chief justices. }
+ \label{number_supporting}
+\vspace{-.25cm}
+\end{figure}
+
+\newpage

### 2018-01-02 Tex/paper.tex
revised results section

@@
 \subsection{c-TERGM Results}
+The results from the c-TERGM are presented in Table \ref{tab:tergmcoef} and Figure \ref{fig:coeftrends}.
+Table \ref{tab:tergmcoef} gives the coefficient estimates and 95\% bootstrap confidence intervals.
+In Figure, \ref{fig:coeftrends}, we depict coefficient estimates and confidence intervals for all of the effects for which we estimated trends (excluding the intercept---edge count---which is typically not interpreted substantively).
+The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one unit increase in the respective variable.
+We first discuss the dependence effect estimates (panels (e)--(h) in Figure \ref{fig:coeftrends}).
+The reciprocity effect is substantial, positive, and statistically significant for the entire period under study.
+Conditional upon cases $i$ and $j$ being considered within the same term, the log of the odds that opinion $i$ cites opinion $j$ increases by approximately 3 when opinion $j$ cites opinion $i$.
+The naive probability that any one case cites another case decided at the same time or previously in the Supreme Court Citation network is approximately 0.0024.
+This probability increases twenty-fold, to approximately 0.046 if the citation would be reciprocal.
+The transitivity (i.e., triangle closure) effect is also substantial, positive, and statistically significant over the entire period studied.
+The probability that case $i$ cites case $j$ increases approximately five-fold if the citation would close a triangle involving cases $i$ and $j$.
+The activity and popularity effects are both estimated to be positive, but trend downward and are more modest in scale (and in the case of activity, not statistically significant later in the time series).
+Interpreting the popularity effect---the log odds that $i$ cites $j$ increases by 0.015--0.030 with every additional citation to case $j$.
+Based on the in-degree distribution, it appears that the popularity of cases can be classified by multiples of 10.
+If case $j$ increases from ten to twenty citations from other cases, the probability that case $i$ cites case $j$ increases goes from approximately 0.0030 to 0.0040.
+Overall, these results represent evidence of complex dependence in Supreme Court citation formation.
 
-INTERPRETATION OF EFFECTS
+The effects of the exogenous covariates included in the c-TERGM (panels (a)--(d) in Figure \ref{fig:coeftrends}) exhibit some surprising dynamics.
+The effects of absolute difference in Martin-Quinn score, ideological breadth of the recipient case, and majority size of the recipient case all flip signs throughout the time period under study.
+This may be an artifact of the simple linear trends we have estimated, but is interesting nonetheless.
+The attitudinal model holds that ideological preferences drive decision-making on the US Supreme Court.
+Ideological selection of precedents would be evidenced by a negative effect of the absolute difference in MQ scores in the c-TERGM.
+We find this effect is negative and statistically significant only after 1975.
+The absolute difference in MQ scores variable is on the scale of approximately 0---3.
+At its most ideological point (i.e.
+2001), the probability that case $i$ cites case $j$ decreases from approximately 0.024 to 0.018---by one third---if the ideological distance between the majority coalitions in the two cases increases by 3.
+The effects of the majority size and ideological breadth of case $j$ are similar in both scale and trend.
+It is not until later in the time series that they both exhibit their expected positive effects.
+The effect of same issue is positive and statistically significant, and substantial for the time period under study.
+The probability that $i$ cites $j$ increases by approximately three-fold over the naive probability of citation if $i$ and $j$ have the same legal issue area.
+
+The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes.
+The scales of covariate and dependence effects are comparable.
+The simple linear trends represent one limitation of the current analysis.
+In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.
 
@@
 A '*' indicates that the 2.5th and 97.5th quantile of the variable does not include '0' and as a result is statistically significant.}
-\label{my-label}
+\label{tab:tergmcoef}
 \end{table}
@@
 \end{tabular}
-\caption{Effects}
+\caption{Trending effects from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
  \label{fig:coeftrends}
@@
 
-Our case for studying legal citaions at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence.
-To quantify this contribution, we use out-of-sample prediction [CITES].
-Predicting out-of-sample offers an unbiased and general purpose way to evaluate the contribution, in terms of model fit, of one or more terms/parameters in a model.
+Our case for studying legal citations at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence.
+To quantify this contribution, we use out-of-sample prediction.
+Predicting out-of-sample offers an unbiased and general purpose way to evaluate the contribution, in terms of model fit, of one or more terms/parameters in a model \citep{jensen2000multiple,ward2010perils}.
 Unlike in-sample measures of model fit, out-of-sample methods are highly robust in avoiding overfitting, and work when we cannot accurately calculate the value of the likelihood function, as in the current case.
-Out-of-sample prediction is a common way to evaluate methods for modeling ties in networks, and has recently been applied to TERGMs in particular [CITE PA AND IEEE].
+Out-of-sample prediction is a common way to evaluate methods for modeling ties in networks, and has recently been applied to TERGMs in particular \citep{desmarais2013forecasting,cranmer2017can}.
 
@@
 We run this experiment for 10 iterations.
-Predictive performance is evaluated with three common and related measures---precision (i.e., the proportion of predicted citations that are actually citations), recall (the proportion of actual citations that are predicted to be citations), and the F1 score (the harmonic mean of precision and recall).
+Predictive performance is evaluated with three common and related measures---precision (i.e., the proportion of predicted citations that are actually citations), recall (the proportion of actual citations that are predicted to be citations), and the F1 score (the harmonic mean of precision and recall) \citep[see, e.g.,][for discussion of these measures]{makhoul1999performance}.
 All three measures are bounded between 0 and 1, with higher scores indicating better performance.
@@
 
-\section{conclusion}
+\section{Conclusion}
 

### 2018-01-02 Tex/paper.tex
added short conclusion

@@
 
-MAJOR LIMITATION---DO NOT CONSIDER SIGN OF CASES
+We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network.
+This methodology---the citation-TERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation.
+We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2001.
+We find, somewhat counterintuitively, that Supreme Court citations are highly reciprocal.
+We also find that citations are driven by dependencies such as triad closure, popularity, and activity.
+The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model.
+The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.
+
+We recognize several limitations of the current analysis, which should be addressed in future iterations of this work or in future studies.
+First, our theoretical claims and analyses treat all citations---positive and negative---as equivalent.
+Both data and methodological limitations will need to be overcome to consider these two citation types.
+Second, there is nothing in the c-TERGM that governs the generation of new cases (i.e., cases are assumed to arise exogenously, unrelated to the history of the citation network).
+As new cases are often taken up by the Supreme Court in order to test, clarify, and/or challenge existing precedents, it is likely inappropriate to assume that new cases arise exogenous to the existing citation network.
+Overcoming this limitation would require methodological innovation in the c-TERGM as well as theoretical development regarding exactly how citation structure might predict case emergence.
+Third, the data used in this study does not cover the most recent history of the Court.
+Our data ends in 2001---missing the tenure of the most recent chief justice, and more than fifteen terms.
+To overcome this limitation we would need to extend the citation data beyond that collected by Fowler and colleagues.
 

### 2018-01-02 Tex/paper.tex
finished analysis

@@
 \usepackage[T1]{fontenc}
+\usepackage{authblk}
 \usepackage{amssymb}
@@
 \title{\Large Analyzing the Supreme Court Citation Network}
+%\author{}
+ % Toggle % below to blind and unblind the manuscript
+  \author[1]{Christian Schmid \thanks{cxs5700@psu.edu}}
+   \author[2]{Ted Hsuan Yun Chen\thanks{thc126@psu.edu}}
+ \author[2]{Bruce A. Desmarais\thanks{bdesmarais@psu.edu}}
+ \author[1]{David R. Hunter \thanks{dhunter@stat.psu.edu}}
+ \affil[2]{Department of Political Science, Pennsylvania State University}
+ \affil[1]{Department of Statistics, Pennsylvania State University}
+
 \maketitle
-
 \begin{abstract}
-\noindent 
+\noindent The significance and influence of US Supreme Court majority opinions derive in large part to opinions' roles as precedents for future opinions.
+A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns.
+We raise two limitations of existing work on Supreme Court citaitons.
+First, dyadic citations are typically aggregated to the case level before they are analyzed.
+Second, citaitons are treated as if they arise independent of one another.
+We make the case that studying Supreme Court citations as a dynamic citation network overcomes both of these limitations.
+We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network.
+This methodology---the citation temporal exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation.
+We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2001.
+We find evidence for dependence processes, including reciprocity, transitivity, popularity, and activity.
+The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model.
+The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.\footnote{
+Prepared for presentation at the ``The Enemy of My Enemy: Judicial Politics meets Network Analysis'' conference within a conference at the 2018 Annual Conference of the Southern Political Science Association.
+This work was supported in part by NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219.}
 \end{abstract}
@@
 Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited by a case.
-We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics}.
+We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{harris1982structural,caldeira1988legal,fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics}.
 
@@
 We are aware of one prior study, \citet{clark2010locating}, in which a statistical model is used to analyze dyadic citations between cases.
-However, citet{clark2010locating} uses a dyadic latent variable model in order to estimate ideal points for Supreme Court Opinions, but does not use any explanatory variables to predict the formation of citation ties between opinions.
+However, \citet{clark2010locating} use a dyadic latent variable model in order to estimate ideal points for Supreme Court Opinions, but does not use any explanatory variables to predict the formation of citation ties between opinions.
 We build upon this literature both methodologically and substantively.
@@
 The formation (or lack thereof) of that relationship cannot be fully understood without considering how the relationship fits into the complex web.
-Analytical designs that account only for covariates in explaining tie formation are incomplete theoretically, and, as a consequence, are subject to a form of omitted variable bias [CITE ISQ].
+Analytical designs that account only for covariates in explaining tie formation are incomplete theoretically, and, as a consequence, are subject to a form of omitted variable bias \citep{cranmer2016critique}.
 Citations in legal opinions are unique in terms of the windows into network dependencies offered by the texts of the opinions.
 A number of common structural dependencies that are found in networks are likely to apply to citations in Supreme Court opinions.
-In this section we present these dependence forms, and document the mechanisms by which they arise through quotations in archetypal example opinions.
+In this section we present these dependence forms, and document the mechanisms by which they arise through archetypal passages in example opinions.
 

### 2018-01-03 Tex/presentations/SPSA2018/DesmaraisSPSATalk.tex
adding SPSA slides

\documentclass[handout]{beamer}
\usepackage{bm}
\usepackage{listings}
\usepackage{algorithmic}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{multirow}
\lstloadlanguages{R}
\lstset{ language=R, basicstyle=\scriptsize\ttfamily, commentstyle=\ttfamily\color{gray}, numbers=left, numberstyle=\ttfamily\color{gray}\footnotesize, stepnumber=1, numbersep=5pt, backgroundcolor=\color{white}, showspaces=false, showstringspaces=false, showtabs=false, frame=single, tabsize=2, captionpos=b, breaklines=true, breakatwhitespace=false, escapeinside={}, keywordstyle={}, morekeywords={} }\renewcommand{\P}{\mathcal{P}}
\newcommand{\bt}{\pmb{\theta}}
\newcommand{\bG}{\pmb{\Gamma}}
\newcommand{\p}{\pause}
\newcommand{\pitem}{\pause \item}
\newcommand{\N}{\mathcal{N}}
\newcommand{\Y}{\bm{\mathcal{Y}}}
\newcommand{\bh}{\bm{h}} 
\DeclareMathOperator*{\argmax}{arg\,max}

\usefonttheme{serif}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

%
\title{Inferential Analysis of the  \\ Supreme Court Citation Network}
%
\definecolor{umassred}{HTML}{001769}
\setbeamercolor{structure}{fg=umassred} 
\author{ \textbf{Bruce A. Desmarais}\\ Collaborators: Christian Schmid, Ted Chen, David Hunter \\~\\ {\bf Pennsylvania State University}}

%\affil[1]{University of Massachusetts Amherst}
%\affil[2]{University of Connecticut}

\begin{document}
%\date{November 16, 2012}

\begin{frame}
\vspace{1cm}
  \titlepage
  \vspace{0cm}
  \begin{center}
   \begin{tabular}{cc}
\hspace*{-.2in} \tiny \begin{minipage}{3.5in}
Work supported by NSF grants SES-1558661, SES-1619644, SES-1637089, and CISE-1320219)\\ ~\\~\\~\\~\\
\end{minipage}
& \includegraphics[scale=.05]{figures/NSF_logo.png}
\end{tabular}
\end{center}
\end{frame}

\begin{frame} \frametitle{Captivating Story/Anecdote}

\end{frame}

\begin{frame} \frametitle{Two Limitations in Case-Level Analysis}

\end{frame}

\begin{frame} \frametitle{Reciprocity}

\end{frame}

\begin{frame} \frametitle{Transitivity}

\end{frame}

\begin{frame} \frametitle{Popularity and Sociality}

\end{frame}

\begin{frame} \frametitle{Citation TERGM}

\end{frame}

\begin{frame} \frametitle{Model Specification}

\end{frame}

\begin{frame} \frametitle{Results Dependence}

\end{frame}

\begin{frame} \frametitle{Results Covariates}

\end{frame}

\begin{frame} \frametitle{Predictive Performance}

\end{frame}

\begin{frame} \frametitle{Conclusion}

\end{frame}

\end{document}

### 2018-01-03 NetworkVisualizations/visualizations.tex
updated slides

--- 2017-12-31 Network Visualizations/visualizations.tex
@@
 
+\newpage
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{{citations.trans}.pdf}
+		\caption{Illustration of transitivity in the Supreme Court citation network.
+This is a subnetwork consisting of cases 377 U.S.
+184, 491 U.S.
+1, and 517 U.S.
+44, highlighted in yellow; and the cases that cite and are cited by these three cases.
+The transitive structure formed by the three named cases is highlighted in pink.}
+	\end{center}
+\end{figure}
+
+\newpage
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{{citations.recip}.pdf}
+		\caption{Illustration of reciprocity in the Supreme Court citation network.
+This is a subnetwork consisting of cases 472 U.S.
+353 and 472 U.S.
+400, highlighted in yellow; and the cases that cite and are cited by these two cases.
+Reciprocal ties are highlighted in pink.}
+	\end{center}
+\end{figure}
+
+\newpage
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{{citations.pop}.pdf}
+		\caption{Illustration of popularity in the Supreme Court citation network.
+This is a subnetwork consisting of all cases contained in volume 369 of the U.S Supreme Court Reporter of Decisions, which are highlighted in yellow, and the cases that cite these cases.
+Case 369 U.S.
+186, discussed in the text, is highlighted in pink.}
+	\end{center}
+\end{figure}
+
+\newpage
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{{citations.soci}.pdf}
+		\caption{Illustration of reciprocity in the Supreme Court citation network.
+This is a subnetwork consisting of all cases contained in volume 466 of the U.S Supreme Court Reporter of Decisions, which are highlighted in yellow, and the cases that are cited by these cases.
+Case 466 U.S.
+789, discussed in the text, is highlighted in pink.}
+	\end{center}
+\end{figure}
+
 \end{document}

### 2018-01-03 Tex/presentations/SPSA2018/DesmaraisSPSATalk.tex
updated slides

@@
 \usepackage{bm}
+\usepackage{colortbl}
 \usepackage{listings}
@@
 \usepackage{rotating}
+\usepackage{booktabs}
 \usepackage{multirow}
@@
 
-%\affil[1]{University of Massachusetts Amherst}
-%\affil[2]{University of Connecticut}
-
 \begin{document}
@@
 \begin{frame} \frametitle{Two Limitations in Case-Level Analysis}
+\Large
+\begin{enumerate}
+\item {\bf Theoretical:} Cannot learn about relational processes
+\\~\\
+\item {\bf Methodological:} Risk bias from omitted structure
+\end{enumerate}
 
@@
 
+\centering
+
+\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{../../../NetworkVisualizations/citations_recip.pdf}
+
 \end{frame}
@@
 \begin{frame} \frametitle{Transitivity}
+\centering
+
+\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{../../../NetworkVisualizations/citations_trans.pdf}
 
@@
 
-\begin{frame} \frametitle{Popularity and Sociality}
+\begin{frame} \frametitle{Popularity and Activity}
+
+\centering
+
+\includegraphics[width = 0.95\textwidth,trim= 0cm 3cm 6cm 8cm,clip=true ]{../../../NetworkVisualizations/citations_pop.pdf}
 
@@
 
+\begin{frame} \frametitle{Data}
+
+\begin{columns}
+\begin{column}{4cm}
+\scalebox{0.8}{
+\begin{tabular}{|
+>{\columncolor[HTML]{C0C0C0}}l |l|l|}
+\hline
+{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Terms} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Cases} \\ \hline
+CE Hughes*              & 1937 - 1941                                          & 629                                                                                                                  \\ \hline
+HF Stone                & 1942 - 1946                                          & 766                                                                                                                \\ \hline
+FM Vinson               & 1946 - 1953                                          & 788                                                                                                                  \\ \hline
+E Warren                & 1954 - 1969                                          & 2159                                                                                                                  \\ \hline
+WE Burger               & 1970 - 1986                                          & 2805                                                                                                                   \\ \hline
+W Rehnquist**            & 1987 - 2001                                          & 1670                                                                                                                 \\ \hline
+\end{tabular}
+}
+\end{column}
+\begin{column}{6cm}
+\includegraphics[scale=0.15,clip=true,trim=.5cm 0cm 0cm 2cm]{../../../NetworkVisualizations/citations1}
+\end{column}
+
+\end{columns}
+
+\end{frame}
+
 \begin{frame} \frametitle{Model Specification}
@@
 
-\begin{frame} \frametitle{Results Dependence}
+\begin{frame} \frametitle{Results: Dependence}
+\begin{changemargin}{-1cm}{-1cm}
+\centering
+ \begin{tabular}{cc}
 
+    Reciprocity &  Transitivity \\
+\includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/mutual_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/triangle_coef_trend.pdf} \\
+
+     Activity & Popularity \\
+\includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/o2star_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/i2star_coef_trend.pdf} \\
+
+\end{tabular}
+
+\end{changemargin}
 \end{frame}
 
-\begin{frame} \frametitle{Results Covariates}
+\begin{frame} \frametitle{Results: Covariates}
+\begin{changemargin}{-1cm}{-1cm}
+\centering
 
+ \begin{tabular}{cc}
+  MQ Score Difference & Ideological Breadth \\
+\includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/mq_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/absdiffmq_coef_trend.pdf} \\
+
+   Same Issue &  Majority Size \\
+\includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/sameissue_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/numberjusticespro_coef_trend.pdf} \\
+
+\end{tabular}
+
+\end{changemargin}
 \end{frame}
@@
 \begin{frame} \frametitle{Predictive Performance}
+
+\begin{tabular}{rllll}
+\hline \hline
+& \multicolumn{2}{c}{Independent Model} & \multicolumn{2}{c}{Full Model} \\
+  \hline
+ & mean & range & mean & range \\ 
+  \hline
+precision & 0.5499 & (0.5384, 0.5619) & 0.8605 & (0.8526, 0.8666) \\ 
+  recall & 0.0827 & (0.0811, 0.0843) & 0.5858 & (0.5817, 0.5891) \\ 
+  F1 score & 0.1438 & (0.141, 0.1463) & 0.6971 & (0.6941, 0.7008) \\ 
+   \hline \hline
+\end{tabular}
 

### 2018-01-04 Tex/presentations/SPSA2018/DesmaraisSPSATalk.tex
added daggish visualization

@@
 
-\begin{frame} \frametitle{Captivating Story/Anecdote}
+\begin{frame}[plain]% \frametitle{The Importance of Precedent}
+\centering
 
+\includegraphics[width = 0.9\textwidth ]{../../images/apsrReference}
+\\
+\includegraphics[width = 0.9\textwidth ]{../../images/apsrTableScreenshot}
 \end{frame}
@@
 
-\begin{frame} \frametitle{Citation TERGM}
+\frame{
+\frametitle{The Exponential Random Graph Model (ERGM)}\vspace{0.5cm}
+
+The probability (likelihood function) of observing network $N$ is:
+\[
+\mathcal{P}(N, \bt) = \frac{\exp \{ \bt' \bh(N)  \}}{\sum_{N^* \in \mathcal{N}} \exp \{\bt ' \bh(N^*)  \}}
+\]
+Decomposition:
+\[
+\underbrace{\bh(N)}_{Net\hspace{3pt} Stats} \qquad \underbrace{\bt}_{Effects} \qquad \underbrace{\exp \{\bt' \bh(N) \}}_{+ \hspace{3pt} Weight} \qquad \underbrace{\sum_{N^* \in \mathcal{N}} \exp \{\bt ' \bh(N^*)  \}}_{Normalizer}
+\]
+
+Flexible: $\bh$ can capture virtually any form of interdependence among the edges $+$ covariates\\
+\vspace{0.5cm}
+Normalizing constant can make estimation difficult
+}
+
+\begin{frame} \frametitle{DAG[gish] Dynamics}
+\centering
+
+\includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 1.5cm,clip=true  ]{../../../Tex/images/daggish.pdf}
 
@@
 \begin{frame} \frametitle{Model Specification}
+
+\begin{columns}
+\begin{column}{5cm}
+{\bf Exogenous Effects}
+\begin{itemize}
+\item Ideological Distance
+\item Ideological Breadth
+\item Same issue
+\item Majority size
+\item Issue area dummies
+\end{itemize}
+\end{column}
+
+\begin{column}{6cm}
+{\bf Dependence Effects}
+
+\begin{itemize}
+\item \# Mutual Dyads
+\item \# Transitive Ties
+\item \# In-two-stars
+\item  \# Out-two-stars
+\end{itemize}
+
+\end{column}
+
+\end{columns}
 
@@
 
+{\bf Key Findings}
+\begin{enumerate}
+\item Citation is characterized by reciprocity and transitivity
+\item Dependence effects improve models of citations
+\item c-TERGM represents comprehensive method
+\end{enumerate}
+~\\
+{\bf Limitations}
+\begin{enumerate}
+\item Do not consider the signs of the citations
+\item We miss recent terms
+\item Case emergence considered exogenous
+\end{enumerate}
+
 \end{frame}

### 2018-01-04 R-Code/daggishIllustration (Autosaved).R

@@ -1,0 +1,44 @@
+# code assumes working directory is ./Supreme_Court_Citation_Network/R-Code
+
+nodes_per_year <- 3
+years <- 3
+
+greyseq <- paste("grey",round(seq(30,90,length=years)),sep="")
+
+cases <- years*nodes_per_year
+pdf("../Tex/images/daggish.pdf",height=5,width=5)
+par(las=1)
+plot(1:(cases-1),1:(cases-1),type="n",xlim=c(0,cases),ylim=c(0,cases),xaxs="i",yaxs="i",xaxt="n",yaxt="n",ylab="sending case",xlab="receiving case")
+
+yrstart=0
+for(yr in 1:years){
+	rect(xleft=0,xright=yrstart+nodes_per_year,ybottom=yrstart,ytop=yrstart+nodes_per_year,col=greyseq[yr])
+	rect(xleft=yrstart+nodes_per_year,xright=yrstart+nodes_per_year,ybottom=yrstart,ytop=yrstart+nodes_per_year,col=greyseq[yr])
+	yrstart=yrstart+nodes_per_year
+	
+}
+
+abline(h=0:cases,v=0:cases)
+
+
+set.seed(1234)
+nzeros <- round(0.7*cases^2)
+zeros.sender <- c(1:cases,sample(1:cases,nzeros,rep=T))
+zeros.receiver <- c(1:cases,sample(1:cases,nzeros,rep=T))
+
+for(i in 1:nzeros){
+	rect(xleft = zeros.receiver[i]-1,xright=zeros.receiver[i],ybottom=zeros.sender[i]-1,zeros.sender[i],col="white")
+}
+
+
+yrstart=0
+for(yr in 1:years){
+	rect(xleft=yrstart+nodes_per_year,xright=cases,ybottom=yrstart,ytop=yrstart+nodes_per_year,col="white")
+	yrstart=yrstart+nodes_per_year
+	
+}
+
+axis(1,at=(1:cases-0.5),lab=as.character(1:cases))
+axis(2,at=(1:cases-0.5),lab=as.character(1:cases))
+
+dev.off()

### 2018-01-04 R-Code/daggishIllustration.R

@@ -1,0 +1,41 @@
+nodes_per_year <- 3
+years <- 3
+
+greyseq <- paste("grey",round(seq(30,90,length=years)),sep="")
+
+cases <- years*nodes_per_year
+pdf()
+par(las=1)
+plot(1:(cases-1),1:(cases-1),type="n",xlim=c(0,cases),ylim=c(0,cases),xaxs="i",yaxs="i",xaxt="n",yaxt="n",ylab="sending case",xlab="receiving case")
+
+yrstart=0
+for(yr in 1:years){
+	rect(xleft=0,xright=yrstart+nodes_per_year,ybottom=yrstart,ytop=yrstart+nodes_per_year,col=greyseq[yr])
+	rect(xleft=yrstart+nodes_per_year,xright=yrstart+nodes_per_year,ybottom=yrstart,ytop=yrstart+nodes_per_year,col=greyseq[yr])
+	yrstart=yrstart+nodes_per_year
+	
+}
+
+abline(h=0:cases,v=0:cases)
+
+
+set.seed(1234)
+nzeros <- round(0.7*cases^2)
+zeros.sender <- c(1:cases,sample(1:cases,nzeros,rep=T))
+zeros.receiver <- c(1:cases,sample(1:cases,nzeros,rep=T))
+
+for(i in 1:nzeros){
+	rect(xleft = zeros.receiver[i]-1,xright=zeros.receiver[i],ybottom=zeros.sender[i]-1,zeros.sender[i],col="white")
+}
+
+
+yrstart=0
+for(yr in 1:years){
+	rect(xleft=yrstart+nodes_per_year,xright=cases,ybottom=yrstart,ytop=yrstart+nodes_per_year,col="white")
+	yrstart=yrstart+nodes_per_year
+	
+}
+
+axis(1,at=(1:cases-0.5),lab=as.character(1:cases))
+axis(2,at=(1:cases-0.5),lab=as.character(1:cases))
+

### 2018-01-04 Tex/presentations/SPSA2018/DesmaraisSPSATalk.tex
finished notation revisions

@@
 
+\begin{frame} \frametitle{DAG[gish] Dynamics}
+\centering
+
+\includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 1.5cm,clip=true  ]{../../../Tex/images/daggish.pdf}
+
+\end{frame}
+
 \frame{
-\frametitle{The Exponential Random Graph Model (ERGM)}\vspace{0.5cm}
+\frametitle{Citation Temporal Exponential Random Graph Model}\vspace{0.5cm}
 
-The probability (likelihood function) of observing network $N$ is:
+The probability  of observing the $N_t \times N_{\leq t}$ adjacency matrix $C_t$ is
 \[
-\mathcal{P}(N, \bt) = \frac{\exp \{ \bt' \bh(N)  \}}{\sum_{N^* \in \mathcal{N}} \exp \{\bt ' \bh(N^*)  \}}
+\mathcal{P}(C_t, \bt) = \frac{\exp \{ \bt' \bh(C_{\leq t})  \}}{\sum_{C^*_t \in \mathcal{C}_t} \exp \{\bt ' \bh(C^*_{\leq t})\}  }
 \]
@@
 \[
-\underbrace{\bh(N)}_{Net\hspace{3pt} Stats} \qquad \underbrace{\bt}_{Effects} \qquad \underbrace{\exp \{\bt' \bh(N) \}}_{+ \hspace{3pt} Weight} \qquad \underbrace{\sum_{N^* \in \mathcal{N}} \exp \{\bt ' \bh(N^*)  \}}_{Normalizer}
+\underbrace{\bh(C_{\leq t})}_{Net\hspace{3pt} Stats} \qquad \underbrace{\bt}_{Effects} \qquad \underbrace{\exp \{\bt' \bh(C_{\leq t}) \}}_{+ \hspace{3pt} Weight} \qquad \underbrace{\sum_{C^*_t \in \mathcal{C}_t} \exp \{\bt ' \bh(C^*_{\leq t})\}}_{Normalizer}
 \]
@@
 }
-
-\begin{frame} \frametitle{DAG[gish] Dynamics}
-\centering
-
-\includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 1.5cm,clip=true  ]{../../../Tex/images/daggish.pdf}
-
-\end{frame}
 

### 2018-01-04 R-Code/daggishIllustration (Autosaved).R

@@ -1,44 +1,0 @@
-# code assumes working directory is ./Supreme_Court_Citation_Network/R-Code
-
-nodes_per_year <- 3
-years <- 3
-
-greyseq <- paste("grey",round(seq(30,90,length=years)),sep="")
-
-cases <- years*nodes_per_year
-pdf("../Tex/images/daggish.pdf",height=5,width=5)
-par(las=1)
-plot(1:(cases-1),1:(cases-1),type="n",xlim=c(0,cases),ylim=c(0,cases),xaxs="i",yaxs="i",xaxt="n",yaxt="n",ylab="sending case",xlab="receiving case")
-
-yrstart=0
-for(yr in 1:years){
-	rect(xleft=0,xright=yrstart+nodes_per_year,ybottom=yrstart,ytop=yrstart+nodes_per_year,col=greyseq[yr])
-	rect(xleft=yrstart+nodes_per_year,xright=yrstart+nodes_per_year,ybottom=yrstart,ytop=yrstart+nodes_per_year,col=greyseq[yr])
-	yrstart=yrstart+nodes_per_year
-	
-}
-
-abline(h=0:cases,v=0:cases)
-
-
-set.seed(1234)
-nzeros <- round(0.7*cases^2)
-zeros.sender <- c(1:cases,sample(1:cases,nzeros,rep=T))
-zeros.receiver <- c(1:cases,sample(1:cases,nzeros,rep=T))
-
-for(i in 1:nzeros){
-	rect(xleft = zeros.receiver[i]-1,xright=zeros.receiver[i],ybottom=zeros.sender[i]-1,zeros.sender[i],col="white")
-}
-
-
-yrstart=0
-for(yr in 1:years){
-	rect(xleft=yrstart+nodes_per_year,xright=cases,ybottom=yrstart,ytop=yrstart+nodes_per_year,col="white")
-	yrstart=yrstart+nodes_per_year
-	
-}
-
-axis(1,at=(1:cases-0.5),lab=as.character(1:cases))
-axis(2,at=(1:cases-0.5),lab=as.character(1:cases))
-
-dev.off()

### 2018-01-04 R-Code/daggishIllustration.R

@@ -1,10 +1,12 @@
+# code assumes working directory is ./Supreme_Court_Citation_Network/R-Code
+
 nodes_per_year <- 3
 years <- 3
 
-greyseq <- paste("grey",round(seq(30,90,length=years)),sep="")
+greyseq <- paste("grey",round(seq(30,80,length=years)),sep="")
 
 cases <- years*nodes_per_year
-pdf()
+pdf("../Tex/images/daggish.pdf",height=5,width=5)
 par(las=1)
 plot(1:(cases-1),1:(cases-1),type="n",xlim=c(0,cases),ylim=c(0,cases),xaxs="i",yaxs="i",xaxt="n",yaxt="n",ylab="sending case",xlab="receiving case")
 
@@ -39,3 +41,4 @@ for(yr in 1:years){
 axis(1,at=(1:cases-0.5),lab=as.character(1:cases))
 axis(2,at=(1:cases-0.5),lab=as.character(1:cases))
 
+dev.off()

### 2018-01-04 R-Code/daggishIllustration.R
added notation to daggish plot

@@ -1,14 +1,46 @@
 # code assumes working directory is ./Supreme_Court_Citation_Network/R-Code
 
+CurlyBraces <- function(x0, x1, y0, y1, pos = 1, direction = 1, depth = 1) {
+
+    a=c(1,2,3,48,50)    # set flexion point for spline
+    b=c(0,.2,.28,.7,.8) # set depth for spline flexion point
+
+    curve = spline(a, b, n = 50, method = "natural")$y * depth
+
+    curve = c(curve,rev(curve))
+
+    if (pos == 1){
+        a_sequence = seq(x0,x1,length=100)
+        b_sequence = seq(y0,y1,length=100)  
+    }
+    if (pos == 2){
+        b_sequence = seq(x0,x1,length=100)
+        a_sequence = seq(y0,y1,length=100)      
+    }
+
+    # direction
+    if(direction==1)
+        a_sequence = a_sequence+curve
+    if(direction==2)
+        a_sequence = a_sequence-curve
+
+    # pos
+    if(pos==1)
+        lines(a_sequence,b_sequence, lwd=1.5,   xpd=NA) # vertical
+    if(pos==2)
+        lines(b_sequence,a_sequence, lwd=1.5, xpd=NA) # horizontal
+
+}
+
 nodes_per_year <- 3
 years <- 3
 
 greyseq <- paste("grey",round(seq(30,80,length=years)),sep="")
 
 cases <- years*nodes_per_year
-pdf("../Tex/images/daggish.pdf",height=5,width=5)
-par(las=1)
-plot(1:(cases-1),1:(cases-1),type="n",xlim=c(0,cases),ylim=c(0,cases),xaxs="i",yaxs="i",xaxt="n",yaxt="n",ylab="sending case",xlab="receiving case")
+pdf("../Tex/images/daggish.pdf",height=5,width=6)
+par(las=1,mar=c(4,8,1,1))
+plot(1:(cases-1),1:(cases-1),type="n",xlim=c(0,cases),ylim=c(0,cases),xaxs="i",yaxs="i",xaxt="n",yaxt="n",ylab="",xlab="")
 
 yrstart=0
 for(yr in 1:years){
@@ -40,5 +72,24 @@ for(yr in 1:years){
 
 axis(1,at=(1:cases-0.5),lab=as.character(1:cases))
[…]

### 2018-01-04 Tex/presentations/SPSA2018/DesmaraisSPSATalk.tex
added notation to daggish plot
finished daggish plot

@@
 
-\begin{frame} \frametitle{DAG[gish] Dynamics}
-\centering
-
-\includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 1.5cm,clip=true  ]{../../../Tex/images/daggish.pdf}
-
-\end{frame}
-
 \frame{
@@
 
-The probability  of observing the $N_t \times N_{\leq t}$ adjacency matrix $C_t$ is
+The probability  of observing the $N_t \times N_{\leq t}$ adjacency matrix $C_t$ given past citations $C_{<t}$, where $C_{\leq t}$ is the network up to time $t$
 \[
-\mathcal{P}(C_t, \bt) = \frac{\exp \{ \bt' \bh(C_{\leq t})  \}}{\sum_{C^*_t \in \mathcal{C}_t} \exp \{\bt ' \bh(C^*_{\leq t})\}  }
+\mathcal{P}(C_t|C_{<t}, \bt) = \frac{\exp \{ \bt' \bh(C_{\leq t})  \}}{\sum_{C^*_t \in \mathcal{C}_t} \exp \{\bt ' \bh(C^*_{\leq t})\}  }
 \]
@@
 }
+
+\begin{frame} \frametitle{DAG[gish] Dynamics}
+\centering
+
+\includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 0cm,clip=true  ]{../../../Tex/images/daggish.pdf}
+
+\end{frame}
 

### 2018-01-04 R-Code/daggishIllustration.R

@@ -74,11 +74,11 @@ axis(1,at=(1:cases-0.5),lab=as.character(1:cases))
 axis(2,at=(1:cases-0.5),lab=as.character(1:cases))
 
 
-CurlyBraces(x0=-.75, x1=-.75, y0=cases-nodes_per_year, y1=cases, pos = 1, direction = 2, depth = .5) 
+CurlyBraces(x0=-.75, x1=-.75, y0=cases-nodes_per_year, y1=cases, pos = 1, direction = 2, depth = .75) 
 
-CurlyBraces(x0=-.75, x1=-.75, y0=0, y1=cases-nodes_per_year, pos = 1, direction = 2, depth = .5) 
+CurlyBraces(x0=-.75, x1=-.75, y0=0, y1=cases-nodes_per_year, pos = 1, direction = 2, depth = .75) 
 
-CurlyBraces(x0=-1.95, x1=-1.95, y0=0, y1=cases, pos = 1, direction = 2, depth = .5) 
+CurlyBraces(x0=-1.95, x1=-1.95, y0=0, y1=cases, pos = 1, direction = 2, depth = .75) 
 
 library(latex2exp)
 
@@ -86,9 +86,9 @@ Ct <- TeX('$C_t$')
 Clt <- TeX('$C_{<t}$')
 Clet <- TeX('$C_{ \\leq t}$')
 
-text(-1.55,(cases-nodes_per_year+cases)/2,Ct,xpd=T,cex=1.5)
-text(-1.55,(cases-nodes_per_year)/2,Clt,xpd=T,cex=1.5)
-text(-2.75,(cases)/2,Clet,xpd=T,cex=1.5)
+text(-1.8,(cases-nodes_per_year+cases)/2,Ct,xpd=T,cex=1.5)
+text(-1.8,(cases-nodes_per_year)/2,Clt,xpd=T,cex=1.5)
+text(-3,(cases)/2,Clet,xpd=T,cex=1.5)
 
 title(xlab="receiving case",line=2.25)
 

### 2018-02-19 R-Code_courtlistener/data_readin.R
R-code to merge CL-Data with SCDB data

@@ -1,0 +1,444 @@
+memory.limit(30000)
+
+#####################################################
+#### read in Washington law data
+#####################################################
+
+wash.law1<- read.csv("SCDB1.csv", stringsAsFactors = F)
+wash.law2<- read.csv("SCDB2.csv", stringsAsFactors = F)
+# combine data sets
+
+wash.law<- rbind(wash.law1, wash.law2)
+
+#####################################################
+#### read in courtlistener citation csv
+#####################################################
+
+library(readr)
+citations <- read_csv("cl_citations.csv")
+citations<- as.data.frame(citations)
+
+#######################################################
+#### read in Courtlistener scotus opinion data
+#######################################################
+
+library(rjson)
+filenames_op <- list.files("C:/Users/Geiler Typ/Desktop/PSU/Bruce Desmarais/Supreme Court Citation Project/scotus_opinion", pattern="*.json", full.names=TRUE) # this should give you a character vector, with each file name represented by an entry
+myJSON_opinion <- lapply(filenames_op, function(x) fromJSON(file=x)) # a list in which each element is one of the original JSON files, takes a while ~ 30 min
+
+######################################################
+#### read in courtlistener scotus cluster-based data
+
+filenames_cluster <- list.files("C:/Users/Geiler Typ/Desktop/PSU/Bruce Desmarais/Supreme Court Citation Project/R-Code_courtlistener_data/all_cluster/scotus", pattern="*.json", full.names=TRUE) # this should give you a character vector, with each file name represented by an entry
+myJSON_cluster <- lapply(filenames_cluster, function(x) fromJSON(file=x)) # a list in which each element is one of the original JSON files, takes a while ~ 30 min
+
+l.cluster<- length(myJSON_cluster) # 63967
+
+#########################
+# create matrix that takes the important information from the JSON cluster files
+
+cluster.mat <- matrix(NA, l.cluster, 7)
[…]

### 2018-02-19 R-Code_courtlistener/data_readin.R
updated tex: empirical analysis

@@ -347,10 +347,31 @@ counts<- table(scc$year)
 barplot(counts, main="Number of new Cases every Year", xlab="Year", ylab="Frequency")
 
 ################# outdegree distribution
-hist(rowSums(adjacency.matrix), breaks=-0.5:165.5, main="Outdegree Distribution", xlab="Outdegree")
+library(ggplot2)
+outdegree.plot<- qplot(rowSums(adjacency.matrix), geom="histogram", binwidth = 1,  
+      main = "Outdegree Distribution", 
+      xlab = "Outdegree", ylab="Frequency",  
+      fill=I("lightblue"), 
+      col=I("black"), ylim=c(0,1000),
+      xlim=c(0,50)) + theme(axis.text=element_text(size=24),axis.title=element_text(size=26),
+                            plot.title=element_text(size=28, face='bold', hjust=0.5))
+
+max(rowSums(adjacency.matrix))
 
 ################# indegree distribution
-hist(colSums(adjacency.matrix), breaks=-0.5:228.5, main="Indegree Distribution", xlab="Indegree")
+indegree.plot<- qplot(colSums(adjacency.matrix), geom="histogram", binwidth = 1,  
+      main = "Indegree Distribution", 
+      xlab = "Indegree", ylab="Frequency",  
+      fill=I("lightblue"), 
+      col=I("black"), ylim=c(0,1000),
+      xlim=c(0,50)) + theme(axis.text=element_text(size=24),axis.title=element_text(size=26),
+                            plot.title=element_text(size=28, face='bold', hjust=0.5))
+
+max(colSums(adjacency.matrix))
+
+library(gridExtra)
+grid.arrange(indegree.plot, outdegree.plot, ncol=1, nrow=2)
+
 
 ## number of ties
 sum(adjacency.matrix) # 110742

### 2018-02-19 Tex/paper.tex
updated title

@@
 \usepackage[T1]{fontenc}
-\usepackage{authblk}
+%\usepackage{authblk}
 \usepackage{amssymb}
@@
 
-\title{\Large Analyzing the Supreme Court Citation Network}
+\title{\Large Inferential Analysis of the Supreme Court Citation Network}
 %\author{}
@@
  \author[1]{David R. Hunter \thanks{dhunter@stat.psu.edu}}
- \affil[2]{Department of Political Science, Pennsylvania State University}
- \affil[1]{Department of Statistics, Pennsylvania State University}
+% \affil[2]{Department of Political Science, Pennsylvania State University}
+ %\affil[1]{Department of Statistics, Pennsylvania State University}
 
@@
 A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns.
-We raise two limitations of existing work on Supreme Court citaitons.
+We raise two limitations of existing work on Supreme Court citations.
 First, dyadic citations are typically aggregated to the case level before they are analyzed.
@@
 For example, case-level analyses can model whether opinions supported by a liberal majority coalition are more likely than those supported by a conservative majority coalition to be cited heavily in the future, but they cannot precisely model whether liberal cases will be cited more by liberal cases than by conservative cases.
-Thus, the first reason for analyzing citations at the dyadic level it to expand the set of hypotheses that can be represented in the model.
+Thus, the first reason for analyzing citations at the dyadic level is to expand the set of hypotheses that can be represented in the model.
 The second reason for studying citations at the directed dyadic level is that, as articulated in the growing literature on legal citation networks, citations form complex networks in which a citation at one point in time may influence future citations.
@@
 We account for this effect following citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
-\citet{clark2010locating} estimates a latent coordinate model of Cupreme Court opinions based on the network of case-to-case citations.
+\citet{clark2010locating} estimates a latent coordinate model of Supreme Court opinions based on the network of case-to-case citations.
 They find that the majority opinion falls at the ideal point of the median member of the majority coalition in the case.
@@
  \section{Empirical Analysis}
-Our three data sources for this study include the SCDB, Martin-Quinn scores, and Supreme Court citation data provided by James Fowler and colleagues \citep{fowler2007network,fowler2008authority}.
-We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2001).
-The Supreme Court citation network from $1937 - 2001$ consists of $8,817$.
+Our three data sources for this study include the SCDB, Martin-Quinn scores, and Supreme Court citation data provided by Courtlistener ?cite.
+We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015).
+The Supreme Court citation network from $1937 - 2015$ consists of $9,841$ cases.
 The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
-The network has a total of $93,263$ ties.
-The number of triangles in the network is $211,855$.
+The network has a total of $110,742$ ties.
+The number of triangles in the network is $246,134$.
 The in- and outdegree distributions (i.e., the distributions of the number of citations sent and received by cases, respectively), are visualized in figures \ref{indegree_dist} and \ref{outdegree_dist}.
-The maximum indegree is $190$ and the maximum outdegree is $159$.
+The maximum indegree is $228$ and the maximum outdegree is $162$.
 
@@
 {\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Terms} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Total Number Cases} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Cases/Term} \\ \hline
-CE Hughes*              & 1937 - 1941                                          & 629                                                               & 125.5                                                     \\ \hline
-HF Stone                & 1942 - 1946                                          & 766                                                               & 153.2                                                     \\ \hline
-FM Vinson               & 1946 - 1953                                          & 788                                                               & 98.5                                                      \\ \hline
-E Warren                & 1954 - 1969                                          & 2159                                                              & 127.0                                                     \\ \hline
-WE Burger               & 1970 - 1986                                          & 2805                                                              & 155.8                                                     \\ \hline
-W Rehnquist**            & 1987 - 2001                                          & 1670                                                              & 83.5                                                      \\ \hline
+CE Hughes*              & 1937 - 1941                                          & 628                                                               & 125.6                                                     \\ \hline
+HF Stone                & 1942 - 1946                                          & 756                                                               & 151.2                                                     \\ \hline
+FM Vinson               & 1946 - 1953                                          & 786                                                               & 98.25                                                      \\ \hline
+E Warren                & 1954 - 1969                                          & 2123                                                              & 124.88                                                     \\ \hline
+WE Burger               & 1970 - 1986                                          & 2755                                                              & 153.06                                                     \\ \hline
+W Rehnquist            & 1987 - 2005                                          & 2004                                                              & 105.47                                                      \\ \hline
+J Roberts **           & 2006 - 2015                                          & 789                                                              & 78.9                                                      \\ \hline
 \end{tabular}
-\caption[caption]{For the time range of interest (1937 - 2001) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
-\\\hspace{\textwidth} ** W Rehnquist served as chief justice from 1987 - 2005.}
+\caption[caption]{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
+\\\hspace{\textwidth} ** J Roberts still serves as chief justice (2/2018).}
 \label{tab:chiefs}

### 2018-02-19 Tex/paper.tex

@@
 The number of triangles in the network is $246,134$.
-The in- and outdegree distributions (i.e., the distributions of the number of citations sent and received by cases, respectively), are visualized in figures \ref{indegree_dist} and \ref{outdegree_dist}.
+The in- and outdegree distributions (i.e., the distributions of the number of citations sent and received by cases, respectively), are visualized in Figure \ref{degree_dist}.
 The maximum indegree is $228$ and the maximum outdegree is $162$.

### 2018-03-11 Tex/paper.tex
holdout experiment results

@@
   \tikz[baseline]\node[draw,shape=circle,anchor=base] {#1} ;}
+  \usepackage{hyperref}
+\newcommand{\footremember}[2]{%
+    \footnote{#2}
+    \newcounter{#1}
+    \setcounter{#1}{\value{footnote}}%
+}
+\newcommand{\footrecall}[1]{%
+    \footnotemark[\value{#1}]%
+}
 
@@
  % Toggle % below to blind and unblind the manuscript
-  \author[1]{Christian Schmid \thanks{cxs5700@psu.edu}}
-   \author[2]{Ted Hsuan Yun Chen\thanks{thc126@psu.edu}}
- \author[2]{Bruce A. Desmarais\thanks{bdesmarais@psu.edu}}
- \author[1]{David R. Hunter \thanks{dhunter@stat.psu.edu}}
+%  \author[1]{Christian Schmid \thanks{cxs5700@psu.edu}}
+%   \author[2]{Ted Hsuan Yun Chen\thanks{thc126@psu.edu}}
+% \author[2]{Bruce A. Desmarais\thanks{bdesmarais@psu.edu}}
+% \author[1]{David R. Hunter \thanks{dhunter@stat.psu.edu}}
 % \affil[2]{Department of Political Science, Pennsylvania State University}
  %\affil[1]{Department of Statistics, Pennsylvania State University}
+
+\author{%
+  Christian S. Schmid\footnote{Department of Statistics, The Pennsylvania State University, schmid@psu.edu}%
+  \and Ted Hsuan Yun Chen \footnote{Department of Political Science, The Pennsylvania State University, thc126@psu.edu}%
+  \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
+  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
+  }
 

### 2018-03-11 Tex/paper.tex
added holdout experiment results to tex

@@
 
-\begin{table}[ht]
+\begin{table}[H]
 \centering
@@
   \hline
-precision & 0.5499 & (0.5384, 0.5619) & 0.8605 & (0.8526, 0.8666) \\ 
-  recall & 0.0827 & (0.0811, 0.0843) & 0.5858 & (0.5817, 0.5891) \\ 
-  F1 score & 0.1438 & (0.141, 0.1463) & 0.6971 & (0.6941, 0.7008) \\ 
+precision & 0.5582 & (0.5509, 0.5751) & 0.8678 & (0.865, 0.8721) \\ 
+  recall & 0.0995 & (0.0953, 0.1043) & 0.5937 & (0.588, 0.6027) \\ 
+  F1 score & 0.1688 & (0.163, 0.1766) & 0.7051 & (0.7009, 0.7117) \\ 
    \hline \hline
 \end{tabular}
-\caption{The predictive performance of the directed dyadic models, over ten 80/20 train/test splits.}
-\label{tab:prediction}
+\caption{The predictive performance of the directed dyadic models with linear trends, over ten 80/20 train/test splits.}
+\label{tab:prediction_linear}
+\begin{tabular}{rllll}
+\hline \hline
+& \multicolumn{2}{c}{Independent Model} & \multicolumn{2}{c}{Full Model} \\
+   \hline
+ & mean & range & mean & range \\ 
+  \hline
+precision & 0.8827 & (0.878, 0.8876) & 0.8672 & (0.8629, 0.8731) \\ 
+  recall & 0.5316 & (0.5255, 0.5408) & 0.5939 & (0.5872, 0.6008) \\ 
+  F1 score & 0.6635 & (0.6594, 0.6709) & 0.705 & (0.7008, 0.7118) \\ 
+   \hline \hline
+\end{tabular}
+\caption{The predictive performance of the directed dyadic models with quadratic trends, over ten 80/20 train/test splits.}
+\label{tab:prediction_quadratic}
+\begin{tabular}{rllll}
+\hline \hline
+& \multicolumn{2}{c}{Independent Model} & \multicolumn{2}{c}{Full Model} \\
+   \hline
+ & mean & range & mean & range \\ 
+  \hline
+precision & 0.8745 & (0.8718, 0.8769) & 0.8667 & (0.8649, 0.8684) \\ 
+  recall & 0.5732 & (0.5693, 0.5783) & 0.5949 & (0.5902, 0.5981) \\ 
+  F1 score & 0.6925 & (0.6892, 0.6969) & 0.7056 & (0.7025, 0.7083) \\ 
+   \hline \hline
+\end{tabular}
+\caption{The predictive performance of the directed dyadic models with cubic trends, over ten 80/20 train/test splits.}
+\label{tab:prediction_cubic}
+\begin{tabular}{rllll}
+\hline \hline
+& \multicolumn{2}{c}{Independent Model} & \multicolumn{2}{c}{Full Model} \\
+   \hline
+ & mean & range & mean & range \\ 
+  \hline
+precision & 0.8769 & (0.8738, 0.881) & 0.8687 & (0.8646, 0.8718) \\ 
+  recall & 0.583 & (0.5754, 0.5896) & 0.5956 & (0.5895, 0.6008) \\ 
+  F1 score & 0.7004 & (0.6946, 0.7042) & 0.7067 & (0.7018, 0.7098) \\ 
+   \hline \hline
+\end{tabular}
+\caption{The predictive performance of the directed dyadic models with quartic trends, over ten 80/20 train/test splits.}
+\label{tab:prediction_quartic}
 \end{table}
 
-We see from Table \ref{tab:prediction} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms.
+We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms.
 The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set.

### 2018-03-20 Tex/paper.tex
new holdout experiment results

@@
  \section{Empirical Analysis}
-Our three data sources for this study include the SCDB, Martin-Quinn scores, and Supreme Court citation data provided by Courtlistener ?cite.
-We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015).
-The Supreme Court citation network from $1937 - 2015$ consists of $9,841$ cases.
+Our three data sources for this study include the Supreme Court Database (SCDB), Martin-Quinn scores, and Supreme Court citation data provided by Courtlistener ?cite.
+We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015)\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the Courtlistener data.
+We decided to exclude these 145 cases from our model.}.
+The Supreme Court citation network from $1937 - 2015$ consists of $9,945$ cases.
 The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
-The network has a total of $110,742$ ties.
-The number of triangles in the network is $246,134$.
+The network has a total of $111,986$ ties.
+The number of triangles in the network is $250,717$ and the number of mutual ties is $56$.
 The in- and outdegree distributions (i.e., the distributions of the number of citations sent and received by cases, respectively), are visualized in Figure \ref{degree_dist}.
-The maximum indegree is $228$ and the maximum outdegree is $162$.
+The maximum indegree is $230$ and the maximum outdegree is $162$.
 
@@
 HF Stone                & 1942 - 1946                                          & 756                                                               & 151.2                                                     \\ \hline
-FM Vinson               & 1946 - 1953                                          & 786                                                               & 98.25                                                      \\ \hline
-E Warren                & 1954 - 1969                                          & 2123                                                              & 124.88                                                     \\ \hline
-WE Burger               & 1970 - 1986                                          & 2755                                                              & 153.06                                                     \\ \hline
-W Rehnquist            & 1987 - 2005                                          & 2004                                                              & 105.47                                                      \\ \hline
-J Roberts **           & 2006 - 2015                                          & 789                                                              & 78.9                                                      \\ \hline
+FM Vinson               & 1946 - 1953                                          & 789                                                               & 98.63                                                      \\ \hline
+E Warren                & 1954 - 1969                                          & 2149                                                              & 126.41                                                     \\ \hline
+WE Burger               & 1970 - 1986                                          & 2805                                                             & 155.83                                                     \\ \hline
+W Rehnquist            & 1987 - 2005                                          & 2022                                                              & 106.42                                                      \\ \hline
+J Roberts **           & 2006 - 2015                                          & 796                                                              & 79.6                                           \\ \hline
 \end{tabular}
 \caption[caption]{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
-\\\hspace{\textwidth} ** J Roberts still serves as chief justice (2/2018).}
+\\\hspace{\textwidth} ** J Roberts still serves as chief justice (3/2018).}
 \label{tab:chiefs}
@@
  (a) MQ Score Difference & (b) Ideological Breadth \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend.pdf} \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\
 
   (c) Same Issue & (d) Majority Size \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend.pdf} \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\
 
-   (e) Reciprocity & (f) Transitivity \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend.pdf} \\
+ (e) Overruled Cases & (f) Difference in Years \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} \\
 
-    (g) Activity & (h) Popularity \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend.pdf} \\
+   (g) Reciprocity & (h) Transitivity \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
+
+    (i) Activity & (j) Popularity \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
 
@@
 \hline \hline
-& \multicolumn{2}{c}{Independent Model} & \multicolumn{2}{c}{Full Model} \\
-   \hline
  & mean & range & mean & range \\ 
   \hline
-precision & 0.5582 & (0.5509, 0.5751) & 0.8678 & (0.865, 0.8721) \\ 
-  recall & 0.0995 & (0.0953, 0.1043) & 0.5937 & (0.588, 0.6027) \\ 
-  F1 score & 0.1688 & (0.163, 0.1766) & 0.7051 & (0.7009, 0.7117) \\ 
+precision & 0.5872 & (0.5781, 0.5962) & 0.8622 & (0.8593, 0.8668) \\ 
+  recall & 0.0565 & (0.0547, 0.0587) & 0.5839 & (0.5777, 0.592) \\ 
+  F1 score & 0.103 & (0.1, 0.1069) & 0.6963 & (0.6915, 0.7027) \\ 
+   \hline \hline
+\end{tabular}
+\caption{The predictive performance of the directed dyadic models with no trends, over ten 80/20 train/test splits.}
+\label{tab:prediction_linear}
+\begin{tabular}{rllll}
+\hline \hline
+ & mean & range & mean & range \\ 
+  \hline
+precision & 0.5589 & (0.5508, 0.5757) & 0.8677 & (0.865, 0.8718) \\ 
+  recall & 0.098 & (0.0943, 0.103) & 0.5943 & (0.5887, 0.6032) \\ 
+  F1 score & 0.1667 & (0.1618, 0.1748) & 0.7054 & (0.7013, 0.7119) \\ 
    \hline \hline
@@
 \hline \hline
-& \multicolumn{2}{c}{Independent Model} & \multicolumn{2}{c}{Full Model} \\
-   \hline
  & mean & range & mean & range \\ 
   \hline
-precision & 0.8827 & (0.878, 0.8876) & 0.8672 & (0.8629, 0.8731) \\ 
-  recall & 0.5316 & (0.5255, 0.5408) & 0.5939 & (0.5872, 0.6008) \\ 
-  F1 score & 0.6635 & (0.6594, 0.6709) & 0.705 & (0.7008, 0.7118) \\ 
+precision & 0.6441 & (0.6302, 0.6536) & 0.8671 & (0.8626, 0.873) \\ 
+  recall & 0.0852 & (0.0832, 0.0863) & 0.594 & (0.5874, 0.6012) \\ 
+  F1 score & 0.1504 & (0.1473, 0.1525) & 0.705 & (0.7008, 0.712) \\ 
    \hline \hline
@@
 \hline \hline
-& \multicolumn{2}{c}{Independent Model} & \multicolumn{2}{c}{Full Model} \\
-   \hline
  & mean & range & mean & range \\ 
   \hline
-precision & 0.8745 & (0.8718, 0.8769) & 0.8667 & (0.8649, 0.8684) \\ 
-  recall & 0.5732 & (0.5693, 0.5783) & 0.5949 & (0.5902, 0.5981) \\ 
-  F1 score & 0.6925 & (0.6892, 0.6969) & 0.7056 & (0.7025, 0.7083) \\ 
+precision & 0.642 & (0.6311, 0.6527) & 0.8664 & (0.8645, 0.8681) \\ 
+  recall & 0.0866 & (0.0835, 0.0882) & 0.5947 & (0.591, 0.5978) \\ 
+  F1 score & 0.1526 & (0.1475, 0.1552) & 0.7053 & (0.7026, 0.7079) \\ 
    \hline \hline
@@
 \hline \hline
-& \multicolumn{2}{c}{Independent Model} & \multicolumn{2}{c}{Full Model} \\
-   \hline
  & mean & range & mean & range \\ 
   \hline
-precision & 0.8769 & (0.8738, 0.881) & 0.8687 & (0.8646, 0.8718) \\ 
-  recall & 0.583 & (0.5754, 0.5896) & 0.5956 & (0.5895, 0.6008) \\ 
-  F1 score & 0.7004 & (0.6946, 0.7042) & 0.7067 & (0.7018, 0.7098) \\ 
+precision & 0.645 & (0.635, 0.6668) & 0.8681 & (0.8639, 0.8712) \\ 
+  recall & 0.0864 & (0.083, 0.0887) & 0.5955 & (0.5899, 0.6004) \\ 
+  F1 score & 0.1524 & (0.1468, 0.1562) & 0.7064 & (0.7018, 0.7094) \\ 
    \hline \hline

### 2018-03-20 Tex/paper.tex

@@
 
-Let $C$ be the case-to-case adjacency matrix such that $C_{ij} \in \{0,1\}$ is a binary indicator of whether case $i$ cites $j$. Furthermore, let
-$$\mathcal{C} =\{C \in \{0,1\}^{(N \cdot (N-1))/2}: C_{ij} \in \{0,1\} \}$$ 
-be the set of all possible adjacency matrices among $N$ cases.
-Note that the cardinality of $\mathcal{C}$ increases exponentially for every newly added case.
-The probability function of the c-TERGM is defined as
+Let $C_t$ be the case-to-case adjacency matrix at time $t$ such that $C_t^{ij} \in \{0,1\}$ is a binary indicator of whether case $i$ cites $j$. Furthermore, let
+$$\mathcal{C}_t =\{C_{\leq t} \in \{0,1\}^{(N_t \cdot (N_t-1))/2}: C_{\leq t}^{ij} \in \{0,1\} \}$$ 
+be the set of all possible adjacency matrices among $N_t$ cases.
+$N_t$ denotes the the number of cases in the network at time $t$.
+Note that the cardinality of $\mathcal{C}_t$ increases exponentially for every newly added case.
+The probability of observing $C_t$ given past citations $C_{<t}$, where $C_{\leq t}$ is the network up to time $t$ is defined as
 \begin{equation}
-P_{\theta}(C)=\cfrac{\exp\big(\theta^T\cdot h\big(C)\big)\big)}{\sum_{Z^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(Z^*)\big)\big)}
+P(C_t | C_{<t}, \theta)=\cfrac{\exp\big(\theta^T\cdot h\big(C_{\leq t})\big)\big)}{\sum_{Z_t^*\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)\big)}
 \label{ercm}
 \end{equation}
-where $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C} \to \mathbb{R}^q$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{Z^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(Z^*)\big)$ is a normalization constant that ensures that Equation (\ref{ercm}) defines a probability function on $\mathcal{C}$.
+where $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C}_t \to \mathbb{R}^q$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{Z^*_t\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)$ is a normalization constant that ensures that Equation (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.
 
@@
 We include the following statistics for the Supreme Court citation network, which are derived from the interdependence hypotheses described in the previous section:
-$$h_{edges}:\mathcal{C} \to \mathbb{R}~~~, ~~~C \to \sum_{ij} C_{ij}, $$
+$$h_{edges}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}, $$
 the number of citations.  $h_{edges}$ performs the function of an intercept, and models the expected value of any given edge.
-$$h_{outstar}:\mathcal{C} \to \mathbb{R}~~~, ~~~C \to \sum_{i=1}^N {\sum_{j \neq i} c_{ij} \choose 2} $$
+$$h_{outstar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{i=1} {\sum_{j \neq i} C_{\leq t}^{ij} \choose 2} $$
 the number of out-two-stars.
@@
 The out-two-stars configuration is commonly used to model activity (sender activation) in ERGMs.
-$$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C \to \sum_{j=1}^N {\sum_{i \neq j} c_{ij} \choose 2} $$
+$$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{j=1} {\sum_{i \neq j} C_{\leq t}^{ij} \choose 2} $$
 the number of in-two-stars.
@@
 \begin{align*}
-h_{triangle}:\mathcal{C} \to \mathbb{R}~~~, ~~~C \to & \sum_{j<i<k} C_{ij}\cdot C_{jk} \cdot C_{ki} + C_{ji}\cdot C_{jk} \cdot C_{ki} + C_{ij}\cdot C_{kj} \cdot C_{ki} \\ & + C_{ji}\cdot C_{kj} \cdot C_{ki} + C_{ij}\cdot C_{jk} \cdot C_{ik} + C_{ji}\cdot C_{jk} \cdot C_{ik}  \\ & + C_{ij}\cdot C_{kj} \cdot C_{ik} + C_{ji}\cdot C_{kj} \cdot C_{ik},
+h_{triangle}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to & \sum_{j<i<k} C^{ij}\cdot C^{jk} \cdot C^{ki} + C^{ji}\cdot C^{jk} \cdot C^{ki} + C^{ij}\cdot C^{kj} \cdot C^{ki} \\ & + C^{ji}\cdot C^{kj} \cdot C^{ki} + C^{ij}\cdot C^{jk} \cdot C^{ik} + C^{ji}\cdot C^{jk} \cdot C^{ik}  \\ & + C^{ij}\cdot C^{kj} \cdot C^{ik} + C^{ji}\cdot C^{kj} \cdot C^{ik},
 \end{align*}
@@
 Triangles measure triad closure, as discussed above.
-$$h_{covariate}:\mathcal{C} \to \mathbb{R}~~~, ~~~C \to \sum_{ij} C_{ij}X_{ij}, $$ the effect of an exogenous covariate ($X$).
+$$h_{covariate}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}X_{ij}, $$ the effect of an exogenous covariate ($X$).
 We include several exogenous covariates based on this standard statistic formulation.
@@
 The structure of this model is very similar to a conventional (T)ERGM setup.
-The main difference is that  $\mathcal{C}$ excludes adjacency matrices in which there are loops that include edges sent at different time points.
+The main difference is that  $\mathcal{C}_t$ excludes adjacency matrices in which there are loops that include edges sent at different time points.
 Loops can exist in the Supreme Court citation network, but only among cases decided in the same term.
@@
 This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with ERGM family models.
-Methods of Monte Carlo approximation are well established \citep{hunter2006inference,van2009framework,hummel2012improving}, but with over 8,000 vertices, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}.
+Methods of Monte Carlo approximation are well established \citep{hunter2006inference,van2009framework,hummel2012improving}, but with almost $10,000$ vertices, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}.
 To estimate the parameters of the c-TERGM, we rely on two features of the model and data that facilitate the application of an approximate inference method that is faster than Monte Carlo approximation---maximum pseudolikelihood.

### 2018-03-22 Tex/paper.tex
add citation of data sources

@@
  \section{Empirical Analysis}
-Our three data sources for this study include the Supreme Court Database (SCDB), Martin-Quinn scores, and Supreme Court citation data provided by Courtlistener ?cite.
+Our three data sources for this study include the Supreme Court Database (SCDB) \citep{SCDB}, Martin-Quinn scores \citep{MartinQuinnScores}, and Supreme Court citation data provided by Courtlistener ?cite.
 We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015)\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the Courtlistener data.

### 2018-03-23 Tex/paper.tex
updated citations in tex

--- 2018-03-20 Tex/paper.tex
@@
  \section{Empirical Analysis}
-Our three data sources for this study include the Supreme Court Database (SCDB), Martin-Quinn scores, and Supreme Court citation data provided by Courtlistener ?cite.
+Our three data sources for this study include the Supreme Court Database (SCDB) \citep{spaeth2014supreme}, Martin-Quinn scores \citep{martin2002dynamic}, and Supreme Court citation data provided by the CourtListener Free Law Project \citep{CourtListener}.
 We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015)\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the Courtlistener data.

### 2018-03-23 Tex/paper.tex
updated result table 2

@@
 This methodology---the citation temporal exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation.
-We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2001.
+We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2015.
 We find evidence for dependence processes, including reciprocity, transitivity, popularity, and activity.
@@
 We incorporate a statistic to model the degree to which cases cite those that are similar in terms of the ideological positions of the justices who supported the decision.
-We account for this effect following citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
+We account for this effect following \citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
 \citet{clark2010locating} estimates a latent coordinate model of Supreme Court opinions based on the network of case-to-case citations.
@@
 
-\citet{benjamin2012standing} Study the propensity for cases to be overruled and cited in other negative ways.
+\citet{benjamin2012standing} study the propensity for cases to be overruled and cited in other negative ways.
 They find that cases with majority coalitions that are large and ideologically broad are less likely to be cited negatively.
@@
 We expect both of these covariates to have positive effects.
+
+\begin{itemize}
+\item mention new variable: overruled cases
+\end{itemize}
 
@@
                                                     \hline
-Edges                                              & -5.533                           & -5.733                              & -5.388                              & *                                    \\ \hline
-Instar(2)                                          & 0.031                            & 0.027                               & 0.036                               & *                                    \\ \hline
-Outstar(2)                                         & 0.022                            & 0.020                               & 0.032                               & *                                    \\ \hline
-Mutual                                             & 3.316                            & 2.622                               & 3.983                               & *                                    \\ \hline
-Triangle                                           & 1.490                            & 1.410                               & 1.560                               & *                                    \\ \hline
-Martin Quinn Score                                 & 0.080                            & 0.022                               & 0.126                               & *                                    \\ \hline
-Same Issue Area Homophily                          & 1.378                            & 1.313                               & 1.451                               & *                                    \\ \hline
-Year Difference                                    & -0.077                           & -0.090                              & -0.064                              & *                                    \\ \hline
-$(\text{Year Differnce})^2$                        & 0.0029                           & 0.0025                              & 0.0032                              & *                                    \\ \hline
-Receiver Abs Diff MQ Score in Majority             & -0.030                           & -0.048                              & -0.014                              & *                                    \\ \hline
-Receiver Number Justices in Majority               & -0.115                           & -0.135                              & -0.089                              & *                                    \\ \hline
-Receiver  Sender Year                              & 0.0007                           & 0.0006                              & 0.0009                              & *                                    \\ \hline
-Sender Same Issue Area 2                           & 0.162                            & 0.105                               & 0.217                               & *                                    \\ \hline
-Sender Same Issue Area 3                           & -0.317                           & -0.403                              & -0.236                              & *                                    \\ \hline
-Sender Same Issue Area 4                           & 0.497                            & 0.426                               & 0.570                               & *                                    \\ \hline
-Sender Same Issue Area 5                           & 0.332                            & 0.174                               & 0.495                               & *                                    \\ \hline
-Sender Same Issue Area 6                           & 0.527                            & 0.402                               & 0.635                               & *                                    \\ \hline
-Sender Same Issue Area 7                           & 0.327                            & 0.251                               & 0.394                               & *                                    \\ \hline
-Sender Same Issue Area 8                           & 0.076                            & 0.021                               & 0.129                               & *                                    \\ \hline
-Sender Same Issue Area 9                           & 0.261                            & 0.209                               & 0.315                               & *                                    \\ \hline
-Sender Same Issue Area 10                          & 0.369                            & 0.301                               & 0.432                               & *                                    \\ \hline
-Sender Same Issue Area 11                          & -0.046                           & -0.233                              & 0.151                               &                                      \\ \hline
-Sender Same Issue Area 12                          & 0.0073                           & -0.075                              & 0.089                               &                                      \\ \hline
-Sender Same Issue Area 13                          & 0.381                            & 0.162                               & 0.564                               & *                                    \\ \hline
-Sender Same Issue Area 14                          & 0.151                            & -0.019                              & 0.300                               &                                      \\ \hline
-Receiver Same Issue Area 2                         & 0.231                            & 0.166                               & 0.300                               & *                                    \\ \hline
-Receiver Same Issue Area 3                         & -0.101                           & -0.209                              & -0.006                              & *                                    \\ \hline
-Receiver Same Issue Area 4                         & 0.516                            & 0.425                               & 0.612                               & *                                    \\ \hline
-Receiver Same Issue Area 5                         & 0.351                            & 0.183                               & 0.534                               & *                                    \\ \hline
-Receiver Same Issue Area 6                         & 0.461                            & 0.284                               & 0.628                               & *                                    \\ \hline
-Receiver Same Issue Area 7                         & 0.361                            & 0.281                               & 0.437                               & *                                    \\ \hline
-Receiver Same Issue Area 8                         & 0.153                            & 0.097                               & 0.221                               & *                                    \\ \hline
-Receiver Same Issue Area 9                         & 0.279                            & 0.221                               & 0.346                               & *                                    \\ \hline
-Receiver Same Issue Area 10                        & 0.492                            & 0.413                               & 0.570                               & *                                    \\ \hline
-Receiver Same Issue Area 11                        & 0.247                            & 0.056                               & 0.411                               & *                                    \\ \hline
-Receiver Same Issue Area 12                        & 0.339                            & 0.246                               & 0.430                               & *                                    \\ \hline
-Receiver Same Issue Area 13                        & 1.060                            & 0.845                               & 1.239                               & *                                    \\ \hline
-Receiver Same Issue Area 14                        & 0.557                            & 0.255                               & 0.793                               & *                                    \\ \hline
-Instar(2) $\times$ Sender Year                     & -0.00025                         & -0.00034                            & -0.00017                            & *                                    \\ \hline
-Outstar(2) $\times$ Sender Year                    & -0.00036                         & -0.00056                            & -0.00028                            & *                                    \\ \hline
-Mutual $\times$ Sender Year                        & -0.010                           & -0.043                              & 0.029                               &                                      \\ \hline
-Triangle $\times$ Sender Year                      & -0.0004                          & -0.0021                             & 0.0016                              &                                      \\ \hline
-Martin Quinn Score $\times$ Sender Year            & -0.0025                          & -0.0035                             & -0.0013                             & *                                    \\ \hline
-Same Issue Area $\times$ Sender Year               & -0.0058                          & -0.0076                             & -0.0043                             & *                                    \\ \hline
-Year Difference $\times$ Sender Year               & 0.0004                           & 0.0002                              & 0.0007                              & *                                    \\ \hline
-$\text{Year Difference}^2 \times$ Sender Year      & -0.000037                        & -0.000044                           & -0.000030                           & *                                    \\ \hline
-MQ Score in Majority $\times$ Sender Year & 0.0013                           & 0.0009                              & 0.0017                              & *                                    \\ \hline
-Justices in Majority $\times$ Sender Year   & 0.0029                           & 0.0023                              & 0.0035                              & *                                    \\ \hline
+Edges                                              & -6.249                           & -6.412                              & -6.110                              & *                                    \\ \hline
+Instar(2)                                          & 0.035                            & 0.032                               & 0.038                               & *                                    \\ \hline
+Outstar(2)                                         & 0.021                            & 0.019                               & 0.028                               & *                                    \\ \hline
+Mutual                                             & 2.071                            & -4.157                               & 4.072                                &                                     \\ \hline
+Triangle                                           & 1.507                            & 1.431                               & 1.559                               & *                                    \\ \hline
+Martin Quinn Score                                 & 0.079                            & 0.029                               & 0.130                               & *                                    \\ \hline
+Same Issue Area Homophily                          & 1.447                            & 1.389                               & 1.515                               & *                                    \\ \hline
+Year Difference                                    & -0.068                           & -0.079                              & -0.058                              & *                                    \\ \hline
+$(\text{Year Differnce})^2$                        & 0.0019                           & 0.0016                              & 0.0022                              & *                                    \\ \hline
+Receiver Abs Diff MQ Score in Majority             & 0.047                           & 0.029                              & 0.063                              & *                                     \\ \hline
+Receiver Number Justices in Majority               & -0.075                           & -0.095                              & -0.055                              & *                                    \\ \hline
+Receiver  Sender Year                              & 0.0039                           & 0.0036                              & 0.0042                              & *                                    \\ \hline
+Overruled Cases   & -1.043                           & -1.624                              & -0.564                              & *                                    \\ \hline
+Sender Same Issue Area 2                           & 0.142                            & 0.100                               & 0.185                               & *                                    \\ \hline
+Sender Same Issue Area 3                           & -0.337                           & -0.403                              & -0.270                              & *                                    \\ \hline
+Sender Same Issue Area 4                           & 0.498                            & 0.438                               & 0.561                               & *                                    \\ \hline
+Sender Same Issue Area 5                           & 0.182                            & 0.031                               & 0.305                               & *                                    \\ \hline
+Sender Same Issue Area 6                           & 0.536                            & 0.422                               & 0.648                               & *                                    \\ \hline
+Sender Same Issue Area 7                           & 0.370                            & 0.316                               & 0.426                               & *                                    \\ \hline
+Sender Same Issue Area 8                           & 0.107                            & 0.059                               & 0.156                               & *                                    \\ \hline
+Sender Same Issue Area 9                           & 0.278                            & 0.238                               & 0.329                               & *                                    \\ \hline
+Sender Same Issue Area 10                          & 0.371                            & 0.317                               & 0.433                               & *                                    \\ \hline
+Sender Same Issue Area 11                          & -0.030                           & -0.151                              & 0.164                               &                                      \\ \hline
+Sender Same Issue Area 12                          & 0.064                           & -0.014                              & 0.130                               &                                      \\ \hline
+Sender Same Issue Area 13                          & 0.608                            & 0.423                               & 0.792                               & *                                    \\ \hline
+Sender Same Issue Area 14                          & 0.232                            & 0.075                              & 0.393                               & *                                      \\ \hline
+Receiver Same Issue Area 2                         & 0.252                            & 0.199                               & 0.310                               & *                                    \\ \hline
+Receiver Same Issue Area 3                         & -0.092                           & -0.194                              & 0.014                              &                                     \\ \hline
+Receiver Same Issue Area 4                         & 0.481                            & 0.397                               & 0.570                               & *                                    \\ \hline
+Receiver Same Issue Area 5                         & 0.374                            & 0.224                               & 0.519                               & *                                    \\ \hline
+Receiver Same Issue Area 6                         & 0.521                            & 0.346                               & 0.652                               & *                                    \\ \hline
+Receiver Same Issue Area 7                         & 0.412                            & 0.339                               & 0.487                               & *                                    \\ \hline
+Receiver Same Issue Area 8                         & 0.174                            & 0.134                               & 0.232                               & *                                    \\ \hline
+Receiver Same Issue Area 9                         & 0.311                            & 0.240                               & 0.361                               & *                                    \\ \hline
+Receiver Same Issue Area 10                        & 0.506                            & 0.429                               & 0.577                               & *                                    \\ \hline
+Receiver Same Issue Area 11                        & 0.299                            & 0.088                               & 0.432                               & *                                    \\ \hline
+Receiver Same Issue Area 12                        & 0.350                            & 0.276                               & 0.445                               & *                                    \\ \hline
+Receiver Same Issue Area 13                        & 0.972                            & 0.780                               & 1.135                               & *                                    \\ \hline
+Receiver Same Issue Area 14                        & 0.517                            & 0.200                               & 0.740                               & *                                    \\ \hline
+Instar(2) $\times$ Sender Year                     & -0.00026                         & -0.00031                            & -0.00021                            & *                                    \\ \hline
+Outstar(2) $\times$ Sender Year                    & -0.00031                         & -0.00045                            & -0.00027                            & *                                    \\ \hline
+Mutual $\times$ Sender Year                        & 0.110                           & 0.010                              & 0.524                               & *                                      \\ \hline
+Triangle $\times$ Sender Year                      & -0.00022                          & -0.0016                             & 0.0013                              &                                      \\ \hline
+Martin Quinn Score $\times$ Sender Year            & -0.0019                          & -0.0030                             & -0.0009                             & *                                    \\ \hline
+Same Issue Area $\times$ Sender Year               & -0.0068                          & -0.0081                             & -0.0055                             & *                                    \\ \hline
+Year Difference $\times$ Sender Year               & 0.00046                           & 0.00028                              & 0.00065                              & *                                    \\ \hline
+$\text{Year Difference}^2 \times$ Sender Year      & -0.000024                        & -0.000028                           & -0.000021                           & *                                    \\ \hline
+MQ Score in Majority $\times$ Sender Year & -0.00053                           & -0.00093                              & -0.0018                              & *                                    \\ \hline
+Justices in Majority $\times$ Sender Year   & 0.0017                           & 0.0013                              & 0.0021                              & *                                    \\ \hline
+Overruled Cases $\times$ Sender Year   & 0.012                           & 0.002                              & 0.023                              & *                                    \\ \hline
 \end{tabular}
-\caption{Bootstrapped MPLE Results for the time period $1937-2001$.
+\caption{Bootstrapped MPLE Results for the time period $1937-2015$.
 A '*' indicates that the 2.5th and 97.5th quantile of the variable does not include '0' and as a result is statistically significant.}

### 2018-03-23 Tex/paper.tex
added ToDos, updated chapter 5

@@
 Finally, we develop and estimate a specification of this model in an analysis of US Supreme Court citations between 1937 and 2001.
-We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's predictive performance.
+We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's predictive performance.\\[0.5cm]
+\textit{Comment from Benjamin Kassov: I think the discussion in the introduction of the paper focusing on a "case-based" versus network based analysis is a bit short. 
+I was left slightly confused by what a "case-based" analysis of citation actually is or what it means. 
+I can see several different conceptions of this, some being more problematic than others in a methodological sense, 
+although none of them are ideal for analyzing networks and all likely suffer some amount of OV bias.}
 
@@
 We include these variables because \citet{cross2010determinants} find that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
+Table \ref{issue_area_coding} states which number represents which issie area.
+
+\begin{table}[]
+\centering
+\begin{tabular}{llll}
+1 & Criminal Procedure & 8  & Economic Activity    \\
+2 & Civil Rights       & 9  & Judicial Power       \\
+3 & First Amendment    & 10 & Federalism           \\
+4 & Due Process        & 11 & Interstate Relations \\
+5 & Privacy            & 12 & Federal Taxation     \\
+6 & Attorneys          & 13 & Miscellaneous        \\
+7 & Unions             & 14 & Private Action      
+\end{tabular}
+\caption{Code Book for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
+\label{issue_area_coding}
+\end{table}
 
@@
 \begin{itemize}
-\item mention new variable: overruled cases
+\item ToDo: mention new variable: overruled cases
 \end{itemize}
 
+\textit{Comment from Benjamin Kassow: In section 5, I think the discussion is mostly good, but could be organized a bit differently. 
+I think the discussion of the random intercepts is a bit confusing early on (it is explained more effectively later). }
 %%%% DONT DELETE THESE COMMENTS, INCLUDE IDEAS FOR ADDITIONAL COVARIATES %%%%%
@@
 \label{tab:prediction_quadratic}
-\begin{tabular}{rllll}
-\hline \hline
- & mean & range & mean & range \\ 
-  \hline
-precision & 0.642 & (0.6311, 0.6527) & 0.8664 & (0.8645, 0.8681) \\ 
-  recall & 0.0866 & (0.0835, 0.0882) & 0.5947 & (0.591, 0.5978) \\ 
-  F1 score & 0.1526 & (0.1475, 0.1552) & 0.7053 & (0.7026, 0.7079) \\ 
-   \hline \hline
-\end{tabular}
-\caption{The predictive performance of the directed dyadic models with cubic trends, over ten 80/20 train/test splits.}
-\label{tab:prediction_cubic}
-\begin{tabular}{rllll}
-\hline \hline
- & mean & range & mean & range \\ 
-  \hline
-precision & 0.645 & (0.635, 0.6668) & 0.8681 & (0.8639, 0.8712) \\ 
-  recall & 0.0864 & (0.083, 0.0887) & 0.5955 & (0.5899, 0.6004) \\ 
-  F1 score & 0.1524 & (0.1468, 0.1562) & 0.7064 & (0.7018, 0.7094) \\ 
-   \hline \hline
-\end{tabular}
-\caption{The predictive performance of the directed dyadic models with quartic trends, over ten 80/20 train/test splits.}
-\label{tab:prediction_quartic}
 \end{table}
@@
 This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.
+
+\textit{Comment from Benjamin Kassow: In section 6.2 (predictive performance), I like what is here, but was hoping to see a bit more discussion about the predictive 
+performance of the model, with some explanation of what the predictive performance means. 
+I understand it (I think), but depending on where this gets sent to, I am not sure if all reviewers will. 
+I also think tying this into the overall model explanation or discussion of model performance overall will be helpful for better 
+integrating this section into the rest of the manuscript.}
 

### 2018-03-26 Tex/paper.tex

@@
 $$\mathcal{C}_t =\{C_{\leq t} \in \{0,1\}^{(N_t \cdot (N_t-1))/2}: C_{\leq t}^{ij} \in \{0,1\} \}$$ 
-be the set of all possible adjacency matrices among $N_t$ cases.
+be the set of all possible adjacency matrices among $N_t$ cases at time $t$.
 $N_t$ denotes the the number of cases in the network at time $t$.
@@
 the decision regarding which network statistics $h(\cdot)$ to incorporate.
-We include the following statistics for the Supreme Court citation network, which are derived from the interdependence hypotheses described in the previous section:
+We include the following network statistics for the Supreme Court citation network, which are derived from the interdependence hypotheses described in the previous section:
 $$h_{edges}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}, $$
@@
 \end{align*}
-The number of triangles in the network.
-Triangles measure triad closure, as discussed above.
+The number of triangles in the network. Triangles measure triad closure, as discussed above.      
+$$h_{reciprocity}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}C_{\leq t}^{ji}, $$
+The number of reciprocal ties in the network.
+Even though the reciprocity of ties is not a meaningful network statistic for most citation networks, it may still play a role in the Supreme Court citation network.
+This statistic comes with the restriction that it only can appear among cases that entered the network at the same time.
+
 $$h_{covariate}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}X_{ij}, $$ the effect of an exogenous covariate ($X$).
@@
 Loops can exist in the Supreme Court citation network, but only among cases decided in the same term.
+Furthermore, a case $i$ can only cite a case $j$ if $j$ entered the network before or at the same time as $i$ did.
 The normalizing constant in the denominator of the probability function (i.e., Equation \ref{ercm}) is computationally intractable.
@@
 \end{tabular}
-\caption{Code Book for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
+\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
 \label{issue_area_coding}
@@
 \caption[caption]{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
-\\\hspace{\textwidth} ** J Roberts still serves as chief justice (3/2018).}
+\\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 3/2018).}
 \label{tab:chiefs}
@@
 \includegraphics[scale=0.35,clip=true,trim=.5cm 0cm 0cm 2cm]{images/citations1}
-\caption{ Supreme Court Citation Network, 1937?2001.
+\caption{ Supreme Court Citation Network, 1937-2001.
 Nodes are Supreme Court cases, with size based on incoming citations.
@@
 \end{figure}
+
+\begin{itemize}
+\item update Figure 2 to time period 1937 - 2015
+\end{itemize}
 
@@
 
- (e) Overruled Cases & (f) Difference in Years \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} \\
+ (e) Difference in Years & (f) Difference in Years$^2$ \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\
 
-   (g) Reciprocity & (h) Transitivity \\
+ (g) Overruled Cases &  \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} &  \\
+
+ \end{tabular}
+\caption{Trending effects for exogenous variables from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
+ \label{fig:coeftrends_exo}
+\vspace{-.25cm}
+\end{figure}
+
+\begin{figure}[H]
+ \begin{tabular}{cc}
+
+   (h) Reciprocity & (i) Transitivity \\
 \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
 
-    (i) Activity & (j) Popularity \\
+    (j) Activity & (k) Popularity \\
 \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
@@
 \end{tabular}
-\caption{Trending effects from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
- \label{fig:coeftrends}
+\caption{Trending effects for endogenous statistics from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
+ \label{fig:coeftrends_endo}
 \vspace{-.25cm}

### 2018-03-26 Tex/paper.tex
updated section: predictive performance

@@
 \subsection{c-TERGM Results}
-The results from the c-TERGM are presented in Table \ref{tab:tergmcoef} and Figure \ref{fig:coeftrends}.
+The results from the c-TERGM are presented in Table \ref{tab:tergmcoef} and Figures \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}.
 Table \ref{tab:tergmcoef} gives the coefficient estimates and 95\% bootstrap confidence intervals.
-In Figure, \ref{fig:coeftrends}, we depict coefficient estimates and confidence intervals for all of the effects for which we estimated trends (excluding the intercept---edge count---which is typically not interpreted substantively).
+In Figures, \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}, we depict coefficient estimates and confidence intervals for all of the effects for which we estimated trends (excluding the intercept---edge count---which is typically not interpreted substantively).
 The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one unit increase in the respective variable.
-We first discuss the dependence effect estimates (panels (e)--(h) in Figure \ref{fig:coeftrends}).
-The reciprocity effect is substantial, positive, and statistically significant for the entire period under study.
-Conditional upon cases $i$ and $j$ being considered within the same term, the log of the odds that opinion $i$ cites opinion $j$ increases by approximately 3 when opinion $j$ cites opinion $i$.
-The naive probability that any one case cites another case decided at the same time or previously in the Supreme Court Citation network is approximately 0.0024.
-This probability increases twenty-fold, to approximately 0.046 if the citation would be reciprocal.
+We first discuss the dependence effect estimates (panels (a)--(d) in Figure \ref{fig:coeftrends_endo}).
+The reciprocity effect is substantial, positive, and statistically significant for most of the period under study.
+Conditional upon cases $i$ and $j$ being considered within the same term, the log of the odds that opinion $i$ cites opinion $j$ increases by approximately $2$ when opinion $j$ cites opinion $i$.
+The naive probability that any one case cites another case decided at the same time or previously in the Supreme Court Citation network is approximately $0.0019$.
+This probability increases eight-fold, to approximately $0.015$ if the citation would be reciprocal.
 The transitivity (i.e., triangle closure) effect is also substantial, positive, and statistically significant over the entire period studied.
@@
 The activity and popularity effects are both estimated to be positive, but trend downward and are more modest in scale (and in the case of activity, not statistically significant later in the time series).
-Interpreting the popularity effect---the log odds that $i$ cites $j$ increases by 0.015--0.030 with every additional citation to case $j$.
+Interpreting the popularity effect---the log odds that $i$ cites $j$ increases by 0.018--0.035 with every additional citation to case $j$.
 Based on the in-degree distribution, it appears that the popularity of cases can be classified by multiples of 10.
-If case $j$ increases from ten to twenty citations from other cases, the probability that case $i$ cites case $j$ increases goes from approximately 0.0030 to 0.0040.
+If case $j$ increases from ten to twenty citations from other cases, the probability that case $i$ cites case $j$ increases from approximately 0.0030 to 0.0040.
 Overall, these results represent evidence of complex dependence in Supreme Court citation formation.
 
-The effects of the exogenous covariates included in the c-TERGM (panels (a)--(d) in Figure \ref{fig:coeftrends}) exhibit some surprising dynamics.
+The effects of the exogenous covariates included in the c-TERGM (panels (e)--(k) in Figure \ref{fig:coeftrends_exo}) exhibit some surprising dynamics.
 The effects of absolute difference in Martin-Quinn score, ideological breadth of the recipient case, and majority size of the recipient case all flip signs throughout the time period under study.
@@
 The scales of covariate and dependence effects are comparable.
-The simple linear trends represent one limitation of the current analysis.
+%The simple linear trends represent one limitation of the current analysis.
 In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.
+
+\begin{itemize}
+\item adjust results/numbers in text
+\end{itemize}
 
@@
  \begin{tabular}{cc}
- (a) MQ Score Difference & (b) Ideological Breadth \\
+
+   (a) Reciprocity & (b) Transitivity \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
+
+    (c) Activity & (d) Popularity \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
+
+\end{tabular}
+\caption{Trending effects for dependence effect estimates from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
+ \label{fig:coeftrends_endo}
+\vspace{-.25cm}
+\end{figure}
+
+  \begin{figure}[H]
+  \begin{tabular}{cc}
+ (e) MQ Score Difference & (f) Ideological Breadth \\
 \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\
 
-  (c) Same Issue & (d) Majority Size \\
+  (g) Same Issue & (h) Majority Size \\
 \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\
 
- (e) Difference in Years & (f) Difference in Years$^2$ \\
+ (i) Difference in Years & (j) Difference in Years$^2$ \\
 \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\
 
- (g) Overruled Cases &  \\
+ (k) Overruled Cases &  \\
 \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} &  \\
@@
  \label{fig:coeftrends_exo}
-\vspace{-.25cm}
-\end{figure}
-
-\begin{figure}[H]
- \begin{tabular}{cc}
-
-   (h) Reciprocity & (i) Transitivity \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
-
-    (j) Activity & (k) Popularity \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
-
-\end{tabular}
-\caption{Trending effects for endogenous statistics from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
- \label{fig:coeftrends_endo}
 \vspace{-.25cm}
@@
 \centering
+
 \begin{tabular}{rllll}
 \hline \hline
- & mean & range & mean & range \\ 
-  \hline
-precision & 0.5872 & (0.5781, 0.5962) & 0.8622 & (0.8593, 0.8668) \\ 
-  recall & 0.0565 & (0.0547, 0.0587) & 0.5839 & (0.5777, 0.592) \\ 
-  F1 score & 0.103 & (0.1, 0.1069) & 0.6963 & (0.6915, 0.7027) \\ 
-   \hline \hline
-\end{tabular}
-\caption{The predictive performance of the directed dyadic models with no trends, over ten 80/20 train/test splits.}
-\label{tab:prediction_linear}
-\begin{tabular}{rllll}
-\hline \hline
+& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
  & mean & range & mean & range \\ 
@@
 \label{tab:prediction_linear}
+
+\end{table}
+
+We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms.
+The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set.
+This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.
+
+\begin{table}[H]
+\centering
 \begin{tabular}{rllll}
 \hline \hline
+& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
+ & mean & range & mean & range \\ 
+  \hline
+precision & 0.5872 & (0.5781, 0.5962) & 0.8622 & (0.8593, 0.8668) \\ 
+  recall & 0.0565 & (0.0547, 0.0587) & 0.5839 & (0.5777, 0.592) \\ 
+  F1 score & 0.103 & (0.1, 0.1069) & 0.6963 & (0.6915, 0.7027) \\ 
+   \hline \hline
+\end{tabular}
+\caption{The predictive performance of the directed dyadic models with no trends, over ten 80/20 train/test splits.}
+\label{tab:prediction_notrend}
+\vspace{0.4cm}
+\begin{tabular}{rllll}
+\hline \hline
+& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
  & mean & range & mean & range \\ 
@@
 \end{table}
-
-We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms.
-The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set.
-This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.
+We also conduct a predective experiment for the model that does not consider any trends and for the model that does consider quadratic instead of linear trends.
+The predictive performance results are given in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic}.
+By comparing the full model results in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic} with the results of the full linear trend model in table \ref{tab:prediction_linear} we can conclude all three models perform similarly well.
+\\[0,5cm]
 
@@
 This methodology---the citation-TERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation.
-We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2001.
+We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2015.
 We find, somewhat counterintuitively, that Supreme Court citations are highly reciprocal.
@@
 Overcoming this limitation would require methodological innovation in the c-TERGM as well as theoretical development regarding exactly how citation structure might predict case emergence.
-Third, the data used in this study does not cover the most recent history of the Court.
+%Third, the data used in this study does not cover the most recent history of the Court.
 Our data ends in 2001---missing the tenure of the most recent chief justice, and more than fifteen terms.

### 2018-04-09 Tex/paper.tex

@@
 First, dyadic citations are typically aggregated to the case level before they are analyzed.
-Second, citaitons are treated as if they arise independent of one another.
+Second, citations are treated as if they arise independent of one another.
 We make the case that studying Supreme Court citations as a dynamic citation network overcomes both of these limitations.
@@
 In this opinion passage Baker v.
-Carr is cited in reference to its role as a landmark precedent, and noted for the number of other cases by which it has been followed.
-which an authoritative opinion is referenced, and even discussed in terms of the number of other cases by which it was followed.
+Carr is cited in reference to its role as a landmark precedent, and noted for the number of other cases by which it has been followed and for which an authoritative opinion is referenced, and even discussed in terms of the number of other cases by which it was followed.
 \begin{quotation}
@@
 The Supreme Court citation network is nearly a directed acyclic graph---a graph in which there can be no loops/cycles (e.g., ties cannot be reciprocated, which would constitute a two-node loop).
-If two cases are decided in different terms, the later case can cite the earlier case, but not the reverse.
+If two cases are decided in different terms, the latter case can cite the earlier case, but not the reverse.
 However, as noted above, if two opinions are written in the same term, they can cite each other reciprocally.
@@
 be the set of all possible adjacency matrices among $N_t$ cases at time $t$.
-$N_t$ denotes the the number of cases in the network at time $t$.
+$N_t$ denotes the number of cases in the network at time $t$.
 Note that the cardinality of $\mathcal{C}_t$ increases exponentially for every newly added case.
@@
 We include these variables because \citet{cross2010determinants} find that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
-Table \ref{issue_area_coding} states which number represents which issie area.
+Table \ref{issue_area_coding} states which number represents which issue area.
 
@@
 Our three data sources for this study include the Supreme Court Database (SCDB) \citep{spaeth2014supreme}, Martin-Quinn scores \citep{martin2002dynamic}, and Supreme Court citation data provided by the CourtListener Free Law Project \citep{CourtListener}.
-We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015)\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the Courtlistener data.
+We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015)\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the CourtListener data.
 We decided to exclude these 145 cases from our model.}.
@@
 The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
-The network has a total of $111,986$ ties.
+In order to only account for positive citations, we exclude $315$ citations that caused the cited case to be overruled.
+As a result, the network has a total of $111,986$ ties.
 The number of triangles in the network is $250,717$ and the number of mutual ties is $56$.
@@
 \item update Figure 2 to time period 1937 - 2015
+\item add data source for overruling citations
 \end{itemize}
@@
 \end{table}
-We also conduct a predective experiment for the model that does not consider any trends and for the model that does consider quadratic instead of linear trends.
+We also conduct a predictive experiment for the model that does not consider any trends and for the model that does consider quadratic instead of linear trends.
 The predictive performance results are given in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic}.

### 2018-04-09 NetworkVisualizations/visualizations.R
R code for network visualization

@@ -1,0 +1,469 @@
+#Preparation
+rm(list=ls())
+#setwd("C:/Users/TC/Dropbox/28_Citation Networks/R")
+
+library(statnet)
+
+load("supreme.RData")
+rm(model20.mle)
+
+#Distinguishing between different courts
+#https://www.supremecourt.gov/about/members_text.aspx
+cj<-list(
+  hughes=seq(which(node_attributes$usid=="280US390"),which(node_attributes$usid=="313US508")), #February 24, 1930	
+  stone=seq(which(node_attributes$usid=="314US1"),which(node_attributes$usid=="328US781")), #July 3, 1941	
+  vinson=seq(which(node_attributes$usid=="329US1"),which(node_attributes$usid=="346US324")), #June 24, 1946	
+  warren=seq(which(node_attributes$usid=="346US325"),which(node_attributes$usid=="395US710")), #October 5, 1953	
+  burger=seq(which(node_attributes$usid=="395US711"),which(node_attributes$usid=="478US1311")), #June 23, 1969	
+  rehnquist=seq(which(node_attributes$usid=="479US1"),which(node_attributes$usid=="534US533")) #September 26, 1986
+  #Roberts: September 29, 2005
+)
+
+court<-rep(0,nrow(node_attributes))
+for(i in 1:length(cj)){
+  court[cj[[i]]]<-i
+}
+
+#Some nodal attributes
+scnetwork%v%"caseid"<-node_attributes$caseid
+scnetwork%v%"court"<-court
+scnetwork%v%"indeg"<-node_attributes$indeg
+scnetwork%v%"auth"<-node_attributes$auth
+scnetwork%v%"iso"<-ifelse(degree(scnetwork)==0,1,0)
+
+#Subsetting the network objects for different figures
+#subset to 1937-2002 (Figure 2)
+net<-scnetwork%s%which(scnetwork%v%"year">1936)
+el.net<-edgelist[which((edgelist[,1]%in%(net%v%"caseid"))&(edgelist[,2]%in%(net%v%"caseid"))),]
[…]

### 2018-04-09 Tex/paper.tex
fixed some typos

@@
 
-$$h_{covariate}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}X_{ij}, $$ the effect of an exogenous covariate ($X$).
+$$h_{covariate}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}X_{\leq t}^{ij}, $$ the effect of an exogenous covariate ($X$).
 We include several exogenous covariates based on this standard statistic formulation.
-These covariates are discussed below.
-
-The structure of this model is very similar to a conventional (T)ERGM setup.
+These covariates are discussed below.\\
+\indent The structure of this model is very similar to a conventional (T)ERGM setup.
 The main difference is that  $\mathcal{C}_t$ excludes adjacency matrices in which there are loops that include edges sent at different time points.
@@
 Furthermore, a case $i$ can only cite a case $j$ if $j$ entered the network before or at the same time as $i$ did.
-The normalizing constant in the denominator of the probability function (i.e., Equation \ref{ercm}) is computationally intractable.
+\\
+\indent The normalizing constant in the denominator of the probability function (i.e., Equation \ref{ercm}) is computationally intractable.
 This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with ERGM family models.
@@
 We include two sets of dummy variables that account for the issue areas of cases.
-In one set of dummy variable---sender intercepts---the variable $X_{ij}$ indicates the issue area of the sending case ($i$).
-In the set of receiver intercepts $X_{ij}$ indicates the issue are of the receiver case.
+In one set of dummy variable---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$).
+In the set of receiver intercepts $X^{ij}$ indicates the issue of the receiver case.
 Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
-We include these variables because \citet{cross2010determinants} find that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
-Table \ref{issue_area_coding} states which number represents which issue area.
+We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
+Table \ref{issue_area_coding} states the different issue areas.
 
@@
 We model the way in which citations to a case depend upon the age of a case.
-For this we use a second-order polynomial in which one covariate $X_{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X_{ij}$ is the squared age of case $j$ at the time that case $i$ is decided.
+For this we use a second-order polynomial in which one covariate $X^{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X^{ij}$ is the squared age of case $j$ at the time that case $i$ is decided.
 We include these covariates because \citet{black2013citation} find that the number of citations to a Supreme Court case over time depends significantly on the age of the case, characterized by a sharp drop off and leveling out with age.
@@
 In our data we do not differentiate between negative and positive citations, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis.
-We include one covariate in which $X_{ij}$ is the number of justices in the majority coalition for case $j$.
-We also include another covariate in which $X_{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$.
+We include one covariate in which $X^{ij}$ is the number of justices in the majority coalition for case $j$.
+We also include another covariate in which $X^{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$.
 We expect both of these covariates to have positive effects.
@@
 We decided to exclude these 145 cases from our model.}.
-The Supreme Court citation network from $1937 - 2015$ consists of $9,945$ cases.
-The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
+The Supreme Court citation network from $1937 - 2015$ consists of $9,945$ cases and the breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
 In order to only account for positive citations, we exclude $315$ citations that caused the cited case to be overruled.
@@
 The effects of the exogenous covariates included in the c-TERGM (panels (e)--(k) in Figure \ref{fig:coeftrends_exo}) exhibit some surprising dynamics.
-The effects of absolute difference in Martin-Quinn score, ideological breadth of the recipient case, and majority size of the recipient case all flip signs throughout the time period under study.
+The effects of absolute difference in Martin-Quinn score, and majority size of the recipient case all flip signs throughout the time period under study.
 This may be an artifact of the simple linear trends we have estimated, but is interesting nonetheless.

### 2018-04-12 Tex/paper.tex

@@
 
+\begin{itemize}
+\item add to the discussion to which degree network statistics reflect causual relation
+\end{itemize}
+
 \section{Network Approaches to Studying Citations}
@@
 \begin{itemize}
-\item ToDo: mention new variable: overruled cases
+\item mention new variable: overruled cases
+\item mention new variable: nodemix(same issue area)
+\item mention new variable:Sender/receiver effect for judge who wrote majority opinion
+\item Same justice homophily with time trend
+\item Argue/justify why certain covariates are included
 \end{itemize}
@@
 
-\subsection{c-TERGM Results}
-The results from the c-TERGM are presented in Table \ref{tab:tergmcoef} and Figures \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}.
-Table \ref{tab:tergmcoef} gives the coefficient estimates and 95\% bootstrap confidence intervals.
-In Figures, \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}, we depict coefficient estimates and confidence intervals for all of the effects for which we estimated trends (excluding the intercept---edge count---which is typically not interpreted substantively).
-The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one unit increase in the respective variable.
-We first discuss the dependence effect estimates (panels (a)--(d) in Figure \ref{fig:coeftrends_endo}).
-The reciprocity effect is substantial, positive, and statistically significant for most of the period under study.
-Conditional upon cases $i$ and $j$ being considered within the same term, the log of the odds that opinion $i$ cites opinion $j$ increases by approximately $2$ when opinion $j$ cites opinion $i$.
-The naive probability that any one case cites another case decided at the same time or previously in the Supreme Court Citation network is approximately $0.0019$.
-This probability increases eight-fold, to approximately $0.015$ if the citation would be reciprocal.
-The transitivity (i.e., triangle closure) effect is also substantial, positive, and statistically significant over the entire period studied.
-The probability that case $i$ cites case $j$ increases approximately five-fold if the citation would close a triangle involving cases $i$ and $j$.
-The activity and popularity effects are both estimated to be positive, but trend downward and are more modest in scale (and in the case of activity, not statistically significant later in the time series).
-Interpreting the popularity effect---the log odds that $i$ cites $j$ increases by 0.018--0.035 with every additional citation to case $j$.
-Based on the in-degree distribution, it appears that the popularity of cases can be classified by multiples of 10.
-If case $j$ increases from ten to twenty citations from other cases, the probability that case $i$ cites case $j$ increases from approximately 0.0030 to 0.0040.
-Overall, these results represent evidence of complex dependence in Supreme Court citation formation.
-
-The effects of the exogenous covariates included in the c-TERGM (panels (e)--(k) in Figure \ref{fig:coeftrends_exo}) exhibit some surprising dynamics.
-The effects of absolute difference in Martin-Quinn score, and majority size of the recipient case all flip signs throughout the time period under study.
-This may be an artifact of the simple linear trends we have estimated, but is interesting nonetheless.
-The attitudinal model holds that ideological preferences drive decision-making on the US Supreme Court.
-Ideological selection of precedents would be evidenced by a negative effect of the absolute difference in MQ scores in the c-TERGM.
-We find this effect is negative and statistically significant only after 1975.
-The absolute difference in MQ scores variable is on the scale of approximately 0---3.
-At its most ideological point (i.e.
-2001), the probability that case $i$ cites case $j$ decreases from approximately 0.024 to 0.018---by one third---if the ideological distance between the majority coalitions in the two cases increases by 3.
-The effects of the majority size and ideological breadth of case $j$ are similar in both scale and trend.
-It is not until later in the time series that they both exhibit their expected positive effects.
-The effect of same issue is positive and statistically significant, and substantial for the time period under study.
-The probability that $i$ cites $j$ increases by approximately three-fold over the naive probability of citation if $i$ and $j$ have the same legal issue area.
-
-The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes.
-The scales of covariate and dependence effects are comparable.
-%The simple linear trends represent one limitation of the current analysis.
-In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.
-
-\begin{itemize}
-\item adjust results/numbers in text
-\end{itemize}
-
-\begin{table}[htp]
-\footnotesize
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{EFEFEF}}l |l|l|l|l|} 
-\hline
-                                                   & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}Lower Bound & \cellcolor[HTML]{EFEFEF}Upper Bound & \cellcolor[HTML]{EFEFEF}Significance \\
-                                                    \hline
-Edges                                              & -6.249                           & -6.412                              & -6.110                              & *                                    \\ \hline
-Instar(2)                                          & 0.035                            & 0.032                               & 0.038                               & *                                    \\ \hline
-Outstar(2)                                         & 0.021                            & 0.019                               & 0.028                               & *                                    \\ \hline
-Mutual                                             & 2.071                            & -4.157                               & 4.072                                &                                     \\ \hline
-Triangle                                           & 1.507                            & 1.431                               & 1.559                               & *                                    \\ \hline
-Martin Quinn Score                                 & 0.079                            & 0.029                               & 0.130                               & *                                    \\ \hline
-Same Issue Area Homophily                          & 1.447                            & 1.389                               & 1.515                               & *                                    \\ \hline
-Year Difference                                    & -0.068                           & -0.079                              & -0.058                              & *                                    \\ \hline
-$(\text{Year Differnce})^2$                        & 0.0019                           & 0.0016                              & 0.0022                              & *                                    \\ \hline
-Receiver Abs Diff MQ Score in Majority             & 0.047                           & 0.029                              & 0.063                              & *                                     \\ \hline
-Receiver Number Justices in Majority               & -0.075                           & -0.095                              & -0.055                              & *                                    \\ \hline
-Receiver  Sender Year                              & 0.0039                           & 0.0036                              & 0.0042                              & *                                    \\ \hline
-Overruled Cases   & -1.043                           & -1.624                              & -0.564                              & *                                    \\ \hline
-Sender Same Issue Area 2                           & 0.142                            & 0.100                               & 0.185                               & *                                    \\ \hline
-Sender Same Issue Area 3                           & -0.337                           & -0.403                              & -0.270                              & *                                    \\ \hline
-Sender Same Issue Area 4                           & 0.498                            & 0.438                               & 0.561                               & *                                    \\ \hline
-Sender Same Issue Area 5                           & 0.182                            & 0.031                               & 0.305                               & *                                    \\ \hline
-Sender Same Issue Area 6                           & 0.536                            & 0.422                               & 0.648                               & *                                    \\ \hline
-Sender Same Issue Area 7                           & 0.370                            & 0.316                               & 0.426                               & *                                    \\ \hline
-Sender Same Issue Area 8                           & 0.107                            & 0.059                               & 0.156                               & *                                    \\ \hline
-Sender Same Issue Area 9                           & 0.278                            & 0.238                               & 0.329                               & *                                    \\ \hline
-Sender Same Issue Area 10                          & 0.371                            & 0.317                               & 0.433                               & *                                    \\ \hline
-Sender Same Issue Area 11                          & -0.030                           & -0.151                              & 0.164                               &                                      \\ \hline
-Sender Same Issue Area 12                          & 0.064                           & -0.014                              & 0.130                               &                                      \\ \hline
-Sender Same Issue Area 13                          & 0.608                            & 0.423                               & 0.792                               & *                                    \\ \hline
-Sender Same Issue Area 14                          & 0.232                            & 0.075                              & 0.393                               & *                                      \\ \hline
-Receiver Same Issue Area 2                         & 0.252                            & 0.199                               & 0.310                               & *                                    \\ \hline
-Receiver Same Issue Area 3                         & -0.092                           & -0.194                              & 0.014                              &                                     \\ \hline
-Receiver Same Issue Area 4                         & 0.481                            & 0.397                               & 0.570                               & *                                    \\ \hline
-Receiver Same Issue Area 5                         & 0.374                            & 0.224                               & 0.519                               & *                                    \\ \hline
-Receiver Same Issue Area 6                         & 0.521                            & 0.346                               & 0.652                               & *                                    \\ \hline
-Receiver Same Issue Area 7                         & 0.412                            & 0.339                               & 0.487                               & *                                    \\ \hline
-Receiver Same Issue Area 8                         & 0.174                            & 0.134                               & 0.232                               & *                                    \\ \hline
-Receiver Same Issue Area 9                         & 0.311                            & 0.240                               & 0.361                               & *                                    \\ \hline
-Receiver Same Issue Area 10                        & 0.506                            & 0.429                               & 0.577                               & *                                    \\ \hline
-Receiver Same Issue Area 11                        & 0.299                            & 0.088                               & 0.432                               & *                                    \\ \hline
-Receiver Same Issue Area 12                        & 0.350                            & 0.276                               & 0.445                               & *                                    \\ \hline
-Receiver Same Issue Area 13                        & 0.972                            & 0.780                               & 1.135                               & *                                    \\ \hline
-Receiver Same Issue Area 14                        & 0.517                            & 0.200                               & 0.740                               & *                                    \\ \hline
-Instar(2) $\times$ Sender Year                     & -0.00026                         & -0.00031                            & -0.00021                            & *                                    \\ \hline
-Outstar(2) $\times$ Sender Year                    & -0.00031                         & -0.00045                            & -0.00027                            & *                                    \\ \hline
-Mutual $\times$ Sender Year                        & 0.110                           & 0.010                              & 0.524                               & *                                      \\ \hline
-Triangle $\times$ Sender Year                      & -0.00022                          & -0.0016                             & 0.0013                              &                                      \\ \hline
-Martin Quinn Score $\times$ Sender Year            & -0.0019                          & -0.0030                             & -0.0009                             & *                                    \\ \hline
-Same Issue Area $\times$ Sender Year               & -0.0068                          & -0.0081                             & -0.0055                             & *                                    \\ \hline
-Year Difference $\times$ Sender Year               & 0.00046                           & 0.00028                              & 0.00065                              & *                                    \\ \hline
-$\text{Year Difference}^2 \times$ Sender Year      & -0.000024                        & -0.000028                           & -0.000021                           & *                                    \\ \hline
-MQ Score in Majority $\times$ Sender Year & -0.00053                           & -0.00093                              & -0.0018                              & *                                    \\ \hline
-Justices in Majority $\times$ Sender Year   & 0.0017                           & 0.0013                              & 0.0021                              & *                                    \\ \hline
-Overruled Cases $\times$ Sender Year   & 0.012                           & 0.002                              & 0.023                              & *                                    \\ \hline
-\end{tabular}
-\caption{Bootstrapped MPLE Results for the time period $1937-2015$.
-A '*' indicates that the 2.5th and 97.5th quantile of the variable does not include '0' and as a result is statistically significant.}
-\label{tab:tergmcoef}
-\end{table}
-\newpage
-
-\begin{figure}[H]
- \begin{tabular}{cc}
-
-   (a) Reciprocity & (b) Transitivity \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
-
-    (c) Activity & (d) Popularity \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
-
-\end{tabular}
-\caption{Trending effects for dependence effect estimates from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
- \label{fig:coeftrends_endo}
-\vspace{-.25cm}
-\end{figure}
-
-  \begin{figure}[H]
-  \begin{tabular}{cc}
- (e) MQ Score Difference & (f) Ideological Breadth \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\
-
-  (g) Same Issue & (h) Majority Size \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\
-
- (i) Difference in Years & (j) Difference in Years$^2$ \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\
-
- (k) Overruled Cases &  \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} &  \\
-
- \end{tabular}
-\caption{Trending effects for exogenous variables from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
- \label{fig:coeftrends_exo}
-\vspace{-.25cm}
-\end{figure}
-
 \subsection{Predictive Performance}
@@
 
+\subsection{c-TERGM Results}
+The results from the c-TERGM are presented in Table \ref{tab:tergmcoef} and Figures \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}.
+Table \ref{tab:tergmcoef} gives the coefficient estimates and 95\% bootstrap confidence intervals.
+In Figures, \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}, we depict coefficient estimates and confidence intervals for all of the effects for which we estimated trends (excluding the intercept---edge count---which is typically not interpreted substantively).
+The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one unit increase in the respective variable.
+We first discuss the dependence effect estimates (panels (a)--(d) in Figure \ref{fig:coeftrends_endo}).
+The reciprocity effect is substantial, positive, and statistically significant for most of the period under study.
+Conditional upon cases $i$ and $j$ being considered within the same term, the log of the odds that opinion $i$ cites opinion $j$ increases by approximately $2$ when opinion $j$ cites opinion $i$.
+The naive probability that any one case cites another case decided at the same time or previously in the Supreme Court Citation network is approximately $0.0019$.
+This probability increases eight-fold, to approximately $0.015$ if the citation would be reciprocal.
+The transitivity (i.e., triangle closure) effect is also substantial, positive, and statistically significant over the entire period studied.
+The probability that case $i$ cites case $j$ increases approximately five-fold if the citation would close a triangle involving cases $i$ and $j$.
+The activity and popularity effects are both estimated to be positive, but trend downward and are more modest in scale (and in the case of activity, not statistically significant later in the time series).
+Interpreting the popularity effect---the log odds that $i$ cites $j$ increases by 0.018--0.035 with every additional citation to case $j$.
+Based on the in-degree distribution, it appears that the popularity of cases can be classified by multiples of 10.
+If case $j$ increases from ten to twenty citations from other cases, the probability that case $i$ cites case $j$ increases from approximately 0.0030 to 0.0040.
+Overall, these results represent evidence of complex dependence in Supreme Court citation formation.
+
+The effects of the exogenous covariates included in the c-TERGM (panels (e)--(k) in Figure \ref{fig:coeftrends_exo}) exhibit some surprising dynamics.
+The effects of absolute difference in Martin-Quinn score, and majority size of the recipient case all flip signs throughout the time period under study.
+This may be an artifact of the simple linear trends we have estimated, but is interesting nonetheless.
+The attitudinal model holds that ideological preferences drive decision-making on the US Supreme Court.
+Ideological selection of precedents would be evidenced by a negative effect of the absolute difference in MQ scores in the c-TERGM.
+We find this effect is negative and statistically significant only after 1975.
+The absolute difference in MQ scores variable is on the scale of approximately 0---3.
+At its most ideological point (i.e.
+2001), the probability that case $i$ cites case $j$ decreases from approximately 0.024 to 0.018---by one third---if the ideological distance between the majority coalitions in the two cases increases by 3.
+The effects of the majority size and ideological breadth of case $j$ are similar in both scale and trend.
+It is not until later in the time series that they both exhibit their expected positive effects.
+The effect of same issue is positive and statistically significant, and substantial for the time period under study.
+The probability that $i$ cites $j$ increases by approximately three-fold over the naive probability of citation if $i$ and $j$ have the same legal issue area.
+
+The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes.
+The scales of covariate and dependence effects are comparable.
+%The simple linear trends represent one limitation of the current analysis.
+In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.
+
+\begin{itemize}
+\item adjust results/numbers in text
+\end{itemize}
+
+\begin{figure}[H]
+ \begin{tabular}{cc}
+
+   (a) Reciprocity & (b) Transitivity \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
+
+    (c) Activity & (d) Popularity \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
+
+\end{tabular}
+\caption{Trending effects for dependence effect estimates from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
+ \label{fig:coeftrends_endo}
+\vspace{-.25cm}
+\end{figure}
+
+  \begin{figure}[H]
+  \begin{tabular}{cc}
+ (e) MQ Score Difference & (f) Ideological Breadth \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\
+
+  (g) Same Issue & (h) Majority Size \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\
+
+ (i) Difference in Years & (j) Difference in Years$^2$ \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\
+
+ (k) Overruled Cases &  \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} &  \\
+
+ \end{tabular}
+\caption{Trending effects for exogenous variables from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
+ \label{fig:coeftrends_exo}
+\vspace{-.25cm}
+\end{figure}
+
+\begin{itemize}
+\item visualize Sender/receiver Issue Area Results
+\item heatplot for issue area nodemix
+\item visualize Sender/receiver for majority opinion writer
+\item add Network effect plots for model without network statistics (add to existing plots?)
+\end{itemize}
+
 \section{Conclusion}
@@
 To overcome this limitation we would need to extend the citation data beyond that collected by Fowler and colleagues.
+
+\section{Appendix}
+The following table presents the Bootstrapped MPLE Results for the time period $1937-2015$.
+
+\begin{itemize}
+\item add robustness check: MCMLE for each year -> plot time trend of coefs -> are results similar to the MPLE time trends presented in the paper?
+\item (investigate which statistic contributes most on recall improvement)
+\end{itemize}
+
+\begin{table}[htp]
+\footnotesize
+\centering
+\begin{tabular}{|
+>{\columncolor[HTML]{EFEFEF}}l |l|l|l|l|} 
+\hline
+                                                   & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}Lower Bound & \cellcolor[HTML]{EFEFEF}Upper Bound & \cellcolor[HTML]{EFEFEF}Significance \\
+                                                    \hline
+Edges                                              & -6.249                           & -6.412                              & -6.110                              & *                                    \\ \hline
+Instar(2)                                          & 0.035                            & 0.032                               & 0.038                               & *                                    \\ \hline
+Outstar(2)                                         & 0.021                            & 0.019                               & 0.028                               & *                                    \\ \hline
+Mutual                                             & 2.071                            & -4.157                               & 4.072                                &                                     \\ \hline
+Triangle                                           & 1.507                            & 1.431                               & 1.559                               & *                                    \\ \hline
+Martin Quinn Score                                 & 0.079                            & 0.029                               & 0.130                               & *                                    \\ \hline
+Same Issue Area Homophily                          & 1.447                            & 1.389                               & 1.515                               & *                                    \\ \hline
+Year Difference                                    & -0.068                           & -0.079                              & -0.058                              & *                                    \\ \hline
+$(\text{Year Differnce})^2$                        & 0.0019                           & 0.0016                              & 0.0022                              & *                                    \\ \hline
+Receiver Abs Diff MQ Score in Majority             & 0.047                           & 0.029                              & 0.063                              & *                                     \\ \hline
+Receiver Number Justices in Majority               & -0.075                           & -0.095                              & -0.055                              & *                                    \\ \hline
+Receiver  Sender Year                              & 0.0039                           & 0.0036                              & 0.0042                              & *                                    \\ \hline
+Overruled Cases   & -1.043                           & -1.624                              & -0.564                              & *                                    \\ \hline
+Sender Same Issue Area 2                           & 0.142                            & 0.100                               & 0.185                               & *                                    \\ \hline
+Sender Same Issue Area 3                           & -0.337                           & -0.403                              & -0.270                              & *                                    \\ \hline
+Sender Same Issue Area 4                           & 0.498                            & 0.438                               & 0.561                               & *                                    \\ \hline
+Sender Same Issue Area 5                           & 0.182                            & 0.031                               & 0.305                               & *                                    \\ \hline
+Sender Same Issue Area 6                           & 0.536                            & 0.422                               & 0.648                               & *                                    \\ \hline
+Sender Same Issue Area 7                           & 0.370                            & 0.316                               & 0.426                               & *                                    \\ \hline
+Sender Same Issue Area 8                           & 0.107                            & 0.059                               & 0.156                               & *                                    \\ \hline
+Sender Same Issue Area 9                           & 0.278                            & 0.238                               & 0.329                               & *                                    \\ \hline
+Sender Same Issue Area 10                          & 0.371                            & 0.317                               & 0.433                               & *                                    \\ \hline
+Sender Same Issue Area 11                          & -0.030                           & -0.151                              & 0.164                               &                                      \\ \hline
+Sender Same Issue Area 12                          & 0.064                           & -0.014                              & 0.130                               &                                      \\ \hline
+Sender Same Issue Area 13                          & 0.608                            & 0.423                               & 0.792                               & *                                    \\ \hline
+Sender Same Issue Area 14                          & 0.232                            & 0.075                              & 0.393                               & *                                      \\ \hline
+Receiver Same Issue Area 2                         & 0.252                            & 0.199                               & 0.310                               & *                                    \\ \hline
+Receiver Same Issue Area 3                         & -0.092                           & -0.194                              & 0.014                              &                                     \\ \hline
+Receiver Same Issue Area 4                         & 0.481                            & 0.397                               & 0.570                               & *                                    \\ \hline
+Receiver Same Issue Area 5                         & 0.374                            & 0.224                               & 0.519                               & *                                    \\ \hline
+Receiver Same Issue Area 6                         & 0.521                            & 0.346                               & 0.652                               & *                                    \\ \hline
+Receiver Same Issue Area 7                         & 0.412                            & 0.339                               & 0.487                               & *                                    \\ \hline
+Receiver Same Issue Area 8                         & 0.174                            & 0.134                               & 0.232                               & *                                    \\ \hline
+Receiver Same Issue Area 9                         & 0.311                            & 0.240                               & 0.361                               & *                                    \\ \hline
+Receiver Same Issue Area 10                        & 0.506                            & 0.429                               & 0.577                               & *                                    \\ \hline
+Receiver Same Issue Area 11                        & 0.299                            & 0.088                               & 0.432                               & *                                    \\ \hline
+Receiver Same Issue Area 12                        & 0.350                            & 0.276                               & 0.445                               & *                                    \\ \hline
+Receiver Same Issue Area 13                        & 0.972                            & 0.780                               & 1.135                               & *                                    \\ \hline
+Receiver Same Issue Area 14                        & 0.517                            & 0.200                               & 0.740                               & *                                    \\ \hline
+Instar(2) $\times$ Sender Year                     & -0.00026                         & -0.00031                            & -0.00021                            & *                                    \\ \hline
+Outstar(2) $\times$ Sender Year                    & -0.00031                         & -0.00045                            & -0.00027                            & *                                    \\ \hline
+Mutual $\times$ Sender Year                        & 0.110                           & 0.010                              & 0.524                               & *                                      \\ \hline
+Triangle $\times$ Sender Year                      & -0.00022                          & -0.0016                             & 0.0013                              &                                      \\ \hline
+Martin Quinn Score $\times$ Sender Year            & -0.0019                          & -0.0030                             & -0.0009                             & *                                    \\ \hline
+Same Issue Area $\times$ Sender Year               & -0.0068                          & -0.0081                             & -0.0055                             & *                                    \\ \hline
+Year Difference $\times$ Sender Year               & 0.00046                           & 0.00028                              & 0.00065                              & *                                    \\ \hline
+$\text{Year Difference}^2 \times$ Sender Year      & -0.000024                        & -0.000028                           & -0.000021                           & *                                    \\ \hline
+MQ Score in Majority $\times$ Sender Year & -0.00053                           & -0.00093                              & -0.0018                              & *                                    \\ \hline
+Justices in Majority $\times$ Sender Year   & 0.0017                           & 0.0013                              & 0.0021                              & *                                    \\ \hline
+Overruled Cases $\times$ Sender Year   & 0.012                           & 0.002                              & 0.023                              & *                                    \\ \hline
+\end{tabular}
+\caption{Bootstrapped MPLE Results for the time period $1937-2015$.
+A '*' indicates that the 2.5th and 97.5th quantile of the variable does not include '0' and as a result is statistically significant.}
+\label{tab:tergmcoef}
+\end{table}
+\newpage
 

### 2018-04-14 NetworkVisualizations/visualizations.R
update to network visualization

@@ -1,469 +1,147 @@
-#Preparation
+#----Code to produce citation network plot----#
+# The 'network' package is required for the force-directed algorithm on line 57
+
+#----Preparation----#
+# Set Directory
 rm(list=ls())
-#setwd("C:/Users/TC/Dropbox/28_Citation Networks/R")
+setwd("C:/Users/TC/Dropbox/28_Citation Networks/R")
 
-library(statnet)
+# Loading Data
+load("SCOTUS.RData")
+rm(mq.matrix,Overruled.matrix,same.issue.area,SCDB_justice,year.diff.matrix)
 
-load("supreme.RData")
-rm(model20.mle)
+# Defining Colour
+blue.c<-rgb(0.42,0.71,1)
 
-#Distinguishing between different courts
-#https://www.supremecourt.gov/about/members_text.aspx
-cj<-list(
-  hughes=seq(which(node_attributes$usid=="280US390"),which(node_attributes$usid=="313US508")), #February 24, 1930	
-  stone=seq(which(node_attributes$usid=="314US1"),which(node_attributes$usid=="328US781")), #July 3, 1941	
-  vinson=seq(which(node_attributes$usid=="329US1"),which(node_attributes$usid=="346US324")), #June 24, 1946	
-  warren=seq(which(node_attributes$usid=="346US325"),which(node_attributes$usid=="395US710")), #October 5, 1953	
-  burger=seq(which(node_attributes$usid=="395US711"),which(node_attributes$usid=="478US1311")), #June 23, 1969	
-  rehnquist=seq(which(node_attributes$usid=="479US1"),which(node_attributes$usid=="534US533")) #September 26, 1986
-  #Roberts: September 29, 2005
-)
+# Preparing Network
+# Remove Isolates (there aren't many (348), and it helps with the rest of the visualization)
+isolates<-which(rowSums(adjacency.matrix)+colSums(adjacency.matrix)==0)
+mat<-adjacency.matrix[-isolates,-isolates]
+rm(adjacency.matrix)
 
-court<-rep(0,nrow(node_attributes))
-for(i in 1:length(cj)){
[…]

### 2018-04-26 Tex/paper.tex

@@
 \caption[caption]{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
-\\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 3/2018).}
+\\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 4/2018).}
 \label{tab:chiefs}
@@
 
- (k) Overruled Cases &  \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} &  \\
+ (k) Overruled Cases & (l) Justice Homophily \\
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} & 
+\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/justicehomophily_coef_trend_linear.pdf} \\
 

### 2018-04-26 Tex/paper.tex
results of new model

@@
 \end{itemize}
+
+  \begin{figure}[H]
+\includegraphics[width=15cm]{justice_popularity}
+\caption{Nodeifactor Majority Opinion Writer. Baseline= LDBrandeis }
+ \label{number_supporting}
+\vspace{-.25cm}
+\end{figure}
+
+ \begin{figure}[H]
+\includegraphics[width=15cm]{justice_activity}
+\caption{Nodeofactor Majority Opinion Writer. Baseline= LDBrandeis }
+ \label{number_supporting}
+\vspace{-.25cm}
+\end{figure}
+
+ \begin{figure}[H]
+  \begin{tabular}{c}
+\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_activity}
+ \\
+\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_popularity}
+\end{tabular}
+\caption{Issue Area Popularity and Activity (nodeifactor and nodeofactor). Baseline is Issue Area 'Criminal Procedure'.}
+\end{figure}
+
+ \begin{figure}[H]
+\includegraphics[width=15cm]{heatplot_issue_area}
+\caption{Nodemix Issue Area. Sender14Receiver14 and Sender14Receiver13 couldnt be estimated.}
+ \label{number_supporting}
+\vspace{-.25cm}
+\end{figure}
 

### 2018-04-26 R-Code_courtlistener/data_merging_and_bootMPLE.R

@@ -37,6 +37,7 @@ mpledata$yeardiffsquare_sendertime<- mpledata[,10]*mpledata[,14]
 mpledata$AbsDiffMQscores_sendertime<- mpledata[,16]*mpledata[,14]
 mpledata$NumberJusticesPro_sendertime<- mpledata[,17]*mpledata[,14]
 mpledata$Overruled_sendertime<- mpledata[,44]*mpledata[,14]
+mpledata$jusitcehomophily_sendertime<- mpledata[,45]*mpledata[,14] 
 
 # 2nd degree
 mpledata$instar2_sendertime2<- mpledata[,3]*mpledata[,14]*mpledata[,14]
@@ -50,6 +51,7 @@ mpledata$yeardiffsquare_sendertime2<- mpledata[,10]*mpledata[,14]*mpledata[,14]
 mpledata$AbsDiffMQscores_sendertime2<- mpledata[,16]*mpledata[,14]*mpledata[,14]
 mpledata$NumberJusticesPro_sendertime2<- mpledata[,17]*mpledata[,14]*mpledata[,14]
 mpledata$Overruled_sendertime2<- mpledata[,44]*mpledata[,14]*mpledata[,14]
+mpledata$jusitcehomophily_sendertime2<- mpledata[,45]*mpledata[,14]*mpledata[,14] 
 
 # 3rd degree, as.numeric() is used to prevent integer overflow
 mpledata$instar2_sendertime3<- mpledata[,3]*mpledata[,14]*mpledata[,14]*mpledata[,14]
@@ -63,6 +65,7 @@ mpledata$yeardiffsquare_sendertime3<- as.numeric(mpledata[,10])*as.numeric(mpled
 mpledata$AbsDiffMQscores_sendertime3<- mpledata[,16]*mpledata[,14]*mpledata[,14]*mpledata[,14]
 mpledata$NumberJusticesPro_sendertime3<- mpledata[,17]*mpledata[,14]*mpledata[,14]*mpledata[,14]
 mpledata$Overruled_sendertime3<- mpledata[,44]*mpledata[,14]*mpledata[,14]*mpledata[,14]
+mpledata$jusitcehomophily_sendertime3<- mpledata[,45]*mpledata[,14]*mpledata[,14]*mpledata[,14] 
 
 # 4th degree
[…]

### 2018-04-26 R-Code_courtlistener/data_readin.R

@@ -1,7 +1,7 @@
 memory.limit(30000)
 
 
-# reading and preprocessing all the data takes a couple of hours (4-5h)
+# loading and preprocessing all the data takes a couple of hours (4-5h)
 
 
 #####################################################

### 2018-04-30 Tex/paper.tex
updated holdout experiment results
revised network theory section

@@
 
-\title{\Large Inferential Analysis of the Supreme Court Citation Network}
+\title{\Large Generative Dynamics of Supreme Court Citations}
 %\author{}
@@
 
-Majority opinions written by the United States Supreme Court exercise their authority and influence, in part, through their roles as precedents in future Supreme Court decisions and opinions.
-The findings regarding the extent and exact nature of the influences of precedent have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,gillman2001s,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained}.
+United States Supreme Court opinions exercise authority and influence, in part, through their roles as precedents affecting future jurisprudence in the US.
+The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,gillman2001s,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained}.
 Despite a considerable body of research that focuses on the way in which relevant precedents shape decision making on the Court, relatively little work has focused on understanding which past opinions are cited by an opinion.
-Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited by a case.
+Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited in an opinion.
 We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{harris1982structural,caldeira1988legal,fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics}.
@@
 We are not the first to ask what predicts the citations in US Supreme Court Opinions.
-Indeed, a voluminous body of work has sought to explain how many times an opinion is cited \citep[e.g.,][]{cross2010determinants,benjamin2012standing}, when in a lifecycle an opinion is cited \citep[e.g.,][]{black2013citation,spriggs2001explaining}, and how many cases are cited by an opinion \citep[e.g.,][]{lupu2013strategic}---all focused on the US Supreme Court.
-One common feature of the research design in all of these studies is that the observations are at the case or case-time level.
+Indeed, a voluminous body of work has sought to explain how many times an opinion is cited \citep[e.g.,][]{cross2010determinants,benjamin2012standing}, when an opinion is cited \citep[e.g.,][]{black2013citation,spriggs2001explaining}, and how many cases are cited by an opinion \citep[e.g.,][]{lupu2013strategic}---all focused on the US Supreme Court.
+One common feature of the research design in all of these studies is that the observations are at the case or case-year level.
 The outcome variables in these analyses are defined as measures of the number of citations to a case over a period of time, the number of citations to a case at a particular time, or a measurement on the cases cited by a case.
-None of these studies treats citation in its micro-level form---as a relationship between two opinions, the citing opinion and the cited opinion.
-We are aware of one prior study, \citet{clark2010locating}, in which a statistical model is used to analyze dyadic citations between cases.
-However, \citet{clark2010locating} use a dyadic latent variable model in order to estimate ideal points for Supreme Court Opinions, but does not use any explanatory variables to predict the formation of citation ties between opinions.
-We build upon this literature both methodologically and substantively.
-Methodologically, we develop a novel statistical framework for modeling directed dyadic citations.
+These are case-level studies in that, based on the unit of analysis, it is impossible to determine both the origin and target case of a citation that contributes to the dependent variable.
+
+An alternative approach to case-level analysis of citations would be to model them in the directed dyadic form through which they arise.
+A case decided at time $t$ can cite (or not cite) each case decided previously, and in the US Supreme Court, each other case decided at time $t$.
+We are aware of one prior study, \citet{clark2010locating}, in which a statistical model is used to analyze directed dyadic citations between cases.
+However, \citet{clark2010locating} use a dyadic latent variable model in order to estimate ideal points for Supreme Court Opinions, but do not use this model to understand the relationships between explanatory variables and the formation of citation ties between opinions.
+We build upon the literature on citation analysis both methodologically and substantively.
+Methodologically, we develop a novel extension of a statistical model for networks, which we adapt to the network structural constraints of court citations.
 Second, we apply this methodology to a half-century of directed dyadic citations between US Supreme Court opinions.
 
-There exist two broad reasons why empirical analyses of citations are best defined on the directed dyad level, not the case level.
+There exist two broad benefits of conducting empirical analyses of citations at the directed dyad level.
 The first is that directed dyadic analyses can test both dyadic and case-level hypotheses.
@@
 
-In what follows we develop a theoretical case that citations on the US Supreme Court are characterized by forms of complex dependence that are common in networks.
-We then develop an extension of a model---the exponential random graph model---that can incorporate both exogenous covariates and complex forms of interdependence into a directed dyadic analysis of citations.
-Finally, we develop and estimate a specification of this model in an analysis of US Supreme Court citations between 1937 and 2001.
-We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's predictive performance.\\[0.5cm]
-\textit{Comment from Benjamin Kassov: I think the discussion in the introduction of the paper focusing on a "case-based" versus network based analysis is a bit short. 
-I was left slightly confused by what a "case-based" analysis of citation actually is or what it means. 
-I can see several different conceptions of this, some being more problematic than others in a methodological sense, 
-although none of them are ideal for analyzing networks and all likely suffer some amount of OV bias.}
-
-\section{Complex Interdependence in Supreme Court Citations}
-
-When it comes to the development and testing of theory, the defining feature of networks is that the fundamental element under study---the relationship between two units (i.e., the citation from one opinion to another) is a piece of a complex web of relations.
-The formation (or lack thereof) of that relationship cannot be fully understood without considering how the relationship fits into the complex web.
+In this paper we develop a theoretical case that citations on the US Supreme Court are characterized by forms of complex dependence that are common in networks.
+We then develop an extension of a model---the exponential random graph model (ERGM)---that can incorporate both exogenous covariates and complex forms of interdependence into a directed dyadic analysis of citations.
+Finally, we develop and estimate a specification of this model in an analysis of US Supreme Court citations between 1937 and 2015.
+We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's predictive performance.
+
+\section{Network Processes in Supreme Court Citations}
+
+When it comes to the development and testing of theory, the defining feature of networks is that the fundamental element under study---the relationship between two units (i.e., the citation from one opinion to another) is a component of a complex system of relations.
+The formation (or lack thereof) of that relationship cannot be fully understood without considering how the relationship fits into the system.
 Analytical designs that account only for covariates in explaining tie formation are incomplete theoretically, and, as a consequence, are subject to a form of omitted variable bias \citep{cranmer2016critique}.
@@
 
-\textbf{Transitivity:} In a network of directed relations (e.g., A cites B, but B doesn't cite A) transitivity refers to the tendency for A to send a tie to C if A sends a tie to B and B sends a tie to C \citep{holland1971transitivity}.
+We should note that we do not distinguish between positive and negative citations in this theoretical framework.
+The dynamics we outline are not specific to a particular type of citation.
+The only distinction we draw (in the empirical analysis) is the instance in which a case has been overruled.
+In the extreme instance of overruled precedents, we assume that (and test whether) a case is much less likely to be cited after it has been overruled.
+
+The first network property that we theorize in the context of Supreme Court citations is transitivity.
+In a network of directed relations (e.g., A cites B, but B doesn't cite A) transitivity refers to the tendency for A to send a tie to C if A sends a tie to B and B sends a tie to C \citep{holland1971transitivity}.
 In undirected networks, transitivity is simply the process by which friends of friends become friends (i.e., a friend of a friend is a friend).
@@
 The two examples presented below illustrate this process.
+
+\begin{figure}[htp]
+\centering
+\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{images/citations_trans.pdf}
+\caption{Illustration of transitive triangle connecting US Supreme Court opinions through citations.}
+\label{fig:transitivity}
+\end{figure}
 
@@
 \end{quotation} %https://supreme.justia.com/cases/federal/us/548/163/dissent2.html
-The second example is a passage from Seminole Tribe of Fla. v. Florida, 
+The second example, which we illustrate visually in Figure \ref{fig:transitivity} is a passage from Seminole Tribe of Fla. v. Florida, 
 (517 U.S. 44 1996)---a case addressing the rights of groups and citizens to sue states in federal court. In this example, Pennsylvania v. Union Gas Co
-(491 U.S. 1, 1989) is the primary precedent under discussion, and several cases are cited and discussed in terms of their roles as precedents in the Union Gas opinion. 
+(491 U.S.
+1, 1989) is the primary precedent being critiqued, and several cases are cited and discussed in terms of their roles as precedents in the Union Gas opinion.
+We highlight one---Parden v.
+Terminal R.
+Co., which is cited and discussed in both the Seminole and Union Gas opinions.
 \begin{quotation}
@@
 
-\noindent \textbf{Reciprocity:} Reciprocity (also referred to as mutuality) is the tendency for node B to send a tie to node A in response to or coordination with A sending a tie to B \citep{garlaschelli2004patterns}.
+The second network property we consider in the context of Supreme Court citations is reciprocity.
+Reciprocity (also referred to as mutuality) is the tendency for node B to send a tie to node A in response to or coordination with A sending a tie to B \citep{garlaschelli2004patterns}.
 It is typically not possible for reciprocated ties to form in legal opinions.
@@
 
+\begin{figure}[htp]
+\centering
+\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{images/citations_recip.pdf}
+\caption{Illustration of a reciprocal tie between two US Supreme Court opinions.}
+\label{fig:reciprocity}
+\end{figure}
+
 The first case in our example reciprocal dyad is a passage from Western Air Lines v.
@@
 353, 1985) is another case considering whether mandatory retirement violates the Age Discrimination in Employment Act.
+The mutual edge connecting these two cases is visualized in Figure \ref{fig:reciprocity}.
 These cases addressed very similar legal questions, which increased the likelihood that they would inform each other, and the opinions were written within the same term, which made it possible for them to cite each other.
@@
 
-\noindent \textbf{Popularity} Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent \citep{barabasi1999emergence}.
+The third network property we consider in the context of Supreme Court citations is popularity.
+Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent \citep{barabasi1999emergence}.
 Citations to an opinion signal both the Court's awareness of the legal reasoning of the case and the Court's evaluation that the opinion is an authoritative precedent.
@@
 In this opinion passage Baker v.
-Carr is cited in reference to its role as a landmark precedent, and noted for the number of other cases by which it has been followed and for which an authoritative opinion is referenced, and even discussed in terms of the number of other cases by which it was followed.
+Carr is cited in reference to its role as a landmark precedent, and noted for the number of other cases by which it has been followed.
+and for which an authoritative opinion is referenced, and even discussed in terms of the number of other cases by which it was followed.
+The language in this passage suggests that the attention to Baker v.
+Carr in previous Court opinions is in part responsible for its authority in Oregon v.
+Mitchell.
+The citations to Baker v.
+Carr are visualized in Figrue \ref{fig:popularity}.
 \begin{quotation}
@@
 
-\noindent \textbf{Activity} Activity, or sender activation, is the tendency for ties sent to beget more ties sent \citep{howard2016understanding}.
+\begin{figure}[htp]
+\centering
+\includegraphics[width = 0.95\textwidth,trim= 0cm 3cm 6cm 8cm,clip=true ]{images/citations_pop.pdf}
+\caption{Illustration of ties sent to a landmark Supreme Court opinion via citations.}
+\label{fig:popularity}
+\end{figure}
+
+The fourth network property we consider in the context of Supreme Court citations is activity.
+Activity, or sender activation, is the tendency for ties sent to beget more ties sent \citep{howard2016understanding}.
 Structurally, this means that there are few nodes that send a moderate number of ties---the number of ties sent is either small or fairly large.
@@
 It is common that opinions distinguish among a variety of legal rules in terms of their applicability to a case, which means that for each rule that is established in an applicable precedent, one or more need to be discussed in terms of why they do not apply.
+We do not present a separate visualization related to activity, as it would look very similar to the visualization for popularity.
 \begin{quotation}
@@
 \end{quotation} % https://supreme.justia.com/cases/federal/us/460/719/
-
-\begin{itemize}
-\item add to the discussion to which degree network statistics reflect causual relation
-\end{itemize}
 
@@
 
+\includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 0cm,clip=true  ]{images/daggish.pdf}
+
 Networks researchers have developed approaches to understanding the mechanisms that drive citations between documents.
@@
 Second, theoretical models are used to account for prevalent graph structures.
-In the next section we describe a methodology that can be used to integrate node attributes and graph structures/motifs into a comprehensive model of US Supreme Court citations.
+In the next section we describe a methodology that can be used to integrate case attributes and graph structures/motifs into a comprehensive model of US Supreme Court citations.
 
@@
 
+Before we transition to the empirical analysis, we should note that there is nothing about the c-TERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another.
+For example, if we find a positive popularity effect, that could be driven by one of two (or some combination of both) dynamics.
+First, it could be the case that a large volume of citations to a case in past opinions draws the attention of the justices to the ``popular'' case.
+Second, it could be that we have omitted a case-level variable that is strongly associated with the number of citations accumulated by a case, and the number of citations in past cases is operating as a proxy for this unmeasured case characteristic.
+We cannot statistically differentiate between these two dynamics.
+However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-TERGM can be specified to account for the dependence in the data.
+
 \section{Correlates of Legal Citations}
@@
   \hline
-precision & 0.5589 & (0.5508, 0.5757) & 0.8677 & (0.865, 0.8718) \\ 
-  recall & 0.098 & (0.0943, 0.103) & 0.5943 & (0.5887, 0.6032) \\ 
-  F1 score & 0.1667 & (0.1618, 0.1748) & 0.7054 & (0.7013, 0.7119) \\ 
+precision & 0.6007 & (0.5934, 0.6074) & 0.8666 & (0.8647, 0.8701) \\ 
+  recall & 0.1522 & (0.1495, 0.1549) & 0.5969 & (0.5945, 0.6019) \\ 
+  F1 score & 0.2429 & (0.2389, 0.2462) & 0.7069 & (0.7048, 0.7107) \\ 
    \hline \hline
@@
   \hline
-precision & 0.5872 & (0.5781, 0.5962) & 0.8622 & (0.8593, 0.8668) \\ 
-  recall & 0.0565 & (0.0547, 0.0587) & 0.5839 & (0.5777, 0.592) \\ 
-  F1 score & 0.103 & (0.1, 0.1069) & 0.6963 & (0.6915, 0.7027) \\ 
+precision & 0.5938 & (0.5844, 0.5996) & 0.8668 & (0.8653, 0.8693) \\ 
+  recall & 0.1551 & (0.1515, 0.1575) & 0.5969 & (0.5937, 0.6005) \\ 
+  F1 score & 0.246 & (0.2406, 0.249) & 0.707 & (0.7045, 0.7095) \\  
    \hline \hline
@@
   \hline
-precision & 0.6441 & (0.6302, 0.6536) & 0.8671 & (0.8626, 0.873) \\ 
-  recall & 0.0852 & (0.0832, 0.0863) & 0.594 & (0.5874, 0.6012) \\ 
-  F1 score & 0.1504 & (0.1473, 0.1525) & 0.705 & (0.7008, 0.712) \\ 
+precision & 0.6209 & (0.611, 0.6351) & 0.8669 & (0.8612, 0.8711) \\ 
+  recall & 0.1511 & (0.1451, 0.1558) & 0.5962 & (0.5884, 0.6011) \\ 
+  F1 score & 0.243 & (0.2347, 0.2486) & 0.7065 & (0.7011, 0.7103) \\ 
    \hline \hline

### 2018-04-30 Tex/presentations/SPSA2018/DesmaraisSPSATalk.tex
revised network theory section

@@
 
-\begin{frame} \frametitle{Two Limitations in Case-Level Analysis}
+\begin{frame} \frametitle{Two Limitations to Case-Level Analysis}
 \Large
@@
 \item {\bf Methodological:} Risk bias from omitted structure
-\end{enumerate}
+\end{enumerate} ~\\
+
+We propose a dyadic, network-analytic approach
 
@@
 
-\frame{
-\frametitle{Citation Temporal Exponential Random Graph Model}\vspace{0.5cm}
-
-The probability  of observing the $N_t \times N_{\leq t}$ adjacency matrix $C_t$ given past citations $C_{<t}$, where $C_{\leq t}$ is the network up to time $t$
-\[
-\mathcal{P}(C_t|C_{<t}, \bt) = \frac{\exp \{ \bt' \bh(C_{\leq t})  \}}{\sum_{C^*_t \in \mathcal{C}_t} \exp \{\bt ' \bh(C^*_{\leq t})\}  }
-\]
-Decomposition:
-\[
-\underbrace{\bh(C_{\leq t})}_{Net\hspace{3pt} Stats} \qquad \underbrace{\bt}_{Effects} \qquad \underbrace{\exp \{\bt' \bh(C_{\leq t}) \}}_{+ \hspace{3pt} Weight} \qquad \underbrace{\sum_{C^*_t \in \mathcal{C}_t} \exp \{\bt ' \bh(C^*_{\leq t})\}}_{Normalizer}
-\]
-
-Flexible: $\bh$ can capture virtually any form of interdependence among the edges $+$ covariates\\
-\vspace{0.5cm}
-Normalizing constant can make estimation difficult
-}
-
-\begin{frame} \frametitle{DAG[gish] Dynamics}
+\begin{frame} \frametitle{ The Citation TERGM: $C_t|C_{<t}$ via $h(C_{\leq t})$}
 \centering
@@
     Reciprocity &  Transitivity \\
-\includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/mutual_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/triangle_coef_trend.pdf} \\
+\includegraphics[width = 0.55\textwidth, trim= .5cm 1cm 0.5cm .45cm,clip=true]{../../images/mutual_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= .5cm 1cm 0.5cm .45cm,clip=true]{../../images/triangle_coef_trend.pdf} \\
 
      Activity & Popularity \\
-\includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/o2star_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/i2star_coef_trend.pdf} \\
+\includegraphics[width = 0.55\textwidth, trim= .5cm 1cm 0.5cm .45cm,clip=true]{../../images/o2star_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= .5cm 1cm 0.5cm .45cm,clip=true]{../../images/i2star_coef_trend.pdf} \\
 
@@
   MQ Score Difference & Ideological Breadth \\
-\includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/mq_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/absdiffmq_coef_trend.pdf} \\
+\includegraphics[width = 0.55\textwidth, trim= .5cm 1cm 0.5cm .45cm,clip=true]{../../images/mq_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= .5cm 1cm 0.5cm .45cm,clip=true]{../../images/absdiffmq_coef_trend.pdf} \\
 
    Same Issue &  Majority Size \\
-\includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/sameissue_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= 1cm 1cm 0.5cm .45cm,clip=true]{../../images/numberjusticespro_coef_trend.pdf} \\
+\includegraphics[width = 0.55\textwidth, trim= .5cm 1cm 0.5cm .45cm,clip=true]{../../images/sameissue_coef_trend.pdf} & \includegraphics[width = 0.55\textwidth, trim= .5cm 1cm 0.5cm .45cm,clip=true]{../../images/numberjusticespro_coef_trend.pdf} \\
 
@@
 \begin{frame} \frametitle{Predictive Performance}
+
+\begin{itemize}
+
+\item Randomly split dyads 80/20 into training and test
+\item Estimate models using training, predict test
+\item Repeat 10 times
+
+\end{itemize}
+~\\
 
@@
 
+\frame{
+\frametitle{Citation Temporal Exponential Random Graph Model}\vspace{0.5cm}
+
+The probability  of observing the $N_t \times N_{\leq t}$ adjacency matrix $C_t$ given past citations $C_{<t}$, where $C_{\leq t}$ is the network up to time $t$
+\[
+\mathcal{P}(C_t|C_{<t}, \bt) = \frac{\exp \{ \bt' \bh(C_{\leq t})  \}}{\sum_{C^*_t \in \mathcal{C}_t} \exp \{\bt ' \bh(C^*_{\leq t})\}  }
+\]
+Decomposition:
+\[
+\underbrace{\bh(C_{\leq t})}_{Net\hspace{3pt} Stats} \qquad \underbrace{\bt}_{Effects} \qquad \underbrace{\exp \{\bt' \bh(C_{\leq t}) \}}_{+ \hspace{3pt} Weight} \qquad \underbrace{\sum_{C^*_t \in \mathcal{C}_t} \exp \{\bt ' \bh(C^*_{\leq t})\}}_{Normalizer}
+\]
+
+Flexible: $\bh$ can capture virtually any form of interdependence among the edges $+$ covariates\\
+\vspace{0.5cm}
+Normalizing constant can make estimation difficult
+}
+
 \end{document}

### 2018-04-30 Tex/paper.tex
finished model definition

@@
 \usepackage{amsthm}
+\usepackage{bm}
 \usepackage{enumerate}
@@
 
+\doublespacing
  \section{Introduction}
@@
 
-\section{Network Approaches to Studying Citations}
-
-The dependence processes discussed in the previous section represent new theoretical claims regarding the factors that account for citation formation among US Supreme Court opinions.
-Before we can test these claims, we consider methods for analyzing citation data as a network.
-Researchers from several fields have used network analysis to analyze citations---legal, patent, and scientific.
-Before describing our approach to modeling Supreme Court citations, we review the methods researchers have previously used to study citations.
-A citation is a directed link between two documents that indicates the citing document attributes the cited to be relevant to the evidentiary basis of the citing document.
-Collective patterns of citations within a certain domain make up citation networks.
-When raw citation data is formally represented as a network, the nodes are usually the documents themselves, and each directed arc is the existence of one or more cites from the sending to the receiving document.
-Given the permanent nature of documents in citations networks, these networks are acyclic \citep{leicht2007large,karrer2009random}.
-Relatedly, citation networks grow as new documents and citations enter the network, but established arcs persist in all but exceptional cases.
-
+ \section{The Citation Temporal Exponential Random Graph Model}
+We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
+To accomplish this, we extend a model that has been developed to jointly model covariate and dependence effects in longitudinal network data---the temporal exponential random graph model (TERGM) \citep[e.g.,][]{cranmer2012complex,clark2013multimember, masket2015polarization, graif2017neighborhood}.
+The TERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
+The longitudinal network structures for which TERGMs are currently designed are insufficient to account for the structure of citation networks.
+
+\subsection{c-TERGM definition}
+
+We develop the citation TERGM (c-TERGM), to account for the structural constraints that apply to the network of Supreme Court citations.
+These structural constraints amount to three departures from the structure of longitudinal networks for which TERGMs are currently designed.
+First, the citation network is partially acyclic.
+If two cases are decided during the same term, they can cite each other, forming a mutual edge (or two-cycle).
+However, if case $i$ is decided before case $j$, case $j$ can cite case $i$, but case $i$ cannot cite case $j$.
+Second, new edges can be created over time, but cannot be eliminated.
+Unlike in, e.g., an alliance network, in which two countries can dissolve an alliance, once a citation exists in a citation network it cannot be dissolved.
+Third, the set of nodes in the network must increase for new edges to be created.
+In conventional TERGMs, the number of nodes in the network can increase or decrease in each time period, and is typically stable over time.
+In a citation network, new edges (and non-edges) are introduced over time via the introduction of new nodes.
+The structure of the citation network is depicted in Figure \ref{fig:ctergm}---a hypothetical citation network established over three time periods, with three cases decided in each time period.
+We denote $C_t$ to be the set of citation and non-citations added to the network at time $t$ (i.e., via the addition of three cases), $C_{ <t}$ to be the citations among cases decided before time $t$, and $C_{ \leq t}$ to denote the entire set of citations and non-citations on which $C_t$ can depend through the c-TERGM specification.
+
+\begin{figure}[htp]
+\centering
 \includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 0cm,clip=true  ]{images/daggish.pdf}
-
-Networks researchers have developed approaches to understanding the mechanisms that drive citations between documents.
-Work in this area generally proceeds by specifying a set of citation behaviors, then proposing a model to capture the combination of these behavioral rules \citep{simkin2007mathematical}.
-Assessment of the model is based on how well it fits citation distributions.
-Researchers working in this area tend to focus on modeling the growth of the citation network as governed by a degree-based mechanism such as preferential attachment \citep{barabasi1999emergence, vazquez2001statistics}, and the age of the paper \citep{jeong2003measuring,eom2011characterizing,wang2013quantifying}.
-Regarding age, the previous standard approach was to treat the probability of citation as function of degree (i.e., the number of citations previously accrued by the document) and age \citep{hajra2005aging,hajra2006modelling}, then examine the distribution of citations.
-More recently, work has been done in relaxing the assumption that the effect of degree is static, instead allowing it to vary with time \citep{wang2008measuring}.
-
-Another approach to understanding the generative process of citation networks is to examine the existence of network motifs, or subnetwork structures, that can be interpreted as measuring different generative mechanisms.
-For example, a triangle in which case A cites both cases B and C, and case B cites case C, is a type of motif---a triadic motif.
-If researchers find that a triangle is a highly prevalent motif in a network, that finding would suggest that triad closure is a prevalent process by which the network is generated.
-The analysis of network motifs is done by comparing the prevalence of the motifs in the observed network to the prevalence of the motifs in networks drawn from a relevant null distribution.
-This null model must also be characterized by features of citation networks discussed earlier, meaning that it must be a directed acyclic graph with unweighted arcs \citep{carstens2016topology}.
-\citet{karrer2009random} proposes the use of null models based on fixed-degree sequences.
-In these models, the nodes in the networks (i.e., opinions) have the same, or close to the same, numbers of ties.
-Conditional on the number of ties accrued by each node, ties are randomly re-wired such that there is no systematic pattern regarding the nodes to which any node connects.
-
-From this brief review we see two broad components of past methods for analyzing citation networks.
-First, node features, age in particular, are used to model the rate at which a document is cited.
-Second, theoretical models are used to account for prevalent graph structures.
-In the next section we describe a methodology that can be used to integrate case attributes and graph structures/motifs into a comprehensive model of US Supreme Court citations.
-
- \section{The Citation Temporal Exponential Random Graph Model}
-We build upon existing methods for network analysis to define a model that fits the constraints to which Supreme Court citation networks are subject.
-The Supreme Court citation network is nearly a directed acyclic graph---a graph in which there can be no loops/cycles (e.g., ties cannot be reciprocated, which would constitute a two-node loop).
-If two cases are decided in different terms, the latter case can cite the earlier case, but not the reverse.
-However, as noted above, if two opinions are written in the same term, they can cite each other reciprocally.
-We implement a version of the temporal exponential random graph model (TERGM)---a model that is commonly used for longitudinal network data \citep[e.g.,][]{cranmer2012complex,desmarais2013forecasting,clark2013multimember,block2018change,graif2017neighborhood}---that is subject to the constraints specific to the Supreme Court citation network.
-The TERGM is a model that can simultaneously incorporate structural dependencies such as transitivity and reciprocity and exogenous covariates  (e.g.
-the issue area of the citing case, the ideological distance between the opinion authors of the two cases in the dyad) to explain tie formation.
-If the ties do not depend upon each other, and the coefficients associated with the structural dependencies are either assumed or estimated to be zero, the TERGM reduces to a logistic regression in which the observation is a directed dyad \citep{cranmer2010inferential}.
-In this section we describe the citation TERGM (c-TERGM), and explain estimation methods.
-
-Let $C_t$ be the case-to-case adjacency matrix at time $t$ such that $C_t^{ij} \in \{0,1\}$ is a binary indicator of whether case $i$ cites $j$. Furthermore, let
-$$\mathcal{C}_t =\{C_{\leq t} \in \{0,1\}^{(N_t \cdot (N_t-1))/2}: C_{\leq t}^{ij} \in \{0,1\} \}$$ 
-be the set of all possible adjacency matrices among $N_t$ cases at time $t$.
-$N_t$ denotes the number of cases in the network at time $t$.
-Note that the cardinality of $\mathcal{C}_t$ increases exponentially for every newly added case.
-The probability of observing $C_t$ given past citations $C_{<t}$, where $C_{\leq t}$ is the network up to time $t$ is defined as
+\caption{Illustration of temporal structure of the Supreme Court Citation Network.
+$C_{\leq t}$ is the entire set of citations (and non-citations) on which citations and non-citations at time $t$ (i.e., $C_t$) depend.
+$C_t$ are conditioned on the citations and non-citations established before time $t$ (i.e., $C_{< t}$).
+The shaded small squares are hypothetical observed citations, and the white small squares are citations that could have been observed but were not.
+The regions of the citation matrix that are represented by large white rectangles are citations that could not have been observed since the citing case would have been decided in a term that preceded the term of the cited case.
+The citing case ID is given in the row and the prospective cited case is given in the column.
+}
+\label{fig:ctergm}
+\end{figure}
+
+The likelihood function of the c-TERGM is given by
 \begin{equation}
-P(C_t | C_{<t}, \theta)=\cfrac{\exp\big(\theta^T\cdot h\big(C_{\leq t})\big)\big)}{\sum_{Z_t^*\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)\big)}
-\label{ercm}
+l(\bm{\theta},C_{\leq T}) =  \prod_{t=1}^T \frac{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right] }{ \sum_{C_t^* \in \mathcal{C}_t} \exp \left[ {\bm{\theta}'\bm{h}(C^*_{t},C_{<t}) }\right]  },
 \end{equation}
-where $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C}_t \to \mathbb{R}^q$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{Z^*_t\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)$ is a normalization constant that ensures that Equation (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.
-
-% Network Statistics
-The c-TERGM is specified through 
-the decision regarding which network statistics $h(\cdot)$ to incorporate.
-We include the following network statistics for the Supreme Court citation network, which are derived from the interdependence hypotheses described in the previous section:
-$$h_{edges}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}, $$
-the number of citations.  $h_{edges}$ performs the function of an intercept, and models the expected value of any given edge.
-$$h_{outstar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{i=1} {\sum_{j \neq i} C_{\leq t}^{ij} \choose 2} $$
-the number of out-two-stars.
-An out-two-star is a configuration in which one node sends to two other nodes.
-The number of these configurations grows quadratically as the origins of ties concentrate on a few highly active/social senders.
-The out-two-stars configuration is commonly used to model activity (sender activation) in ERGMs.
-$$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{j=1} {\sum_{i \neq j} C_{\leq t}^{ij} \choose 2} $$
-the number of in-two-stars.
-An in-two-star is a configuration in which one node receives ties from two other nodes.
-The number of these configurations grows quadratically as the destinations of ties concentrate on a few highly popular recipients.
-The out-two-stars configuration is commonly used to model popularity in ERGMs.
-\begin{align*}
-h_{triangle}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to & \sum_{j<i<k} C^{ij}\cdot C^{jk} \cdot C^{ki} + C^{ji}\cdot C^{jk} \cdot C^{ki} + C^{ij}\cdot C^{kj} \cdot C^{ki} \\ & + C^{ji}\cdot C^{kj} \cdot C^{ki} + C^{ij}\cdot C^{jk} \cdot C^{ik} + C^{ji}\cdot C^{jk} \cdot C^{ik}  \\ & + C^{ij}\cdot C^{kj} \cdot C^{ik} + C^{ji}\cdot C^{kj} \cdot C^{ik},
-\end{align*}
-The number of triangles in the network. Triangles measure triad closure, as discussed above.      
-$$h_{reciprocity}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}C_{\leq t}^{ji}, $$
-The number of reciprocal ties in the network.
-Even though the reciprocity of ties is not a meaningful network statistic for most citation networks, it may still play a role in the Supreme Court citation network.
-This statistic comes with the restriction that it only can appear among cases that entered the network at the same time.
-
-$$h_{covariate}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}X_{\leq t}^{ij}, $$ the effect of an exogenous covariate ($X$).
-We include several exogenous covariates based on this standard statistic formulation.
-These covariates are discussed below.\\
-\indent The structure of this model is very similar to a conventional (T)ERGM setup.
-The main difference is that  $\mathcal{C}_t$ excludes adjacency matrices in which there are loops that include edges sent at different time points.
-Loops can exist in the Supreme Court citation network, but only among cases decided in the same term.
-Furthermore, a case $i$ can only cite a case $j$ if $j$ entered the network before or at the same time as $i$ did.
-\\
+where $\bm{\theta}$ is a vector of real-valued parameters, and $\bm{h}(C_{t},C_{<t})$ is a vector of scalar-valued functions that each quantify a feature of the citation network (e.g., the relationship between citation ties and a case attribute, the number of mutual edges in the citation network).
+$\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right]$ is a positive weight that is proportional to the probability of observing any particular form of the citations and non-citations added to the network at time $t$.
+The denominator of Equation 1 represents a normalizing constant, in which the positive weight is summed over all possible configurations of $C_{t}$, from the network in which the cases added to the network at time $t$ send no citations at all, to the network in which the cases added at time $t$ cite every possible case, and everything in between.
+The probability of observing the citation network up to time $T$---the final time point in the dataset---is given by the product over the probability of adding the citations at each time point $t$, conditional on the citations that were added to the network previously.
+
+Though the likelihood function of the c-TERGM may appear quite different from that of statistical models conventionally used in political science, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients.
+$$ P(C_{ij,t} = 1 | C_{-ij,t}, C_{ < t}) = \frac{\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right]}{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right] + \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 0) } \right]}, $$ $$ \text{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~} = \frac{1}{ 1 + \exp \left[ - {\bm{\theta}'\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)} \right]}, $$
+where $C_{ij,t} = 1$ indicates that case $i$ cites case $j$, $C_{ij,t} = 0$ indicates that case $i$ does not cite case $j$, $C_{-ij,t}$ is the observed elements of $C_{t}$ except $C_{ij,t}$, and $\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)$ is the change in $\bm{h}(C_{t},C_{<t})$ that results from toggling $C_{ij,t} = 0$ to $C_{ij,t} = 1$.
+This re-arrangement illustrates that the parameters can be interpreted in terms of the change in the log odds of a citation from $i$ to $j$ given a one-unit increase in the corresponding element of $\bm{h}$, conditional on the other citations observed in the network.
+For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log odds of observing $C_{ij,t} = 1$ increase by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$).
+
+\subsection{c-TERGM Estimation}
+
+The normalizing constant in Equation 1 is intractable.
+For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure \ref{fig:ctergm}---there are 16,777,216 unique configurations of $C_t$ that could be observed.
+The typical Supreme Court term involves adding hundreds of cases to a network that
+
 \indent The normalizing constant in the denominator of the probability function (i.e., Equation \ref{ercm}) is computationally intractable.
@@
 \newpage
+
+Let $C_t$ be the case-to-case adjacency matrix at time $t$ such that $C_t^{ij} \in \{0,1\}$ is a binary indicator of whether case $i$ cites $j$. Furthermore, let
+$$\mathcal{C}_t =\{C_{\leq t} \in \{0,1\}^{(N_t \cdot (N_t-1))/2}: C_{\leq t}^{ij} \in \{0,1\} \}$$ 
+be the set of all possible adjacency matrices among $N_t$ cases at time $t$.
+$N_t$ denotes the number of cases in the network at time $t$.
+Note that the cardinality of $\mathcal{C}_t$ increases exponentially for every newly added case.
+The probability of observing $C_t$ given past citations $C_{<t}$, where $C_{\leq t}$ is the network up to time $t$ is defined as
+\begin{equation}
+P(C_t | C_{<t}, \theta)=\cfrac{\exp\big(\theta^T\cdot h\big(C_{\leq t})\big)\big)}{\sum_{Z_t^*\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)\big)}
+\label{ercm}
+\end{equation}
+where $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C}_t \to \mathbb{R}^q$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{Z^*_t\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)$ is a normalization constant that ensures that Equation (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.
+
+% Network Statistics
+The c-TERGM is specified through 
+the decision regarding which network statistics $h(\cdot)$ to incorporate.
+We include the following network statistics for the Supreme Court citation network, which are derived from the interdependence hypotheses described in the previous section:
+$$h_{edges}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}, $$
+the number of citations.  $h_{edges}$ performs the function of an intercept, and models the expected value of any given edge.
+$$h_{outstar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{i=1} {\sum_{j \neq i} C_{\leq t}^{ij} \choose 2} $$
+the number of out-two-stars.
+An out-two-star is a configuration in which one node sends to two other nodes.
+The number of these configurations grows quadratically as the origins of ties concentrate on a few highly active/social senders.
+The out-two-stars configuration is commonly used to model activity (sender activation) in ERGMs.
+$$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{j=1} {\sum_{i \neq j} C_{\leq t}^{ij} \choose 2} $$
+the number of in-two-stars.
+An in-two-star is a configuration in which one node receives ties from two other nodes.
+The number of these configurations grows quadratically as the destinations of ties concentrate on a few highly popular recipients.
+The out-two-stars configuration is commonly used to model popularity in ERGMs.
+\begin{align*}
+h_{triangle}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to & \sum_{j<i<k} C^{ij}\cdot C^{jk} \cdot C^{ki} + C^{ji}\cdot C^{jk} \cdot C^{ki} + C^{ij}\cdot C^{kj} \cdot C^{ki} \\ & + C^{ji}\cdot C^{kj} \cdot C^{ki} + C^{ij}\cdot C^{jk} \cdot C^{ik} + C^{ji}\cdot C^{jk} \cdot C^{ik}  \\ & + C^{ij}\cdot C^{kj} \cdot C^{ik} + C^{ji}\cdot C^{kj} \cdot C^{ik},
+\end{align*}
+The number of triangles in the network. Triangles measure triad closure, as discussed above.      
+$$h_{reciprocity}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}C_{\leq t}^{ji}, $$
+The number of reciprocal ties in the network.
+Even though the reciprocity of ties is not a meaningful network statistic for most citation networks, it may still play a role in the Supreme Court citation network.
+This statistic comes with the restriction that it only can appear among cases that entered the network at the same time.
+
+$$h_{covariate}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}X_{\leq t}^{ij}, $$ the effect of an exogenous covariate ($X$).
+We include several exogenous covariates based on this standard statistic formulation.
+These covariates are discussed below.\\
+\indent The structure of this model is very similar to a conventional (T)ERGM setup.
+The main difference is that  $\mathcal{C}_t$ excludes adjacency matrices in which there are loops that include edges sent at different time points.
+Loops can exist in the Supreme Court citation network, but only among cases decided in the same term.
+Furthermore, a case $i$ can only cite a case $j$ if $j$ entered the network before or at the same time as $i$ did.
+\\
+
+\section{Network Approaches to Studying Citations}
+
+The dependence processes discussed in the previous section represent new theoretical claims regarding the factors that account for citation formation among US Supreme Court opinions.
+Before we can test these claims, we consider methods for analyzing citation data as a network.
+Researchers from several fields have used network analysis to analyze citations---legal, patent, and scientific.
+Before describing our approach to modeling Supreme Court citations, we review the methods researchers have previously used to study citations.
+A citation is a directed link between two documents that indicates the citing document attributes the cited to be relevant to the evidentiary basis of the citing document.
+Collective patterns of citations within a certain domain make up citation networks.
+When raw citation data is formally represented as a network, the nodes are usually the documents themselves, and each directed arc is the existence of one or more cites from the sending to the receiving document.
+Given the permanent nature of documents in citations networks, these networks are acyclic \citep{leicht2007large,karrer2009random}.
+Relatedly, citation networks grow as new documents and citations enter the network, but established arcs persist in all but exceptional cases.
+
+Networks researchers have developed approaches to understanding the mechanisms that drive citations between documents.
+Work in this area generally proceeds by specifying a set of citation behaviors, then proposing a model to capture the combination of these behavioral rules \citep{simkin2007mathematical}.
+Assessment of the model is based on how well it fits citation distributions.
+Researchers working in this area tend to focus on modeling the growth of the citation network as governed by a degree-based mechanism such as preferential attachment \citep{barabasi1999emergence, vazquez2001statistics}, and the age of the paper \citep{jeong2003measuring,eom2011characterizing,wang2013quantifying}.
+Regarding age, the previous standard approach was to treat the probability of citation as function of degree (i.e., the number of citations previously accrued by the document) and age \citep{hajra2005aging,hajra2006modelling}, then examine the distribution of citations.
+More recently, work has been done in relaxing the assumption that the effect of degree is static, instead allowing it to vary with time \citep{wang2008measuring}.
+
+Another approach to understanding the generative process of citation networks is to examine the existence of network motifs, or subnetwork structures, that can be interpreted as measuring different generative mechanisms.
+For example, a triangle in which case A cites both cases B and C, and case B cites case C, is a type of motif---a triadic motif.
+If researchers find that a triangle is a highly prevalent motif in a network, that finding would suggest that triad closure is a prevalent process by which the network is generated.
+The analysis of network motifs is done by comparing the prevalence of the motifs in the observed network to the prevalence of the motifs in networks drawn from a relevant null distribution.
+This null model must also be characterized by features of citation networks discussed earlier, meaning that it must be a directed acyclic graph with unweighted arcs \citep{carstens2016topology}.
+\citet{karrer2009random} proposes the use of null models based on fixed-degree sequences.
+In these models, the nodes in the networks (i.e., opinions) have the same, or close to the same, numbers of ties.
+Conditional on the number of ties accrued by each node, ties are randomly re-wired such that there is no systematic pattern regarding the nodes to which any node connects.
+
+From this brief review we see two broad components of past methods for analyzing citation networks.
+First, node features, age in particular, are used to model the rate at which a document is cited.
+Second, theoretical models are used to account for prevalent graph structures.
+In the next section we describe a methodology that can be used to integrate case attributes and graph structures/motifs into a comprehensive model of US Supreme Court citations.

### 2018-04-30 Tex/paper.tex
finished estimation

@@
 For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log odds of observing $C_{ij,t} = 1$ increase by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$).
-
-\subsection{c-TERGM Estimation}
-
-The normalizing constant in Equation 1 is intractable.
-For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure \ref{fig:ctergm}---there are 16,777,216 unique configurations of $C_t$ that could be observed.
-The typical Supreme Court term involves adding hundreds of cases to a network that
-
-\indent The normalizing constant in the denominator of the probability function (i.e., Equation \ref{ercm}) is computationally intractable.
-This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with ERGM family models.
-Methods of Monte Carlo approximation are well established \citep{hunter2006inference,van2009framework,hummel2012improving}, but with almost $10,000$ vertices, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}.
-To estimate the parameters of the c-TERGM, we rely on two features of the model and data that facilitate the application of an approximate inference method that is faster than Monte Carlo approximation---maximum pseudolikelihood.
-First, our sample size is very large.
-Second, under the c-TERGM we assume that citations are formed in each time period conditional upon the citations that exist in the past.
-We estimate the c-TERGM via bootstrapped pseudolikelihood \citep{desmarais2012statistical,desmarais2010consistent}.
-
-Before we transition to the empirical analysis, we should note that there is nothing about the c-TERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another.
-For example, if we find a positive popularity effect, that could be driven by one of two (or some combination of both) dynamics.
-First, it could be the case that a large volume of citations to a case in past opinions draws the attention of the justices to the ``popular'' case.
+The logit form conditional probability is not unique to c-TERGM.
+It is well known for the ERGM family \citep{goodreau2009birds}.
+
+We should note that there is nothing about the c-TERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another.
+For example, if we find a positive popularity effect, the term for that dynamic would indicate that there is an increase in the log odds that case $i$ cites case $j$ with each other case that has cited case $j$.
+That finding could be driven by one of two (or some combination of both) dynamics.
+First, it could be the case that a large volume of citations to a case in past opinions draws the attention of the justices to the ``popular'' case---as in the Baker v.
+Carr passage cited above.
 Second, it could be that we have omitted a case-level variable that is strongly associated with the number of citations accumulated by a case, and the number of citations in past cases is operating as a proxy for this unmeasured case characteristic.
@@
 However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-TERGM can be specified to account for the dependence in the data.
+
+\subsection{c-TERGM Estimation}
+
+The normalizing constant in Equation 1 is intractable.
+For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure \ref{fig:ctergm}---there are 16,777,216 unique configurations of $C_t$ that could be observed.
+The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases.
+This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the c-TERGM.
+Methods of Monte Carlo approximation are well established in the ERGM literature, \citep{hunter2006inference,van2009framework,hummel2012improving}, but with a number of nodes on the order of $10,000$, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}.
+However, because the network is large, we can rely on a fast estimation method---maximum pseudo-likelihood estimation (MPLE), which is consistent for discrete exponential family models \citep{nguyen2017near}, and has long been used with ERGM \citep{strauss1990pseudolikelihood}.
+The MPLE is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
+The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated above, assume a logit form.
+
+One challenge that arises with the use of MPLE is that the usual estimate of the asymptotic covariance matrix (i.e., inverting the negated Hessian of the objective function at the maximum), results in standard errors that are biased downward.
+This leads to a high Type-1 error rate in hypothesis testing with MPLE if conventional standard errors are used \citep{van2009framework}.
+A form of non-parametric cluster bootstrap results in confidence intervals with correct coverage probability with MPLE \citep{desmarais2012statistical,desmarais2010consistent}.
+In the case of c-TERGM, the clusters are defined by $t$---the time points in which the cases are decided, and $T$ time-specific citation matrices $C_{t}$ are re-sampled with replacement to form a non-parametric cluster bootstrap sample.
+Confidence intervals are defined by the quantiles in the empirical distributions of estimates that result from re-estimating the MPLE on non-parametric cluster bootstrap samples.
 

### 2018-05-01 Tex/paper.tex
now revising model specification section

@@
 Prepared for presentation at the ``The Enemy of My Enemy: Judicial Politics meets Network Analysis'' conference within a conference at the 2018 Annual Conference of the Southern Political Science Association.
-This work was supported in part by NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219.}
+This work was supported in part by NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219.
+We would like to thank Rachael Hinkle and Benjamin Kassow.}
 \end{abstract}
@@
 
-\section{Correlates of Legal Citations}
-
-Theorizing that legal citations arise through network interdependence processes represents a potential contribution to the literature.
-To demonstrate the value of our contribution we must also incorporate established non-network processes that have been found to affect legal citations.
-We incorporate these processes through the specification of $h_{covariate}$  terms to be included in the c-TERGM.
-
-We incorporate a statistic to model the degree to which cases cite those that are similar in terms of the ideological positions of the justices who supported the decision.
-We account for this effect following \citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
-\citet{clark2010locating} estimates a latent coordinate model of Supreme Court opinions based on the network of case-to-case citations.
-They find that the majority opinion falls at the ideal point of the median member of the majority coalition in the case.
-We include a covariate term in which $X_{ij}$ is the absolute difference between the Martin-Quinn scores \citep{martin2002dynamic} of the median justices in the majority coalitions for cases $i$ and $j$.
-We expect this variable to have a negative effect, which would correspond to cases citing those to which they are ideologically similar.%WE COULD USE THIS DATASET AS A SECOND APPLICATION
-
-We include two sets of dummy variables that account for the issue areas of cases.
-In one set of dummy variable---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$).
-In the set of receiver intercepts $X^{ij}$ indicates the issue of the receiver case.
-Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
-We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
-Table \ref{issue_area_coding} states the different issue areas.
-
-\begin{table}[]
-\centering
-\begin{tabular}{llll}
-1 & Criminal Procedure & 8  & Economic Activity    \\
-2 & Civil Rights       & 9  & Judicial Power       \\
-3 & First Amendment    & 10 & Federalism           \\
-4 & Due Process        & 11 & Interstate Relations \\
-5 & Privacy            & 12 & Federal Taxation     \\
-6 & Attorneys          & 13 & Miscellaneous        \\
-7 & Unions             & 14 & Private Action      
-\end{tabular}
-\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
-\label{issue_area_coding}
-\end{table}
-
-We model the way in which citations to a case depend upon the age of a case.
-For this we use a second-order polynomial in which one covariate $X^{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X^{ij}$ is the squared age of case $j$ at the time that case $i$ is decided.
-We include these covariates because \citet{black2013citation} find that the number of citations to a Supreme Court case over time depends significantly on the age of the case, characterized by a sharp drop off and leveling out with age.
-
-\citet{benjamin2012standing} study the propensity for cases to be overruled and cited in other negative ways.
-They find that cases with majority coalitions that are large and ideologically broad are less likely to be cited negatively.
-In our data we do not differentiate between negative and positive citations, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis.
-We include one covariate in which $X^{ij}$ is the number of justices in the majority coalition for case $j$.
-We also include another covariate in which $X^{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$.
-We expect both of these covariates to have positive effects.
-
-\begin{itemize}
-\item mention new variable: overruled cases
-\item mention new variable: nodemix(same issue area)
-\item mention new variable:Sender/receiver effect for judge who wrote majority opinion
-\item Same justice homophily with time trend
-\item Argue/justify why certain covariates are included
-\end{itemize}
-
-\textit{Comment from Benjamin Kassow: In section 5, I think the discussion is mostly good, but could be organized a bit differently. 
-I think the discussion of the random intercepts is a bit confusing early on (it is explained more effectively later). }
-%%%% DONT DELETE THESE COMMENTS, INCLUDE IDEAS FOR ADDITIONAL COVARIATES %%%%%
-
-%\citet{cross2010determinants} find that the number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
-(case level).
-Found case issue area, opinion ideology, opinion author, vote margin was not significant.
-Negative citations decreased future positive citations.
-
-%\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
-Show that the rate of citations to a case goes down very quickly.
-THE EFFECT OF AN INDEPENDENT VARIABLE ON A PRECEDENT IS DEPENDENT UPON THE AGE OF THE CASE.
-
-% \item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
-
-%\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
-Lower if stautory cases.
-Increases with ideological variance of the majority coalition.
-Increases with the number of concurring opinions.
-
  \section{Empirical Analysis}
 Our three data sources for this study include the Supreme Court Database (SCDB) \citep{spaeth2014supreme}, Martin-Quinn scores \citep{martin2002dynamic}, and Supreme Court citation data provided by the CourtListener Free Law Project \citep{CourtListener}.
+In the next section we explain the variables we construct using these data sources.
 We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015)\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the CourtListener data.
-We decided to exclude these 145 cases from our model.}.
-The Supreme Court citation network from $1937 - 2015$ consists of $9,945$ cases and the breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
-In order to only account for positive citations, we exclude $315$ citations that caused the cited case to be overruled.
-As a result, the network has a total of $111,986$ ties.
-The number of triangles in the network is $250,717$ and the number of mutual ties is $56$.
-The in- and outdegree distributions (i.e., the distributions of the number of citations sent and received by cases, respectively), are visualized in Figure \ref{degree_dist}.
-The maximum indegree is $230$ and the maximum outdegree is $162$.
+We decided to exclude these 145 cases from our analysis}.
+The Supreme Court citation network from $1937 - 2015$ consists of $9,945$ cases.
+The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
+The network has a total of $111,986$ citation ties.\footnote{In order to focus on citation actions that were not intended to totally invalidate an opinion \textbf{ HERE, SOURCE FOR OVERRULED, MAJORITY ONLY?}, we exclude $315$ citations that caused the cited case to be overruled.
+The data on overruling citation came from \textbf{CITE}.
+Our results are virtually unchanged if we include the overruling citations.} The in- and outdegree distributions (i.e., the distributions of the number of citations sent and received by cases, respectively), are visualized in Figure \ref{degree_dist}.
+The maximum indegree (i.e., number of cases citing to a case) is $230$ and the maximum outdegree (i.e., number of cases cited by a case) is $162$.
+The majority of cases cite to and/or are cited by twenty or fewer other cases.
 
@@
 The degree distributions indicate that there is a long tail to both the number of citations sent and received.
-These long-tailed distributions provide preliminary evidence of both popularity and activity dynamics.
+These long-tailed (i.e., high kurtosis) distributions provide preliminary evidence of both popularity and activity dynamics \citep{strogatz2001exploring}.
 Figure \ref{fig:networkviz} displays the complete network.
@@
 
+\subsection{Correlates of Legal Citations}
+
+Theorizing that legal citations arise through network interdependence processes represents a potential contribution to the literature.
+To demonstrate the value of our contribution we must also incorporate established non-network processes that have been found to affect legal citations.
+We incorporate these processes through the specification of $h_{covariate}$  terms to be included in the c-TERGM.
+
+We incorporate a statistic to model the degree to which cases cite those that are similar in terms of the ideological positions of the justices who supported the decision.
+We account for this effect following \citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
+\citet{clark2010locating} estimates a latent coordinate model of Supreme Court opinions based on the network of case-to-case citations.
+They find that the majority opinion falls at the ideal point of the median member of the majority coalition in the case.
+We include a covariate term in which $X_{ij}$ is the absolute difference between the Martin-Quinn scores \citep{martin2002dynamic} of the median justices in the majority coalitions for cases $i$ and $j$.
+We expect this variable to have a negative effect, which would correspond to cases citing those to which they are ideologically similar.%WE COULD USE THIS DATASET AS A SECOND APPLICATION
+
+We include two sets of dummy variables that account for the issue areas of cases.
+In one set of dummy variable---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$).
+In the set of receiver intercepts $X^{ij}$ indicates the issue of the receiver case.
+Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
+We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
+Table \ref{issue_area_coding} states the different issue areas.
+
+\begin{table}[]
+\centering
+\begin{tabular}{llll}
+1 & Criminal Procedure & 8  & Economic Activity    \\
+2 & Civil Rights       & 9  & Judicial Power       \\
+3 & First Amendment    & 10 & Federalism           \\
+4 & Due Process        & 11 & Interstate Relations \\
+5 & Privacy            & 12 & Federal Taxation     \\
+6 & Attorneys          & 13 & Miscellaneous        \\
+7 & Unions             & 14 & Private Action      
+\end{tabular}
+\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
+\label{issue_area_coding}
+\end{table}
+
+We model the way in which citations to a case depend upon the age of a case.
+For this we use a second-order polynomial in which one covariate $X^{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X^{ij}$ is the squared age of case $j$ at the time that case $i$ is decided.
+We include these covariates because \citet{black2013citation} find that the number of citations to a Supreme Court case over time depends significantly on the age of the case, characterized by a sharp drop off and leveling out with age.
+
+\citet{benjamin2012standing} study the propensity for cases to be overruled and cited in other negative ways.
+They find that cases with majority coalitions that are large and ideologically broad are less likely to be cited negatively.
+In our data we do not differentiate between negative and positive citations, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis.
+We include one covariate in which $X^{ij}$ is the number of justices in the majority coalition for case $j$.
+We also include another covariate in which $X^{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$.
+We expect both of these covariates to have positive effects.
+
 \begin{itemize}
-\item update Figure 2 to time period 1937 - 2015
-\item add data source for overruling citations
+\item mention new variable: overruled cases
+\item mention new variable: nodemix(same issue area)
+\item mention new variable:Sender/receiver effect for judge who wrote majority opinion
+\item Same justice homophily with time trend
+\item Argue/justify why certain covariates are included
 \end{itemize}
+
+\textit{Comment from Benjamin Kassow: In section 5, I think the discussion is mostly good, but could be organized a bit differently. 
+I think the discussion of the random intercepts is a bit confusing early on (it is explained more effectively later). }
+%%%% DONT DELETE THESE COMMENTS, INCLUDE IDEAS FOR ADDITIONAL COVARIATES %%%%%
+
+%\citet{cross2010determinants} find that the number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
+(case level).
+Found case issue area, opinion ideology, opinion author, vote margin was not significant.
+Negative citations decreased future positive citations.
+
+%\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
+Show that the rate of citations to a case goes down very quickly.
+THE EFFECT OF AN INDEPENDENT VARIABLE ON A PRECEDENT IS DEPENDENT UPON THE AGE OF THE CASE.
+
+% \item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
+
+%\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
+Lower if stautory cases.
+Increases with ideological variance of the majority coalition.
+Increases with the number of concurring opinions.
 

### 2018-05-02 Tex/paper.tex
revised the covariate discussion

@@
 The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
-The network has a total of $111,986$ citation ties.\footnote{In order to focus on citation actions that were not intended to totally invalidate an opinion \textbf{ HERE, SOURCE FOR OVERRULED, MAJORITY ONLY?}, we exclude $315$ citations that caused the cited case to be overruled.
-The data on overruling citation came from \textbf{CITE}.
+The network has a total of $111,986$ citation ties.\footnote{In order to focus on citation actions that were not intended to totally invalidate an opinion, we exclude $315$ citations that caused the cited case to be overruled.
+The data on overruling citation came from \citet{senate2016constitution}.
 Our results are virtually unchanged if we include the overruling citations.} The in- and outdegree distributions (i.e., the distributions of the number of citations sent and received by cases, respectively), are visualized in Figure \ref{degree_dist}.
@@
 
-\subsection{Correlates of Legal Citations}
-
-Theorizing that legal citations arise through network interdependence processes represents a potential contribution to the literature.
-To demonstrate the value of our contribution we must also incorporate established non-network processes that have been found to affect legal citations.
-We incorporate these processes through the specification of $h_{covariate}$  terms to be included in the c-TERGM.
-
-We incorporate a statistic to model the degree to which cases cite those that are similar in terms of the ideological positions of the justices who supported the decision.
+\subsection{c-TERGM specification}
+
+The c-TERGM we estimate includes two classes of terms---one that captures the effects of covariates on tie formation, and another that captures the complex dependence processes that we expect to observe in the Supreme Court Citation Network.
+We describe our model specification by defining the terms within these two classes.
+
+\subsubsection{Covariate terms}
+
+Covariate effects are accounted for in the c-TERGM via the term that is used to specify the effect of covariates in other ERGM family models, as
+$$h_{covariate}(C_t,C_{<t},X) =  \sum_{ij} C_{ij,t}X_{ij}.$$ Since $C_{ij,t}$ is a binary indicator of whether case $i$ cites case $j$, $h_{covariate}$ amounts to the sum of covariate values among directed dyads for which we observed a citation.
+The dyadic interpretation of the coefficient attributed to this term is the change in the log odds of a tie from $i$ to $j$ given a unit increase in $X_{ij}$ (i.e., exactly the interpretation of the effect of a covariate in logistic regression).
+We include several exogenous covariates based on this standard statistic formulation.
+
+The first covariate (i.e., $X$) we incorporate into the model accounts for the degree to which cases cite those that are similar in terms of the ideological positions of the justices who supported the decision.
 We account for this effect following \citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
@@
 
-We include two sets of dummy variables that account for the issue areas of cases.
-In one set of dummy variable---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$).
-In the set of receiver intercepts $X^{ij}$ indicates the issue of the receiver case.
-Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
+We include three sets of dummy variables that account for the issue areas of cases.
+In the first set of dummy variables---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$) (e.g., for the Criminal Procedure indicator, $X_{ij} = 1$ if case $i$ is a criminal procedure case, and $0$ otherwise).
+The second set of dummy variables represent receiver issue area intercepts in which $X^{ij}$ indicates the issue of the receiver case.
+The third set of dummy variables account for the rate of citation within each issue area and between each combination of issue areas.
+There is a separate dummy variable for each directed pair of issue areas.
+For example the A$\rightarrow$B dummy variable $X_{ij} = 1$ if case $i$ is issue area A and case $j$ is issue area B, and 0 otherwise.
+Note that, given the high dimension of these dummy variables, we omit one each of the sender and recipient indicators, and 28 of the 196 mixing rate indicators.
+Unfortunately, this leaves these indicators largely uninterpretable, but these variables are nonetheless important to absorb variation in citations that is attributable to the legal domains of two cases in a dyad.
+The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
 We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
-Table \ref{issue_area_coding} states the different issue areas.
+Table \ref{issue_area_coding} reports the issue areas covered in our data.
+
+We also include a set of covariates that account for justice-specific variation in citations.
+We include two types of terms at the justice level.
+First, we include an indicator variable that models the rate at which justices cite themselves.
+This effect is modeled with a single indicator variable in which $X_{ij} = 1$ if the majority opinions in cases $i$ and $j$ were written by the same justice.
+Second, we use two sets of dummy variables to account for differences across justices both in terms of their tendencies to include citations in the opinions they write and to be cited by other opinions.
+The justice sender dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $i$.
+The justice recipient dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $j$.
+This set of variables is motivated by the use of citations by \citet{kosma1998measuring} to measure the influence of individual Supreme Court justices.
 
@@
 
-\begin{itemize}
-\item mention new variable: overruled cases
-\item mention new variable: nodemix(same issue area)
-\item mention new variable:Sender/receiver effect for judge who wrote majority opinion
-\item Same justice homophily with time trend
-\item Argue/justify why certain covariates are included
-\end{itemize}
-
-\textit{Comment from Benjamin Kassow: In section 5, I think the discussion is mostly good, but could be organized a bit differently. 
-I think the discussion of the random intercepts is a bit confusing early on (it is explained more effectively later). }
+The final control variable we include in the model is one in which $X_{ij} = 1$ if case $j$ was overruled prior to the term in which case $i$ was decided.
+This variable, quite simply, models the effect of being overruled on the rate of citation to a case after the overruling citation.
+\citet{fowler2008authority} find that citations to a case drop off quickly after the case has been overruled.
+
+\subsubsection{Dependence terms}
+
+%\textit{Comment from Benjamin Kassow: In section 5, I think the discussion is mostly good, but could be organized a bit differently. 
+%I think the discussion of the random intercepts is a bit confusing early on (it is explained more effectively later). }
 %%%% DONT DELETE THESE COMMENTS, INCLUDE IDEAS FOR ADDITIONAL COVARIATES %%%%%
@@
 \caption{Nodeofactor Majority Opinion Writer. Baseline= LDBrandeis }
- \label{number_supporting}
-\vspace{-.25cm}
-\end{figure}
-
- \begin{figure}[H]
-  \begin{tabular}{c}
-\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_activity}
- \\
-\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_popularity}
-\end{tabular}
-\caption{Issue Area Popularity and Activity (nodeifactor and nodeofactor). Baseline is Issue Area 'Criminal Procedure'.}
-\end{figure}
-
- \begin{figure}[H]
-\includegraphics[width=15cm]{heatplot_issue_area}
-\caption{Nodemix Issue Area. Sender14Receiver14 and Sender14Receiver13 couldnt be estimated.}
  \label{number_supporting}
@@
 
+ \begin{figure}[H]
+  \begin{tabular}{c}
+\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_activity}
+ \\
+\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_popularity}
+\end{tabular}
+\caption{Issue Area Popularity and Activity (nodeifactor and nodeofactor). Baseline is Issue Area 'Criminal Procedure'.}
+\end{figure}
+
+ \begin{figure}[H]
+\includegraphics[width=15cm]{heatplot_issue_area}
+\caption{Nodemix Issue Area. Sender14Receiver14 and Sender14Receiver13 couldnt be estimated.}
+ \label{number_supporting}
+\vspace{-.25cm}
+\end{figure}
+
 \end{document}

### 2018-05-08 Tex/paper.tex
added information about agreement between Fowler and Courtlistener

@@
 We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015)\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the CourtListener data.
-We decided to exclude these 145 cases from our analysis}.
+We decided to exclude these 145 cases from our analysis.
+Additionally, since the most commonly used data on Supreme Court citations in political science come from \citet{fowler2007network}, we compared the Fowler data to the CourtListener data in the time interval that they overlap (1937--2001).
+We found considerable agreement---over 95\% of the citations recorded in the CourtListener dataset were also in the Fowler dataset, and over 96\% of the citations recorded in the Fowler data were also in the CourtListener dataset.}.
 The Supreme Court citation network from $1937 - 2015$ consists of $9,945$ cases.
@@
 
-%\textit{Comment from Benjamin Kassow: In section 5, I think the discussion is mostly good, but could be organized a bit differently. 
-%I think the discussion of the random intercepts is a bit confusing early on (it is explained more effectively later). }
-%%%% DONT DELETE THESE COMMENTS, INCLUDE IDEAS FOR ADDITIONAL COVARIATES %%%%%
-
-%\citet{cross2010determinants} find that the number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
-(case level).
-Found case issue area, opinion ideology, opinion author, vote margin was not significant.
-Negative citations decreased future positive citations.
-
-%\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
-Show that the rate of citations to a case goes down very quickly.
-THE EFFECT OF AN INDEPENDENT VARIABLE ON A PRECEDENT IS DEPENDENT UPON THE AGE OF THE CASE.
-
-% \item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
-
-%\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
-Lower if stautory cases.
-Increases with ideological variance of the majority coalition.
-Increases with the number of concurring opinions.
+We include a dependence term to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, popularity, and activity.
+We expect that the coefficients associated with each of these dependence effects will be positive.
+The dependence effect included in the model to account for transitivity is the count of the number of triangles in $C_{\leq t}$.
+The log odds interpretation of the triangle statistic is that  $\theta$ gives the change in the log odds of case $i$ citing case $j$ associated  with each additional triangle that would be created by case $i$ citing case $j$ (e.g., if opinions in cases $i$ and $j$ cite one common case $k$, the log odds that case $i$ cites case $j$ change by theta relative to the configuration in which cases $i$ and $j$ do not cite any of the same cases).
+
+The statistic included in the model for reciprocity counts the number of mutual dyads (i.e., dyads in which cases $i$ cites case $j$, and case $j$ cites case $i$) in $C_{\leq t}$.
+In practice, this is the number of mutual dyads in $C_{\leq t}$, since mutual dyads must emerge among opinions that were written in the same term.The log odds interpretation of the reciprocity term is that $\theta$ gives the change in the log odds of a tie
+
+  \subsubsection{Dynamics in effects}
 
@@
 In the next section we describe a methodology that can be used to integrate case attributes and graph structures/motifs into a comprehensive model of US Supreme Court citations.
+
+%\textit{Comment from Benjamin Kassow: In section 5, I think the discussion is mostly good, but could be organized a bit differently. 
+%I think the discussion of the random intercepts is a bit confusing early on (it is explained more effectively later). }
+%%%% DONT DELETE THESE COMMENTS, INCLUDE IDEAS FOR ADDITIONAL COVARIATES %%%%%
+
+%\citet{cross2010determinants} find that the number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
+(case level).
+Found case issue area, opinion ideology, opinion author, vote margin was not significant.
+Negative citations decreased future positive citations.
+
+%\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
+Show that the rate of citations to a case goes down very quickly.
+THE EFFECT OF AN INDEPENDENT VARIABLE ON A PRECEDENT IS DEPENDENT UPON THE AGE OF THE CASE.
+
+% \item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
+
+%\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
+Lower if stautory cases.
+Increases with ideological variance of the majority coalition.
+Increases with the number of concurring opinions.

### 2018-08-27 Tex/paper.tex
added maggie's grant
added APSA version

@@
 The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.\footnote{
-Prepared for presentation at the ``The Enemy of My Enemy: Judicial Politics meets Network Analysis'' conference within a conference at the 2018 Annual Conference of the Southern Political Science Association.
-This work was supported in part by NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219.
-We would like to thank Rachael Hinkle and Benjamin Kassow.}
+Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. This work was supported in part by NIH, 
+R01 AI36664-01. NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219. We would like to thank Rachael Hinkle and Benjamin Kassow for detailed comments on an earlier draft of this paper.}
 \end{abstract}
@@
 
-\textit{Comment from Benjamin Kassow: In section 6.2 (predictive performance), I like what is here, but was hoping to see a bit more discussion about the predictive 
-performance of the model, with some explanation of what the predictive performance means. 
-I understand it (I think), but depending on where this gets sent to, I am not sure if all reviewers will. 
-I also think tying this into the overall model explanation or discussion of model performance overall will be helpful for better 
-integrating this section into the rest of the manuscript.}
+%\textit{Comment from Benjamin Kassow: In section 6.2 (predictive performance), I like what is here, but was hoping to see a bit more discussion about the predictive 
+%performance of the model, with some explanation of what the predictive performance means. 
+%I understand it (I think), but depending on where this gets sent to, I am not sure if all reviewers will. 
+%I also think tying this into the overall model explanation or discussion of model performance overall will be helpful for better 
+%integrating this section into the rest of the manuscript.}
 
@@
 
-The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes.
+%The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes.
 The scales of covariate and dependence effects are comparable.
@@
 
-\begin{itemize}
-\item adjust results/numbers in text
-\end{itemize}
+{\em {\bf Note:} The results section of this paper is incomplete and in flux.
+We recently discovered evidence of non-convergence in the C-TERGM results currently presented in the paper, and have changed the models and estimation settings to address this problem.
+Most of the substantive conclusions look like they will hold using the updated models, but we are not sure that we will be able to include the updated models in the APSA presentation.}
 
@@
 
-\begin{itemize}
-\item visualize Sender/receiver Issue Area Results
-\item heatplot for issue area nodemix
-\item visualize Sender/receiver for majority opinion writer
-\item add Network effect plots for model without network statistics (add to existing plots?)
-\end{itemize}
+%\begin{itemize}
+%\item visualize Sender/receiver Issue Area Results
+%\item heatplot for issue area nodemix
+%\item visualize Sender/receiver for majority opinion writer
+%\item add Network effect plots for model without network statistics (add to existing plots?)
+%\end{itemize}
 
@@
 
-\begin{itemize}
-\item add robustness check: MCMLE for each year -> plot time trend of coefs -> are results similar to the MPLE time trends presented in the paper?
-\item (investigate which statistic contributes most on recall improvement)
-\end{itemize}
+%\begin{itemize}
+%\item add robustness check: MCMLE for each year -> plot time trend of coefs -> are results similar to the MPLE time trends presented in the paper?
+%\item (investigate which statistic contributes most on recall improvement)
+%\end{itemize}
 

### 2020-03-31 Tex/paper.tex
Updated Section 4.1.1 Covariate Terms

@@
   \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
-  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
+ % \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
   }
@@
  \section{The Citation Temporal Exponential Random Graph Model}
+{\color{blue} This section needs to be completely revised. Model is not a TERGM anymore. - CS} 
 We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
@@
 
-We include three sets of dummy variables that account for the issue areas of cases.
+%We include three sets of dummy variables that account for the issue areas of cases.
 In the first set of dummy variables---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$) (e.g., for the Criminal Procedure indicator, $X_{ij} = 1$ if case $i$ is a criminal procedure case, and $0$ otherwise).
@@
 
-We also include a set of covariates that account for justice-specific variation in citations.
+We include one covariate that accounts for the rate at which cases that fall under the same issue area cite each other.
+The variable $X^{ij}$ indicates whether cases $i$ and $j$ share the same issue area, e.g., $X^{ij}=1$ if cases $i$ and $j$ are classified into the same issue area, and $0$ otherwise.
+The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
+We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
+Table \ref{issue_area_coding} reports the issue areas covered in our data.
+
+%We also include a set of covariates that account for justice-specific variation in citations.
 We include two types of terms at the justice level.
@@
 This set of variables is motivated by the use of citations by \citet{kosma1998measuring} to measure the influence of individual Supreme Court justices.
+
+We also include an indicator variable that models the rate at which justices cite themselves.
+Similar to the same issue area variable, this effect is modeled with a single indicator variable in which $X_{ij} = 1$ if the majority opinions in cases $i$ and $j$ were written by the same justice.
+The data is taken from the SCDB \citep{spaeth2014supreme}.
+{\color{blue} Justification for including this variable?
+}
 
@@
 
-  \subsubsection{Dynamics in effects}
+ % \subsubsection{Dynamics in effects}
+
+% For the current model this section is not relevant anymore
 

### 2020-04-01 Tex/paper.tex
started writing section 4.1.2 Dependence Terms

@@
 We expect that the coefficients associated with each of these dependence effects will be positive.
-The dependence effect included in the model to account for transitivity is the count of the number of triangles in $C_{\leq t}$.
-The log odds interpretation of the triangle statistic is that  $\theta$ gives the change in the log odds of case $i$ citing case $j$ associated  with each additional triangle that would be created by case $i$ citing case $j$ (e.g., if opinions in cases $i$ and $j$ cite one common case $k$, the log odds that case $i$ cites case $j$ change by theta relative to the configuration in which cases $i$ and $j$ do not cite any of the same cases).
+
+The common dependence effects for these four dynamics are the number of triangles (case $i$ cites cases $j$ and $k$, and case $j$ cites case $k$), the numnber of mutual dyads (case $i$ cites $j$ and vice versa), the numnber of out-2-stars (case $i$ cites $j$ and $k$), and the number of in-2-stars (case $i$ is being cited by cases $j$ and $k$) in $C_{\leq t}$.
+Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy \citep{Handcock.2003}.
+Model degeneracy occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network.
+In this scenario, the simulated networks are too different from the observed network, making the underlying distribution defined by the model extremly unrealistc.
 
 The statistic included in the model for reciprocity counts the number of mutual dyads (i.e., dyads in which cases $i$ cites case $j$, and case $j$ cites case $i$) in $C_{\leq t}$.
-In practice, this is the number of mutual dyads in $C_{\leq t}$, since mutual dyads must emerge among opinions that were written in the same term.The log odds interpretation of the reciprocity term is that $\theta$ gives the change in the log odds of a tie
+In practice, this is the number of mutual dyads in $C_{\leq t}$, since mutual dyads must emerge among opinions that were written in the same term.
+The log odds interpretation of the reciprocity statistic is that if opinions in case $i$ cite case $j$, the log odds that case $j$ also cites $i$ change by the theta relative to the configuration in which case $j$ does not cite case $i$.
+
+Accouting for the popularity effect through the in-2-star statistic is prone to cause model degeneracy.
+In an attempt to stabilize the model against model degeneracy as well as to still capture the popularity effect, \citet{SnijdersTomA.B..2006} introduces the \textit{alternating-in-k-star} statistic, which was shown to be equivalent to the \textit{geometrically weighted in-degree distribution} (gwidegree) statistic introduced by \citet{hunter2006inference}.
+Define $D_r(C_t, C_{<t})$ as the number of nodes with indegree $r, r \in \{0, \dots , N_t -1 \}$, where $N_t$ is the number of cases in the network at time $t$, then gwidegree represents a network's indegree distribution in a single statistic by geometrically weighting the degree distribution
+$$h_{gwidegree}(C_{t}, C_{<t}, \lambda)= \lambda \sum_{r=1}^{N_t-1}\biggl(1-\bigl(1-\frac{1}{\lambda}\bigr)^r\biggr)D_r(C_t, C_{<t}). $$
+The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on $D_r(C_t, C_{<t})$ as $r$ increases.
+Adding one single edge to the network changes the indegree distribution of the network so that a case with an indegree of $s$ turns into a case with an indegree of $s + 1$.
+With our notation this means that $D_r$ and $D_{r+1}$ get replaced by
+$D_r -1$ and $D_{r+1} +1$, while no changes for $D_\ell,
+ \ell \in \{0, \dots , N_t -1 \} \setminus \{r, r-1\} $ are
+made. This changes the log odds in the following way
+$$\text{logit} P(C_{ij,t}=1 | C_{-ij,t}, C_{<t}) = \exp \bigl(~\theta \cdot( 1- \frac{1}{\lambda})^r \bigr).$$
+
+%In order to account for the popularity effect, we add the \textit{geometrically weighted in-degree distribution} (gwidegree) statistic to the model.
+The geometrically weighted distribution statistics were introduced by \citet{Hunter.2006} to counteract model degeneracy.
+
+%The log odds interpretation of the triangle statistic is that  $\theta$ gives the change in the log odds of case $i$ citing case $j$ associated  with each additional triangle that would be created by case $i$ citing case $j$ (e.g., if opinions in cases $i$ and $j$ cite one common case $k$, the log odds that case $i$ cites case $j$ change by theta relative to the configuration in which cases $i$ and $j$ do not cite any of the same cases).
 

### 2020-04-02 Tex/paper.tex
Wrote paragraphs about gwidegree and nodeicov(odegree) in Section 4.1.2 Dependence Terms

@@
 
+% reciprocity
 The statistic included in the model for reciprocity counts the number of mutual dyads (i.e., dyads in which cases $i$ cites case $j$, and case $j$ cites case $i$) in $C_{\leq t}$.
@@
 
+% popularity
 Accouting for the popularity effect through the in-2-star statistic is prone to cause model degeneracy.
@@
 Define $D_r(C_t, C_{<t})$ as the number of nodes with indegree $r, r \in \{0, \dots , N_t -1 \}$, where $N_t$ is the number of cases in the network at time $t$, then gwidegree represents a network's indegree distribution in a single statistic by geometrically weighting the degree distribution
-$$h_{gwidegree}(C_{t}, C_{<t}, \lambda)= \lambda \sum_{r=1}^{N_t-1}\biggl(1-\bigl(1-\frac{1}{\lambda}\bigr)^r\biggr)D_r(C_t, C_{<t}). $$
-The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on $D_r(C_t, C_{<t})$ as $r$ increases.
+\begin{equation}
+h_{gwidegree}(C_{t}, C_{<t}, \lambda)= \lambda \sum_{r=1}^{N_t-1}\biggl(1-\bigl(1-\frac{1}{\lambda}\bigr)^r\biggr)D_r(C_t, C_{<t}). 
+\label{gwidegree}
+\end{equation}
+The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on $D_r(C_t, C_{<t})$ as $r$ increases which means that the higher $\lambda$ the more the statistic responds to changes for cases with a high indegree.
+We fixed $\lambda=1$.
 Adding one single edge to the network changes the indegree distribution of the network so that a case with an indegree of $s$ turns into a case with an indegree of $s + 1$.
@@
 $$\text{logit} P(C_{ij,t}=1 | C_{-ij,t}, C_{<t}) = \exp \bigl(~\theta \cdot( 1- \frac{1}{\lambda})^r \bigr).$$
+As stated by \citet{Levy2016} this means for the interpretation that a negative parameter value indicates the centralization of edges while a positive parameter indicates the dispersion of edges.
+Since we expect highly cited cases to be more likely to be cited again, we expect the gwidegree parameter to be negative.
+
+% Activity
+In order to account for the activity effect, the straightforward thing to do is to add the \textit{geometrically weighted outdegree} (gwodegree) statistic, which is defined as gwidegree with the difference that $D_r(C_t, C_{<t})$ represents the number of nodes with outdegree $r$.
+However, adding the gwodegree statistic causes the model to degenerate again.
+We circumvent degeneracy by adding the outdegree of every node that entered the network prior time $t$ as a receiver effect.
+In other words, this statistic adds up the outdegree of cases that are being cited at time $t$ and examines whether cases that show a high degree of activity are being more likely to be cited in the future.
+
+% Transitivity
 

### 2020-04-09 Tex/paper.tex
Wrote about Transitivity terms in section 4.1.2 Dependence Terms

@@
 
-We include a dependence term to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, popularity, and activity.
-We expect that the coefficients associated with each of these dependence effects will be positive.
+We include dependence terms to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, popularity, and activity.
+%We expect that the coefficients associated with each of these dependence effects will be positive.
 
@@
 Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy \citep{Handcock.2003}.
-Model degeneracy occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network.
+Model degeneracy is a common obstacle when fitting ERGMs and occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network.
 In this scenario, the simulated networks are too different from the observed network, making the underlying distribution defined by the model extremly unrealistc.
@@
 % Transitivity
+We include two statistics to account for different types of transitivity.
+The first transitivity statistic calculates the number of transitive ties, $i \to j$, where there is a directed path of length two from $i$ to $j$ through at least one case $k$, and $j$ and $k$ were written during a different term than $i$.
+We will refer to this statistic as the \textit{different term transitivity} statistic.
+As illustrated in chapter 2 we theorize Supreme Court citations to pattern a high degree of transitivity.
+As a result, we expect positive parameter estimates for the different term statistics.
+
+The second transitivity statistic is the \textit{geometrically weighted edgewise shared partners} (gwesp) statistic introduced by \citet{hunter2006inference}.
+For directed networks the edgewise shared $r$-partners statistic ($ESP_r$) can take multiple forms, but we integrate the definition where cases $i$ and $j$ that are connected by a citation, also cite $r$ common other cases.
+Similar to the gwidegree-statistic the gwesp-statistic incorporates a network's edgewise shared partners distribution by replacing $D_r$ by $ESP_r$.
+Note that unlike the different term transitivity statistic the gwesp statistic does not require the cited cases to be written in a previous term.
+
+ \textit{GOF-plot?}\\
 

### 2020-04-13 Tex/paper.tex
Updated section 4.2 Predictive Performance

@@
 
-\subsubsection{Covariate terms}
+\subsubsection{Covariate terms}\label{covariate_terms}
 
@@
 
-\subsubsection{Dependence terms}
+\subsubsection{Dependence terms}\label{dependence_terms}
 
@@
 Our case for studying legal citations at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence.
-To quantify this contribution, we use out-of-sample prediction.
+To quantify this contribution, we use AIC and BIC out-of-sample prediction.
 Predicting out-of-sample offers an unbiased and general purpose way to evaluate the contribution, in terms of model fit, of one or more terms/parameters in a model \citep{jensen2000multiple,ward2010perils}.
@@
 
-In our prediction experiment we randomly split the directed dyads into an 80\% training set and a 20\% test set.
+  \begin{figure}[]
+\includegraphics[width=15cm]{SCC_AIC_BIC}
+\caption{AIC and BIC for the full and the independent model for the time frame 1950-2015 }
+ \label{AIC_BIC}
+\vspace{-.25cm}
+\end{figure}
+
+Figure \ref{AIC_BIC} depicts the AIC and BIC comparison for two different models for each term from 1950 to 2015.
+The full model is defined as the model that includes all covariate terms and dependence terms described in \ref{covariate_terms} and \ref{dependence_terms}.
+The independent model only consists of covariate terms, not accounting for any structural dependencies in the networks.
+We see that the AIC and BIC is considerably smaller for the full model throughout the period of consideration.
+On average the independent model yields a 5525 point higher AIC and an average 5582 point higher BIC.
+The maximum difference is reached in 1983 with a difference of 10,990 points for the AIC and 10,930  points for the BIC.
+
+%In our prediction experiment we randomly split the directed dyads into an 80\% training set and a 20\% test set.
 The parameters of the model are estimated using the directed dyads in the training set, and the parameters are used to form the conditional probability of a tie for all of the directed dyads in the test set.
@@
 
-\begin{table}[H]
-\centering
-
-\begin{tabular}{rllll}
-\hline \hline
-& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
- & mean & range & mean & range \\ 
-  \hline
-precision & 0.6007 & (0.5934, 0.6074) & 0.8666 & (0.8647, 0.8701) \\ 
-  recall & 0.1522 & (0.1495, 0.1549) & 0.5969 & (0.5945, 0.6019) \\ 
-  F1 score & 0.2429 & (0.2389, 0.2462) & 0.7069 & (0.7048, 0.7107) \\ 
-   \hline \hline
-\end{tabular}
-\caption{The predictive performance of the directed dyadic models with linear trends, over ten 80/20 train/test splits.}
-\label{tab:prediction_linear}
-
-\end{table}
-
-We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms.
+%\begin{table}[H]
+%\centering
+
+%\begin{tabular}{rllll}
+%\hline \hline
+%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
+% & mean & range & mean & range \\ 
+%  \hline
+%precision & 0.6007 & (0.5934, 0.6074) & 0.8666 & (0.8647, 0.8701) \\ 
+%  recall & 0.1522 & (0.1495, 0.1549) & 0.5969 & (0.5945, 0.6019) \\ 
+%  F1 score & 0.2429 & (0.2389, 0.2462) & 0.7069 & (0.7048, 0.7107) \\ 
+%   \hline \hline
+%\end{tabular}
+%\caption{The predictive performance of the directed dyadic models with linear trends, over ten 80/20 train/test splits.}
+%\label{tab:prediction_linear}
+
+%\end{table}
+
+%We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms.
 The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set.
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{rllll}
-\hline \hline
-& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
- & mean & range & mean & range \\ 
-  \hline
-precision & 0.5938 & (0.5844, 0.5996) & 0.8668 & (0.8653, 0.8693) \\ 
-  recall & 0.1551 & (0.1515, 0.1575) & 0.5969 & (0.5937, 0.6005) \\ 
-  F1 score & 0.246 & (0.2406, 0.249) & 0.707 & (0.7045, 0.7095) \\  
-   \hline \hline
-\end{tabular}
-\caption{The predictive performance of the directed dyadic models with no trends, over ten 80/20 train/test splits.}
-\label{tab:prediction_notrend}
-\vspace{0.4cm}
-\begin{tabular}{rllll}
-\hline \hline
-& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
- & mean & range & mean & range \\ 
-  \hline
-precision & 0.6209 & (0.611, 0.6351) & 0.8669 & (0.8612, 0.8711) \\ 
-  recall & 0.1511 & (0.1451, 0.1558) & 0.5962 & (0.5884, 0.6011) \\ 
-  F1 score & 0.243 & (0.2347, 0.2486) & 0.7065 & (0.7011, 0.7103) \\ 
-   \hline \hline
-\end{tabular}
-\caption{The predictive performance of the directed dyadic models with quadratic trends, over ten 80/20 train/test splits.}
-\label{tab:prediction_quadratic}
-\end{table}
-We also conduct a predictive experiment for the model that does not consider any trends and for the model that does consider quadratic instead of linear trends.
+%\begin{table}[H]
+%\centering
+%\begin{tabular}{rllll}
+%\hline \hline
+%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
+% & mean & range & mean & range \\ 
+%  \hline
+%precision & 0.5938 & (0.5844, 0.5996) & 0.8668 & (0.8653, 0.8693) \\ 
+%  recall & 0.1551 & (0.1515, 0.1575) & 0.5969 & (0.5937, 0.6005) \\ 
+%  F1 score & 0.246 & (0.2406, 0.249) & 0.707 & (0.7045, 0.7095) \\  
+%   \hline \hline
+%\end{tabular}
+%\caption{The predictive performance of the directed dyadic models with no trends, over ten 80/20 train/test splits.}
+%\label{tab:prediction_notrend}
+%\vspace{0.4cm}
+%\begin{tabular}{rllll}
+%\hline \hline
+%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
+% & mean & range & mean & range \\ 
+%  \hline
+%precision & 0.6209 & (0.611, 0.6351) & 0.8669 & (0.8612, 0.8711) \\ 
+%  recall & 0.1511 & (0.1451, 0.1558) & 0.5962 & (0.5884, 0.6011) \\ 
+%  F1 score & 0.243 & (0.2347, 0.2486) & 0.7065 & (0.7011, 0.7103) \\ 
+%   \hline \hline
+%\end{tabular}
+%\caption{The predictive performance of the directed dyadic models with quadratic trends, over ten 80/20 train/test splits.}
+%\label{tab:prediction_quadratic}
+%\end{table}
+%We also conduct a predictive experiment for the model that does not consider any trends and for the model that does consider quadratic instead of linear trends.
 The predictive performance results are given in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic}.
@@
 
-\begin{figure}[H]
- \begin{tabular}{cc}
-
-   (a) Reciprocity & (b) Transitivity \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
-
-    (c) Activity & (d) Popularity \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
-
-\end{tabular}
-\caption{Trending effects for dependence effect estimates from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
- \label{fig:coeftrends_endo}
-\vspace{-.25cm}
-\end{figure}
-
-  \begin{figure}[H]
-  \begin{tabular}{cc}
- (e) MQ Score Difference & (f) Ideological Breadth \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\
-
-  (g) Same Issue & (h) Majority Size \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\
-
- (i) Difference in Years & (j) Difference in Years$^2$ \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\
-
- (k) Overruled Cases & (l) Justice Homophily \\
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} & 
-\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/justicehomophily_coef_trend_linear.pdf} \\
-
- \end{tabular}
-\caption{Trending effects for exogenous variables from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
- \label{fig:coeftrends_exo}
-\vspace{-.25cm}
-\end{figure}
+%\begin{figure}[H]
+% \begin{tabular}{cc}
+
+%   (a) Reciprocity & (b) Transitivity \\
+%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
+
+%    (c) Activity & (d) Popularity \\
+%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
+
+%\end{tabular}
+%\caption{Trending effects for dependence effect estimates from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
+% \label{fig:coeftrends_endo}
+%\vspace{-.25cm}
+%\end{figure}
+
+%  \begin{figure}[H]
+%  \begin{tabular}{cc}
+% (e) MQ Score Difference & (f) Ideological Breadth \\
+%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\
+
+%  (g) Same Issue & (h) Majority Size \\
+%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\
+
+% (i) Difference in Years & (j) Difference in Years$^2$ \\
+%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\
+
+% (k) Overruled Cases & (l) Justice Homophily \\
+%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} & 
+%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/justicehomophily_coef_trend_linear.pdf} \\
+
+% \end{tabular}
+%\caption{Trending effects for exogenous variables from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
+% \label{fig:coeftrends_exo}
+%\vspace{-.25cm}
+%\end{figure}
 
@@
   \begin{figure}[H]
-\includegraphics[width=15cm]{justice_popularity}
-\caption{Nodeifactor Majority Opinion Writer. Baseline= LDBrandeis }
- \label{number_supporting}
+\includegraphics[width=16cm]{SCC_Results_1}
+\caption{ERGM results for the dependence terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1. }
+ \label{SCC_results_1}
 \vspace{-.25cm}
@@
  \begin{figure}[H]
-\includegraphics[width=15cm]{justice_activity}
-\caption{Nodeofactor Majority Opinion Writer. Baseline= LDBrandeis }
- \label{number_supporting}
+\includegraphics[width=16cm]{SCC_Results_2}
+\caption{ERGM results for the covariate terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1.}
+ \label{SCC_results_2}
 \vspace{-.25cm}

### 2020-04-14 Tex/paper.tex
Started Updating Results-section

@@
 \subsection{c-TERGM Results}
-The results from the c-TERGM are presented in Table \ref{tab:tergmcoef} and Figures \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}.
-Table \ref{tab:tergmcoef} gives the coefficient estimates and 95\% bootstrap confidence intervals.
-In Figures, \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}, we depict coefficient estimates and confidence intervals for all of the effects for which we estimated trends (excluding the intercept---edge count---which is typically not interpreted substantively).
+The results from the c-TERGM are presented Figures \ref{SCC_results_1} and \ref{SCC_results_2}.
+In these figures we depict coefficient estimates and 95\% confidence intervals for all of the effects for all the variables we discussed in sections \ref{covariate_terms} and \ref{dependence_terms}.
 The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one unit increase in the respective variable.
-We first discuss the dependence effect estimates (panels (a)--(d) in Figure \ref{fig:coeftrends_endo}).
-The reciprocity effect is substantial, positive, and statistically significant for most of the period under study.
+For this reason, the number of edges in a network corresponds to the intercept in an ERGM.
+We first discuss the dependence effect estimates in Figure \ref{SCC_results_1}.
+A circle indicates that the estimate was significant at an $\alpha=0.05$ level, while and square translates to a p-value between $0.05$ and $0.1$.
+A triangle signals that the estimate for a given year was not significant at an $\alpha=0.1$ level.
+
+%The reciprocity effect is substantial, positive, and statistically significant for most of the period under study.
 Conditional upon cases $i$ and $j$ being considered within the same term, the log of the odds that opinion $i$ cites opinion $j$ increases by approximately $2$ when opinion $j$ cites opinion $i$.
@@
 
-The effects of the exogenous covariates included in the c-TERGM (panels (e)--(k) in Figure \ref{fig:coeftrends_exo}) exhibit some surprising dynamics.
+%The effects of the exogenous covariates included in the c-TERGM (panels (e)--(k) in Figure \ref{fig:coeftrends_exo}) exhibit some surprising dynamics.
 The effects of absolute difference in Martin-Quinn score, and majority size of the recipient case all flip signs throughout the time period under study.
@@
 
-{\em {\bf Note:} The results section of this paper is incomplete and in flux.
+%{\em {\bf Note:} The results section of this paper is incomplete and in flux.
 We recently discovered evidence of non-convergence in the C-TERGM results currently presented in the paper, and have changed the models and estimation settings to address this problem.
@@
 
-\begin{table}[htp]
-\footnotesize
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{EFEFEF}}l |l|l|l|l|} 
-\hline
-                                                   & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}Lower Bound & \cellcolor[HTML]{EFEFEF}Upper Bound & \cellcolor[HTML]{EFEFEF}Significance \\
-                                                    \hline
-Edges                                              & -6.249                           & -6.412                              & -6.110                              & *                                    \\ \hline
-Instar(2)                                          & 0.035                            & 0.032                               & 0.038                               & *                                    \\ \hline
-Outstar(2)                                         & 0.021                            & 0.019                               & 0.028                               & *                                    \\ \hline
-Mutual                                             & 2.071                            & -4.157                               & 4.072                                &                                     \\ \hline
-Triangle                                           & 1.507                            & 1.431                               & 1.559                               & *                                    \\ \hline
-Martin Quinn Score                                 & 0.079                            & 0.029                               & 0.130                               & *                                    \\ \hline
-Same Issue Area Homophily                          & 1.447                            & 1.389                               & 1.515                               & *                                    \\ \hline
-Year Difference                                    & -0.068                           & -0.079                              & -0.058                              & *                                    \\ \hline
-$(\text{Year Differnce})^2$                        & 0.0019                           & 0.0016                              & 0.0022                              & *                                    \\ \hline
-Receiver Abs Diff MQ Score in Majority             & 0.047                           & 0.029                              & 0.063                              & *                                     \\ \hline
-Receiver Number Justices in Majority               & -0.075                           & -0.095                              & -0.055                              & *                                    \\ \hline
-Receiver  Sender Year                              & 0.0039                           & 0.0036                              & 0.0042                              & *                                    \\ \hline
-Overruled Cases   & -1.043                           & -1.624                              & -0.564                              & *                                    \\ \hline
-Sender Same Issue Area 2                           & 0.142                            & 0.100                               & 0.185                               & *                                    \\ \hline
-Sender Same Issue Area 3                           & -0.337                           & -0.403                              & -0.270                              & *                                    \\ \hline
-Sender Same Issue Area 4                           & 0.498                            & 0.438                               & 0.561                               & *                                    \\ \hline
-Sender Same Issue Area 5                           & 0.182                            & 0.031                               & 0.305                               & *                                    \\ \hline
-Sender Same Issue Area 6                           & 0.536                            & 0.422                               & 0.648                               & *                                    \\ \hline
-Sender Same Issue Area 7                           & 0.370                            & 0.316                               & 0.426                               & *                                    \\ \hline
-Sender Same Issue Area 8                           & 0.107                            & 0.059                               & 0.156                               & *                                    \\ \hline
-Sender Same Issue Area 9                           & 0.278                            & 0.238                               & 0.329                               & *                                    \\ \hline
-Sender Same Issue Area 10                          & 0.371                            & 0.317                               & 0.433                               & *                                    \\ \hline
-Sender Same Issue Area 11                          & -0.030                           & -0.151                              & 0.164                               &                                      \\ \hline
-Sender Same Issue Area 12                          & 0.064                           & -0.014                              & 0.130                               &                                      \\ \hline
-Sender Same Issue Area 13                          & 0.608                            & 0.423                               & 0.792                               & *                                    \\ \hline
-Sender Same Issue Area 14                          & 0.232                            & 0.075                              & 0.393                               & *                                      \\ \hline
-Receiver Same Issue Area 2                         & 0.252                            & 0.199                               & 0.310                               & *                                    \\ \hline
-Receiver Same Issue Area 3                         & -0.092                           & -0.194                              & 0.014                              &                                     \\ \hline
-Receiver Same Issue Area 4                         & 0.481                            & 0.397                               & 0.570                               & *                                    \\ \hline
-Receiver Same Issue Area 5                         & 0.374                            & 0.224                               & 0.519                               & *                                    \\ \hline
-Receiver Same Issue Area 6                         & 0.521                            & 0.346                               & 0.652                               & *                                    \\ \hline
-Receiver Same Issue Area 7                         & 0.412                            & 0.339                               & 0.487                               & *                                    \\ \hline
-Receiver Same Issue Area 8                         & 0.174                            & 0.134                               & 0.232                               & *                                    \\ \hline
-Receiver Same Issue Area 9                         & 0.311                            & 0.240                               & 0.361                               & *                                    \\ \hline
-Receiver Same Issue Area 10                        & 0.506                            & 0.429                               & 0.577                               & *                                    \\ \hline
-Receiver Same Issue Area 11                        & 0.299                            & 0.088                               & 0.432                               & *                                    \\ \hline
-Receiver Same Issue Area 12                        & 0.350                            & 0.276                               & 0.445                               & *                                    \\ \hline
-Receiver Same Issue Area 13                        & 0.972                            & 0.780                               & 1.135                               & *                                    \\ \hline
-Receiver Same Issue Area 14                        & 0.517                            & 0.200                               & 0.740                               & *                                    \\ \hline
-Instar(2) $\times$ Sender Year                     & -0.00026                         & -0.00031                            & -0.00021                            & *                                    \\ \hline
-Outstar(2) $\times$ Sender Year                    & -0.00031                         & -0.00045                            & -0.00027                            & *                                    \\ \hline
-Mutual $\times$ Sender Year                        & 0.110                           & 0.010                              & 0.524                               & *                                      \\ \hline
-Triangle $\times$ Sender Year                      & -0.00022                          & -0.0016                             & 0.0013                              &                                      \\ \hline
-Martin Quinn Score $\times$ Sender Year            & -0.0019                          & -0.0030                             & -0.0009                             & *                                    \\ \hline
-Same Issue Area $\times$ Sender Year               & -0.0068                          & -0.0081                             & -0.0055                             & *                                    \\ \hline
-Year Difference $\times$ Sender Year               & 0.00046                           & 0.00028                              & 0.00065                              & *                                    \\ \hline
-$\text{Year Difference}^2 \times$ Sender Year      & -0.000024                        & -0.000028                           & -0.000021                           & *                                    \\ \hline
-MQ Score in Majority $\times$ Sender Year & -0.00053                           & -0.00093                              & -0.0018                              & *                                    \\ \hline
-Justices in Majority $\times$ Sender Year   & 0.0017                           & 0.0013                              & 0.0021                              & *                                    \\ \hline
-Overruled Cases $\times$ Sender Year   & 0.012                           & 0.002                              & 0.023                              & *                                    \\ \hline
-\end{tabular}
-\caption{Bootstrapped MPLE Results for the time period $1937-2015$.
-A '*' indicates that the 2.5th and 97.5th quantile of the variable does not include '0' and as a result is statistically significant.}
-\label{tab:tergmcoef}
-\end{table}
 \newpage

### 2020-04-20 Tex/paper.tex
Updated Section 3 The citation ERGM

@@
 
- \section{The Citation Temporal Exponential Random Graph Model}
-{\color{blue} This section needs to be completely revised. Model is not a TERGM anymore. - CS} 
+\section{The Citation Exponential Random Graph Model}
 We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
-To accomplish this, we extend a model that has been developed to jointly model covariate and dependence effects in longitudinal network data---the temporal exponential random graph model (TERGM) \citep[e.g.,][]{cranmer2012complex,clark2013multimember, masket2015polarization, graif2017neighborhood}.
-The TERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
-The longitudinal network structures for which TERGMs are currently designed are insufficient to account for the structure of citation networks.
-
-\subsection{c-TERGM definition}
-
-We develop the citation TERGM (c-TERGM), to account for the structural constraints that apply to the network of Supreme Court citations.
-These structural constraints amount to three departures from the structure of longitudinal networks for which TERGMs are currently designed.
+To accomplish this, we extend a model that has been developed to jointly model covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{cranmer2012complex,clark2013multimember, masket2015polarization, graif2017neighborhood} {\color{blue} Citations need to be adjsuted to ERGM, not TERGM- CS}.
+%The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
+The network structures for which ERGMs are currently designed are insufficient to account for the structure of citation networks.
+
+\subsection{c-ERGM definition}
+
+We develop the citation ERGM (c-ERGM), to account for the structural constraints that apply to the network of Supreme Court citations.
+These structural constraints amount to three departures from the structure of networks for which ERGMs are currently designed.
 First, the citation network is partially acyclic.
@@
 Third, the set of nodes in the network must increase for new edges to be created.
-In conventional TERGMs, the number of nodes in the network can increase or decrease in each time period, and is typically stable over time.
+In conventional ERGMs, the number of nodes in the network can increase or decrease in each time period, and is typically stable over time.
 In a citation network, new edges (and non-edges) are introduced over time via the introduction of new nodes.
 The structure of the citation network is depicted in Figure \ref{fig:ctergm}---a hypothetical citation network established over three time periods, with three cases decided in each time period.
-We denote $C_t$ to be the set of citation and non-citations added to the network at time $t$ (i.e., via the addition of three cases), $C_{ <t}$ to be the citations among cases decided before time $t$, and $C_{ \leq t}$ to denote the entire set of citations and non-citations on which $C_t$ can depend through the c-TERGM specification.
+We denote $C_t$ to be the set of citation and non-citations added to the network at time $t$ (i.e., via the addition of three cases), $C_{ <t}$ to be the citations among cases decided before time $t$, and $C_{ \leq t}$ to denote the entire set of citations and non-citations on which $C_t$ can depend through the c-ERGM specification.
 
@@
 
-The likelihood function of the c-TERGM is given by
+The likelihood function of the c-ERGM is given by
 \begin{equation}
-l(\bm{\theta},C_{\leq T}) =  \prod_{t=1}^T \frac{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right] }{ \sum_{C_t^* \in \mathcal{C}_t} \exp \left[ {\bm{\theta}'\bm{h}(C^*_{t},C_{<t}) }\right]  },
+l(\bm{\theta},C_{\leq t}) =  \frac{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right] }{ \sum_{C_t^* \in \mathcal{C}_t} \exp \left[ {\bm{\theta}'\bm{h}(C^*_{t},C_{<t}) }\right]  },
 \end{equation}
@@
 The denominator of Equation 1 represents a normalizing constant, in which the positive weight is summed over all possible configurations of $C_{t}$, from the network in which the cases added to the network at time $t$ send no citations at all, to the network in which the cases added at time $t$ cite every possible case, and everything in between.
-The probability of observing the citation network up to time $T$---the final time point in the dataset---is given by the product over the probability of adding the citations at each time point $t$, conditional on the citations that were added to the network previously.
-
-Though the likelihood function of the c-TERGM may appear quite different from that of statistical models conventionally used in political science, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients.
+%The probability of observing the citation network up to time $T$---the final time point in the dataset---is given by the product over the probability of adding the citations at each time point $t$, conditional on the citations that were added to the network previously.
+
+Though the likelihood function of the c-ERGM may appear quite different from that of statistical models conventionally used in political science, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients.
 $$ P(C_{ij,t} = 1 | C_{-ij,t}, C_{ < t}) = \frac{\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right]}{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right] + \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 0) } \right]}, $$ $$ \text{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~} = \frac{1}{ 1 + \exp \left[ - {\bm{\theta}'\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)} \right]}, $$
@@
 For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log odds of observing $C_{ij,t} = 1$ increase by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$).
-The logit form conditional probability is not unique to c-TERGM.
-It is well known for the ERGM family \citep{goodreau2009birds}.
-
-We should note that there is nothing about the c-TERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another.
+The logit form conditional probability is well known for the ERGM family \citep{goodreau2009birds}.
+
+We should note that there is nothing about the c-ERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another.
 For example, if we find a positive popularity effect, the term for that dynamic would indicate that there is an increase in the log odds that case $i$ cites case $j$ with each other case that has cited case $j$.
@@
 We cannot statistically differentiate between these two dynamics.
-However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-TERGM can be specified to account for the dependence in the data.
-
-\subsection{c-TERGM Estimation}
+However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-ERGM can be specified to account for the dependence in the data.
+
+\subsection{c-ERGM Estimation}
 
@@
 The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases.
-This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the c-TERGM.
-Methods of Monte Carlo approximation are well established in the ERGM literature, \citep{hunter2006inference,van2009framework,hummel2012improving}, but with a number of nodes on the order of $10,000$, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}.
+This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the c-ERGM.
+
+The common alternative relates on Monte Carlo methods to approximate the normalizing constant by simulating a large set of networks \citep{hunter2006inference,hummel2012improving}.
+The resulting estimator, the Monte Carlo MLE (MCMLE), is approximately consistent, meaning that it converges to the MLE as the sample size, i.e., the number of simulated networks, increases.
+However, one drawback is that with the number of nodes in the Supreme Court citation network being in the order of $10,000$, obtaining the MCMLE is computationally expensive \citep{schmid2017exponential} and the success of the algorithm heavily relies on the starting parameter vector $\theta_0$, which ideally is in the proximity of the MLE \citep{hummel2012improving}.
+The prevailing choice for $\theta_0$ is the maximum pseudo-likelihood estimation (MPLE) \citep{strauss1990pseudolikelihood}, a fast estimation method that is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
+The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated above, assume a logit form.
+The MPLE is simple to obtain, but does not guarantee a starting value close to the MLE \citep{SchmidHunter2020}.
+
+We estimate the MLE for every term of the c-ERGM between 1950 and 2015.
+While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows.
+The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant.
+However, the estimation of most networks in the 90s with the MPLE as starting valus was not obtainable in an reasonable time frame anymore.
+Instead, we fixed the MCMLE of the previous term $t-1$ as starting value to succesfully obtain the MCMLE at time $t$.
+
+But even this approach started to fail for networks at the turn of the millennium.
+Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}.
+The MCMLE for these large citation networks was finally obtained by setting the starting value according to an approach introduced by \citet{SchmidHunter2020}.
+This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$.
+Instead of starting the MCMLE algorithm at the MPLE, \citet{SchmidHunter2020} propose searching for a network $C_t^*$ using simulated annealing \citep{Kirkpatrick83}, with $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$ and a weak dependence structure among unfixed ties.
+The MPLE of such a network is similar to the MLE, making it an effective starting value for the MCMLE algorithm.
+
+%Methods of Monte Carlo approximation are well established in the ERGM literature, \citep{hunter2006inference,van2009framework,hummel2012improving}, but with a number of nodes on the order of $10,000$, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}.
 However, because the network is large, we can rely on a fast estimation method---maximum pseudo-likelihood estimation (MPLE), which is consistent for discrete exponential family models \citep{nguyen2017near}, and has long been used with ERGM \citep{strauss1990pseudolikelihood}.
@@
 
-One challenge that arises with the use of MPLE is that the usual estimate of the asymptotic covariance matrix (i.e., inverting the negated Hessian of the objective function at the maximum), results in standard errors that are biased downward.
+%One challenge that arises with the use of MPLE is that the usual estimate of the asymptotic covariance matrix (i.e., inverting the negated Hessian of the objective function at the maximum), results in standard errors that are biased downward.
 This leads to a high Type-1 error rate in hypothesis testing with MPLE if conventional standard errors are used \citep{van2009framework}.

### 2020-04-21 Tex/paper.tex
Started writing the appendix

@@
 The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.\footnote{
-Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. This work was supported in part by NIH, 
+%Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. 
+This work was supported in part by NIH, 
 R01 AI36664-01. NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219. We would like to thank Rachael Hinkle and Benjamin Kassow for detailed comments on an earlier draft of this paper.}
@@
 We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
-To accomplish this, we extend a model that has been developed to jointly model covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{cranmer2012complex,clark2013multimember, masket2015polarization, graif2017neighborhood} {\color{blue} Citations need to be adjsuted to ERGM, not TERGM- CS}.
+To accomplish this, we extend a model that has been developed to jointly model covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{cranmer2012complex,clark2013multimember, masket2015polarization, graif2017neighborhood} {\color{blue} Citations need to be adjusted to ERGM, not TERGM- CS}.
 %The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
@@
 We cannot statistically differentiate between these two dynamics.
-However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-ERGM can be specified to account for the dependence in the data.
+However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables and the c-ERGM can be specified to account for the dependence in the data.
 
@@
 The resulting estimator, the Monte Carlo MLE (MCMLE), is approximately consistent, meaning that it converges to the MLE as the sample size, i.e., the number of simulated networks, increases.
-However, one drawback is that with the number of nodes in the Supreme Court citation network being in the order of $10,000$, obtaining the MCMLE is computationally expensive \citep{schmid2017exponential} and the success of the algorithm heavily relies on the starting parameter vector $\theta_0$, which ideally is in the proximity of the MLE \citep{hummel2012improving}.
+However, one drawback is that with the number of nodes in the Supreme Court citation network being in the order of $10,000$, obtaining the MCMLE is computationally expensive \citep{schmid2017exponential} and the success of the algorithm heavily relies on the starting parameter vector $\theta_0$, which is ideally chosen in the proximity of the unknown MLE \citep{hummel2012improving}.
 The prevailing choice for $\theta_0$ is the maximum pseudo-likelihood estimation (MPLE) \citep{strauss1990pseudolikelihood}, a fast estimation method that is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
@@
 The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant.
-However, the estimation of most networks in the 90s with the MPLE as starting valus was not obtainable in an reasonable time frame anymore.
-Instead, we fixed the MCMLE of the previous term $t-1$ as starting value to succesfully obtain the MCMLE at time $t$.
-
-But even this approach started to fail for networks at the turn of the millennium.
+However, the estimation of most networks in the 90s with the MPLE as starting values was not feasible in an reasonable time frame anymore.
+Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$.
+
+But even this approach started to fail for networks around the turn of the millennium.
 Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}.
@@
 This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$.
-Instead of starting the MCMLE algorithm at the MPLE, \citet{SchmidHunter2020} propose searching for a network $C_t^*$ using simulated annealing \citep{Kirkpatrick83}, with $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$ and a weak dependence structure among unfixed ties.
+However, the MPLE of networks with the same sufficient statistics is not necessarily the same.
+Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a network $C_t^*$ using simulated annealing \citep{Kirkpatrick83}, with $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$ and a weak dependence structure among unfixed ties.
 The MPLE of such a network is similar to the MLE, making it an effective starting value for the MCMLE algorithm.
@@
 
-The common dependence effects for these four dynamics are the number of triangles (case $i$ cites cases $j$ and $k$, and case $j$ cites case $k$), the numnber of mutual dyads (case $i$ cites $j$ and vice versa), the numnber of out-2-stars (case $i$ cites $j$ and $k$), and the number of in-2-stars (case $i$ is being cited by cases $j$ and $k$) in $C_{\leq t}$.
+The common dependence effects for these four dynamics are the number of triangles (case $i$ cites cases $j$ and $k$, and case $j$ cites case $k$), the number of mutual dyads (case $i$ cites $j$ and vice versa), the number of out-2-stars (case $i$ cites $j$ and $k$), and the number of in-2-stars (case $i$ is being cited by cases $j$ and $k$) in $C_{\leq t}$.
 Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy \citep{Handcock.2003}.
 Model degeneracy is a common obstacle when fitting ERGMs and occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network.
-In this scenario, the simulated networks are too different from the observed network, making the underlying distribution defined by the model extremly unrealistc.
+In this scenario, the simulated networks are too different from the observed network, making the underlying distribution defined by the model extremely unrealistic.
 
@@
 % popularity
-Accouting for the popularity effect through the in-2-star statistic is prone to cause model degeneracy.
-In an attempt to stabilize the model against model degeneracy as well as to still capture the popularity effect, \citet{SnijdersTomA.B..2006} introduces the \textit{alternating-in-k-star} statistic, which was shown to be equivalent to the \textit{geometrically weighted in-degree distribution} (gwidegree) statistic introduced by \citet{hunter2006inference}.
+Accounting for the popularity effect through the in-2-star statistic is prone to cause model degeneracy.
+In an attempt to stabilize the model against model degeneracy as well as to still capture the popularity effect, \citet{SnijdersTomA.B..2006} introduces the \textit{alternating-in-k-star} statistic, which was shown to be equivalent to the \textit{geometrically weighted indegree distribution} (gwidegree) statistic introduced by \citet{hunter2006inference}.
 Define $D_r(C_t, C_{<t})$ as the number of nodes with indegree $r, r \in \{0, \dots , N_t -1 \}$, where $N_t$ is the number of cases in the network at time $t$, then gwidegree represents a network's indegree distribution in a single statistic by geometrically weighting the degree distribution
@@
 For directed networks the edgewise shared $r$-partners statistic ($ESP_r$) can take multiple forms, but we integrate the definition where cases $i$ and $j$ that are connected by a citation, also cite $r$ common other cases.
+We follow the notation of \citet{Butts.2008} and call this ESP specification \textit{outgoing shared partners} (OSP).
 Similar to the gwidegree-statistic the gwesp-statistic incorporates a network's edgewise shared partners distribution by replacing $D_r$ by $ESP_r$.
@@
 
- \textit{GOF-plot?}\\
-
-%In order to account for the popularity effect, we add the \textit{geometrically weighted in-degree distribution} (gwidegree) statistic to the model.
+%In order to account for the popularity effect, we add the \textit{geometrically weighted indegree distribution} (gwidegree) statistic to the model.
 The geometrically weighted distribution statistics were introduced by \citet{Hunter.2006} to counteract model degeneracy.
@@
 
-  \begin{figure}[]
+  \begin{figure}[t]
 \includegraphics[width=15cm]{SCC_AIC_BIC}
@@
 Interpreting the popularity effect---the log odds that $i$ cites $j$ increases by 0.018--0.035 with every additional citation to case $j$.
-Based on the in-degree distribution, it appears that the popularity of cases can be classified by multiples of 10.
+Based on the indegree distribution, it appears that the popularity of cases can be classified by multiples of 10.
 If case $j$ increases from ten to twenty citations from other cases, the probability that case $i$ cites case $j$ increases from approximately 0.0030 to 0.0040.
@@
 
+%\newpage
+%   \thispagestyle{empty}
+
   \begin{figure}[H]
@@
  \label{SCC_results_1}
+
 \vspace{-.25cm}
@@
 
+% \thispagestyle{empty}
  \begin{figure}[H]
@@
 We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network.
-This methodology---the citation-TERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation.
+This methodology---the citation-ERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation.
 We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2015.
@@
 Both data and methodological limitations will need to be overcome to consider these two citation types.
-Second, there is nothing in the c-TERGM that governs the generation of new cases (i.e., cases are assumed to arise exogenously, unrelated to the history of the citation network).
+Second, there is nothing in the c-ERGM that governs the generation of new cases (i.e., cases are assumed to arise exogenously, unrelated to the history of the citation network).
 As new cases are often taken up by the Supreme Court in order to test, clarify, and/or challenge existing precedents, it is likely inappropriate to assume that new cases arise exogenous to the existing citation network.
-Overcoming this limitation would require methodological innovation in the c-TERGM as well as theoretical development regarding exactly how citation structure might predict case emergence.
+Overcoming this limitation would require methodological innovation in the c-ERGM as well as theoretical development regarding exactly how citation structure might predict case emergence.
 %Third, the data used in this study does not cover the most recent history of the Court.
@@
 \section{Appendix}
-The following table presents the Bootstrapped MPLE Results for the time period $1937-2015$.
-
+
+\subsection{Goodness-of-Fit}
+We evaluate the goodness-of-fit of our model following \citet{Hunter.2008} by examining the distribution of four hyper statistics, e.g., the out- and indegree distribution and the distribution for two different edgewise-shared partners statistics.
+OTP stands for \textit{outgoing two-paths} and refers to the number of cases $r$ that are cited by case $i$ and that cite case $j$, while $j$ is also directly cited by $i$.
+The second ESP statistic is the OSP specification that has been introduced in section \ref{dependence_terms}.
+Figure \ref{GOF} visualizes the goodness-of-fit results for the citation network for the 1950 (top) and 2015 (bottom) term.
+The solid black line indicates the statistic's distribution in the Supreme Court citation network for that given term and the boxplots depict the statistic's distribution of $1000$ networks that have been simulated from the ERGM defined by the MCMLE.
+This means that in the ideal case the solid black line passes through ever single boxplot.
+
+Talk about results. Wait for 2015 plots to finish
+
+ \begin{figure}[H]
+ \begin{center}
+\includegraphics[width=13.8cm]{GOF_1950}\\
+%\caption{Goodness-of-fit diagnostic for the 1950 network.}
+% \label{GOF_1950}
+\vspace{1cm}
+\includegraphics[width=13.8cm]{GOF_1950}
+\vspace{0.1cm}
+\caption{Goodness-of-fit diagnostic for the 1950 network (top) and the 2015 network (bottom).}
+ \label{GOF}
+\vspace{-.25cm}
+\end{center}
+\end{figure}
+
+\subsection{Checking for Model Degeneracy}
+
+A common challenge when fitting ERGMs is model degeneracy.
+Model degeneracy occurs when the probability distribution defined by the parameter vector does not predominantly yield networks with similar statistics as the observed network.
+Generally, model degeneracy results in simulated networks with no ties or all possible ties.
+In a non-degenerate model the statistics of the simulated networks fall in the proximity of the observed network's statistics.
+Figure \ref{mcmcdiagnostics_1950} depicts the degeneracy check for the dependence terms in the 1950 term citation network.
+The histograms on the left visualize a statistic's density from $1000$ simulated networks, while the right side shows the statistic's trace plot of the same $1000$ networks.
+The solid black line indicates the statistic's value in the actual citation network.
+Figure \ref{mcmcdiagnostics_1950} shows that this model is non-degenerate and that the simulated network's statistics fall almost evenly around the observed statistic.
+The density and trace plots for the ERGM of other terms do not indicate a degenerate model either.
+
+ \begin{figure}[H]
+ \begin{center}
+\includegraphics[width=14.5cm]{Deg_1950}
+\caption{Density and trace plots for the dependency terms of the 1950 term citation network.}
+ \label{mcmcdiagnostics_1950}
+\vspace{-.25cm}
+\end{center}
+\end{figure} 
 %\begin{itemize}

### 2020-04-23 Tex/paper.tex
Finished writing the Appendix

@@
 
+The panel showing the results of the reciprocity statistic does not provide estimates for ever term.
+In these missing terms there were no two cases that would cite each other, making the estimation of the reciprocity statistic impossible.
+A similar scenario is visible for the \textit{Overruled Cases} variable.
+In the time period from 1953-1956 no cases that have been overruled have been cited which results in a non-estimable effect.
+
 %The reciprocity effect is substantial, positive, and statistically significant for most of the period under study.
@@
 \subsection{Goodness-of-Fit}
-We evaluate the goodness-of-fit of our model following \citet{Hunter.2008} by examining the distribution of four hyper statistics, e.g., the out- and indegree distribution and the distribution for two different edgewise-shared partners statistics.
+We evaluate the goodness-of-fit of the model following \citet{Hunter.2008} by examining the distribution of four hyper statistics, e.g., the out- and indegree distribution and the distribution for two different edgewise shared partners statistics.
 OTP stands for \textit{outgoing two-paths} and refers to the number of cases $r$ that are cited by case $i$ and that cite case $j$, while $j$ is also directly cited by $i$.
@@
 Figure \ref{GOF} visualizes the goodness-of-fit results for the citation network for the 1950 (top) and 2015 (bottom) term.
-The solid black line indicates the statistic's distribution in the Supreme Court citation network for that given term and the boxplots depict the statistic's distribution of $1000$ networks that have been simulated from the ERGM defined by the MCMLE.
+The solid black line indicates the statistic's distribution in the Supreme Court citation network of that given term and the boxplots depict the statistic's distribution of $1000$ networks that have been simulated from the ERGM defined by the MCMLE.
 This means that in the ideal case the solid black line passes through ever single boxplot.
 
-Talk about results. Wait for 2015 plots to finish
+We see that our models do a good job capturing the out and indegree distribution of the citation network, since the black line falls almost exclusively within the ranges spanned by the boxplots.
+For the ESP distributions we can observe that the number of ties with $r=0$ shared partners is captured well for both the OTP as well as for the OSP statistic.
+However, the model overestimates the number of $r=1$ shared partners and then, especially in the 2015 term network, underestimates the number of ties with more than $r=1$ shared partners.
 
@@
 \vspace{1cm}
-\includegraphics[width=13.8cm]{GOF_1950}
+\includegraphics[width=13.8cm]{GOF_2015}
 \vspace{0.1cm}
@@
 Generally, model degeneracy results in simulated networks with no ties or all possible ties.
-In a non-degenerate model the statistics of the simulated networks fall in the proximity of the observed network's statistics.
-Figure \ref{mcmcdiagnostics_1950} depicts the degeneracy check for the dependence terms in the 1950 term citation network.
+In a non-degenerate model the statistics of the networks that were simulated from the probability distribution defined by the MCMLE fall in the proximity of the observed network's statistics.
+Figures \ref{mcmcdiagnostics_1950} and \ref{mcmcdiagnostics_1950} depict trace and density plots for the dependence terms in the 1950 and 2015 term citation network.
 The histograms on the left visualize a statistic's density from $1000$ simulated networks, while the right side shows the statistic's trace plot of the same $1000$ networks.
 The solid black line indicates the statistic's value in the actual citation network.
-Figure \ref{mcmcdiagnostics_1950} shows that this model is non-degenerate and that the simulated network's statistics fall almost evenly around the observed statistic.
-The density and trace plots for the ERGM of other terms do not indicate a degenerate model either.
+Both figures indicate that this model is non-degenerate and that the simulated network's statistics fall almost evenly around the observed statistic.
+The density and trace plots for the ERGM of the terms not depicted provide similar results.
 
@@
  \label{mcmcdiagnostics_1950}
+\vspace{-.25cm}
+\end{center}
+\end{figure}
+
+ \begin{figure}[H]
+ \begin{center}
+\includegraphics[width=14.5cm]{Deg_2015}
+\caption{Density and trace plots for the dependency terms of the 2015 term citation network.}
+ \label{mcmcdiagnostics_2015}
 \vspace{-.25cm}

### 2020-04-24 R-Code/ERCM.R
Updated R-Code

@@ -1,287 +1,0 @@
-library(lubridate)
-
-scc1<-scc
-# first sort data frame by decision date
-scc1<-scc1[order(as.Date(scc1$dateDecision, format="%m/%d/%Y")),]
-
-# time is day-discrete -> turn date into number of days that have passed since the first case
-
-x <- scc1$dateDecision
-#transform date into yyyy-mm-dd
-date <- mdy(x)
-
-days <-as.numeric(date)+64970
-scc1$time_t<- days
-
-# assign same date the same number, save in column 'id'
-scc1 <- transform(scc1,id=as.numeric(factor(time_t)))
-
-
-# create a large matrix where every column refers to a time t and every row refers to a scc case
-citation.matrix<- matrix(0, 26803, 5251) # matrix for only the first case, we add the columns one by one
-colnames(citation.matrix)<-1:5251
-
-
-
-for(i in 1:209946){
-  if(i %% 1000 == 0) cat("Starting iteration", i, "\n")
-  # which line can the information of case edgelist[i,1] be found in?
-  l<- which(edgelist[i,1]==scc1[,1])
-  # which line can the information of case edgelist[i,2] be found in? used to count how often a case is being cited
-  q<- which(edgelist[i,2]==scc1[,1])
-  # which time_t was citation made?
-  t<- scc1[l, 71]
-  citation.matrix[q,t]<-1
-}
-
-
-
-#### Statistics
-
-## down-weighted triangle statistic 1787
-
-triangles<- function(c_t, time_t){
-  # get right colum of time t
-  #col.is.a.match <- apply(citation.matrix, 2, identical, c_t)
-  #time_t<- which(col.is.a.match)
-  
-  # which rows have an entry 1
-  citation.row<- which(c_t == 1)
-  length.citation.row <- length(citation.row)
-  
-  #set tri.count =0
-  tri.count<- 0
-  # count number of triangles
-  for(i in 1:(length.citation.row-1)){
-    for(j in (i+1):length.citation.row){
-      # what is time_t of later citation?
-      late_cit<- scc1[citation.row[j], 71]
-      #### year difference between cases
-      # which row is case in
-      year.row<- which(scc1[,71]==time_t)
-      # save dates
-      year.c_t <- scc1[year.row[1], 4]
[…]

### 2020-04-24 R-Code/ERCM3.R

@@ -1,648 +1,0 @@
-# requires scc_data_merging
-library(combinat)
-
-
-#scc1<- scc
-# scc1 only considers cases from term 1937 on
-
-# create empty adjacency matrix (full AM for all cases from the beginning)
-memory.limit(30000)
-adjacency.matrix<- matrix(0, 26803, 26803)
-
-# fill adjacency matrix
-for(i in 1:209946){
-  #print(i)
-  if(i %% 1000 == 0) cat("Starting iteration", i, "\n")
-  # which line can the information of case edgelist[i,1] be found in?
-  l<- which(edgelist[i,1]==scc1[,1])
-  # which line can the information of case edgelist[i,2] be found in? used to count how often a case is being cited
-  q<- which(edgelist[i,2]==scc1[,1])
-  # indicate tie in adjacency matrix
-  adjacency.matrix[l,q]<- 1
-}
-
-# only keep the part of the adjacency matrix for which we have Martin Quinn scores
-# 17987 is the first row in scc and the first row id of a case for which we were able to calculate the mean MQ score
-adjacency.matrix<- adjacency.matrix[17987:26803, 17987:26803]
-
-
-# write function that cuts out the necessary adjacency matrix for a time t
-
-#AM.t<- function(time.t){ # first id with MQ scores is 3136, last 5251
-  # what is the id of the case/cases that entered the network at time t
-#  latest.case<- max(which(time.t==scc1[, 71]))
-#  AM<- adjacency.matrix[1:(latest.case-17986), 1:(latest.case-17986)]
-#  return(AM)
-#}
-
-
-#### statistics
-time.t<- 3500
-latest.case<- max(which(time.t==scc1[, 71]))
-year.latest.case <- scc1[latest.case, 4]
-
-decayFun <- function(age,beta){2/(1+exp(beta*age))}
-par(las=1)
-age = 0:100
-plot(age,decayFun(age,1),type="l",lwd=2,ylab="weight")
-lines(age,decayFun(age,0.25),col="red",lwd=2)
-lines(age,decayFun(age,0.1),col="green",lwd=2)
-lines(age,decayFun(age,0.05),col="blue",lwd=2)
-lines(age,decayFun(age,0.025),col="grey60",lwd=2)
-legend("topright",legend=c("1","0.25","0.10","0.05","0.025"),lwd=2,col=c("black","red","green","blue","grey60"),title="Beta")
-
-#######################################
[…]

### 2020-04-24 R-Code/ERCM_MPLE.R

@@ -1,472 +1,0 @@
-# ERCM bootstrapped MPLE
-
-
-# requires scc1 from scc_data_merging
-
-
-########################################################
-# network statistics
-########################################################
-
-
-# requires functions count.outstars and count.triangles
-count.outstars.sim<- function(A, age, beta, i){
-  # decay vector
-  decay<-  decayFun(age, 0.025)
-  # set number of triangles = 0
-  outstars <-0
-  
-  # which cases are cited by case i
-  ones<- which(A[i,]==1)
-  # what are all the twostars that case i has formed
-  if(length(ones)>= 2){
-    twostars<- t(combn(ones,2))
-    l<- dim(twostars)[1]
-    # loop through twostars to count weighted outstars
-    for(k in 1:l){
-      #add outstars
-      outstars<- outstars +sqrt(decay[twostars[k,1]]*decay[twostars[k,2]])
-    } # end for k
-    
-  } # end if
-  return(outstars)
-} # end function
-
-# triangles
-count.triangles.sim<- function(A, age, beta, i){
-  # decay vector
-  decay<- decayFun(age, 0.025)
-  # set number of triangles = 0
-  triangles <-0
-  
-  # which cases are cited by case i
-  ones<- which(A[i,]==1)
-  # what are all the twostars that case i has formed
-  if(length(ones)>= 2){
-    twostars<- t(combn(ones,2))
-    l<- dim(twostars)[1]
-    # loop through twostars to count triangles
-    for(k in 1:l){
-      #add triangles
-      triangles<- triangles +A[twostars[k,2],twostars[k,1]]*sqrt(decay[twostars[k,1]]*decay[twostars[k,2]])
-    } # end for k
-    
-  } # end if
-  
-  return(triangles)
-} # end function
-
-
-#######################################
-
-# for the optimization we require network statistics that do not only calculate the change statistic, but the entire statistic
-
-count.citation.triangles.sim <- function(A,age,beta){
-  # decay vector
-  decay<- decayFun(age, 0.025)
-  # set number of triangles = 0
-  triangles <-0
-  
-  d<- dim(A)[1]
-  for(i in 2:d){
-    # which cases are cited by case i
-    ones<- which(A[i,]==1)
[…]

### 2020-04-24 R-Code/ERCM_simulation_study.R

@@ -1,532 +1,0 @@
-# simulation study for supreme court citation project
-
-# requires functions count.outstars and count.triangles
-count.outstars.sim<- function(A, age, beta, i){
-  # decay vector
-  decay<-  decayFun(age, 0.025)
-  # set number of triangles = 0
-  outstars <-0
-  
-    # which cases are cited by case i
-    ones<- which(A[i,]==1)
-    # what are all the twostars that case i has formed
-    if(length(ones)>= 2){
-      twostars<- t(combn(ones,2))
-      l<- dim(twostars)[1]
-      # loop through twostars to count weighted outstars
-      for(k in 1:l){
-        #add outstars
-        outstars<- outstars +sqrt(decay[twostars[k,1]]*decay[twostars[k,2]])
-      } # end for k
-      
-    } # end if
-  return(outstars)
-} # end function
-
-# triangles
-count.triangles.sim<- function(A, age, beta, i){
-  # decay vector
-  decay<- decayFun(age, 0.025)
-  # set number of triangles = 0
-  triangles <-0
-
-    # which cases are cited by case i
-    ones<- which(A[i,]==1)
-    # what are all the twostars that case i has formed
-    if(length(ones)>= 2){
-      twostars<- t(combn(ones,2))
-      l<- dim(twostars)[1]
-      # loop through twostars to count triangles
-      for(k in 1:l){
-        #add triangles
-        triangles<- triangles +A[twostars[k,2],twostars[k,1]]*sqrt(decay[twostars[k,1]]*decay[twostars[k,2]])
-      } # end for k
-      
-    } # end if
-  
-  return(triangles)
-} # end function
-
-
-#######################################
-
-# for the optimization we require network statistics that do not only calculate the change statistic, but the entire statistic
-
-count.citation.triangles.sim <- function(A,age,beta){
-  # decay vector
-  decay<- decayFun(age, 0.025)
-  # set number of triangles = 0
-  triangles <-0
-  
-  d<- dim(A)[1]
-  for(i in 2:d){
-    # which cases are cited by case i
-    ones<- which(A[i,]==1)
-    # what are all the twostars that case i has formed
-    if(length(ones)>= 2){
[…]

### 2020-04-24 R-Code/Labeling_bootresults.R

@@ -1,41 +1,0 @@
-
-library(readr)
-
-mpledata <- read_csv("bootresults.csv")
-mpledata<- mpledata[,-1]
-d<- c("edges","istar2","ostar2","mutual","triangle", "edgecov.mq.t", "edgecov.same.issue.area.t", "edgecov.year.diff.t" ,"edgecov.year.diff.square.t" ,  
-                   "nodeicov.AbsDiffMQscores", "nodeicov.NumberJusticesPro", "nodeocov.sender.time" , "nodeifactor.SameIssueArea.2", "nodeifactor.SameIssueArea.3",
-                      "nodeifactor.SameIssueArea.4","nodeifactor.SameIssueArea.5", "nodeifactor.SameIssueArea.6", "nodeifactor.SameIssueArea.7", 
-                      "nodeifactor.SameIssueArea.8","nodeifactor.SameIssueArea.9",  "nodeifactor.SameIssueArea.10",
-                      "nodeifactor.SameIssueArea.11", "nodeifactor.SameIssueArea.12", "nodeifactor.SameIssueArea.13","nodeifactor.SameIssueArea.14", 
-                      "nodeofactor.SameIssueArea.2", "nodeofactor.SameIssueArea.3", "nodeofactor.SameIssueArea.4","nodeofactor.SameIssueArea.5", 
-                      "nodeofactor.SameIssueArea.6","nodeofactor.SameIssueArea.7",
-                      "nodeofactor.SameIssueArea.8","nodeofactor.SameIssueArea.9","nodeofactor.SameIssueArea.10", "nodeofactor.SameIssueArea.11", 
-                      "nodeofactor.SameIssueArea.12", "nodeofactor.SameIssueArea.13","nodeofactor.SameIssueArea.14",
-                      "instar2_sendertime","outstar2_sendertime", "mutual_sendertime", "triangle_sendertime","mq_sendertime" ,"sameissuearea_sendertime",
-                      "yeardiff_sendertime" , "yeardiffsquare_sendertime", "AbsDiffMQscores_sendertime" , "NumberJusticesPro_sendertime")
-colnames(mpledata)<- d
-
-
-write.csv(mpledata, file = "bootresults.csv")
-
-########################################
-##### results for model without endogenous statistics
-
[…]

### 2020-04-24 R-Code/bootstrap_coefficient_summaries.R

@@ -1,278 +1,0 @@
-# code assumes that working directory is set
-# to top-level project
-bootstrapResults <- read.csv("bootresults.csv",stringsAsFactors=F,row.names=1)
-
-# Christian, can you correct this? The sender time values should be set to map onto the range
-# of sender times in our data
-sender_times <- 1:65
-sender_years <- 1937:2001
-
-density_coefficients <- c("edges","nodeocov.sender.time")
-i2star_coefficients <- c("istar2","instar2_sendertime")
-o2star_coefficients <- c("ostar2","outstar2_sendertime")
-mutual_coefficients <- c("mutual","mutual_sendertime")
-triangle_coefficients <- c("triangle","triangle_sendertime")
-mq_coefficients <- c("edgecov.mq.t","mq_sendertime")
-sameissue_coefficients <- c("edgecov.same.issue.area.t","sameissuearea_sendertime")
-absdiffmq_coefficients <- c("nodeicov.AbsDiffMQscores","AbsDiffMQscores_sendertime")
-numberjusticespro_coefficients <- c("nodeicov.NumberJusticesPro","NumberJusticesPro_sendertime")
-
-# example trend CI plot for MQ coef
-mq_coef_trend <- matrix(NA,length(sender_times),3)
-colnames(mq_coef_trend) <- c("mean","lower","upper")
-for(t in sender_times){
-  effect.t <- bootstrapResults[,mq_coefficients[1]]+t*bootstrapResults[,mq_coefficients[2]]
-  mq_coef_trend[t,] <- c(mean(effect.t),quantile(effect.t,prob=c(0.025,.975)))
-}
-
-pdf("./images/mq_coef_trend.pdf",width=6,height=3.5)
-par(las=1,mar=c(4,4,1,1))
-plot(sender_years,mq_coef_trend[,1],type="n",ylab="Effect",xlab="Year", main="Martin-Quinn Score" ,ylim=c(min(c(mq_coef_trend)),max(c(mq_coef_trend))))
-polygon(c(sender_years,rev(sender_years)),c(mq_coef_trend[,2],rev(mq_coef_trend[,3])),col="lightblue",border=NA)
-lines(sender_years,mq_coef_trend[,1],lwd=2,col="grey50")
-abline(h=0,lty=2,lwd=2)
-dev.off()
-
-
-# trend CI plot for edges coef
[…]

### 2020-04-24 R-Code/data_merging_and_bootMPLE.R

@@ -1,418 +1,0 @@
-# merge together csv data that was obtained from the aci.ics cluster
- 
-# rm(list= ls()[!(ls() %in% c('scc1','adjacency.matrix', 'mq.matrix', 'same.issue.area', 'year.diff.matrix'))])
-memory.limit(30000)
-
-library(readr)
-
-mpledata1 <- read_csv("mpledata.csv")
-mpledata2 <- read_csv("mpledata2.csv")
-mpledata3 <- read_csv("mpledata3.csv")
-mpledata4 <- read_csv("mpledata4.csv")
-mpledata5 <- read_csv("mpledata5.csv")
-
-mpledata1<- as.data.frame(mpledata1)
-mpledata2<- as.data.frame(mpledata2)
-mpledata3<- as.data.frame(mpledata3)
-mpledata4<- as.data.frame(mpledata4)
-mpledata5<- as.data.frame(mpledata5)
-
-mpledata<- rbind(mpledata1, mpledata2, mpledata3, mpledata4, mpledata5)
-mpledata<- mpledata[,-1]
-
-rm(mpledata1, mpledata2, mpledata3, mpledata4, mpledata5)
-
-
-# next, I create new columns to add interaction effects
-
-# first degree
-mpledata$instar2_sendertime<- mpledata[,3]*mpledata[,14]
-mpledata$mutual_sendertime<- mpledata[,4]*mpledata[,14]
-mpledata$outstar2_sendertime<- mpledata[,5]*mpledata[,14]
-mpledata$triangle_sendertime<- mpledata[,6]*mpledata[,14]
-mpledata$mq_sendertime<- mpledata[,7]*mpledata[,14]
-mpledata$sameissuearea_sendertime<- mpledata[,8]*mpledata[,14]
-mpledata$yeardiff_sendertime<- mpledata[,9]*mpledata[,14]
-mpledata$yeardiffsquare_sendertime<- mpledata[,10]*mpledata[,14]
-mpledata$AbsDiffMQscores_sendertime<- mpledata[,16]*mpledata[,14]
-mpledata$NumberJusticesPro_sendertime<- mpledata[,17]*mpledata[,14]
-
-# 2nd degree
-mpledata$instar2_sendertime2<- mpledata[,3]*mpledata[,14]*mpledata[,14]
-mpledata$mutual_sendertime2<- mpledata[,4]*mpledata[,14]*mpledata[,14]
-mpledata$outstar2_sendertime2<- mpledata[,5]*mpledata[,14]*mpledata[,14]
[…]

### 2020-04-24 R-Code/holdOutExperiment.R

@@ -1,106 +1,0 @@
-scdata <- read.csv("mpledata_all.csv",stringsAsFactors=F)
-
-set.seed(123)
-
-niter = 10
-predict.results <- matrix(0,niter,6)
-
-for(i in 1:niter){
-
-train.obs <- sample(1:nrow(scdata),round(0.8*nrow(scdata)))
-
-train.data <- scdata[train.obs,]
-
-library(biglm)
-
-# MPLE
-mple.est <- bigglm(edgeij ~ istar2+ostar2+mutual+triangle+ edgecov.mq.t+ edgecov.same.issue.area.t+ edgecov.year.diff.t +edgecov.year.diff.square.t +  
-                nodeicov.AbsDiffMQscores+ nodeicov.NumberJusticesPro+ nodeocov.sender.time + nodeifactor.SameIssueArea.2+ nodeifactor.SameIssueArea.3+ 
-                nodeifactor.SameIssueArea.4+nodeifactor.SameIssueArea.5+ nodeifactor.SameIssueArea.6+ nodeifactor.SameIssueArea.7+ 
-                nodeifactor.SameIssueArea.8+nodeifactor.SameIssueArea.9+  nodeifactor.SameIssueArea.10+
-                nodeifactor.SameIssueArea.11+ nodeifactor.SameIssueArea.12+ nodeifactor.SameIssueArea.13+nodeifactor.SameIssueArea.14+ 
-                nodeofactor.SameIssueArea.2+ nodeofactor.SameIssueArea.3+ nodeofactor.SameIssueArea.4+nodeofactor.SameIssueArea.5+ 
-                nodeofactor.SameIssueArea.6+nodeofactor.SameIssueArea.7+
-                nodeofactor.SameIssueArea.8+nodeofactor.SameIssueArea.9+nodeofactor.SameIssueArea.10+ nodeofactor.SameIssueArea.11+ 
-                nodeofactor.SameIssueArea.12+ nodeofactor.SameIssueArea.13+nodeofactor.SameIssueArea.14+
-                instar2_sendertime+outstar2_sendertime+ mutual_sendertime+ triangle_sendertime+mq_sendertime +sameissuearea_sendertime+
-                yeardiff_sendertime + yeardiffsquare_sendertime+ AbsDiffMQscores_sendertime + NumberJusticesPro_sendertime,family=binomial(logit),data=train.data)
-summary(mple.est)
-
-save(list="mple.est",file="training_full.RData")
-
-
-# MPLE
[…]

### 2020-04-24 R-Code/scc_data_merging.R

@@ -1,319 +1,0 @@
-### Edgelist gives case ids of citing and cited cases
-edgelist <- read.table("allcites.txt",header=F)
-## Need to add characters to ids so that they are not used as integers by network functions
-edgelist[,1] <- paste("c",edgelist[,1],sep="")
-edgelist[,2] <- paste("c",edgelist[,2],sep="")
-
-### Read in node attributes
-node_attributes <- read.csv("judicial.csv",stringsAsFactors=F)
-node_attributes[,1] <- paste("c",node_attributes$caseid,sep="")
-
-wash.law1<- read.csv("SCDB1.csv", stringsAsFactors = F)
-wash.law2<- read.csv("SCDB2.csv", stringsAsFactors = F)
-
-# remove space and dots in "1 U.S. 239" to match it with citeid in node_attributes
-uscite1<- gsub(" ", "", wash.law1[,7], fixed = TRUE)
-uscite1<- gsub(".", "", uscite1, fixed=TRUE)
-
-uscite2<- gsub(" ", "", wash.law2[,7], fixed = TRUE)
-uscite2<- gsub(".", "", uscite2, fixed=TRUE)
-
-wash.law1[,7]<- uscite1
-wash.law2[,7]<- uscite2
-
-# merge both data sets together
-
-# create empty matrix, so we can merge data from all three sets together
-scc<- cbind(node_attributes, matrix(0, 30288, 53))
-colnames(scc)<- c(colnames(node_attributes), colnames(wash.law1))
-
-# loop that merges data from wash.law1 into scc
-
-# merge wash.law1
-for(i in 1:19861){
-  if(i %% 1000 == 0) cat("Starting iteration", i, "\n")
-  scc.row<- which(wash.law1[i,7]==scc[,2])
-  scc[scc.row,17:69] <- wash.law1[i,]
-}
-
-# loop that merges data from wash.law2 into scc
-
-# merge wash.law2
-for(i in 1:8809){
-  if(i %% 1000 == 0) cat("Starting iteration", i, "\n")
-  scc.row<- which(wash.law2[i,7]==scc[,2])
-  scc[scc.row,17:69] <- wash.law2[i,]
-}
-
-
-# delete all cases where we don't have data
-for(i in 30188:1){
-  if(i %% 1000 == 0) cat("Less than", i, "\n")
-  if(scc[i, 23]==0){
-    scc<- scc[-i,]
-  }
-}
-
[…]

### 2020-04-24 R-Code/supreme_court_code.R

@@ -1,54 +1,0 @@
-### Edgelist gives case ids of citing and cited cases
-edgelist <- read.table("allcites.txt",header=F)
-## Need to add characters to ids so that they are not used as integers by network functions
-edgelist[,1] <- paste("c",edgelist[,1],sep="")
-edgelist[,2] <- paste("c",edgelist[,2],sep="")
-
-### Read in node attributes
-node_attributes <- read.csv("judicial.csv",stringsAsFactors=F)
-node_attributes[,1] <- paste("c",node_attributes$caseid,sep="")
-
-### Create a netowrk object
-## The following sequence assures that edges are matched to the correct nodes
-library(network)
-## this is a directed acyclic graph, so it makes most senese to analyze it as undirected (i.e., since there can be no cycling)
-scnetwork <- network.initialize(nrow(node_attributes),dir=F)
-## set name to case ids, which will make it easy to add edges
-network.vertex.names(scnetwork) <- node_attributes[,1]
-
-# convert edgelist to a matrix
-edgelist <- as.matrix(edgelist)
-
-# match node indices to numeric place
-edge_row <- match(edgelist[,1],network.vertex.names(scnetwork))
-edge_col <- match(edgelist[,2],network.vertex.names(scnetwork))
-
-# estimate timing
-test_time <- system.time( for(i in 2:200){ scnetwork[cbind(edge_row[(i-1):i], edge_col[(i-1):i])] <- 1} )
-
-# convert to hours
-estimated_hours <- test_time/199*length(edge_row)/60/60
-
-# view estimated timing
-print(estimated_hours)
-
-# run all rows to store edges
-for(i in 98500:length(edge_row)){ #stopped at 98500
-  if(i %% 250 == 0) cat("Starting iteration", i, "\n")
-  scnetwork[cbind(edge_row[(i-1):i], edge_col[(i-1):i])] <- 1}
-
-### Lets use the age of the case and the salience indicator
-set.vertex.attribute(scnetwork,c("year","salience"),node_attributes[,c("year","oxford")])
-
-
-# only use vertexes created after 1953
-sc<- scnetwork
-
-scnetwork54<- delete.vertices(sc, 1:21065)
-rm(sc)
-
-
-# fit mple
-library(statnet)
-mple <- try(ergm(scnetwork54~edges+ nodecov('salience')+ nodecov('year.center') + absdiff('year.center') + absdiff('year.center.square') + dsp(0) + esp(0), estimate = "MPLE" )) #
-summary(mple)

### 2020-04-24 R-Code_courtlistener/bootstrap_coefficient_summaries.R

@@ -1,1303 +1,0 @@
-
-#### code to create polynomial plots
-
-#####################################################
-### degree 3 polynomials
-
-# code assumes that working directory is set
-# to top-level project
-bootstrapResults <- read.csv("bootresults_3.csv",stringsAsFactors=F,row.names=1)
-
-d<- c("edges","istar2","ostar2","mutual","triangle", "edgecov.mq.t", "edgecov.same.issue.area.t", "edgecov.year.diff.t" ,"edgecov.year.diff.square.t" ,  
-      "nodeicov.AbsDiffMQscores", "nodeicov.NumberJusticesPro", "edgecov.year" , "nodeifactor.SameIssueArea.2", "nodeifactor.SameIssueArea.3",
-      "nodeifactor.SameIssueArea.4","nodeifactor.SameIssueArea.5", "nodeifactor.SameIssueArea.6", "nodeifactor.SameIssueArea.7", 
-      "nodeifactor.SameIssueArea.8","nodeifactor.SameIssueArea.9",  "nodeifactor.SameIssueArea.10",
-      "nodeifactor.SameIssueArea.11", "nodeifactor.SameIssueArea.12", "nodeifactor.SameIssueArea.13","nodeifactor.SameIssueArea.14", 
-      "nodeofactor.SameIssueArea.2", "nodeofactor.SameIssueArea.3", "nodeofactor.SameIssueArea.4","nodeofactor.SameIssueArea.5", 
-      "nodeofactor.SameIssueArea.6","nodeofactor.SameIssueArea.7",
-      "nodeofactor.SameIssueArea.8","nodeofactor.SameIssueArea.9","nodeofactor.SameIssueArea.10", "nodeofactor.SameIssueArea.11", 
-      "nodeofactor.SameIssueArea.12", "nodeofactor.SameIssueArea.13","nodeofactor.SameIssueArea.14", "nodeicov.Overruled",
-      "instar2_sendertime","outstar2_sendertime", "mutual_sendertime", "triangle_sendertime","mq_sendertime" ,"sameissuearea_sendertime",
-      "yeardiff_sendertime" , "yeardiffsquare_sendertime", "AbsDiffMQscores_sendertime" , "NumberJusticesPro_sendertime", "Overruled_sendertime",
[…]

### 2020-04-24 R-Code_courtlistener/data_merging_and_bootMPLE.R

@@ -1,220 +1,0 @@
-# merge together csv data that was obtained from the aci.ics cluster
- 
-# rm(list= ls()[!(ls() %in% c('scc1','adjacency.matrix', 'mq.matrix', 'same.issue.area', 'year.diff.matrix'))])
-memory.limit(30000)
-
-library(readr)
-
-mpledata1 <- read_csv("mpledata1.csv")
-mpledata2 <- read_csv("mpledata2.csv")
-mpledata3 <- read_csv("mpledata3.csv")
-mpledata4 <- read_csv("mpledata4.csv")
-mpledata5 <- read_csv("mpledata5.csv")
-
-mpledata1<- as.data.frame(mpledata1)
-mpledata2<- as.data.frame(mpledata2)
-mpledata3<- as.data.frame(mpledata3)
-mpledata4<- as.data.frame(mpledata4)
-mpledata5<- as.data.frame(mpledata5)
-
-mpledata<- rbind(mpledata1, mpledata2, mpledata3, mpledata4, mpledata5)
-mpledata<- mpledata[,-1]
-
-rm(mpledata1, mpledata2, mpledata3, mpledata4, mpledata5)
-
-
-# next, I create new columns to add interaction effects
-
-# first degree
-mpledata$instar2_sendertime<- mpledata[,3]*mpledata[,14]
-mpledata$mutual_sendertime<- mpledata[,4]*mpledata[,14]
-mpledata$outstar2_sendertime<- mpledata[,5]*mpledata[,14]
-mpledata$triangle_sendertime<- mpledata[,6]*mpledata[,14]
-mpledata$mq_sendertime<- mpledata[,7]*mpledata[,14]
-mpledata$sameissuearea_sendertime<- mpledata[,8]*mpledata[,14]
-mpledata$yeardiff_sendertime<- mpledata[,9]*mpledata[,14]
-mpledata$yeardiffsquare_sendertime<- mpledata[,10]*mpledata[,14]
-mpledata$AbsDiffMQscores_sendertime<- mpledata[,16]*mpledata[,14]
-mpledata$NumberJusticesPro_sendertime<- mpledata[,17]*mpledata[,14]
-mpledata$Overruled_sendertime<- mpledata[,44]*mpledata[,14]
-mpledata$jusitcehomophily_sendertime<- mpledata[,45]*mpledata[,14] 
-
-# 2nd degree
[…]

### 2020-04-24 R-Code_courtlistener/data_readin.R

@@ -1,641 +1,0 @@
-memory.limit(30000)
-
-
-# loading and preprocessing all the data takes a couple of hours (4-5h)
-
-
-#####################################################
-#### read in Washington law data
-#####################################################
-
-wash.law1<- read.csv("SCDB1.csv", stringsAsFactors = F)
-wash.law2<- read.csv("SCDB2.csv", stringsAsFactors = F)
-# combine data sets
-
-wash.law<- rbind(wash.law1, wash.law2)
-
-#####################################################
-#### read in courtlistener citation csv
-#####################################################
-
-library(readr)
-citations <- read_csv("cl_citations.csv")
-citations<- as.data.frame(citations)
-
-#######################################################
-#### read in Courtlistener scotus opinion data
-#######################################################
-
-library(rjson)
-# takes about 5 mins
-# filenames_op <- list.files("C:/Users/Geiler Typ/Desktop/PSU/Bruce Desmarais/Supreme Court Citation Project/scotus_opinion", pattern="*.json", full.names=TRUE) # this should give you a character vector, with each file name represented by an entry
-# takes about 30 mins
-# myJSON_opinion <- lapply(filenames_op, function(x) fromJSON(file=x)) # a list in which each element is one of the original JSON files, takes a while ~ 30 min
-
-######################################################
-#### read in courtlistener scotus cluster-based data
-
-filenames_cluster <- list.files("C:/Users/Geiler Typ/Desktop/PSU/Bruce Desmarais/Supreme Court Citation Project/R-Code_courtlistener_data/all_cluster/scotus", pattern="*.json", full.names=TRUE) # this should give you a character vector, with each file name represented by an entry
-# takes about 30 mins
-myJSON_cluster <- lapply(filenames_cluster, function(x) fromJSON(file=x)) # a list in which each element is one of the original JSON files, takes a while ~ 30 min
-
[…]

### 2020-04-24 RCode/Plot Figures/GOF.R

@@ -1,0 +1,193 @@
+
+#################################################
+#################################################
+###             GOF for 1950 and 2015 for OSP and OST
+#################################################
+#################################################
+
+
+
+library(statnet) 
+
+# calculate indegree, outdegree and esp of unfixed nodes of observed network
+
+year.total <- scc1[,11]-1936 #-1936 => the first year 1937 is 1, 1938 is 2 aso, term column 
+
+set.seed(12123)
+#sampled.years <- c(14, 24, 34, 44, 54, 64, 74)
+
+t<-  79 # 79 stands for 2015, 14 for 1950
+print(t+1936)
+
+
+cases <- max(which(year.total==t)) 
+
+# simple assignment of time periods to cases
+years <- year.total[1:cases] 
+
+# going to need a sender time matrix covariate
+sender.time <- matrix(years,length(years),length(years),byrow=F)
+year <- matrix(years, length(years),length(years),byrow=F)
+
+# extract the network up to time t
+AM <- adjacency.matrix[which(years <= t),which(years <= t)]
+
+# determine the unfixed nodes
+unfixed<- which(years==t)
+
+# get outdegree of these nodes
+unfixed.odeg <- rowSums(AM[unfixed,])
+#hist(unfixed.odeg)
+
+# get indegree of these nodes
+unfixed.ideg <- colSums(AM[unfixed,])
+#hist(unfixed.ideg)
+
+net.t <- network(AM) 
+
+# get edgewise shared partner distribution
+full.esp <- summary(net.t ~ esp(0:20))
+full.esp   # this is esp dist if entire network is unfixed
+#plot(full.esp)
+
+# get esp dist of t-1
+AM.red <- AM[-unfixed, -unfixed]
+net.r <- network(AM.red) 
+
+# get edgewise shared partner distribution
+red.esp <- summary(net.r ~ esp(0:20))
+red.esp
+
+unfixed.esp.ost<- full.esp-red.esp
+#plot(unfixed.esp)
+
+
+w<- which(sender.time[,1]==t)  
+
+esp <-rep(0,21)
+for(u in w){
+  
[…]

### 2020-04-24 RCode/Plot Figures/Plot_SCC.R

@@ -1,0 +1,403 @@
+
+
+
+############################################
+#### Plot SCC Results
+############################################
+
+load(file="1950_2.RData")
+load(file="1951_2.RData")
+load(file="1952_2.RData")
+load(file="1953_2.RData")
+load(file="1954_2.RData")
+load(file="1955_2.RData")
+load(file="1956_2.RData")
+load(file="1957_2.RData")
+load(file="1958_2.RData")
+load(file="1959_2.RData")
+load(file="1960_2.RData")
+load(file="1961_2.RData")
+load(file="1962_2.RData")
+load(file="1963_2.RData")
+load(file="1964_2.RData")
+load(file="1965_2.RData")
+load(file="1966_2.RData")
+load(file="1967_2.RData")
+load(file="1968_2.RData")
+load(file="1969_2.RData")
+load(file="1970_2.RData")
+load(file="1971_2.RData")
+load(file="1972_2.RData")
+load(file="1973_2.RData")
+load(file="1974_2.RData")
+load(file="1975_2.RData")
+load(file="1976_2.RData")
+load(file="1977_2.RData")
+load(file="1978_2.RData")
+load(file="1979_2.RData")
+load(file="1980_2.RData")
+load(file="1981_2.RData")
+load(file="1982_2.RData")
+load(file="1983_2.RData")
+load(file="1984_2.RData")
+load(file="1985_2.RData")
+load(file="1986_2.RData")
+load(file="1987_2.RData")
+load(file="1988_2.RData")
+load(file="1989_2.RData")
+load(file="1990_2.RData")
+load(file="1991_2.RData")
+load(file="1992_2.RData")
+load(file="1993_2.RData")
+load(file="1994_2.RData")
+load(file="1995_2.RData")
+load(file="1996_2.RData")
+load(file="1997_2.RData")
+load(file="1998_2.RData")
+load(file="1999_2.RData")
+load(file="2000_2.RData")
+load(file="2001_san.RData")
+load(file="2002_2.RData")
+load(file="2003_2.RData")
+load(file="2004_san.RData")
+load(file="2005_2.RData")
+load(file="2006_san.RData")
+load(file="2007_2.RData")
+load(file="2008_2.RData")
+load(file="2009_2.RData")
[…]

### 2020-04-24 RCode/data_readin.R

@@ -1,0 +1,825 @@
+memory.limit(30000)
+
+
+# loading and preprocessing all the data takes a couple of hours (4-5h)
+
+
+#####################################################
+#### read in Washington law data
+#####################################################
+
+wash.law1<- read.csv("SCDB1.csv", stringsAsFactors = F)
+wash.law2<- read.csv("SCDB2.csv", stringsAsFactors = F)
+# combine data sets
+
+wash.law<- rbind(wash.law1, wash.law2)
+
+#####################################################
+#### read in courtlistener citation csv
+##################################################
+library(readr)
+citations.old <- read_csv("cl_citations.csv")
+citations.old<- as.data.frame(citations)
+
+# 10/4/2019
+citations<- read_csv("cl_citations2.csv")
+citations<- as.data.frame(citations2)
+
+#######################################################
+#### read in Courtlistener scotus opinion data
+#######################################################
+
+library(rjson)
+# takes about 5 mins
+# filenames_op <- list.files("C:/Users/Geiler Typ/Desktop/PSU/Bruce Desmarais/Supreme Court Citation Project/scotus_opinion", pattern="*.json", full.names=TRUE) # this should give you a character vector, with each file name represented by an entry
+# takes about 30 mins
+# myJSON_opinion <- lapply(filenames_op, function(x) fromJSON(file=x)) # a list in which each element is one of the original JSON files, takes a while ~ 30 min
+
+######################################################
+#### read in courtlistener scotus cluster-based data
+
+filenames_cluster <- list.files("all_cluster/scotus", pattern="*.json", full.names=TRUE) # data is not on git. this should give you a character vector, with each file name represented by an entry
+# takes about 30 mins
+myJSON_cluster <- lapply(filenames_cluster, function(x) fromJSON(file=x)) # a list in which each element is one of the original JSON files, takes a while ~ 30 min
+
[…]

### 2020-04-29 Tex/paper.tex
revised up through c-ERGM specification

@@
 
-\title{\Large Generative Dynamics of Supreme Court Citations}
+\title{\Large Generative Dynamics of Supreme Court Citations: \\ Analysis with a New Statistical Model}
 %\author{}
@@
 \begin{abstract}
-\noindent The significance and influence of US Supreme Court majority opinions derive in large part to opinions' roles as precedents for future opinions.
+\noindent The significance and influence of US Supreme Court majority opinions derive in large part from opinions' roles as precedents for future opinions.
 A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns.
@@
 This methodology---the citation temporal exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation.
-We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2015.
+We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015.
 We find evidence for dependence processes, including reciprocity, transitivity, popularity, and activity.
@@
 United States Supreme Court opinions exercise authority and influence, in part, through their roles as precedents affecting future jurisprudence in the US.
-The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,gillman2001s,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained}.
+The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,gillman2001s,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained,hitt2016measuring}.
 Despite a considerable body of research that focuses on the way in which relevant precedents shape decision making on the Court, relatively little work has focused on understanding which past opinions are cited by an opinion.
 Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited in an opinion.
-We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{harris1982structural,caldeira1988legal,fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics}.
+We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{harris1982structural,caldeira1988legal,fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics,ethayarajh2018rose}.
 
 We are not the first to ask what predicts the citations in US Supreme Court Opinions.
-Indeed, a voluminous body of work has sought to explain how many times an opinion is cited \citep[e.g.,][]{cross2010determinants,benjamin2012standing}, when an opinion is cited \citep[e.g.,][]{black2013citation,spriggs2001explaining}, and how many cases are cited by an opinion \citep[e.g.,][]{lupu2013strategic}---all focused on the US Supreme Court.
+Indeed, a voluminous body of work has sought to explain how many times an opinion is cited \citep[e.g.,][]{cross2010determinants,benjamin2012standing,fix2019effect}, when an opinion is cited \citep[e.g.,][]{black2013citation,spriggs2001explaining}, and how many cases are cited by an opinion \citep[e.g.,][]{lupu2013strategic}---all focused on the US Supreme Court.
 One common feature of the research design in all of these studies is that the observations are at the case or case-year level.
@@
 Methodologically, we develop a novel extension of a statistical model for networks, which we adapt to the network structural constraints of court citations.
-Second, we apply this methodology to a half-century of directed dyadic citations between US Supreme Court opinions.
+Second, we apply this methodology to a half-century of directed dyadic citations between U.S.
+Supreme Court opinions.
 
@@
 This phenomenon of complex dependence is very common in networks of many types, but processes specific to Supreme Court citations create interdependence in citations.
-For example, if opinion $i$ relies heavily on opinion $j$ as precedent, opinion $i$ is likely to discuss the legal basis for opinion $j$, and as a consequence cite some of the opinions cited by opinion $j$.
+For example, if opinion $i$ relies heavily on opinion $j$ as precedent, opinion $i$ is likely to discuss the legal basis for opinion $j$, and as a consequence, cite some of the opinions cited by opinion $j$.
 Suppose opinion $k$ is cited by opinion $j$.
@@
 We then develop an extension of a model---the exponential random graph model (ERGM)---that can incorporate both exogenous covariates and complex forms of interdependence into a directed dyadic analysis of citations.
-Finally, we develop and estimate a specification of this model in an analysis of US Supreme Court citations between 1937 and 2015.
-We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's predictive performance.
+Finally, we develop and estimate a specification of this model in an analysis of US Supreme Court citations between 1950 and 2015.
+We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's fit.
 
@@
 
-When it comes to the development and testing of theory, the defining feature of networks is that the fundamental element under study---the relationship between two units (i.e., the citation from one opinion to another) is a component of a complex system of relations.
+When it comes to the development and testing of theory, the defining feature of networks is that the micro-level unit of analysis---the relationship between two entities (i.e., the citation from one opinion to another) is a component of a complex system of relations.
 The formation (or lack thereof) of that relationship cannot be fully understood without considering how the relationship fits into the system.
@@
 
-The third network property we consider in the context of Supreme Court citations is popularity.
+The third, and final, network property we consider in the context of Supreme Court citations is popularity.
 Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent \citep{barabasi1999emergence}.
@@
 
-The fourth network property we consider in the context of Supreme Court citations is activity.
-Activity, or sender activation, is the tendency for ties sent to beget more ties sent \citep{howard2016understanding}.
-Structurally, this means that there are few nodes that send a moderate number of ties---the number of ties sent is either small or fairly large.
-In terms of Supreme Court opinions, for each opinion discussed, that discussion is likely to raise other tangential issues on which the justices will want to draw upon past opinions.
-Furthermore, for each opinion that applies to and is cited by the current opinion, there is often a case/opinion that needs to be discussed in terms of why it does not apply.
-Justices often clarify not only those legal principles that apply, but often those that do not.
-The example passage that illustrates activity is from Kush v.
-Rutledge (460 U.S.
-719, 1983)---a case addressing witness intimidation and due process protections.
-In the Kush opinion, Griffin v.
-Beckenridge (i.e., Griffin) is discussed at length in terms of its lack of applicability.
-It is common that opinions distinguish among a variety of legal rules in terms of their applicability to a case, which means that for each rule that is established in an applicable precedent, one or more need to be discussed in terms of why they do not apply.
-We do not present a separate visualization related to activity, as it would look very similar to the visualization for popularity.
-\begin{quotation}
-No allegations of racial or class-based invidiously discriminatory animus are required to establish a cause of action under the first part of 1985(2).
-The statutory provisions now codified at \S 1985 were originally enacted as \S 2 of the Civil Rights Act of 1871, and the substantive meaning of the 1871 Act has not been changed.
-The provisions relating to institutions and processes of the Federal Government (including the first part of \S 1985(2)) -- unlike those encompassing activity that is usually of primary state concern (including the second part of \S 1985(2) and the part of \P 1985(3) involved in Griffin, supra -- contain no language requiring that the conspirators act with intent to deprive their victims of the equal protection of the laws.
-Thus, the reasoning of Griffin is not applicable here, and, given the structure of \S 2 of the 1871 Act, it is clear''
-\end{quotation} % https://supreme.justia.com/cases/federal/us/460/719/
+These three properties---transitivity, reciprocity, and popularity form the core of our network theory of U.S.
+Supreme Court citations.
+We also seek to model the effects of covariates (i.e., case features) on citation formation.
+In the next section, we introduce a modeling framework that can incorporate all of these effects into a single model, and adapt to the peculiar structure of citation matrices.
 
@@
 We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
-To accomplish this, we extend a model that has been developed to jointly model covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{cranmer2012complex,clark2013multimember, masket2015polarization, graif2017neighborhood} {\color{blue} Citations need to be adjusted to ERGM, not TERGM- CS}.
-%The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
+To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{welles2015individual,duque2018recognizing, graif2017neighborhood,osei2018elite}  %The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
 The network structures for which ERGMs are currently designed are insufficient to account for the structure of citation networks.
@@
 This re-arrangement illustrates that the parameters can be interpreted in terms of the change in the log odds of a citation from $i$ to $j$ given a one-unit increase in the corresponding element of $\bm{h}$, conditional on the other citations observed in the network.
-For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log odds of observing $C_{ij,t} = 1$ increase by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$).
+For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log odds of observing $C_{ij,t} = 1$ increases by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$).
 The logit form conditional probability is well known for the ERGM family \citep{goodreau2009birds}.
@@
 We cannot statistically differentiate between these two dynamics.
-However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables and the c-ERGM can be specified to account for the dependence in the data.
+However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-ERGM can be specified to account for the dependence in the data.
 
@@
 
-The common alternative relates on Monte Carlo methods to approximate the normalizing constant by simulating a large set of networks \citep{hunter2006inference,hummel2012improving}.
+The common alternative relies on Monte Carlo methods to approximate the normalizing constant by simulating a large set of networks \citep{hunter2006inference,hummel2012improving}.
 The resulting estimator, the Monte Carlo MLE (MCMLE), is approximately consistent, meaning that it converges to the MLE as the sample size, i.e., the number of simulated networks, increases.
@@
 
-We estimate the MLE for every term of the c-ERGM between 1950 and 2015.
+We estimate the MCMLE for every term of the c-ERGM between 1950 and 2015.
 While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows.
+The growth in the network also serves as our reason for estimating a separate model for each term.
+First, the network is large, with each term introducing the potential for hundreds-of-thousands of new citations.
+This means we have the statistical power necessary to estimate a separate set of parameter values for each term.
+Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network.
+This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term.
+
 The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant.
@@
 Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$.
-
 But even this approach started to fail for networks around the turn of the millennium.
 Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}.
-The MCMLE for these large citation networks was finally obtained by setting the starting value according to an approach introduced by \citet{SchmidHunter2020}.
+The MCMLE for these large citation networks was obtained by setting the starting value according to an approach introduced by \citet{SchmidHunter2020}.
 This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$.
@@
 The degree distributions indicate that there is a long tail to both the number of citations sent and received.
-These long-tailed (i.e., high kurtosis) distributions provide preliminary evidence of both popularity and activity dynamics \citep{strogatz2001exploring}.
-Figure \ref{fig:networkviz} displays the complete network.
+These long-tailed (i.e., high kurtosis) distributions provide preliminary evidence of substantial heterogeneity in the features that drive citations to and from cases \citep{strogatz2001exploring}.
+Figure \ref{fig:networkviz} displays the citation network network from 1937--2001.
 We see here that the densest rates of tie formation tend to be between consecutive courts (e.g., the Stone Court is much more tightly tied to the Hughes Court than the Rehnquist Court is to the Hughes Court).
@@
 
-\subsection{c-TERGM specification}
-
-The c-TERGM we estimate includes two classes of terms---one that captures the effects of covariates on tie formation, and another that captures the complex dependence processes that we expect to observe in the Supreme Court Citation Network.
+\subsection{c-ERGM specification}
+
+The c-ERGM we estimate includes two classes of terms---one that captures the effects of covariates on tie formation, and another that captures the complex dependence processes that we expect to observe in the Supreme Court Citation Network.
 We describe our model specification by defining the terms within these two classes.
@@
 
-Covariate effects are accounted for in the c-TERGM via the term that is used to specify the effect of covariates in other ERGM family models, as
+Covariate effects are accounted for in the c-ERGM via the term that is used to specify the effect of covariates in other ERGM family models, as
 $$h_{covariate}(C_t,C_{<t},X) =  \sum_{ij} C_{ij,t}X_{ij}.$$ Since $C_{ij,t}$ is a binary indicator of whether case $i$ cites case $j$, $h_{covariate}$ amounts to the sum of covariate values among directed dyads for which we observed a citation.

### 2020-04-30 Tex/paper.tex
revised covariate terms

@@
 $$h_{covariate}(C_t,C_{<t},X) =  \sum_{ij} C_{ij,t}X_{ij}.$$ Since $C_{ij,t}$ is a binary indicator of whether case $i$ cites case $j$, $h_{covariate}$ amounts to the sum of covariate values among directed dyads for which we observed a citation.
-The dyadic interpretation of the coefficient attributed to this term is the change in the log odds of a tie from $i$ to $j$ given a unit increase in $X_{ij}$ (i.e., exactly the interpretation of the effect of a covariate in logistic regression).
-We include several exogenous covariates based on this standard statistic formulation.
-
-The first covariate (i.e., $X$) we incorporate into the model accounts for the degree to which cases cite those that are similar in terms of the ideological positions of the justices who supported the decision.
+The dyadic interpretation of the coefficient attributed to this term is the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in $X_{ij}$ (i.e., exactly the interpretation of the effect of a covariate in logistic regression).
+We include several exogenous covariates based on this standard formulation.
+
+The first covariate we incorporate into the model accounts for the degree to which cases cite those that are similar in terms of the ideological positions of the justices who supported the decision.
 We account for this effect following \citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
@@
 We include a covariate term in which $X_{ij}$ is the absolute difference between the Martin-Quinn scores \citep{martin2002dynamic} of the median justices in the majority coalitions for cases $i$ and $j$.
-We expect this variable to have a negative effect, which would correspond to cases citing those to which they are ideologically similar.%WE COULD USE THIS DATASET AS A SECOND APPLICATION
+We expect this variable to have a negative effect, which would correspond to cases citing those to which they are ideologically similar.
+In addition to the absolute difference variable, we also include a linear effect of the Martin-Quinn score of the median justice in the majority coalition for case $j$, which adjusts for constant ideological biases exhibited within a given term.
+%WE COULD USE THIS DATASET AS A SECOND APPLICATION
 

### 2020-05-01 Tex/paper.tex
revising covariate terms
revising dependence terms

@@
 We expect this variable to have a negative effect, which would correspond to cases citing those to which they are ideologically similar.
-In addition to the absolute difference variable, we also include a linear effect of the Martin-Quinn score of the median justice in the majority coalition for case $j$, which adjusts for constant ideological biases exhibited within a given term.
 %WE COULD USE THIS DATASET AS A SECOND APPLICATION
@@
 Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy \citep{Handcock.2003}.
-Model degeneracy is a common obstacle when fitting ERGMs and occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network.
+Model degeneracy is a common obstacle when fitting ERGMs, and occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network.
 In this scenario, the simulated networks are too different from the observed network, making the underlying distribution defined by the model extremely unrealistic.
@@
 In practice, this is the number of mutual dyads in $C_{\leq t}$, since mutual dyads must emerge among opinions that were written in the same term.
-The log odds interpretation of the reciprocity statistic is that if opinions in case $i$ cite case $j$, the log odds that case $j$ also cites $i$ change by the theta relative to the configuration in which case $j$ does not cite case $i$.
+The log odds interpretation of the reciprocity statistic is that if the opinion in case $i$ cites case $j$, the log odds that case $j$ also cites $i$ change by the estimated coefficient relative to the configuration in which case $j$ does not cite case $i$.
+We expect the reciprocity effect to be positive.
 
@@
 \end{equation}
-The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on $D_r(C_t, C_{<t})$ as $r$ increases which means that the higher $\lambda$ the more the statistic responds to changes for cases with a high indegree.
+The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on $D_r(C_t, C_{<t})$ as $r$ increases, which means that the higher $\lambda$ the more the statistic responds to changes for cases with a high indegree.
 We fixed $\lambda=1$.
@@
 $$\text{logit} P(C_{ij,t}=1 | C_{-ij,t}, C_{<t}) = \exp \bigl(~\theta \cdot( 1- \frac{1}{\lambda})^r \bigr).$$
-As stated by \citet{Levy2016} this means for the interpretation that a negative parameter value indicates the centralization of edges while a positive parameter indicates the dispersion of edges.
+As stated by \citet{Levy2016}, this means for the interpretation that a negative parameter value indicates the centralization of edges (i.e., popularity) while a positive parameter indicates the dispersion of edges (i.e., new ties going to less popular nodes).
 Since we expect highly cited cases to be more likely to be cited again, we expect the gwidegree parameter to be negative.
+
+% Transitivity
+We include two statistics to account for different types of transitivity.
+The first transitivity statistic calculates the number of transitive ties, $i \to j$, where there is a directed path of length two from $i$ to $j$ through at least one case $k$, and $j$ and $k$ were written during a different term than $i$.
+We will refer to this statistic as the \textit{different term transitivity} statistic.
+As illustrated in Section 2, we theorize Supreme Court citations to pattern a high degree of transitivity.
+As a result, we expect positive parameter estimates for the \textit{different term transitivity} statistic.
+
+The second transitivity statistic, intended to capture clustering that includes more than one tie formed in the same term, is the \textit{geometrically weighted edgewise shared partners} (gwesp) statistic introduced by \citet{hunter2006inference}.
+For directed networks the edgewise shared $r$-partners statistic ($ESP_r$) can take multiple forms, but we integrate the definition where cases $i$ and $j$ that are connected by a citation, also cite $r$ common other cases.
+We follow the notation of \citet{Butts.2008} and call this ESP specification \textit{outgoing shared partners} (OSP).
+Similar to the gwidegree-statistic the gwesp-statistic incorporates a network's edgewise shared partners distribution by replacing $D_r$ by $ESP_r$.
+Triangles that are incident to the outgoing shared partners configuration arise when there is a cite between two cases decided at the same time, and those two cases cite one or more of the same third cases (decided either previously or also during the same term).
+This would happen if citations were clustering around legal issues that are not completely captured by our same-issue-area effects.
+As we do not expect that the issue area variable will account for all of the clustering around legal issues, we expect the gwesp statistic to also carry a positive coefficient value.
 
@@
 In other words, this statistic adds up the outdegree of cases that are being cited at time $t$ and examines whether cases that show a high degree of activity are being more likely to be cited in the future.
-
-% Transitivity
-We include two statistics to account for different types of transitivity.
-The first transitivity statistic calculates the number of transitive ties, $i \to j$, where there is a directed path of length two from $i$ to $j$ through at least one case $k$, and $j$ and $k$ were written during a different term than $i$.
-We will refer to this statistic as the \textit{different term transitivity} statistic.
-As illustrated in chapter 2 we theorize Supreme Court citations to pattern a high degree of transitivity.
-As a result, we expect positive parameter estimates for the different term statistics.
-
-The second transitivity statistic is the \textit{geometrically weighted edgewise shared partners} (gwesp) statistic introduced by \citet{hunter2006inference}.
-For directed networks the edgewise shared $r$-partners statistic ($ESP_r$) can take multiple forms, but we integrate the definition where cases $i$ and $j$ that are connected by a citation, also cite $r$ common other cases.
-We follow the notation of \citet{Butts.2008} and call this ESP specification \textit{outgoing shared partners} (OSP).
-Similar to the gwidegree-statistic the gwesp-statistic incorporates a network's edgewise shared partners distribution by replacing $D_r$ by $ESP_r$.
-Note that unlike the different term transitivity statistic the gwesp statistic does not require the cited cases to be written in a previous term.
 

### 2020-05-04 Tex/paper.tex
completed dependence terms

@@
 % Activity
-In order to account for the activity effect, the straightforward thing to do is to add the \textit{geometrically weighted outdegree} (gwodegree) statistic, which is defined as gwidegree with the difference that $D_r(C_t, C_{<t})$ represents the number of nodes with outdegree $r$.
-However, adding the gwodegree statistic causes the model to degenerate again.
-We circumvent degeneracy by adding the outdegree of every node that entered the network prior time $t$ as a receiver effect.
-In other words, this statistic adds up the outdegree of cases that are being cited at time $t$ and examines whether cases that show a high degree of activity are being more likely to be cited in the future.
+We include one more structural network effect to adjust for inherent differences across cases in terms of the overall scope of the legal issues addressed in a case.
+\citet{fowler2008authority} finds that cases vary in the degree to which they serve as ``hubs''---citing to many other cases.
+We expect that new cases will be more likely to cite cases that themselves cited many cases, because cases that themselves cite a large number of cases address a broader array of legal issues.
+We include the outdegree of every node that entered the network prior time $t$ as a receiver effect.
+We expect the coefficient on this term to be positive.
+This is not technically a dependence term, as it is formulated as a covariate, but we discuss it in the current section because it is purely a network structure effect.
 
@@
 
-\subsection{Predictive Performance}
+\subsection{Model Fit}
 

### 2020-05-04 Tex/paper.tex
completed editing pass

@@
 We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015.
-We find evidence for dependence processes, including reciprocity, transitivity, popularity, and activity.
+We find evidence for dependence processes, including reciprocity, transitivity, and popularity.
 The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model.
@@
 
-We include dependence terms to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, popularity, and activity.
+We include dependence terms to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, and popularity.
 %We expect that the coefficients associated with each of these dependence effects will be positive.
@@
 Our case for studying legal citations at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence.
-To quantify this contribution, we use AIC and BIC out-of-sample prediction.
-Predicting out-of-sample offers an unbiased and general purpose way to evaluate the contribution, in terms of model fit, of one or more terms/parameters in a model \citep{jensen2000multiple,ward2010perils}.
-Unlike in-sample measures of model fit, out-of-sample methods are highly robust in avoiding overfitting, and work when we cannot accurately calculate the value of the likelihood function, as in the current case.
-Out-of-sample prediction is a common way to evaluate methods for modeling ties in networks, and has recently been applied to TERGMs in particular \citep{desmarais2013forecasting,cranmer2017can}.
+To quantify this contribution, we fit one model that incorporates covariate effects only, which we term the {\em Independent Model}---excluding the reciprocity, transitivity, and popularity terms.
+The full model is the one we present below.
+We use both AIC and BIC to compare the fit of the two models, with lower values indicating better fit \citep{ripplinger2008does}.
 
@@
 
-Figure \ref{AIC_BIC} depicts the AIC and BIC comparison for two different models for each term from 1950 to 2015.
-The full model is defined as the model that includes all covariate terms and dependence terms described in \ref{covariate_terms} and \ref{dependence_terms}.
-The independent model only consists of covariate terms, not accounting for any structural dependencies in the networks.
+Figure \ref{AIC_BIC} depicts the AIC and BIC comparison for the two different models for each term from 1950 to 2015.
 We see that the AIC and BIC is considerably smaller for the full model throughout the period of consideration.
-On average the independent model yields a 5525 point higher AIC and an average 5582 point higher BIC.
+On average the independent model yields a 5,525 point higher AIC and an average 5,582 point higher BIC.
 The maximum difference is reached in 1983 with a difference of 10,990 points for the AIC and 10,930 points for the BIC.
+This fit comparison provides robust evidence that the dependence terms contribute significantly to the explanatory power of the model.\footnote{In the appendix we present conventional goodness-of-fit plots that are used to assess the structural fit for ERGM \citep{hunter2008goodness}, as well as degeneracy diagnostics \citep{mukherjee2020degeneracy}.
+We find that the full model fits the degree and edgewise shared partner distributions well, and is not degenerate.}
 
@@
 
-\subsection{c-TERGM Results}
-The results from the c-TERGM are presented Figures \ref{SCC_results_1} and \ref{SCC_results_2}.
+\subsection{c-ERGM Coefficient Estimates}
+The estimates from the c-ERGM are presented in Figures \ref{SCC_results_1} and \ref{SCC_results_2}.
 In these figures we depict coefficient estimates and 95\% confidence intervals for all of the effects for all the variables we discussed in sections \ref{covariate_terms} and \ref{dependence_terms}.
-The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one unit increase in the respective variable.
-For this reason, the number of edges in a network corresponds to the intercept in an ERGM.
+The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in the respective variable.
 We first discuss the dependence effect estimates in Figure \ref{SCC_results_1}.
@@
 A triangle signals that the estimate for a given year was not significant at an $\alpha=0.1$ level.
-
-The panel showing the results of the reciprocity statistic does not provide estimates for ever term.
-In these missing terms there were no two cases that would cite each other, making the estimation of the reciprocity statistic impossible.
+The panel showing the results of the reciprocity statistic does not provide estimates for every term.
+In these missing terms there were no two cases that cited each other, making the estimate of the reciprocity statistic term equivalent to $-\infty$.
+For those years, the reciprocity term is included in the model, but the estimate of $-\infty$.
 A similar scenario is visible for the \textit{Overruled Cases} variable.
@@
 
-%The reciprocity effect is substantial, positive, and statistically significant for most of the period under study.
-Conditional upon cases $i$ and $j$ being considered within the same term, the log of the odds that opinion $i$ cites opinion $j$ increases by approximately $2$ when opinion $j$ cites opinion $i$.
-The naive probability that any one case cites another case decided at the same time or previously in the Supreme Court Citation network is approximately $0.0019$.
-This probability increases eight-fold, to approximately $0.015$ if the citation would be reciprocal.
-The transitivity (i.e., triangle closure) effect is also substantial, positive, and statistically significant over the entire period studied.
-The probability that case $i$ cites case $j$ increases approximately five-fold if the citation would close a triangle involving cases $i$ and $j$.
-The activity and popularity effects are both estimated to be positive, but trend downward and are more modest in scale (and in the case of activity, not statistically significant later in the time series).
-Interpreting the popularity effect---the log odds that $i$ cites $j$ increases by 0.018--0.035 with every additional citation to case $j$.
-Based on the indegree distribution, it appears that the popularity of cases can be classified by multiples of 10.
-If case $j$ increases from ten to twenty citations from other cases, the probability that case $i$ cites case $j$ increases from approximately 0.0030 to 0.0040.
-Overall, these results represent evidence of complex dependence in Supreme Court citation formation.
-
-%The effects of the exogenous covariates included in the c-TERGM (panels (e)--(k) in Figure \ref{fig:coeftrends_exo}) exhibit some surprising dynamics.
-The effects of absolute difference in Martin-Quinn score, and majority size of the recipient case all flip signs throughout the time period under study.
-This may be an artifact of the simple linear trends we have estimated, but is interesting nonetheless.
-The attitudinal model holds that ideological preferences drive decision-making on the US Supreme Court.
-Ideological selection of precedents would be evidenced by a negative effect of the absolute difference in MQ scores in the c-TERGM.
-We find this effect is negative and statistically significant only after 1975.
-The absolute difference in MQ scores variable is on the scale of approximately 0---3.
-At its most ideological point (i.e.
-2001), the probability that case $i$ cites case $j$ decreases from approximately 0.024 to 0.018---by one third---if the ideological distance between the majority coalitions in the two cases increases by 3.
-The effects of the majority size and ideological breadth of case $j$ are similar in both scale and trend.
-It is not until later in the time series that they both exhibit their expected positive effects.
-The effect of same issue is positive and statistically significant, and substantial for the time period under study.
-The probability that $i$ cites $j$ increases by approximately three-fold over the naive probability of citation if $i$ and $j$ have the same legal issue area.
+We first discuss our findings regarding the dependence terms.
+In most of the years that it is estimable, the reciprocity effect is statistically significantly positive, and substantial in magnitude.
+In a typical year, the presence of a citation from case $i$ to case $j$ increases the log-odds of a citation from case $j$ to case $i$ by approximately 0.50.
+Both transitivity effects---different term transitivity, and GWESP---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect.
+The log-odds of a citation from case $i$ to case $j$ increases by more than 2 if there is at least one third case, $k$ that is cited by $i$, and cites to $j$.
+Results regarding the popularity effect (i.e., GWIdegree) are generally supportive of the hypothesis of a popularity effect in Supreme Court citations, as the GWIdegree estimate is negative in every term.
+There are some terms for which the result is not statistically significant at conventional levels.
+The popularity effect looks to have grown substantially stronger after the 2000 term.
+Contrary to expectations, we find that the hub effect is negative.
+That is, each additional citation sent by a case reduces the likelihood that the case is cited.
+For every citation sent, the log-odds of any citation to a case decreases by approximately 0.025---a result that is statistically significant in each term.
+
+The effects of the exogenous covariates included in the c-ERGM are presented in Figure \ref{SCC_results_2}.
+Most of the exogenous covariate effects align with expectations.
+We find that cases are more likely to cite those that (1) have been decided most recently, (2) were decided by large majorities, (3) have not been overruled, (3) were authored by the same justice, and (3) are in the same issue area.
+Surprisingly, we do not find consistent effects of either of the covariates that involve justice ideology.
+The ideological distance between two cases, as measured by the absolute difference in the median of the majority coalitions; and the ideological breadth of the cited cases, as measured by the absolute difference between the most liberal and most conservative justices on the potential receiving case, both exhibit signs and significance levels that are highly volatile.
+Our results leave open the question of whether, and if so, how, citations are shaped by ideological factors.
 
@@
 In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.
-
-%{\em {\bf Note:} The results section of this paper is incomplete and in flux.
-We recently discovered evidence of non-convergence in the C-TERGM results currently presented in the paper, and have changed the models and estimation settings to address this problem.
-Most of the substantive conclusions look like they will hold using the updated models, but we are not sure that we will be able to include the updated models in the APSA presentation.}
 
@@
 This methodology---the citation-ERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation.
-We apply this methodology to a network that includes all Supreme Court cases decided between 1937 and 2015.
+We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015.
 We find, somewhat counterintuitively, that Supreme Court citations are highly reciprocal.
-We also find that citations are driven by dependencies such as triad closure, popularity, and activity.
+We also find that citations are driven by dependencies such as triad closure and popularity.
 The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model.
@@
 
-We recognize several limitations of the current analysis, which should be addressed in future iterations of this work or in future studies.
+We recognize two important limitations of the current analysis, which should be addressed in future iterations of this work or in future studies.
 First, our theoretical claims and analyses treat all citations---positive and negative---as equivalent.

### 2020-05-04 Tex/paper.tex
minor tergm edit
revised abstract

@@
    \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
- % \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
+  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
   }
@@
 First, dyadic citations are typically aggregated to the case level before they are analyzed.
-Second, citations are treated as if they arise independent of one another.
-We make the case that studying Supreme Court citations as a dynamic citation network overcomes both of these limitations.
-We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network.
-This methodology---the citation temporal exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation.
-We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015.
+Second, citations are treated as if they arise independently.
+We present a methodology for studying citations between Supreme Court opinions at the dyadic level, as a network, that overcomes these limitations.
+This methodology---the citation exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation.
+We then analyze a network that includes all Supreme Court cases decided between 1950 and 2015.
 We find evidence for dependence processes, including reciprocity, transitivity, and popularity.
-The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model.
-The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.\footnote{
-%Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. 
+The dependence effects are as substantively and statistically significant as the effects of the exogenous covariates we include in the model, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.\footnote{
+Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. 
 This work was supported in part by NIH, 
@@
 % Network Statistics
-The c-TERGM is specified through 
+The c-ERGM is specified through 
 the decision regarding which network statistics $h(\cdot)$ to incorporate. We include the following statistics for the Supreme Court citation network:

### 2020-05-05 Tex/appendix.tex
Final Changes

--- 2020-05-04 Tex/paper.tex
@@
 \documentclass[headsepline=true, abstracton]{scrartcl}
-%\usepackage[ngerman]{babel}
+
 \usepackage[utf8]{inputenc}
 \usepackage[T1]{fontenc}
-%\usepackage{authblk}
+
 \usepackage{amssymb}
@@
 \usepackage{bm}
-\usepackage{enumerate}
-\usepackage{verbatim}
-\usepackage[a4paper,text={160mm,215mm},centering,headsep=10mm,footskip=10mm]{geometry}
-\usepackage[urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black,bookmarks]{hyperref}
-\usepackage{aliascnt}
-\usepackage{lmodern}
-\usepackage{mdwlist}
+
 \usepackage{natbib}
-\usepackage{listings}
+
 \usepackage[table,xcdraw]{xcolor}
-\usepackage[activate]{pdfcprot}
+
 \usepackage{graphicx}
-\usepackage{slashed}
-\usepackage{mathrsfs}
+
 \usepackage{geometry}
 \usepackage{float}
-\usepackage{oldgerm}
+
 \usepackage{setspace}
-\usepackage{dsfont}
-%\usepackage{mathtools}
-\usepackage[all]{xy}
-\usepackage{cite}
-%\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{algorithm2e}
+
 \usepackage{url}
-\pagestyle{headings}
-\newcommand{\myclearpage}{\clearpage}
-\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
-\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
-\newenvironment{gelaber}{}{}
-\newenvironment{preamble}{}{}
-\newcommand{\tostar}{\overset{*}{\lower0.5em\hbox{$\smash{\scriptscriptstyle\rightharpoonup}$}}}
-\newtheorem{mydef}{Definition}
-\newtheorem{bem}{Bemerkung}
-\usepackage{tikz}
-\newcommand\circlearound[1]{%
-  \tikz[baseline]\node[draw,shape=circle,anchor=base] {#1} ;}
-  \usepackage{hyperref}
-\newcommand{\footremember}[2]{%
-    \footnote{#2}
-    \newcounter{#1}
-    \setcounter{#1}{\value{footnote}}%
-}
-\newcommand{\footrecall}[1]{%
-    \footnotemark[\value{#1}]%
-}
 
@@
 
-\title{\Large Generative Dynamics of Supreme Court Citations: \\ Analysis with a New Statistical Model}
+\title{\Large \textbf{Appendix:} \\ Generative Dynamics of Supreme Court Citations: \\ Analysis with a New Statistical Model}
 %\author{}
@@
 \maketitle
-\begin{abstract}
-\noindent The significance and influence of US Supreme Court majority opinions derive in large part from opinions' roles as precedents for future opinions.
-A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns.
-We raise two limitations of existing work on Supreme Court citations.
-First, dyadic citations are typically aggregated to the case level before they are analyzed.
-Second, citations are treated as if they arise independent of one another.
-We make the case that studying Supreme Court citations as a dynamic citation network overcomes both of these limitations.
-We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network.
-This methodology---the citation temporal exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation.
-We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015.
-We find evidence for dependence processes, including reciprocity, transitivity, and popularity.
-The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model.
-The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.\footnote{
-%Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. 
-This work was supported in part by NIH, 
-R01 AI36664-01. NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219. We would like to thank Rachael Hinkle and Benjamin Kassow for detailed comments on an earlier draft of this paper.}
-\end{abstract}
 
-\doublespacing
- \section{Introduction}
-
-United States Supreme Court opinions exercise authority and influence, in part, through their roles as precedents affecting future jurisprudence in the US.
-The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,gillman2001s,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained,hitt2016measuring}.
-Despite a considerable body of research that focuses on the way in which relevant precedents shape decision making on the Court, relatively little work has focused on understanding which past opinions are cited by an opinion.
-Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited in an opinion.
-We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{harris1982structural,caldeira1988legal,fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics,ethayarajh2018rose}.
-
-We are not the first to ask what predicts the citations in US Supreme Court Opinions.
-Indeed, a voluminous body of work has sought to explain how many times an opinion is cited \citep[e.g.,][]{cross2010determinants,benjamin2012standing,fix2019effect}, when an opinion is cited \citep[e.g.,][]{black2013citation,spriggs2001explaining}, and how many cases are cited by an opinion \citep[e.g.,][]{lupu2013strategic}---all focused on the US Supreme Court.
-One common feature of the research design in all of these studies is that the observations are at the case or case-year level.
-The outcome variables in these analyses are defined as measures of the number of citations to a case over a period of time, the number of citations to a case at a particular time, or a measurement on the cases cited by a case.
-These are case-level studies in that, based on the unit of analysis, it is impossible to determine both the origin and target case of a citation that contributes to the dependent variable.
-
-An alternative approach to case-level analysis of citations would be to model them in the directed dyadic form through which they arise.
-A case decided at time $t$ can cite (or not cite) each case decided previously, and in the US Supreme Court, each other case decided at time $t$.
-We are aware of one prior study, \citet{clark2010locating}, in which a statistical model is used to analyze directed dyadic citations between cases.
-However, \citet{clark2010locating} use a dyadic latent variable model in order to estimate ideal points for Supreme Court Opinions, but do not use this model to understand the relationships between explanatory variables and the formation of citation ties between opinions.
-We build upon the literature on citation analysis both methodologically and substantively.
-Methodologically, we develop a novel extension of a statistical model for networks, which we adapt to the network structural constraints of court citations.
-Second, we apply this methodology to a half-century of directed dyadic citations between U.S.
-Supreme Court opinions.
-
-There exist two broad benefits of conducting empirical analyses of citations at the directed dyad level.
-The first is that directed dyadic analyses can test both dyadic and case-level hypotheses.
-For example, case-level analyses can model whether opinions supported by a liberal majority coalition are more likely than those supported by a conservative majority coalition to be cited heavily in the future, but they cannot precisely model whether liberal cases will be cited more by liberal cases than by conservative cases.
-Thus, the first reason for analyzing citations at the dyadic level is to expand the set of hypotheses that can be represented in the model.
-The second reason for studying citations at the directed dyadic level is that, as articulated in the growing literature on legal citation networks, citations form complex networks in which a citation at one point in time may influence future citations.
-This phenomenon of complex dependence is very common in networks of many types, but processes specific to Supreme Court citations create interdependence in citations.
-For example, if opinion $i$ relies heavily on opinion $j$ as precedent, opinion $i$ is likely to discuss the legal basis for opinion $j$, and as a consequence, cite some of the opinions cited by opinion $j$.
-Suppose opinion $k$ is cited by opinion $j$.
-Opinion $k$ is more likely to be cited by opinion $i$ because opinion $i$ relies heavily on $j$, and opinion $j$ cites $k$.
-This is a special case of a very common process on networks referred to as ``triad closure''.
-Complex dependence is theoretically interesting on its own merits, but the effects of covariates cannot be reliably identified---either in terms of coefficient values or standard errors---without accounting for the interdependence inherent in networks \citep{desmaraisstatistical}.
-
-In this paper we develop a theoretical case that citations on the US Supreme Court are characterized by forms of complex dependence that are common in networks.
-We then develop an extension of a model---the exponential random graph model (ERGM)---that can incorporate both exogenous covariates and complex forms of interdependence into a directed dyadic analysis of citations.
-Finally, we develop and estimate a specification of this model in an analysis of US Supreme Court citations between 1950 and 2015.
-We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's fit.
-
-\section{Network Processes in Supreme Court Citations}
-
-When it comes to the development and testing of theory, the defining feature of networks is that the micro-level unit of analysis---the relationship between two entities (i.e., the citation from one opinion to another) is a component of a complex system of relations.
-The formation (or lack thereof) of that relationship cannot be fully understood without considering how the relationship fits into the system.
-Analytical designs that account only for covariates in explaining tie formation are incomplete theoretically, and, as a consequence, are subject to a form of omitted variable bias \citep{cranmer2016critique}.
-Citations in legal opinions are unique in terms of the windows into network dependencies offered by the texts of the opinions.
-A number of common structural dependencies that are found in networks are likely to apply to citations in Supreme Court opinions.
-In this section we present these dependence forms, and document the mechanisms by which they arise through archetypal passages in example opinions.
-
-We should note that we do not distinguish between positive and negative citations in this theoretical framework.
-The dynamics we outline are not specific to a particular type of citation.
-The only distinction we draw (in the empirical analysis) is the instance in which a case has been overruled.
-In the extreme instance of overruled precedents, we assume that (and test whether) a case is much less likely to be cited after it has been overruled.
-
-The first network property that we theorize in the context of Supreme Court citations is transitivity.
-In a network of directed relations (e.g., A cites B, but B doesn't cite A) transitivity refers to the tendency for A to send a tie to C if A sends a tie to B and B sends a tie to C \citep{holland1971transitivity}.
-In undirected networks, transitivity is simply the process by which friends of friends become friends (i.e., a friend of a friend is a friend).
-The term, ``transitive closure'' refers to a tie forming from A to C in response to extant ties from A to B and B to C.
-When writing opinions, Supreme Court justices present the legal bases for their rulings, which often involves discussing the most primary/relevant precedents underpinning these legal bases, but also the precedents and legal rules on which the primary precedents were based.
-This process of presenting several layers/levels of precedent in an opinion follows the structure of transitive closure exactly---opinion A cites opinion B as a primary precedent, and then cites opinion C because opinion B cites opinion C.
-The two examples presented below illustrate this process.
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{images/citations_trans.pdf}
-\caption{Illustration of transitive triangle connecting US Supreme Court opinions through citations.}
-\label{fig:transitivity}
-\end{figure}
-
-In the first example, a passage from Kansas v.
-Marsh (548 U.S.
-163, 2006)---a case considering the constitutionality of a death sentence statute in Kansas.
-In this example, the case Stringer v.
-Black is cited by Kansas v.
-Marsh as a case that is quoted by Sochor v.
-Florida.
-The primary precedent under discussion in this passage of the opinion is Sochor v.
-Florida, but Stringer v.
-Black is cited as a result of its role in the Sochor v.
-Florida opinion.
-\begin{quotation}
-The statute thus addresses the risk of a morally unjustifiable death sentence, not by minimizing it as precedent unmistakably requires, but by guaranteeing that in equipoise cases the risk will be realized, by ``placing a `thumb [on] death's side of the scale,' '' Sochor v.
-Florida, 504 U.
-S.
-527, 532 (1992) (quoting Stringer v.
-Black, 503 U.
-S.
-222, 232 (1992); alteration in original).
-\end{quotation} %https://supreme.justia.com/cases/federal/us/548/163/dissent2.html
-The second example, which we illustrate visually in Figure \ref{fig:transitivity} is a passage from Seminole Tribe of Fla. v. Florida, 
-(517 U.S. 44 1996)---a case addressing the rights of groups and citizens to sue states in federal court. In this example, Pennsylvania v. Union Gas Co
-(491 U.S.
-1, 1989) is the primary precedent being critiqued, and several cases are cited and discussed in terms of their roles as precedents in the Union Gas opinion.
-We highlight one---Parden v.
-Terminal R.
-Co., which is cited and discussed in both the Seminole and Union Gas opinions.
-\begin{quotation}
-Never before the decision in Union Gas had we suggested that the bounds of Article III could be expanded by Congress operating pursuant to any constitutional provision other than the Fourteenth Amendment.
-Indeed, it had seemed fundamental that Congress could not expand the jurisdiction of the federal courts beyond the bounds of Article III.
-Marbury v.
-Madison, 1 Cranch 137 (1803).
-The plurality's citation of prior decisions for support was based upon what we believe to be a misreading of precedent.
-See Union Gas, 491 U.
-S., at 40-41 (SCALIA, J., dissenting).
-The plurality claimed support for its decision from a case holding the unremarkable, and completely unrelated, proposition that the States may waive their sovereign immunity, see id., at 14-15 (citing Parden v.
-Terminal Railway of Ala.
-Docks Dept., 377 U.
-S.
-184 (1964)), and cited as precedent propositions that had been merely assumed for the sake of argument in earlier cases, see 491 U.
-S., at 15 (citing Welch v.
-Texas Dept.
-of Highways and Public Transp., 483 U.
-S., at 475-476, and n.
-5, and County of Oneida v.
-Oneida Indian Nation of N.
-Y., 470 U.
-S., at 252).'
-\end{quotation}% https://supreme.justia.com/cases/federal/us/517/44/case.html
-
-The second network property we consider in the context of Supreme Court citations is reciprocity.
-Reciprocity (also referred to as mutuality) is the tendency for node B to send a tie to node A in response to or coordination with A sending a tie to B \citep{garlaschelli2004patterns}.
-It is typically not possible for reciprocated ties to form in legal opinions.
-Most citations reference past opinions that were issued before the citing opinion's case was even argued before the Court.
-However, opinions written within the same Supreme Court term are often drafted in tandem, and can cite each other reciprocally.
-Opinion A citing opinion B within the same term represents a signal that opinion B is relevant to the legal reasoning underpinning opinion A.
-Unlike the citations themselves, the applicability of legal rules or lines of reasoning across cases is not directed---if A is relevant to B, B is highly likely to be relevant to A.
-We expect opinions written within the same term to exhibit a high degree of reciprocity.
-Below we provide two example passages from opinions that illustrate the phenomenon of within-term reciprocity.
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{images/citations_recip.pdf}
-\caption{Illustration of a reciprocal tie between two US Supreme Court opinions.}
-\label{fig:reciprocity}
-\end{figure}
-
-The first case in our example reciprocal dyad is a passage from Western Air Lines v.
-Criswell (472 U.S.
-400, 1985)---a case considering mandatory retirement in the context of age discrimination laws.
-The second case in the dyad, Johnson v.
-Mayor of Baltimore (472 U.S.
-353, 1985) is another case considering whether mandatory retirement violates the Age Discrimination in Employment Act.
-The mutual edge connecting these two cases is visualized in Figure \ref{fig:reciprocity}.
-These cases addressed very similar legal questions, which increased the likelihood that they would inform each other, and the opinions were written within the same term, which made it possible for them to cite each other.
-\begin{quotation}
-{\em From Western Air Lines:} On a more specific level, Western argues that flight engineers must meet the same stringent qualifications as pilots, and that it was therefore quite logical to extend to flight engineers the FAA's age 60 retirement rule for pilots.
-Although the FAA's rule for pilots, adopted for safety reasons, is relevant evidence in the airline's BFOQ defense, it is not to be accorded conclusive weight.
-Johnson v.
-Mayor and City Council of Baltimore, ante at 472 U.
-S.
-370-371.
-The extent to which the rule is probative varies with the weight of the evidence supporting its safety rationale and "the congruity between the .
-.
-.
-occupations at issue." Ante at 472 U.
-S.
-371.
-In this case, the evidence clearly established that the FAA, Western, and other airlines all recognized that the qualifications for a flight engineer were less rigorous than those required for a pilot.
-\\~\\
-{\em From Johnson:} The city, supported by several amici, argues for affirmance nonetheless.
-It asserts first that the federal civil service statute is not just a federal retirement provision unrelated to the ADEA, but in fact establishes age as a BFOQ for federal firefighters based on factors that properly go into that determination under the ADEA, see Western Air Lines, Inc.
-v.
-Criswell, post p.
-472 U.
-S.
-400.
-Second, the city asserts, a congressional finding that age is a BFOQ for a certain occupation is dispositive of that determination with respect to nonfederal employees in that occupation.
-\end{quotation} % https://supreme.justia.com/cases/federal/us/472/400/case.html \url{https://supreme.justia.com/cases/federal/us/472/353/case.html#370}
-
-The third, and final, network property we consider in the context of Supreme Court citations is popularity.
-Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent \citep{barabasi1999emergence}.
-Citations to an opinion signal both the Court's awareness of the legal reasoning of the case and the Court's evaluation that the opinion is an authoritative precedent.
-The more citations, the stronger this signal.
-Landmark cases, or those that establish new legal rules, are particularly authoritative and accrue citations from most future opinions that follow the respective line of reasoning.
-The passage below, from Oregon v.
-Mitchell (400 U.S.
-112, 1970)---a case on the legality of state age restrictions on voting in federal elections---illustrates this popularity dynamic.
-In this opinion passage Baker v.
-Carr is cited in reference to its role as a landmark precedent, and noted for the number of other cases by which it has been followed.
-and for which an authoritative opinion is referenced, and even discussed in terms of the number of other cases by which it was followed.
-The language in this passage suggests that the attention to Baker v.
-Carr in previous Court opinions is in part responsible for its authority in Oregon v.
-Mitchell.
-The citations to Baker v.
-Carr are visualized in Figrue \ref{fig:popularity}.
-\begin{quotation}
-The first case in which this Court struck down a statute under the Equal Protection Clause of the Fourteenth Amendment was Strauder v.
-West Virginia, 100 U.
-S.
-303, decided in the 1879 Term.
-[Footnote 2/1] In the 1961 Term, we squarely held that the manner of apportionment of members of a state legislature raised a justiciable question under the Equal Protection Clause, Baker v.
-Carr, 369 U.
-S.
-186.
-That case was followed by numerous others, e.g.: that one person could not be given twice or 10 times the voting power of another person in a state-wide election merely because he lived in a rural area..."
-\end{quotation} %https://supreme.justia.com/cases/federal/us/400/112/case.html
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width = 0.95\textwidth,trim= 0cm 3cm 6cm 8cm,clip=true ]{images/citations_pop.pdf}
-\caption{Illustration of ties sent to a landmark Supreme Court opinion via citations.}
-\label{fig:popularity}
-\end{figure}
-
-These three properties---transitivity, reciprocity, and popularity form the core of our network theory of U.S.
-Supreme Court citations.
-We also seek to model the effects of covariates (i.e., case features) on citation formation.
-In the next section, we introduce a modeling framework that can incorporate all of these effects into a single model, and adapt to the peculiar structure of citation matrices.
-
-\section{The Citation Exponential Random Graph Model}
-We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
-To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{welles2015individual,duque2018recognizing, graif2017neighborhood,osei2018elite}  %The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
-The network structures for which ERGMs are currently designed are insufficient to account for the structure of citation networks.
-
-\subsection{c-ERGM definition}
-
-We develop the citation ERGM (c-ERGM), to account for the structural constraints that apply to the network of Supreme Court citations.
-These structural constraints amount to three departures from the structure of networks for which ERGMs are currently designed.
-First, the citation network is partially acyclic.
-If two cases are decided during the same term, they can cite each other, forming a mutual edge (or two-cycle).
-However, if case $i$ is decided before case $j$, case $j$ can cite case $i$, but case $i$ cannot cite case $j$.
-Second, new edges can be created over time, but cannot be eliminated.
-Unlike in, e.g., an alliance network, in which two countries can dissolve an alliance, once a citation exists in a citation network it cannot be dissolved.
-Third, the set of nodes in the network must increase for new edges to be created.
-In conventional ERGMs, the number of nodes in the network can increase or decrease in each time period, and is typically stable over time.
-In a citation network, new edges (and non-edges) are introduced over time via the introduction of new nodes.
-The structure of the citation network is depicted in Figure \ref{fig:ctergm}---a hypothetical citation network established over three time periods, with three cases decided in each time period.
-We denote $C_t$ to be the set of citation and non-citations added to the network at time $t$ (i.e., via the addition of three cases), $C_{ <t}$ to be the citations among cases decided before time $t$, and $C_{ \leq t}$ to denote the entire set of citations and non-citations on which $C_t$ can depend through the c-ERGM specification.
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 0cm,clip=true  ]{images/daggish.pdf}
-\caption{Illustration of temporal structure of the Supreme Court Citation Network.
-$C_{\leq t}$ is the entire set of citations (and non-citations) on which citations and non-citations at time $t$ (i.e., $C_t$) depend.
-$C_t$ are conditioned on the citations and non-citations established before time $t$ (i.e., $C_{< t}$).
-The shaded small squares are hypothetical observed citations, and the white small squares are citations that could have been observed but were not.
-The regions of the citation matrix that are represented by large white rectangles are citations that could not have been observed since the citing case would have been decided in a term that preceded the term of the cited case.
-The citing case ID is given in the row and the prospective cited case is given in the column.
-}
-\label{fig:ctergm}
-\end{figure}
-
-The likelihood function of the c-ERGM is given by
-\begin{equation}
-l(\bm{\theta},C_{\leq t}) =  \frac{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right] }{ \sum_{C_t^* \in \mathcal{C}_t} \exp \left[ {\bm{\theta}'\bm{h}(C^*_{t},C_{<t}) }\right]  },
-\end{equation}
-where $\bm{\theta}$ is a vector of real-valued parameters, and $\bm{h}(C_{t},C_{<t})$ is a vector of scalar-valued functions that each quantify a feature of the citation network (e.g., the relationship between citation ties and a case attribute, the number of mutual edges in the citation network).
-$\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right]$ is a positive weight that is proportional to the probability of observing any particular form of the citations and non-citations added to the network at time $t$.
-The denominator of Equation 1 represents a normalizing constant, in which the positive weight is summed over all possible configurations of $C_{t}$, from the network in which the cases added to the network at time $t$ send no citations at all, to the network in which the cases added at time $t$ cite every possible case, and everything in between.
-%The probability of observing the citation network up to time $T$---the final time point in the dataset---is given by the product over the probability of adding the citations at each time point $t$, conditional on the citations that were added to the network previously.
-
-Though the likelihood function of the c-ERGM may appear quite different from that of statistical models conventionally used in political science, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients.
-$$ P(C_{ij,t} = 1 | C_{-ij,t}, C_{ < t}) = \frac{\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right]}{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right] + \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 0) } \right]}, $$ $$ \text{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~} = \frac{1}{ 1 + \exp \left[ - {\bm{\theta}'\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)} \right]}, $$
-where $C_{ij,t} = 1$ indicates that case $i$ cites case $j$, $C_{ij,t} = 0$ indicates that case $i$ does not cite case $j$, $C_{-ij,t}$ is the observed elements of $C_{t}$ except $C_{ij,t}$, and $\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)$ is the change in $\bm{h}(C_{t},C_{<t})$ that results from toggling $C_{ij,t} = 0$ to $C_{ij,t} = 1$.
-This re-arrangement illustrates that the parameters can be interpreted in terms of the change in the log odds of a citation from $i$ to $j$ given a one-unit increase in the corresponding element of $\bm{h}$, conditional on the other citations observed in the network.
-For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log odds of observing $C_{ij,t} = 1$ increases by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$).
-The logit form conditional probability is well known for the ERGM family \citep{goodreau2009birds}.
-
-We should note that there is nothing about the c-ERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another.
-For example, if we find a positive popularity effect, the term for that dynamic would indicate that there is an increase in the log odds that case $i$ cites case $j$ with each other case that has cited case $j$.
-That finding could be driven by one of two (or some combination of both) dynamics.
-First, it could be the case that a large volume of citations to a case in past opinions draws the attention of the justices to the ``popular'' case---as in the Baker v.
-Carr passage cited above.
-Second, it could be that we have omitted a case-level variable that is strongly associated with the number of citations accumulated by a case, and the number of citations in past cases is operating as a proxy for this unmeasured case characteristic.
-We cannot statistically differentiate between these two dynamics.
-However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-ERGM can be specified to account for the dependence in the data.
-
-\subsection{c-ERGM Estimation}
-
-The normalizing constant in Equation 1 is intractable.
-For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure \ref{fig:ctergm}---there are 16,777,216 unique configurations of $C_t$ that could be observed.
-The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases.
-This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the c-ERGM.
-
-The common alternative relies on Monte Carlo methods to approximate the normalizing constant by simulating a large set of networks \citep{hunter2006inference,hummel2012improving}.
-The resulting estimator, the Monte Carlo MLE (MCMLE), is approximately consistent, meaning that it converges to the MLE as the sample size, i.e., the number of simulated networks, increases.
-However, one drawback is that with the number of nodes in the Supreme Court citation network being in the order of $10,000$, obtaining the MCMLE is computationally expensive \citep{schmid2017exponential} and the success of the algorithm heavily relies on the starting parameter vector $\theta_0$, which is ideally chosen in the proximity of the unknown MLE \citep{hummel2012improving}.
-The prevailing choice for $\theta_0$ is the maximum pseudo-likelihood estimation (MPLE) \citep{strauss1990pseudolikelihood}, a fast estimation method that is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
-The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated above, assume a logit form.
-The MPLE is simple to obtain, but does not guarantee a starting value close to the MLE \citep{SchmidHunter2020}.
-
-We estimate the MCMLE for every term of the c-ERGM between 1950 and 2015.
-While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows.
-The growth in the network also serves as our reason for estimating a separate model for each term.
-First, the network is large, with each term introducing the potential for hundreds-of-thousands of new citations.
-This means we have the statistical power necessary to estimate a separate set of parameter values for each term.
-Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network.
-This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term.
-
-The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant.
-However, the estimation of most networks in the 90s with the MPLE as starting values was not feasible in an reasonable time frame anymore.
-Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$.
-But even this approach started to fail for networks around the turn of the millennium.
-Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}.
-The MCMLE for these large citation networks was obtained by setting the starting value according to an approach introduced by \citet{SchmidHunter2020}.
-This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$.
-However, the MPLE of networks with the same sufficient statistics is not necessarily the same.
-Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a network $C_t^*$ using simulated annealing \citep{Kirkpatrick83}, with $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$ and a weak dependence structure among unfixed ties.
-The MPLE of such a network is similar to the MLE, making it an effective starting value for the MCMLE algorithm.
-
-%Methods of Monte Carlo approximation are well established in the ERGM literature, \citep{hunter2006inference,van2009framework,hummel2012improving}, but with a number of nodes on the order of $10,000$, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}.
-However, because the network is large, we can rely on a fast estimation method---maximum pseudo-likelihood estimation (MPLE), which is consistent for discrete exponential family models \citep{nguyen2017near}, and has long been used with ERGM \citep{strauss1990pseudolikelihood}.
-The MPLE is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
-The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated above, assume a logit form.
-
-%One challenge that arises with the use of MPLE is that the usual estimate of the asymptotic covariance matrix (i.e., inverting the negated Hessian of the objective function at the maximum), results in standard errors that are biased downward.
-This leads to a high Type-1 error rate in hypothesis testing with MPLE if conventional standard errors are used \citep{van2009framework}.
-A form of non-parametric cluster bootstrap results in confidence intervals with correct coverage probability with MPLE \citep{desmarais2012statistical,desmarais2010consistent}.
-In the case of c-TERGM, the clusters are defined by $t$---the time points in which the cases are decided, and $T$ time-specific citation matrices $C_{t}$ are re-sampled with replacement to form a non-parametric cluster bootstrap sample.
-Confidence intervals are defined by the quantiles in the empirical distributions of estimates that result from re-estimating the MPLE on non-parametric cluster bootstrap samples.
-
- \section{Empirical Analysis}
-Our three data sources for this study include the Supreme Court Database (SCDB) \citep{spaeth2014supreme}, Martin-Quinn scores \citep{martin2002dynamic}, and Supreme Court citation data provided by the CourtListener Free Law Project \citep{CourtListener}.
-In the next section we explain the variables we construct using these data sources.
-We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015)\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the CourtListener data.
-We decided to exclude these 145 cases from our analysis.
-Additionally, since the most commonly used data on Supreme Court citations in political science come from \citet{fowler2007network}, we compared the Fowler data to the CourtListener data in the time interval that they overlap (1937--2001).
-We found considerable agreement---over 95\% of the citations recorded in the CourtListener dataset were also in the Fowler dataset, and over 96\% of the citations recorded in the Fowler data were also in the CourtListener dataset.}.
-The Supreme Court citation network from $1937 - 2015$ consists of $9,945$ cases.
-The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
-The network has a total of $111,986$ citation ties.\footnote{In order to focus on citation actions that were not intended to totally invalidate an opinion, we exclude $315$ citations that caused the cited case to be overruled.
-The data on overruling citation came from \citet{senate2016constitution}.
-Our results are virtually unchanged if we include the overruling citations.} The in- and outdegree distributions (i.e., the distributions of the number of citations sent and received by cases, respectively), are visualized in Figure \ref{degree_dist}.
-The maximum indegree (i.e., number of cases citing to a case) is $230$ and the maximum outdegree (i.e., number of cases cited by a case) is $162$.
-The majority of cases cite to and/or are cited by twenty or fewer other cases.
-
-\begin{table}[htp]
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
-\hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Terms} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Total Number Cases} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Cases/Term} \\ \hline
-CE Hughes*              & 1937 - 1941                                          & 628                                                               & 125.6                                                     \\ \hline
-HF Stone                & 1942 - 1946                                          & 756                                                               & 151.2                                                     \\ \hline
-FM Vinson               & 1946 - 1953                                          & 789                                                               & 98.63                                                      \\ \hline
-E Warren                & 1954 - 1969                                          & 2149                                                              & 126.41                                                     \\ \hline
-WE Burger               & 1970 - 1986                                          & 2805                                                             & 155.83                                                     \\ \hline
-W Rehnquist            & 1987 - 2005                                          & 2022                                                              & 106.42                                                      \\ \hline
-J Roberts **           & 2006 - 2015                                          & 796                                                              & 79.6                                           \\ \hline
-\end{tabular}
-\caption[caption]{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
-\\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 4/2018).}
-\label{tab:chiefs}
-\end{table}
-
-The degree distributions indicate that there is a long tail to both the number of citations sent and received.
-These long-tailed (i.e., high kurtosis) distributions provide preliminary evidence of substantial heterogeneity in the features that drive citations to and from cases \citep{strogatz2001exploring}.
-Figure \ref{fig:networkviz} displays the citation network network from 1937--2001.
-We see here that the densest rates of tie formation tend to be between consecutive courts (e.g., the Stone Court is much more tightly tied to the Hughes Court than the Rehnquist Court is to the Hughes Court).
-This pattern lends preliminary/descriptive support to the hypothesis that the rate of citations to a case decrease over time.
-
-\begin{figure}[htp]
-\includegraphics[scale=0.5]{degree_distribution}
-\caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2001. There are cases with an indegree >50, but they are not captured in this figure.}
- \label{degree_dist}
-\vspace{-.25cm}
-\end{figure}
-
-\begin{figure}[htp]
-\includegraphics[scale=0.35,clip=true,trim=.5cm 0cm 0cm 2cm]{images/citations1}
-\caption{ Supreme Court Citation Network, 1937-2001.
-Nodes are Supreme Court cases, with size based on incoming citations.
-Salient cases ({\em Oxford Guide}) are in yellow.
-Each network layer contains cases decided under a different chief justice.
-Because our data is temporally bound, cases for the Hughes and Rehnquist courts are not complete.
-}
- \label{fig:networkviz}
-\vspace{-.25cm}
-\end{figure}
-
-\subsection{c-ERGM specification}
-
-The c-ERGM we estimate includes two classes of terms---one that captures the effects of covariates on tie formation, and another that captures the complex dependence processes that we expect to observe in the Supreme Court Citation Network.
-We describe our model specification by defining the terms within these two classes.
-
-\subsubsection{Covariate terms}\label{covariate_terms}
-
-Covariate effects are accounted for in the c-ERGM via the term that is used to specify the effect of covariates in other ERGM family models, as
-$$h_{covariate}(C_t,C_{<t},X) =  \sum_{ij} C_{ij,t}X_{ij}.$$ Since $C_{ij,t}$ is a binary indicator of whether case $i$ cites case $j$, $h_{covariate}$ amounts to the sum of covariate values among directed dyads for which we observed a citation.
-The dyadic interpretation of the coefficient attributed to this term is the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in $X_{ij}$ (i.e., exactly the interpretation of the effect of a covariate in logistic regression).
-We include several exogenous covariates based on this standard formulation.
-
-The first covariate we incorporate into the model accounts for the degree to which cases cite those that are similar in terms of the ideological positions of the justices who supported the decision.
-We account for this effect following \citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
-\citet{clark2010locating} estimates a latent coordinate model of Supreme Court opinions based on the network of case-to-case citations.
-They find that the majority opinion falls at the ideal point of the median member of the majority coalition in the case.
-We include a covariate term in which $X_{ij}$ is the absolute difference between the Martin-Quinn scores \citep{martin2002dynamic} of the median justices in the majority coalitions for cases $i$ and $j$.
-We expect this variable to have a negative effect, which would correspond to cases citing those to which they are ideologically similar.
-%WE COULD USE THIS DATASET AS A SECOND APPLICATION
-
-%We include three sets of dummy variables that account for the issue areas of cases.
-In the first set of dummy variables---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$) (e.g., for the Criminal Procedure indicator, $X_{ij} = 1$ if case $i$ is a criminal procedure case, and $0$ otherwise).
-The second set of dummy variables represent receiver issue area intercepts in which $X^{ij}$ indicates the issue of the receiver case.
-The third set of dummy variables account for the rate of citation within each issue area and between each combination of issue areas.
-There is a separate dummy variable for each directed pair of issue areas.
-For example the A$\rightarrow$B dummy variable $X_{ij} = 1$ if case $i$ is issue area A and case $j$ is issue area B, and 0 otherwise.
-Note that, given the high dimension of these dummy variables, we omit one each of the sender and recipient indicators, and 28 of the 196 mixing rate indicators.
-Unfortunately, this leaves these indicators largely uninterpretable, but these variables are nonetheless important to absorb variation in citations that is attributable to the legal domains of two cases in a dyad.
-The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
-We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
-Table \ref{issue_area_coding} reports the issue areas covered in our data.
-
-We include one covariate that accounts for the rate at which cases that fall under the same issue area cite each other.
-The variable $X^{ij}$ indicates whether cases $i$ and $j$ share the same issue area, e.g., $X^{ij}=1$ if cases $i$ and $j$ are classified into the same issue area, and $0$ otherwise.
-The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
-We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
-Table \ref{issue_area_coding} reports the issue areas covered in our data.
-
-%We also include a set of covariates that account for justice-specific variation in citations.
-We include two types of terms at the justice level.
-First, we include an indicator variable that models the rate at which justices cite themselves.
-This effect is modeled with a single indicator variable in which $X_{ij} = 1$ if the majority opinions in cases $i$ and $j$ were written by the same justice.
-Second, we use two sets of dummy variables to account for differences across justices both in terms of their tendencies to include citations in the opinions they write and to be cited by other opinions.
-The justice sender dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $i$.
-The justice recipient dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $j$.
-This set of variables is motivated by the use of citations by \citet{kosma1998measuring} to measure the influence of individual Supreme Court justices.
-
-We also include an indicator variable that models the rate at which justices cite themselves.
-Similar to the same issue area variable, this effect is modeled with a single indicator variable in which $X_{ij} = 1$ if the majority opinions in cases $i$ and $j$ were written by the same justice.
-The data is taken from the SCDB \citep{spaeth2014supreme}.
-{\color{blue} Justification for including this variable?
-}
-
-\begin{table}[]
-\centering
-\begin{tabular}{llll}
-1 & Criminal Procedure & 8  & Economic Activity    \\
-2 & Civil Rights       & 9  & Judicial Power       \\
-3 & First Amendment    & 10 & Federalism           \\
-4 & Due Process        & 11 & Interstate Relations \\
-5 & Privacy            & 12 & Federal Taxation     \\
-6 & Attorneys          & 13 & Miscellaneous        \\
-7 & Unions             & 14 & Private Action      
-\end{tabular}
-\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
-\label{issue_area_coding}
-\end{table}
-
-We model the way in which citations to a case depend upon the age of a case.
-For this we use a second-order polynomial in which one covariate $X^{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X^{ij}$ is the squared age of case $j$ at the time that case $i$ is decided.
-We include these covariates because \citet{black2013citation} find that the number of citations to a Supreme Court case over time depends significantly on the age of the case, characterized by a sharp drop off and leveling out with age.
-
-\citet{benjamin2012standing} study the propensity for cases to be overruled and cited in other negative ways.
-They find that cases with majority coalitions that are large and ideologically broad are less likely to be cited negatively.
-In our data we do not differentiate between negative and positive citations, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis.
-We include one covariate in which $X^{ij}$ is the number of justices in the majority coalition for case $j$.
-We also include another covariate in which $X^{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$.
-We expect both of these covariates to have positive effects.
-
-The final control variable we include in the model is one in which $X_{ij} = 1$ if case $j$ was overruled prior to the term in which case $i$ was decided.
-This variable, quite simply, models the effect of being overruled on the rate of citation to a case after the overruling citation.
-\citet{fowler2008authority} find that citations to a case drop off quickly after the case has been overruled.
-
-\subsubsection{Dependence terms}\label{dependence_terms}
-
-We include dependence terms to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, and popularity.
-%We expect that the coefficients associated with each of these dependence effects will be positive.
-
-The common dependence effects for these four dynamics are the number of triangles (case $i$ cites cases $j$ and $k$, and case $j$ cites case $k$), the number of mutual dyads (case $i$ cites $j$ and vice versa), the number of out-2-stars (case $i$ cites $j$ and $k$), and the number of in-2-stars (case $i$ is being cited by cases $j$ and $k$) in $C_{\leq t}$.
-Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy \citep{Handcock.2003}.
-Model degeneracy is a common obstacle when fitting ERGMs, and occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network.
-In this scenario, the simulated networks are too different from the observed network, making the underlying distribution defined by the model extremely unrealistic.
-
-% reciprocity
-The statistic included in the model for reciprocity counts the number of mutual dyads (i.e., dyads in which cases $i$ cites case $j$, and case $j$ cites case $i$) in $C_{\leq t}$.
-In practice, this is the number of mutual dyads in $C_{\leq t}$, since mutual dyads must emerge among opinions that were written in the same term.
-The log odds interpretation of the reciprocity statistic is that if the opinion in case $i$ cites case $j$, the log odds that case $j$ also cites $i$ change by the estimated coefficient relative to the configuration in which case $j$ does not cite case $i$.
-We expect the reciprocity effect to be positive.
-
-% popularity
-Accounting for the popularity effect through the in-2-star statistic is prone to cause model degeneracy.
-In an attempt to stabilize the model against model degeneracy as well as to still capture the popularity effect, \citet{SnijdersTomA.B..2006} introduces the \textit{alternating-in-k-star} statistic, which was shown to be equivalent to the \textit{geometrically weighted indegree distribution} (gwidegree) statistic introduced by \citet{hunter2006inference}.
-Define $D_r(C_t, C_{<t})$ as the number of nodes with indegree $r, r \in \{0, \dots , N_t -1 \}$, where $N_t$ is the number of cases in the network at time $t$, then gwidegree represents a network's indegree distribution in a single statistic by geometrically weighting the degree distribution
-\begin{equation}
-h_{gwidegree}(C_{t}, C_{<t}, \lambda)= \lambda \sum_{r=1}^{N_t-1}\biggl(1-\bigl(1-\frac{1}{\lambda}\bigr)^r\biggr)D_r(C_t, C_{<t}). 
-\label{gwidegree}
-\end{equation}
-The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on $D_r(C_t, C_{<t})$ as $r$ increases, which means that the higher $\lambda$ the more the statistic responds to changes for cases with a high indegree.
-We fixed $\lambda=1$.
-Adding one single edge to the network changes the indegree distribution of the network so that a case with an indegree of $s$ turns into a case with an indegree of $s + 1$.
-With our notation this means that $D_r$ and $D_{r+1}$ get replaced by
-$D_r -1$ and $D_{r+1} +1$, while no changes for $D_\ell,
- \ell \in \{0, \dots , N_t -1 \} \setminus \{r, r-1\} $ are
-made. This changes the log odds in the following way
-$$\text{logit} P(C_{ij,t}=1 | C_{-ij,t}, C_{<t}) = \exp \bigl(~\theta \cdot( 1- \frac{1}{\lambda})^r \bigr).$$
-As stated by \citet{Levy2016}, this means for the interpretation that a negative parameter value indicates the centralization of edges (i.e., popularity) while a positive parameter indicates the dispersion of edges (i.e., new ties going to less popular nodes).
-Since we expect highly cited cases to be more likely to be cited again, we expect the gwidegree parameter to be negative.
-
-% Transitivity
-We include two statistics to account for different types of transitivity.
-The first transitivity statistic calculates the number of transitive ties, $i \to j$, where there is a directed path of length two from $i$ to $j$ through at least one case $k$, and $j$ and $k$ were written during a different term than $i$.
-We will refer to this statistic as the \textit{different term transitivity} statistic.
-As illustrated in Section 2, we theorize Supreme Court citations to pattern a high degree of transitivity.
-As a result, we expect positive parameter estimates for the \textit{different term transitivity} statistic.
-
-The second transitivity statistic, intended to capture clustering that includes more than one tie formed in the same term, is the \textit{geometrically weighted edgewise shared partners} (gwesp) statistic introduced by \citet{hunter2006inference}.
-For directed networks the edgewise shared $r$-partners statistic ($ESP_r$) can take multiple forms, but we integrate the definition where cases $i$ and $j$ that are connected by a citation, also cite $r$ common other cases.
-We follow the notation of \citet{Butts.2008} and call this ESP specification \textit{outgoing shared partners} (OSP).
-Similar to the gwidegree-statistic the gwesp-statistic incorporates a network's edgewise shared partners distribution by replacing $D_r$ by $ESP_r$.
-Triangles that are incident to the outgoing shared partners configuration arise when there is a cite between two cases decided at the same time, and those two cases cite one or more of the same third cases (decided either previously or also during the same term).
-This would happen if citations were clustering around legal issues that are not completely captured by our same-issue-area effects.
-As we do not expect that the issue area variable will account for all of the clustering around legal issues, we expect the gwesp statistic to also carry a positive coefficient value.
-
-% Activity
-We include one more structural network effect to adjust for inherent differences across cases in terms of the overall scope of the legal issues addressed in a case.
-\citet{fowler2008authority} finds that cases vary in the degree to which they serve as ``hubs''---citing to many other cases.
-We expect that new cases will be more likely to cite cases that themselves cited many cases, because cases that themselves cite a large number of cases address a broader array of legal issues.
-We include the outdegree of every node that entered the network prior time $t$ as a receiver effect.
-We expect the coefficient on this term to be positive.
-This is not technically a dependence term, as it is formulated as a covariate, but we discuss it in the current section because it is purely a network structure effect.
-
-%In order to account for the popularity effect, we add the \textit{geometrically weighted indegree distribution} (gwidegree) statistic to the model.
-The geometrically weighted distribution statistics were introduced by \citet{Hunter.2006} to counteract model degeneracy.
-
-%The log odds interpretation of the triangle statistic is that  $\theta$ gives the change in the log odds of case $i$ citing case $j$ associated  with each additional triangle that would be created by case $i$ citing case $j$ (e.g., if opinions in cases $i$ and $j$ cite one common case $k$, the log odds that case $i$ cites case $j$ change by theta relative to the configuration in which cases $i$ and $j$ do not cite any of the same cases).
-
- % \subsubsection{Dynamics in effects}
-
-% For the current model this section is not relevant anymore
-
-\subsection{Model Fit}
-
-Our case for studying legal citations at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence.
-To quantify this contribution, we fit one model that incorporates covariate effects only, which we term the {\em Independent Model}---excluding the reciprocity, transitivity, and popularity terms.
-The full model is the one we present below.
-We use both AIC and BIC to compare the fit of the two models, with lower values indicating better fit \citep{ripplinger2008does}.
-
-  \begin{figure}[t]
-\includegraphics[width=15cm]{SCC_AIC_BIC}
-\caption{AIC and BIC for the full and the independent model for the time frame 1950-2015 }
- \label{AIC_BIC}
-\vspace{-.25cm}
-\end{figure}
-
-Figure \ref{AIC_BIC} depicts the AIC and BIC comparison for the two different models for each term from 1950 to 2015.
-We see that the AIC and BIC is considerably smaller for the full model throughout the period of consideration.
-On average the independent model yields a 5,525 point higher AIC and an average 5,582 point higher BIC.
-The maximum difference is reached in 1983 with a difference of 10,990 points for the AIC and 10,930 points for the BIC.
-This fit comparison provides robust evidence that the dependence terms contribute significantly to the explanatory power of the model.\footnote{In the appendix we present conventional goodness-of-fit plots that are used to assess the structural fit for ERGM \citep{hunter2008goodness}, as well as degeneracy diagnostics \citep{mukherjee2020degeneracy}.
-We find that the full model fits the degree and edgewise shared partner distributions well, and is not degenerate.}
-
-%In our prediction experiment we randomly split the directed dyads into an 80\% training set and a 20\% test set.
-The parameters of the model are estimated using the directed dyads in the training set, and the parameters are used to form the conditional probability of a tie for all of the directed dyads in the test set.
-Directed dyads for which the conditional probability of a tie exceeds 0.5 are predicted to be citations.
-The experiment is run with the full model, and with a model that excludes all of the dependence terms (i.e., all terms involving reciprocity, in and out stars, and triangles)---the independent dyads model.
-We run this experiment for 10 iterations.
-Predictive performance is evaluated with three common and related measures---precision (i.e., the proportion of predicted citations that are actually citations), recall (the proportion of actual citations that are predicted to be citations), and the F1 score (the harmonic mean of precision and recall) \citep[see, e.g.,][for discussion of these measures]{makhoul1999performance}.
-All three measures are bounded between 0 and 1, with higher scores indicating better performance.
-
-%\begin{table}[H]
-%\centering
-
-%\begin{tabular}{rllll}
-%\hline \hline
-%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
-% & mean & range & mean & range \\ 
-%  \hline
-%precision & 0.6007 & (0.5934, 0.6074) & 0.8666 & (0.8647, 0.8701) \\ 
-%  recall & 0.1522 & (0.1495, 0.1549) & 0.5969 & (0.5945, 0.6019) \\ 
-%  F1 score & 0.2429 & (0.2389, 0.2462) & 0.7069 & (0.7048, 0.7107) \\ 
-%   \hline \hline
-%\end{tabular}
-%\caption{The predictive performance of the directed dyadic models with linear trends, over ten 80/20 train/test splits.}
-%\label{tab:prediction_linear}
-
-%\end{table}
-
-%We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms.
-The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set.
-This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.
-
-%\begin{table}[H]
-%\centering
-%\begin{tabular}{rllll}
-%\hline \hline
-%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
-% & mean & range & mean & range \\ 
-%  \hline
-%precision & 0.5938 & (0.5844, 0.5996) & 0.8668 & (0.8653, 0.8693) \\ 
-%  recall & 0.1551 & (0.1515, 0.1575) & 0.5969 & (0.5937, 0.6005) \\ 
-%  F1 score & 0.246 & (0.2406, 0.249) & 0.707 & (0.7045, 0.7095) \\  
-%   \hline \hline
-%\end{tabular}
-%\caption{The predictive performance of the directed dyadic models with no trends, over ten 80/20 train/test splits.}
-%\label{tab:prediction_notrend}
-%\vspace{0.4cm}
-%\begin{tabular}{rllll}
-%\hline \hline
-%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
-% & mean & range & mean & range \\ 
-%  \hline
-%precision & 0.6209 & (0.611, 0.6351) & 0.8669 & (0.8612, 0.8711) \\ 
-%  recall & 0.1511 & (0.1451, 0.1558) & 0.5962 & (0.5884, 0.6011) \\ 
-%  F1 score & 0.243 & (0.2347, 0.2486) & 0.7065 & (0.7011, 0.7103) \\ 
-%   \hline \hline
-%\end{tabular}
-%\caption{The predictive performance of the directed dyadic models with quadratic trends, over ten 80/20 train/test splits.}
-%\label{tab:prediction_quadratic}
-%\end{table}
-%We also conduct a predictive experiment for the model that does not consider any trends and for the model that does consider quadratic instead of linear trends.
-The predictive performance results are given in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic}.
-By comparing the full model results in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic} with the results of the full linear trend model in table \ref{tab:prediction_linear} we can conclude all three models perform similarly well.
-\\[0,5cm]
-
-%\textit{Comment from Benjamin Kassow: In section 6.2 (predictive performance), I like what is here, but was hoping to see a bit more discussion about the predictive 
-%performance of the model, with some explanation of what the predictive performance means. 
-%I understand it (I think), but depending on where this gets sent to, I am not sure if all reviewers will. 
-%I also think tying this into the overall model explanation or discussion of model performance overall will be helpful for better 
-%integrating this section into the rest of the manuscript.}
-
-\subsection{c-ERGM Coefficient Estimates}
-The estimates from the c-ERGM are presented in Figures \ref{SCC_results_1} and \ref{SCC_results_2}.
-In these figures we depict coefficient estimates and 95\% confidence intervals for all of the effects for all the variables we discussed in sections \ref{covariate_terms} and \ref{dependence_terms}.
-The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in the respective variable.
-We first discuss the dependence effect estimates in Figure \ref{SCC_results_1}.
-A circle indicates that the estimate was significant at an $\alpha=0.05$ level, while and square translates to a p-value between $0.05$ and $0.1$.
-A triangle signals that the estimate for a given year was not significant at an $\alpha=0.1$ level.
-The panel showing the results of the reciprocity statistic does not provide estimates for every term.
-In these missing terms there were no two cases that cited each other, making the estimate of the reciprocity statistic term equivalent to $-\infty$.
-For those years, the reciprocity term is included in the model, but the estimate of $-\infty$.
-A similar scenario is visible for the \textit{Overruled Cases} variable.
-In the time period from 1953-1956 no cases that have been overruled have been cited which results in a non-estimable effect.
-
-We first discuss our findings regarding the dependence terms.
-In most of the years that it is estimable, the reciprocity effect is statistically significantly positive, and substantial in magnitude.
-In a typical year, the presence of a citation from case $i$ to case $j$ increases the log-odds of a citation from case $j$ to case $i$ by approximately 0.50.
-Both transitivity effects---different term transitivity, and GWESP---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect.
-The log-odds of a citation from case $i$ to case $j$ increases by more than 2 if there is at least one third case, $k$ that is cited by $i$, and cites to $j$.
-Results regarding the popularity effect (i.e., GWIdegree) are generally supportive of the hypothesis of a popularity effect in Supreme Court citations, as the GWIdegree estimate is negative in every term.
-There are some terms for which the result is not statistically significant at conventional levels.
-The popularity effect looks to have grown substantially stronger after the 2000 term.
-Contrary to expectations, we find that the hub effect is negative.
-That is, each additional citation sent by a case reduces the likelihood that the case is cited.
-For every citation sent, the log-odds of any citation to a case decreases by approximately 0.025---a result that is statistically significant in each term.
-
-The effects of the exogenous covariates included in the c-ERGM are presented in Figure \ref{SCC_results_2}.
-Most of the exogenous covariate effects align with expectations.
-We find that cases are more likely to cite those that (1) have been decided most recently, (2) were decided by large majorities, (3) have not been overruled, (3) were authored by the same justice, and (3) are in the same issue area.
-Surprisingly, we do not find consistent effects of either of the covariates that involve justice ideology.
-The ideological distance between two cases, as measured by the absolute difference in the median of the majority coalitions; and the ideological breadth of the cited cases, as measured by the absolute difference between the most liberal and most conservative justices on the potential receiving case, both exhibit signs and significance levels that are highly volatile.
-Our results leave open the question of whether, and if so, how, citations are shaped by ideological factors.
-
-%The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes.
-The scales of covariate and dependence effects are comparable.
-%The simple linear trends represent one limitation of the current analysis.
-In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.
-
-%\begin{figure}[H]
-% \begin{tabular}{cc}
-
-%   (a) Reciprocity & (b) Transitivity \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
-
-%    (c) Activity & (d) Popularity \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
-
-%\end{tabular}
-%\caption{Trending effects for dependence effect estimates from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
-% \label{fig:coeftrends_endo}
-%\vspace{-.25cm}
-%\end{figure}
-
-%  \begin{figure}[H]
-%  \begin{tabular}{cc}
-% (e) MQ Score Difference & (f) Ideological Breadth \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\
-
-%  (g) Same Issue & (h) Majority Size \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\
-
-% (i) Difference in Years & (j) Difference in Years$^2$ \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\
-
-% (k) Overruled Cases & (l) Justice Homophily \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} & 
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/justicehomophily_coef_trend_linear.pdf} \\
-
-% \end{tabular}
-%\caption{Trending effects for exogenous variables from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
-% \label{fig:coeftrends_exo}
-%\vspace{-.25cm}
-%\end{figure}
-
-%\begin{itemize}
-%\item visualize Sender/receiver Issue Area Results
-%\item heatplot for issue area nodemix
-%\item visualize Sender/receiver for majority opinion writer
-%\item add Network effect plots for model without network statistics (add to existing plots?)
-%\end{itemize}
-
-%\newpage
-%   \thispagestyle{empty}
-
-  \begin{figure}[H]
-\includegraphics[width=16cm]{SCC_Results_1}
-\caption{ERGM results for the dependence terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1. }
- \label{SCC_results_1}
-
-\vspace{-.25cm}
-\end{figure}
-
-% \thispagestyle{empty}
- \begin{figure}[H]
-\includegraphics[width=16cm]{SCC_Results_2}
-\caption{ERGM results for the covariate terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1.}
- \label{SCC_results_2}
-\vspace{-.25cm}
-\end{figure}
-
-\section{Conclusion}
-
-We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network.
-This methodology---the citation-ERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation.
-We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015.
-We find, somewhat counterintuitively, that Supreme Court citations are highly reciprocal.
-We also find that citations are driven by dependencies such as triad closure and popularity.
-The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model.
-The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.
-
-We recognize two important limitations of the current analysis, which should be addressed in future iterations of this work or in future studies.
-First, our theoretical claims and analyses treat all citations---positive and negative---as equivalent.
-Both data and methodological limitations will need to be overcome to consider these two citation types.
-Second, there is nothing in the c-ERGM that governs the generation of new cases (i.e., cases are assumed to arise exogenously, unrelated to the history of the citation network).
-As new cases are often taken up by the Supreme Court in order to test, clarify, and/or challenge existing precedents, it is likely inappropriate to assume that new cases arise exogenous to the existing citation network.
-Overcoming this limitation would require methodological innovation in the c-ERGM as well as theoretical development regarding exactly how citation structure might predict case emergence.
-%Third, the data used in this study does not cover the most recent history of the Court.
-Our data ends in 2001---missing the tenure of the most recent chief justice, and more than fifteen terms.
-To overcome this limitation we would need to extend the citation data beyond that collected by Fowler and colleagues.
-
-\section{Appendix}
-
-\subsection{Goodness-of-Fit}
-We evaluate the goodness-of-fit of the model following \citet{Hunter.2008} by examining the distribution of four hyper statistics, e.g., the out- and indegree distribution and the distribution for two different edgewise shared partners statistics.
+\section{Goodness-of-Fit}
+We evaluate the goodness-of-fit of the model following \citet{hunter2008goodness} by examining the distribution of four hyper statistics, e.g., the out- and indegree distribution and the distribution for two different edgewise shared partners statistics.
 OTP stands for \textit{outgoing two-paths} and refers to the number of cases $r$ that are cited by case $i$ and that cite case $j$, while $j$ is also directly cited by $i$.
@@
 
-\subsection{Checking for Model Degeneracy}
+\section{Checking for Model Degeneracy}
 
@@
 In a non-degenerate model the statistics of the networks that were simulated from the probability distribution defined by the MCMLE fall in the proximity of the observed network's statistics.
-Figures \ref{mcmcdiagnostics_1950} and \ref{mcmcdiagnostics_1950} depict trace and density plots for the dependence terms in the 1950 and 2015 term citation network.
+Figures \ref{mcmcdiagnostics_1950} and \ref{mcmcdiagnostics_2015} depict trace and density plots for the dependence terms in the 1950 and 2015 term citation network.
 The histograms on the left visualize a statistic's density from $1000$ simulated networks, while the right side shows the statistic's trace plot of the same $1000$ networks.
@@
 The density and trace plots for the ERGM of the terms not depicted provide similar results.
+
+\bibliography{bib} 
+\bibliographystyle{apsr}
 
@@
 \end{figure}
-%\begin{itemize}
-%\item add robustness check: MCMLE for each year -> plot time trend of coefs -> are results similar to the MPLE time trends presented in the paper?
-%\item (investigate which statistic contributes most on recall improvement)
-%\end{itemize}
-
-\newpage
-
-\bibliography{bib} 
-\bibliographystyle{apsr}
-
- \begin{figure}[H]
-  \begin{tabular}{c}
-\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_activity}
- \\
-\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_popularity}
-\end{tabular}
-\caption{Issue Area Popularity and Activity (nodeifactor and nodeofactor). Baseline is Issue Area 'Criminal Procedure'.}
-\end{figure}
-
- \begin{figure}[H]
-\includegraphics[width=15cm]{heatplot_issue_area}
-\caption{Nodemix Issue Area. Sender14Receiver14 and Sender14Receiver13 couldnt be estimated.}
- \label{number_supporting}
-\vspace{-.25cm}
-\end{figure}
 
 \end{document}
-
-% Network Statistics
-The c-TERGM is specified through 
-the decision regarding which network statistics $h(\cdot)$ to incorporate. We include the following statistics for the Supreme Court citation network:
-$$h_{edges}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{i=1}^Nc_i(t)$$
-the number of citations made at time $t$.  $h_{edges}$ is like an intercept, and models the expected value of any given edge.
-$$h_{outstar}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
-the number of weighted outstars occuring at time $t$. We argue that it should be more likely to cite more recent cases than cases that have been decided further in the past. For the weight 
-$$w(a,b):= \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
-we define $a$ and $b$ as the elapsed time since case $i$ and $j$ have been introduced to the network.
-$$h_{triangle}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot
-c_j(t_{-i}) \cdot w(a,b)$$
-where $c_j(t_{-i})$ indicates whether case $j$ was cited at the time case $i$ was introduced into the network.
-Just as for the outstar statistic, we include a weighting factor to favor more recent cases.
-\\[0.3cm]
-
-%Change Statistic
-The individual entries $c_i(t)$ can be taken as a manifestation of single Bernoulli variables $C_i(t)$.
-This interpretation allows the following calculation regarding the conditional distribution of $C_i(t)$:
-%
-\begin{eqnarray*}
-\dfrac{P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~|~ C_i(t)^c=c_i(t)^c)} &=&
-\dfrac{P_{\theta}(C_i(t)=1 ~,~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~,~ C_i(t)^c=c_i(t)^c)} \\
-                           &=&\dfrac{P_{\theta}(C(t)= c_i^+(t))}{P_{\theta}(C(t)=c_i^-(t))}\\
-                           &=&\dfrac{\exp(\theta^T \cdot h(c_i^+(t)~|~c(t-)))}{\exp(\theta^T \cdot h(c_i^-(t)~|~c(t-)))}\\
-                           &=& \exp(\theta^T \cdot (h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-)))
-\end{eqnarray*}
-%
-This implies the following equation:
-%
-\begin{equation}
-\text{logit}(P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c))= \theta^T \cdot (h(c_i^+(t)~~|c(t-)) - h(c_i^-(t)~|~c(t-)))
-\label{Logit}
-\end{equation}
-In the equation above the following notations were used:
-%
-\begin{itemize}
-\item $c_i^+(t)$ emerges from $c(t)$, while assuming $c_i(t)=1$
-\item $c_i^-(t)$ emerges from $c(t)$, while assuming $c_i(t)=0$
-\item The condition $C_i(t)^c=c_i(t)^c$ is short for: $C_j(t)=c_j(t)$ for all $j\in \{1,\dots,N\}$ with $i \neq j$
-\item The expression $(\Delta c_i)(t):=h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-))$ is called the \textit{change statistic}.
-The $k$th component of $(\Delta c_i)(t)$ captures the difference between citation networks $c_i^+(t)$ and $c_i^-(t)$ on the $k$th integrated statistic in the model
-\end{itemize}
-
-\section{Estimation}
-
-% Pseudo Likelihood
-\subsection*{Maximum Pseudo-Likelihood Estimator}
-One can assume that the dyads are independent of each other, which means that
-the random variables $C_i(t)$ inside the random vector $C(t)$ are independent of each other.
-In this case, the equation (\ref{Logit}) reduces to
-$$logit(P_{\theta}(C_i(t) = 1)) = \theta^T \cdot (\Delta c_i)(t)$$
-This corresponds with the logistic regression approach, where the observations of
-the dependent variables are simply edge values of the observed citation vector,
-and the observations of the covariate values are given as the scores of every single
-change statistic. Therefore, the resulting likelihood function is of the following form:
-\begin{equation}
-\text{lik}(\theta)= P_{\theta}(C(t)=c(t))= \prod_{i} \dfrac{ \exp \left(\theta^T \Delta(c_i)(t) \right)}{1+\exp \left(\theta^T \Delta(c_i)(t) \right)}
-\label{PseudoLik}
-\end{equation}
-
-\subsection*{Maximum Likelihood Estimator}
-The more rigorous technique is to estimate the parameters directly with the log-likelihood function derived from (\ref{ercm}), which has the following form:
-%
-\begin{equation}
-\text{loglik}(\theta)=\theta^T \cdot h(c(t)| c(t-))-\log(\kappa(\theta))
-\label{loglik}
-\end{equation}
-%
-The problem resulting from estimating the parameters with (\ref{loglik}) is that the term
-%
-$$\kappa(\theta):= \sum_{c(t)^* \in \mathcal{C}(N)} \exp(\theta^T \cdot h(c(t)^*|c(t-)))$$ 
-%
-which sums up the weighted statistics of all possible binar vectors of length $N$, has to be evaluated.
-However, the cardinality of $\mathcal{C}(N)$ (\#$(\mathcal{C})=2^N$) is incredibly large and a direkt calculation of this sum is for already small $N$ not feasible.
-\\[0.3cm]
-An solution for this limitation is based on the following consideration: Fix a vector of parameters $\theta_0 \in \Theta$ from the underlying parameter range $\Theta$ and compute for $\theta \in \Theta$ the expected value
-%
-\begin{eqnarray*}
-\mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right]&=& 
-\sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot \mathbb{P}_{\theta_0}(C(t)=c(t))\\
-&=& \sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot 
-\frac{\exp(\theta_0^T \cdot \Gamma(c(t)))}{\kappa(\theta_0)}\\
-&=& \frac{1}{\kappa(\theta_0)} \sum_{c(t) \in \mathcal{C}(N)}\exp\left(\theta^T \cdot \Gamma(c(t))\right)\\
-&=&\frac{\kappa(\theta)}{\kappa(\theta_0)}
-\end{eqnarray*}
-%
-This equation offers the following possibility: If one draws $L$ random vectors $c^{(1)}(t), \dots ,c^{(L)}(t)$ out of a distribution $\mathbb{P}_{\theta_0}$ appropriately, one gets with the law of big numbers and a big enough sample $L$ the following relation:
-%
-\begin{equation}
-\frac{1}{L}\cdot \sum_{i=1}^{L}  \exp\left((\theta - \theta_0)^T \cdot \Gamma(c^{(i)}(t))\right)
-~~\approx~~ \mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right] = \frac{\kappa(\theta)}{\kappa(\theta_0)}
-\label{konver}
-\end{equation}
-%
-This approximate can then be used to approximate the log likelihood function.\\[0.4cm]
-Next, we will discuss how a sufficient number of suitable drawings $c^{(1)}(t), \dots ,c^{(L)}(t)$ can be sampled from the distribution $\mathbb{P}_{\theta_0}$. \\
-For this purpose, the Markov Chain Monte Carlo (MCMC) methods can be used.
-\subsection*{Gibbs sampling for the ERCM}
-To be able to compute the approximate likelihood function one needs a sufficiently large number of random vectors from the distribution $\mathbb{P}_{\theta_0}$.
-Snijders \citet{Snijders.2002b} introduces an approach to sample random networks for the ERGM framework by using \textit{MCMC methods}.
-We adapt this approach for sampling appropriate binary vectors for the ERCM.
-\\[0.4cm]
-\textit{Gibbs sampling}\\
-%Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g.
-observed vector).
-Afterwards, the length $L$ of the respective sub-sequence is determined.
-For $k \in \{0,...,L-1\}$ execute the following steps recursively (here the vector in its $k$th iteration is denoted as $c^{(k)}(t)$):\\
-
-\begin{algorithm}[H]
- Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g. observed vector)\\
- \For{i in 1:N}{
-  Compute $\pi:= \cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$\\
-  Draw a random number Z from Bin(1,$\pi$)\\
-  \eIf{Z=1}{
-   set $c^{(k+1)}_i = 1$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
-   }{
-   set $c^{(k+1)}_i = 0$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
-  }
- }
- Start all over using $c^{(k+1)}$\\[0.3cm]
- \caption{Simulation of vectors of $\mathbb{P}_\theta$ using Gibbs sampling}
-\end{algorithm}
-\vspace{0.5cm}
-%\begin{enumerate}
-%\item Randomly choose a number $i \in \{1,\dots, N\}$
-%\item Compute using the likelihood the value 
-%$$\pi:= \mathbb{P}_{\theta}(C_{i}(t)=1 | C_{i}^c(t)=(c_{i}^{(k)}(t))^c)=\cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$$
-%\item Draw a random number $Z$ from Bin$(1, \pi)$. %If
-%\begin{itemize}
-%\item $Z=0$, define $c^{(k+1)}(t)$ via
-%$$c_{p}^{(k+1)}(t)=\begin{cases}
-%0& \text{if}~ p=i \\
-%c_{p}^{(k)}(t) &\text{if}~ p \neq i 
-%\end{cases}$$
-%\item $Z=1$, define $c^{(k+1)}(t)$ via
-%$$c_p^{(k+1)}(t)=\begin{cases}
-%1& \text{if}~ p=i \\
-%c_p^{(k)}(t) &\text{if}~ p \neq i 
-%\end{cases}$$
-%\end{itemize}
-%\item Start at step 1 with $c^{(k+1)}(t)$.
-%\end{enumerate}
-\noindent Using the depicted algorithm, a sequence of random vectors $c^{(0)}(t),...,c^{(L)}(t)$ can be simulated.
-Since the orignial vector was chosen randomly and the first simulated vectors are very dependent on the chosen mvector (only one entry is changed per iteration!), usually the first $B$ vectors, where $N \ll B \ll L$, are discarded as the so called \textit{Burn-In}.
-\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
-Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ to start with (e.g., the observed vector). For $k \in \{0,...,L-1\}$ recursively proceed as follows:\\
-\begin{enumerate}
-\item Randomly choose a number $i\in \{1,\dots, N\}$
-\item Compute, using the equation (\ref{Logit}) the value
-$$\pi := \dfrac{\mathbb{P}_{\theta}(C_{i}(t) \neq c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}{\mathbb{P}_{\theta}(C_{i}(t) = c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}$$
-\item Define $\delta:= \min\{1, \pi\}$ and draw a random number $Z$ from Bin$(1, \delta)$. If
-\begin{itemize}
-\item $Z=0$, let $c^{(k+1)}(t) := c^{(k)}(t)$ 
-\item $Z=1$, define $c^{(k+1)}(t)$ as
-$$c_{p}^{(k+1)}(t)=\begin{cases}
-1-c_{p}^{(k)}(t)& \text{if}~ p=i \\
-c_{p}^{(k)}(t) &\text{if}~ p \neq i 
-\end{cases}$$
-\end{itemize}
-\item Start at step 1 with $c^{(k+1)}(t)$.
-\end{enumerate}
-The first $B \ll L$ vectors are discarded as Burn-In.
-
-First, as these arcs are observed patterns of behaviour, they can be used to infer latent characteristics of nodes or the relations between nodes that produced them \citep{batagelj2003efficient}.
-Moreover, since each judgment is embedded in the broader network of citations, which influences the judgment beyond the characteristics of the two directly linked documents, inferences about the neighborhood of these two documents are possible as well.
-A common application of this general logic is to obtain local or global network indices, such as Kleinberg centrality at the node level \citep{kleinberg1999authoritative,fowler2007network} or global hierarchy \citep{mones2014universal}, then use them to draw conclusions about networks or compare networks across contexts such as time periods \citep[e.g.][]{vazquez2001statistics,fowler2008authority,greenberg2009citation,lupu2012precedent,lupu2013strategic,dawson2014current,jaffe2017patent}.
-
-	Still within the category of using observed network constellation to make inferences about components of the system, identification of communities using methods for clustering, such as modularity maximization \citep[e.g.][]{kajikawa2007creating,shibata2011detecting,chen2010community} and stochastic block modeling \citep{jo2009citing}; and subnetwork identification \citep[e.g.][]{batagelj2017emergence} have been used to analyze relevance of topics within an academic field or to determine trends in technological advancement \citep{verspagen2007mapping,erdi2013prediction}.
-Main path analysis \citep{hummon1989connectivity}, which determines the main path through an acyclic network, is a particularly useful way to examine trends as it allows researchers identify structures of knowledge flow.
-Other methods include classifying documents into groups based on similarities in their citation profiles of cites over time, which can be examined as a function of time to see if there are temporal patterns \citep{leicht2007large}.
-The identification of communities is often coupled with classification of these communities by the researcher using historical knowledge.
-
-	\section{Extra Notes}
-	This section contains some of the extra notes I cut out from the draft that might be useful.
-	\subsection{Broader definitions of citation networks}
-	More broadly, citation networks can have as nodes the document producers, including scientists \citep{ji2016coauthorship}, judges \citep{landes1998judicial}, journals \citep{rice1988citation}, or aggregated units \citep{gelter2012networks,pan2012world}.
-In these cases, which are essentially exercises in aggregation, the resulting network can take on different characteristics such as weighted or cyclic edges.
-Other networks derived from the raw citation network is the co-citation or co-cited-by networks whereby two documents are connected if they co-cite or are co-cited by another document \citep{van2005reference}.
-
-	\subsection{Weighted edges and multiplexity}
-	Recognition, and subsequent quantification, of different levels of complexity in citation arcs allows for a host of advancements in citation network analysis.
-Weights can indicate relevance \citep{liu2014citations} or be used to incorporate the temporal dimension into citation networks \citep{fujita2014detecting}.
-Moving to multiplexity leads to further advancements.
-For example, \citet{greenberg2009citation} accounted for level of support and through simulations based on arc typed switching demonstrated how citation bias against critical articles yielded differences in beliefs within a medical discipline.
-\citet{bommarito2010distance} argued that articles contain different types of information and accounting for the specific piece of information that resulted in the cite can yield clustering that are more interpretable.
-In all cases however, substantial work has to be done into reclassifying the weightedness or multiplexity of citation arcs \citep{zhang2007semantics}.
-
-	\subsection{Misc. notes}
-	\begin{itemize}
-		\item To control for age, \citet{clough2015transitive} propose transitive reduction of citation networks (i.e.
-removal of redundant information ties) as it will primarily remove citation arcs that are disparate in age.
-		\item Most citation networks are directed acyclic graphs, but not all.
-Different versions of the same document can also cause problems (such as strong network components).
-The ``preprint transformation'' is a solution to small strong components \citep{batagelj2017emergence}.
-		\item Date of publication can be assigned as level to nodes in acyclic networks \citep{batagelj2017emergence}.
-Depending on the citation network at hand, temporal ordering can be difficult, as documents can share publication dates.
-To overcome this, begin each iteration by sampling from one of the possible orderings.
-\citep{carstens2016topology}.
-	\end{itemize}
-
-	\begin{figure}[H]
-\includegraphics[scale=0.5]{number_cases}
-\caption{Number of cases in each term. Different colors indicate different chief justices.}
- \label{number_cases}
-\vspace{-.25cm}
-\end{figure}
-
- \begin{figure}[H]
-\includegraphics[scale=0.5]{number_citations}
-\caption{Number of citations for the 1937-2001 time period. Citations for cases prior 1937 are not considered in this figure. Different colors indicate different chief justices.}
- \label{number_citations}
-\vspace{-.25cm}
-\end{figure}
-
-  \begin{figure}[H]
-\includegraphics[scale=0.5]{number_votes_supporting}
-\caption{Number of Votes that were supporting cases between 1937-2001. Different colors indicate terms with different chief justices. }
- \label{number_supporting}
-\vspace{-.25cm}
-\end{figure}
-
-\newpage
-
-Let $C_t$ be the case-to-case adjacency matrix at time $t$ such that $C_t^{ij} \in \{0,1\}$ is a binary indicator of whether case $i$ cites $j$. Furthermore, let
-$$\mathcal{C}_t =\{C_{\leq t} \in \{0,1\}^{(N_t \cdot (N_t-1))/2}: C_{\leq t}^{ij} \in \{0,1\} \}$$ 
-be the set of all possible adjacency matrices among $N_t$ cases at time $t$.
-$N_t$ denotes the number of cases in the network at time $t$.
-Note that the cardinality of $\mathcal{C}_t$ increases exponentially for every newly added case.
-The probability of observing $C_t$ given past citations $C_{<t}$, where $C_{\leq t}$ is the network up to time $t$ is defined as
-\begin{equation}
-P(C_t | C_{<t}, \theta)=\cfrac{\exp\big(\theta^T\cdot h\big(C_{\leq t})\big)\big)}{\sum_{Z_t^*\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)\big)}
-\label{ercm}
-\end{equation}
-where $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C}_t \to \mathbb{R}^q$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{Z^*_t\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)$ is a normalization constant that ensures that Equation (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.
-
-% Network Statistics
-The c-TERGM is specified through 
-the decision regarding which network statistics $h(\cdot)$ to incorporate.
-We include the following network statistics for the Supreme Court citation network, which are derived from the interdependence hypotheses described in the previous section:
-$$h_{edges}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}, $$
-the number of citations.  $h_{edges}$ performs the function of an intercept, and models the expected value of any given edge.
-$$h_{outstar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{i=1} {\sum_{j \neq i} C_{\leq t}^{ij} \choose 2} $$
-the number of out-two-stars.
-An out-two-star is a configuration in which one node sends to two other nodes.
-The number of these configurations grows quadratically as the origins of ties concentrate on a few highly active/social senders.
-The out-two-stars configuration is commonly used to model activity (sender activation) in ERGMs.
-$$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{j=1} {\sum_{i \neq j} C_{\leq t}^{ij} \choose 2} $$
-the number of in-two-stars.
-An in-two-star is a configuration in which one node receives ties from two other nodes.
-The number of these configurations grows quadratically as the destinations of ties concentrate on a few highly popular recipients.
-The out-two-stars configuration is commonly used to model popularity in ERGMs.
-\begin{align*}
-h_{triangle}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to & \sum_{j<i<k} C^{ij}\cdot C^{jk} \cdot C^{ki} + C^{ji}\cdot C^{jk} \cdot C^{ki} + C^{ij}\cdot C^{kj} \cdot C^{ki} \\ & + C^{ji}\cdot C^{kj} \cdot C^{ki} + C^{ij}\cdot C^{jk} \cdot C^{ik} + C^{ji}\cdot C^{jk} \cdot C^{ik}  \\ & + C^{ij}\cdot C^{kj} \cdot C^{ik} + C^{ji}\cdot C^{kj} \cdot C^{ik},
-\end{align*}
-The number of triangles in the network. Triangles measure triad closure, as discussed above.      
-$$h_{reciprocity}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}C_{\leq t}^{ji}, $$
-The number of reciprocal ties in the network.
-Even though the reciprocity of ties is not a meaningful network statistic for most citation networks, it may still play a role in the Supreme Court citation network.
-This statistic comes with the restriction that it only can appear among cases that entered the network at the same time.
-
-$$h_{covariate}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}X_{\leq t}^{ij}, $$ the effect of an exogenous covariate ($X$).
-We include several exogenous covariates based on this standard statistic formulation.
-These covariates are discussed below.\\
-\indent The structure of this model is very similar to a conventional (T)ERGM setup.
-The main difference is that  $\mathcal{C}_t$ excludes adjacency matrices in which there are loops that include edges sent at different time points.
-Loops can exist in the Supreme Court citation network, but only among cases decided in the same term.
-Furthermore, a case $i$ can only cite a case $j$ if $j$ entered the network before or at the same time as $i$ did.
-\\
-
-\section{Network Approaches to Studying Citations}
-
-The dependence processes discussed in the previous section represent new theoretical claims regarding the factors that account for citation formation among US Supreme Court opinions.
-Before we can test these claims, we consider methods for analyzing citation data as a network.
-Researchers from several fields have used network analysis to analyze citations---legal, patent, and scientific.
-Before describing our approach to modeling Supreme Court citations, we review the methods researchers have previously used to study citations.
-A citation is a directed link between two documents that indicates the citing document attributes the cited to be relevant to the evidentiary basis of the citing document.
-Collective patterns of citations within a certain domain make up citation networks.
-When raw citation data is formally represented as a network, the nodes are usually the documents themselves, and each directed arc is the existence of one or more cites from the sending to the receiving document.
-Given the permanent nature of documents in citations networks, these networks are acyclic \citep{leicht2007large,karrer2009random}.
-Relatedly, citation networks grow as new documents and citations enter the network, but established arcs persist in all but exceptional cases.
-
-Networks researchers have developed approaches to understanding the mechanisms that drive citations between documents.
-Work in this area generally proceeds by specifying a set of citation behaviors, then proposing a model to capture the combination of these behavioral rules \citep{simkin2007mathematical}.
-Assessment of the model is based on how well it fits citation distributions.
-Researchers working in this area tend to focus on modeling the growth of the citation network as governed by a degree-based mechanism such as preferential attachment \citep{barabasi1999emergence, vazquez2001statistics}, and the age of the paper \citep{jeong2003measuring,eom2011characterizing,wang2013quantifying}.
-Regarding age, the previous standard approach was to treat the probability of citation as function of degree (i.e., the number of citations previously accrued by the document) and age \citep{hajra2005aging,hajra2006modelling}, then examine the distribution of citations.
-More recently, work has been done in relaxing the assumption that the effect of degree is static, instead allowing it to vary with time \citep{wang2008measuring}.
-
-Another approach to understanding the generative process of citation networks is to examine the existence of network motifs, or subnetwork structures, that can be interpreted as measuring different generative mechanisms.
-For example, a triangle in which case A cites both cases B and C, and case B cites case C, is a type of motif---a triadic motif.
-If researchers find that a triangle is a highly prevalent motif in a network, that finding would suggest that triad closure is a prevalent process by which the network is generated.
-The analysis of network motifs is done by comparing the prevalence of the motifs in the observed network to the prevalence of the motifs in networks drawn from a relevant null distribution.
-This null model must also be characterized by features of citation networks discussed earlier, meaning that it must be a directed acyclic graph with unweighted arcs \citep{carstens2016topology}.
-\citet{karrer2009random} proposes the use of null models based on fixed-degree sequences.
-In these models, the nodes in the networks (i.e., opinions) have the same, or close to the same, numbers of ties.
-Conditional on the number of ties accrued by each node, ties are randomly re-wired such that there is no systematic pattern regarding the nodes to which any node connects.
-
-From this brief review we see two broad components of past methods for analyzing citation networks.
-First, node features, age in particular, are used to model the rate at which a document is cited.
-Second, theoretical models are used to account for prevalent graph structures.
-In the next section we describe a methodology that can be used to integrate case attributes and graph structures/motifs into a comprehensive model of US Supreme Court citations.
-
-%\textit{Comment from Benjamin Kassow: In section 5, I think the discussion is mostly good, but could be organized a bit differently. 
-%I think the discussion of the random intercepts is a bit confusing early on (it is explained more effectively later). }
-%%%% DONT DELETE THESE COMMENTS, INCLUDE IDEAS FOR ADDITIONAL COVARIATES %%%%%
-
-%\citet{cross2010determinants} find that the number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
-(case level).
-Found case issue area, opinion ideology, opinion author, vote margin was not significant.
-Negative citations decreased future positive citations.
-
-%\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
-Show that the rate of citations to a case goes down very quickly.
-THE EFFECT OF AN INDEPENDENT VARIABLE ON A PRECEDENT IS DEPENDENT UPON THE AGE OF THE CASE.
-
-% \item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
-
-%\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
-Lower if stautory cases.
-Increases with ideological variance of the majority coalition.
-Increases with the number of concurring opinions.

### 2020-05-05 Tex/paper.tex
Final Changes

@@
 186.
-That case was followed by numerous others, e.g.: that one person could not be given twice or 10 times the voting power of another person in a state-wide election merely because he lived in a rural area..."
+That case was followed by numerous others, e.g.: that one person could not be given twice or 10 time the voting power of another person in a state-wide election merely because he lived in a rural area..."
 \end{quotation} %https://supreme.justia.com/cases/federal/us/400/112/case.html
@@
 We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
-To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{welles2015individual,duque2018recognizing, graif2017neighborhood,osei2018elite}  %The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
+To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{welles2015individual,duque2018recognizing, graif2017neighborhood,osei2018elite}.
+%The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
 The network structures for which ERGMs are currently designed are insufficient to account for the structure of citation networks.
@@
 The data is taken from the SCDB \citep{spaeth2014supreme}.
-{\color{blue} Justification for including this variable?
-}
 
@@
 
-The common dependence effects for these four dynamics are the number of triangles (case $i$ cites cases $j$ and $k$, and case $j$ cites case $k$), the number of mutual dyads (case $i$ cites $j$ and vice versa), the number of out-2-stars (case $i$ cites $j$ and $k$), and the number of in-2-stars (case $i$ is being cited by cases $j$ and $k$) in $C_{\leq t}$.
+The common dependence effects for these three dynamics are the number of triangles (case $i$ cites cases $j$ and $k$, and case $j$ cites case $k$), the number of mutual dyads (case $i$ cites $j$ and vice versa), the number of out-2-stars (case $i$ cites $j$ and $k$), and the number of in-2-stars (case $i$ is being cited by cases $j$ and $k$) in $C_{\leq t}$.
 Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy \citep{Handcock.2003}.
@@
 Similar to the gwidegree-statistic the gwesp-statistic incorporates a network's edgewise shared partners distribution by replacing $D_r$ by $ESP_r$.
+We fix the decay parameter at $\lambda=0.25$.
 Triangles that are incident to the outgoing shared partners configuration arise when there is a cite between two cases decided at the same time, and those two cases cite one or more of the same third cases (decided either previously or also during the same term).
@@
 In these missing terms there were no two cases that cited each other, making the estimate of the reciprocity statistic term equivalent to $-\infty$.
-For those years, the reciprocity term is included in the model, but the estimate of $-\infty$.
-A similar scenario is visible for the \textit{Overruled Cases} variable.
-In the time period from 1953-1956 no cases that have been overruled have been cited which results in a non-estimable effect.
+%For those years, the reciprocity term is included in the model, but the estimate of $-\infty$.
+A similar scenario is visible for the \textit{Overruled Cases} variable. In the time period from 1953-1956 no cases that have been overruled have been cited which results in a non-estimable effect.
 
@@
 In a typical year, the presence of a citation from case $i$ to case $j$ increases the log-odds of a citation from case $j$ to case $i$ by approximately 0.50.
-Both transitivity effects---different term transitivity, and GWESP---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect.
+Both transitivity effects---different term transitivity, and gwesp---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect.
 The log-odds of a citation from case $i$ to case $j$ increases by more than 2 if there is at least one third case, $k$ that is cited by $i$, and cites to $j$.
-Results regarding the popularity effect (i.e., GWIdegree) are generally supportive of the hypothesis of a popularity effect in Supreme Court citations, as the GWIdegree estimate is negative in every term.
+Results regarding the popularity effect (i.e., gwidegree) are generally supportive of the hypothesis of a popularity effect in Supreme Court citations, as the gwidegree estimate is negative in every term.
 There are some terms for which the result is not statistically significant at conventional levels.
@@
 Most of the exogenous covariate effects align with expectations.
-We find that cases are more likely to cite those that (1) have been decided most recently, (2) were decided by large majorities, (3) have not been overruled, (3) were authored by the same justice, and (3) are in the same issue area.
+We find that cases are more likely to cite those that (1) have been decided most recently, (2) were decided by large majorities, (3) have not been overruled, (4) were authored by the same justice, and (5) are in the same issue area.
 Surprisingly, we do not find consistent effects of either of the covariates that involve justice ideology.
@@
 Overcoming this limitation would require methodological innovation in the c-ERGM as well as theoretical development regarding exactly how citation structure might predict case emergence.
-%Third, the data used in this study does not cover the most recent history of the Court.
-Our data ends in 2001---missing the tenure of the most recent chief justice, and more than fifteen terms.
-To overcome this limitation we would need to extend the citation data beyond that collected by Fowler and colleagues.
-
-\section{Appendix}
-
-\subsection{Goodness-of-Fit}
-We evaluate the goodness-of-fit of the model following \citet{Hunter.2008} by examining the distribution of four hyper statistics, e.g., the out- and indegree distribution and the distribution for two different edgewise shared partners statistics.
-OTP stands for \textit{outgoing two-paths} and refers to the number of cases $r$ that are cited by case $i$ and that cite case $j$, while $j$ is also directly cited by $i$.
-The second ESP statistic is the OSP specification that has been introduced in section \ref{dependence_terms}.
-Figure \ref{GOF} visualizes the goodness-of-fit results for the citation network for the 1950 (top) and 2015 (bottom) term.
-The solid black line indicates the statistic's distribution in the Supreme Court citation network of that given term and the boxplots depict the statistic's distribution of $1000$ networks that have been simulated from the ERGM defined by the MCMLE.
-This means that in the ideal case the solid black line passes through ever single boxplot.
-
-We see that our models do a good job capturing the out and indegree distribution of the citation network, since the black line falls almost exclusively within the ranges spanned by the boxplots.
-For the ESP distributions we can observe that the number of ties with $r=0$ shared partners is captured well for both the OTP as well as for the OSP statistic.
-However, the model overestimates the number of $r=1$ shared partners and then, especially in the 2015 term network, underestimates the number of ties with more than $r=1$ shared partners.
-
- \begin{figure}[H]
- \begin{center}
-\includegraphics[width=13.8cm]{GOF_1950}\\
-%\caption{Goodness-of-fit diagnostic for the 1950 network.}
-% \label{GOF_1950}
-\vspace{1cm}
-\includegraphics[width=13.8cm]{GOF_2015}
-\vspace{0.1cm}
-\caption{Goodness-of-fit diagnostic for the 1950 network (top) and the 2015 network (bottom).}
- \label{GOF}
-\vspace{-.25cm}
-\end{center}
-\end{figure}
-
-\subsection{Checking for Model Degeneracy}
-
-A common challenge when fitting ERGMs is model degeneracy.
-Model degeneracy occurs when the probability distribution defined by the parameter vector does not predominantly yield networks with similar statistics as the observed network.
-Generally, model degeneracy results in simulated networks with no ties or all possible ties.
-In a non-degenerate model the statistics of the networks that were simulated from the probability distribution defined by the MCMLE fall in the proximity of the observed network's statistics.
-Figures \ref{mcmcdiagnostics_1950} and \ref{mcmcdiagnostics_1950} depict trace and density plots for the dependence terms in the 1950 and 2015 term citation network.
-The histograms on the left visualize a statistic's density from $1000$ simulated networks, while the right side shows the statistic's trace plot of the same $1000$ networks.
-The solid black line indicates the statistic's value in the actual citation network.
-Both figures indicate that this model is non-degenerate and that the simulated network's statistics fall almost evenly around the observed statistic.
-The density and trace plots for the ERGM of the terms not depicted provide similar results.
-
-\begin{figure}[H]
- \begin{center}
-\includegraphics[width=14.5cm]{Deg_1950}
-\caption{Density and trace plots for the dependency terms of the 1950 term citation network.}
- \label{mcmcdiagnostics_1950}
-\vspace{-.25cm}
-\end{center}
-\end{figure}
-
- \begin{figure}[H]
- \begin{center}
-\includegraphics[width=14.5cm]{Deg_2015}
-\caption{Density and trace plots for the dependency terms of the 2015 term citation network.}
- \label{mcmcdiagnostics_2015}
-\vspace{-.25cm}
-\end{center}
-\end{figure} 
-%\begin{itemize}
-%\item add robustness check: MCMLE for each year -> plot time trend of coefs -> are results similar to the MPLE time trends presented in the paper?
-%\item (investigate which statistic contributes most on recall improvement)
-%\end{itemize}
 
@@
 
- \begin{figure}[H]
-  \begin{tabular}{c}
-\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_activity}
- \\
-\includegraphics[width = 1\textwidth, trim= 0.1cm 1cm 0.1cm .1cm,clip=true]{issuearea_popularity}
-\end{tabular}
-\caption{Issue Area Popularity and Activity (nodeifactor and nodeofactor). Baseline is Issue Area 'Criminal Procedure'.}
-\end{figure}
-
- \begin{figure}[H]
-\includegraphics[width=15cm]{heatplot_issue_area}
-\caption{Nodemix Issue Area. Sender14Receiver14 and Sender14Receiver13 couldnt be estimated.}
- \label{number_supporting}
-\vspace{-.25cm}
-\end{figure}
-
 \end{document}
-
-% Network Statistics
-The c-ERGM is specified through 
-the decision regarding which network statistics $h(\cdot)$ to incorporate. We include the following statistics for the Supreme Court citation network:
-$$h_{edges}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{i=1}^Nc_i(t)$$
-the number of citations made at time $t$.  $h_{edges}$ is like an intercept, and models the expected value of any given edge.
-$$h_{outstar}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
-the number of weighted outstars occuring at time $t$. We argue that it should be more likely to cite more recent cases than cases that have been decided further in the past. For the weight 
-$$w(a,b):= \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
-we define $a$ and $b$ as the elapsed time since case $i$ and $j$ have been introduced to the network.
-$$h_{triangle}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot
-c_j(t_{-i}) \cdot w(a,b)$$
-where $c_j(t_{-i})$ indicates whether case $j$ was cited at the time case $i$ was introduced into the network.
-Just as for the outstar statistic, we include a weighting factor to favor more recent cases.
-\\[0.3cm]
-
-%Change Statistic
-The individual entries $c_i(t)$ can be taken as a manifestation of single Bernoulli variables $C_i(t)$.
-This interpretation allows the following calculation regarding the conditional distribution of $C_i(t)$:
-%
-\begin{eqnarray*}
-\dfrac{P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~|~ C_i(t)^c=c_i(t)^c)} &=&
-\dfrac{P_{\theta}(C_i(t)=1 ~,~ C_i(t)^c=c_i(t)^c)}{P_{\theta}(C_i(t)=0 ~,~ C_i(t)^c=c_i(t)^c)} \\
-                           &=&\dfrac{P_{\theta}(C(t)= c_i^+(t))}{P_{\theta}(C(t)=c_i^-(t))}\\
-                           &=&\dfrac{\exp(\theta^T \cdot h(c_i^+(t)~|~c(t-)))}{\exp(\theta^T \cdot h(c_i^-(t)~|~c(t-)))}\\
-                           &=& \exp(\theta^T \cdot (h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-)))
-\end{eqnarray*}
-%
-This implies the following equation:
-%
-\begin{equation}
-\text{logit}(P_{\theta}(C_i(t)=1 ~|~ C_i(t)^c=c_i(t)^c))= \theta^T \cdot (h(c_i^+(t)~~|c(t-)) - h(c_i^-(t)~|~c(t-)))
-\label{Logit}
-\end{equation}
-In the equation above the following notations were used:
-%
-\begin{itemize}
-\item $c_i^+(t)$ emerges from $c(t)$, while assuming $c_i(t)=1$
-\item $c_i^-(t)$ emerges from $c(t)$, while assuming $c_i(t)=0$
-\item The condition $C_i(t)^c=c_i(t)^c$ is short for: $C_j(t)=c_j(t)$ for all $j\in \{1,\dots,N\}$ with $i \neq j$
-\item The expression $(\Delta c_i)(t):=h(c_i^+(t)~|~c(t-)) - h(c_i^-(t)~|~c(t-))$ is called the \textit{change statistic}.
-The $k$th component of $(\Delta c_i)(t)$ captures the difference between citation networks $c_i^+(t)$ and $c_i^-(t)$ on the $k$th integrated statistic in the model
-\end{itemize}
-
-\section{Estimation}
-
-% Pseudo Likelihood
-\subsection*{Maximum Pseudo-Likelihood Estimator}
-One can assume that the dyads are independent of each other, which means that
-the random variables $C_i(t)$ inside the random vector $C(t)$ are independent of each other.
-In this case, the equation (\ref{Logit}) reduces to
-$$logit(P_{\theta}(C_i(t) = 1)) = \theta^T \cdot (\Delta c_i)(t)$$
-This corresponds with the logistic regression approach, where the observations of
-the dependent variables are simply edge values of the observed citation vector,
-and the observations of the covariate values are given as the scores of every single
-change statistic. Therefore, the resulting likelihood function is of the following form:
-\begin{equation}
-\text{lik}(\theta)= P_{\theta}(C(t)=c(t))= \prod_{i} \dfrac{ \exp \left(\theta^T \Delta(c_i)(t) \right)}{1+\exp \left(\theta^T \Delta(c_i)(t) \right)}
-\label{PseudoLik}
-\end{equation}
-
-\subsection*{Maximum Likelihood Estimator}
-The more rigorous technique is to estimate the parameters directly with the log-likelihood function derived from (\ref{ercm}), which has the following form:
-%
-\begin{equation}
-\text{loglik}(\theta)=\theta^T \cdot h(c(t)| c(t-))-\log(\kappa(\theta))
-\label{loglik}
-\end{equation}
-%
-The problem resulting from estimating the parameters with (\ref{loglik}) is that the term
-%
-$$\kappa(\theta):= \sum_{c(t)^* \in \mathcal{C}(N)} \exp(\theta^T \cdot h(c(t)^*|c(t-)))$$ 
-%
-which sums up the weighted statistics of all possible binar vectors of length $N$, has to be evaluated.
-However, the cardinality of $\mathcal{C}(N)$ (\#$(\mathcal{C})=2^N$) is incredibly large and a direkt calculation of this sum is for already small $N$ not feasible.
-\\[0.3cm]
-An solution for this limitation is based on the following consideration: Fix a vector of parameters $\theta_0 \in \Theta$ from the underlying parameter range $\Theta$ and compute for $\theta \in \Theta$ the expected value
-%
-\begin{eqnarray*}
-\mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right]&=& 
-\sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot \mathbb{P}_{\theta_0}(C(t)=c(t))\\
-&=& \sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot 
-\frac{\exp(\theta_0^T \cdot \Gamma(c(t)))}{\kappa(\theta_0)}\\
-&=& \frac{1}{\kappa(\theta_0)} \sum_{c(t) \in \mathcal{C}(N)}\exp\left(\theta^T \cdot \Gamma(c(t))\right)\\
-&=&\frac{\kappa(\theta)}{\kappa(\theta_0)}
-\end{eqnarray*}
-%
-This equation offers the following possibility: If one draws $L$ random vectors $c^{(1)}(t), \dots ,c^{(L)}(t)$ out of a distribution $\mathbb{P}_{\theta_0}$ appropriately, one gets with the law of big numbers and a big enough sample $L$ the following relation:
-%
-\begin{equation}
-\frac{1}{L}\cdot \sum_{i=1}^{L}  \exp\left((\theta - \theta_0)^T \cdot \Gamma(c^{(i)}(t))\right)
-~~\approx~~ \mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right] = \frac{\kappa(\theta)}{\kappa(\theta_0)}
-\label{konver}
-\end{equation}
-%
-This approximate can then be used to approximate the log likelihood function.\\[0.4cm]
-Next, we will discuss how a sufficient number of suitable drawings $c^{(1)}(t), \dots ,c^{(L)}(t)$ can be sampled from the distribution $\mathbb{P}_{\theta_0}$. \\
-For this purpose, the Markov Chain Monte Carlo (MCMC) methods can be used.
-\subsection*{Gibbs sampling for the ERCM}
-To be able to compute the approximate likelihood function one needs a sufficiently large number of random vectors from the distribution $\mathbb{P}_{\theta_0}$.
-Snijders \citet{Snijders.2002b} introduces an approach to sample random networks for the ERGM framework by using \textit{MCMC methods}.
-We adapt this approach for sampling appropriate binary vectors for the ERCM.
-\\[0.4cm]
-\textit{Gibbs sampling}\\
-%Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g.
-observed vector).
-Afterwards, the length $L$ of the respective sub-sequence is determined.
-For $k \in \{0,...,L-1\}$ execute the following steps recursively (here the vector in its $k$th iteration is denoted as $c^{(k)}(t)$):\\
-
-\begin{algorithm}[H]
- Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g. observed vector)\\
- \For{i in 1:N}{
-  Compute $\pi:= \cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$\\
-  Draw a random number Z from Bin(1,$\pi$)\\
-  \eIf{Z=1}{
-   set $c^{(k+1)}_i = 1$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
-   }{
-   set $c^{(k+1)}_i = 0$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
-  }
- }
- Start all over using $c^{(k+1)}$\\[0.3cm]
- \caption{Simulation of vectors of $\mathbb{P}_\theta$ using Gibbs sampling}
-\end{algorithm}
-\vspace{0.5cm}
-%\begin{enumerate}
-%\item Randomly choose a number $i \in \{1,\dots, N\}$
-%\item Compute using the likelihood the value 
-%$$\pi:= \mathbb{P}_{\theta}(C_{i}(t)=1 | C_{i}^c(t)=(c_{i}^{(k)}(t))^c)=\cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$$
-%\item Draw a random number $Z$ from Bin$(1, \pi)$. %If
-%\begin{itemize}
-%\item $Z=0$, define $c^{(k+1)}(t)$ via
-%$$c_{p}^{(k+1)}(t)=\begin{cases}
-%0& \text{if}~ p=i \\
-%c_{p}^{(k)}(t) &\text{if}~ p \neq i 
-%\end{cases}$$
-%\item $Z=1$, define $c^{(k+1)}(t)$ via
-%$$c_p^{(k+1)}(t)=\begin{cases}
-%1& \text{if}~ p=i \\
-%c_p^{(k)}(t) &\text{if}~ p \neq i 
-%\end{cases}$$
-%\end{itemize}
-%\item Start at step 1 with $c^{(k+1)}(t)$.
-%\end{enumerate}
-\noindent Using the depicted algorithm, a sequence of random vectors $c^{(0)}(t),...,c^{(L)}(t)$ can be simulated.
-Since the orignial vector was chosen randomly and the first simulated vectors are very dependent on the chosen mvector (only one entry is changed per iteration!), usually the first $B$ vectors, where $N \ll B \ll L$, are discarded as the so called \textit{Burn-In}.
-\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
-Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ to start with (e.g., the observed vector). For $k \in \{0,...,L-1\}$ recursively proceed as follows:\\
-\begin{enumerate}
-\item Randomly choose a number $i\in \{1,\dots, N\}$
-\item Compute, using the equation (\ref{Logit}) the value
-$$\pi := \dfrac{\mathbb{P}_{\theta}(C_{i}(t) \neq c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}{\mathbb{P}_{\theta}(C_{i}(t) = c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}$$
-\item Define $\delta:= \min\{1, \pi\}$ and draw a random number $Z$ from Bin$(1, \delta)$. If
-\begin{itemize}
-\item $Z=0$, let $c^{(k+1)}(t) := c^{(k)}(t)$ 
-\item $Z=1$, define $c^{(k+1)}(t)$ as
-$$c_{p}^{(k+1)}(t)=\begin{cases}
-1-c_{p}^{(k)}(t)& \text{if}~ p=i \\
-c_{p}^{(k)}(t) &\text{if}~ p \neq i 
-\end{cases}$$
-\end{itemize}
-\item Start at step 1 with $c^{(k+1)}(t)$.
-\end{enumerate}
-The first $B \ll L$ vectors are discarded as Burn-In.
-
-First, as these arcs are observed patterns of behaviour, they can be used to infer latent characteristics of nodes or the relations between nodes that produced them \citep{batagelj2003efficient}.
-Moreover, since each judgment is embedded in the broader network of citations, which influences the judgment beyond the characteristics of the two directly linked documents, inferences about the neighborhood of these two documents are possible as well.
-A common application of this general logic is to obtain local or global network indices, such as Kleinberg centrality at the node level \citep{kleinberg1999authoritative,fowler2007network} or global hierarchy \citep{mones2014universal}, then use them to draw conclusions about networks or compare networks across contexts such as time periods \citep[e.g.][]{vazquez2001statistics,fowler2008authority,greenberg2009citation,lupu2012precedent,lupu2013strategic,dawson2014current,jaffe2017patent}.
-
-	Still within the category of using observed network constellation to make inferences about components of the system, identification of communities using methods for clustering, such as modularity maximization \citep[e.g.][]{kajikawa2007creating,shibata2011detecting,chen2010community} and stochastic block modeling \citep{jo2009citing}; and subnetwork identification \citep[e.g.][]{batagelj2017emergence} have been used to analyze relevance of topics within an academic field or to determine trends in technological advancement \citep{verspagen2007mapping,erdi2013prediction}.
-Main path analysis \citep{hummon1989connectivity}, which determines the main path through an acyclic network, is a particularly useful way to examine trends as it allows researchers identify structures of knowledge flow.
-Other methods include classifying documents into groups based on similarities in their citation profiles of cites over time, which can be examined as a function of time to see if there are temporal patterns \citep{leicht2007large}.
-The identification of communities is often coupled with classification of these communities by the researcher using historical knowledge.
-
-	\section{Extra Notes}
-	This section contains some of the extra notes I cut out from the draft that might be useful.
-	\subsection{Broader definitions of citation networks}
-	More broadly, citation networks can have as nodes the document producers, including scientists \citep{ji2016coauthorship}, judges \citep{landes1998judicial}, journals \citep{rice1988citation}, or aggregated units \citep{gelter2012networks,pan2012world}.
-In these cases, which are essentially exercises in aggregation, the resulting network can take on different characteristics such as weighted or cyclic edges.
-Other networks derived from the raw citation network is the co-citation or co-cited-by networks whereby two documents are connected if they co-cite or are co-cited by another document \citep{van2005reference}.
-
-	\subsection{Weighted edges and multiplexity}
-	Recognition, and subsequent quantification, of different levels of complexity in citation arcs allows for a host of advancements in citation network analysis.
-Weights can indicate relevance \citep{liu2014citations} or be used to incorporate the temporal dimension into citation networks \citep{fujita2014detecting}.
-Moving to multiplexity leads to further advancements.
-For example, \citet{greenberg2009citation} accounted for level of support and through simulations based on arc typed switching demonstrated how citation bias against critical articles yielded differences in beliefs within a medical discipline.
-\citet{bommarito2010distance} argued that articles contain different types of information and accounting for the specific piece of information that resulted in the cite can yield clustering that are more interpretable.
-In all cases however, substantial work has to be done into reclassifying the weightedness or multiplexity of citation arcs \citep{zhang2007semantics}.
-
-	\subsection{Misc. notes}
-	\begin{itemize}
-		\item To control for age, \citet{clough2015transitive} propose transitive reduction of citation networks (i.e.
-removal of redundant information ties) as it will primarily remove citation arcs that are disparate in age.
-		\item Most citation networks are directed acyclic graphs, but not all.
-Different versions of the same document can also cause problems (such as strong network components).
-The ``preprint transformation'' is a solution to small strong components \citep{batagelj2017emergence}.
-		\item Date of publication can be assigned as level to nodes in acyclic networks \citep{batagelj2017emergence}.
-Depending on the citation network at hand, temporal ordering can be difficult, as documents can share publication dates.
-To overcome this, begin each iteration by sampling from one of the possible orderings.
-\citep{carstens2016topology}.
-	\end{itemize}
-
-	\begin{figure}[H]
-\includegraphics[scale=0.5]{number_cases}
-\caption{Number of cases in each term. Different colors indicate different chief justices.}
- \label{number_cases}
-\vspace{-.25cm}
-\end{figure}
-
- \begin{figure}[H]
-\includegraphics[scale=0.5]{number_citations}
-\caption{Number of citations for the 1937-2001 time period. Citations for cases prior 1937 are not considered in this figure. Different colors indicate different chief justices.}
- \label{number_citations}
-\vspace{-.25cm}
-\end{figure}
-
-  \begin{figure}[H]
-\includegraphics[scale=0.5]{number_votes_supporting}
-\caption{Number of Votes that were supporting cases between 1937-2001. Different colors indicate terms with different chief justices. }
- \label{number_supporting}
-\vspace{-.25cm}
-\end{figure}
-
-\newpage
-
-Let $C_t$ be the case-to-case adjacency matrix at time $t$ such that $C_t^{ij} \in \{0,1\}$ is a binary indicator of whether case $i$ cites $j$. Furthermore, let
-$$\mathcal{C}_t =\{C_{\leq t} \in \{0,1\}^{(N_t \cdot (N_t-1))/2}: C_{\leq t}^{ij} \in \{0,1\} \}$$ 
-be the set of all possible adjacency matrices among $N_t$ cases at time $t$.
-$N_t$ denotes the number of cases in the network at time $t$.
-Note that the cardinality of $\mathcal{C}_t$ increases exponentially for every newly added case.
-The probability of observing $C_t$ given past citations $C_{<t}$, where $C_{\leq t}$ is the network up to time $t$ is defined as
-\begin{equation}
-P(C_t | C_{<t}, \theta)=\cfrac{\exp\big(\theta^T\cdot h\big(C_{\leq t})\big)\big)}{\sum_{Z_t^*\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)\big)}
-\label{ercm}
-\end{equation}
-where $\theta \in \mathbb{R}^q$ is a $q$-dimensional vector of parameters,  $h: \mathcal{C}_t \to \mathbb{R}^q$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{Z^*_t\in \mathcal{C}_t}\exp\big(\theta^T\cdot h\big(Z^*_{\leq t})\big)$ is a normalization constant that ensures that Equation (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.
-
-% Network Statistics
-The c-TERGM is specified through 
-the decision regarding which network statistics $h(\cdot)$ to incorporate.
-We include the following network statistics for the Supreme Court citation network, which are derived from the interdependence hypotheses described in the previous section:
-$$h_{edges}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}, $$
-the number of citations.  $h_{edges}$ performs the function of an intercept, and models the expected value of any given edge.
-$$h_{outstar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{i=1} {\sum_{j \neq i} C_{\leq t}^{ij} \choose 2} $$
-the number of out-two-stars.
-An out-two-star is a configuration in which one node sends to two other nodes.
-The number of these configurations grows quadratically as the origins of ties concentrate on a few highly active/social senders.
-The out-two-stars configuration is commonly used to model activity (sender activation) in ERGMs.
-$$h_{instar}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{j=1} {\sum_{i \neq j} C_{\leq t}^{ij} \choose 2} $$
-the number of in-two-stars.
-An in-two-star is a configuration in which one node receives ties from two other nodes.
-The number of these configurations grows quadratically as the destinations of ties concentrate on a few highly popular recipients.
-The out-two-stars configuration is commonly used to model popularity in ERGMs.
-\begin{align*}
-h_{triangle}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to & \sum_{j<i<k} C^{ij}\cdot C^{jk} \cdot C^{ki} + C^{ji}\cdot C^{jk} \cdot C^{ki} + C^{ij}\cdot C^{kj} \cdot C^{ki} \\ & + C^{ji}\cdot C^{kj} \cdot C^{ki} + C^{ij}\cdot C^{jk} \cdot C^{ik} + C^{ji}\cdot C^{jk} \cdot C^{ik}  \\ & + C^{ij}\cdot C^{kj} \cdot C^{ik} + C^{ji}\cdot C^{kj} \cdot C^{ik},
-\end{align*}
-The number of triangles in the network. Triangles measure triad closure, as discussed above.      
-$$h_{reciprocity}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}C_{\leq t}^{ji}, $$
-The number of reciprocal ties in the network.
-Even though the reciprocity of ties is not a meaningful network statistic for most citation networks, it may still play a role in the Supreme Court citation network.
-This statistic comes with the restriction that it only can appear among cases that entered the network at the same time.
-
-$$h_{covariate}:\mathcal{C}_t \to \mathbb{R}~~~, ~~~C_{\leq t} \to \sum_{ij} C_{\leq t}^{ij}X_{\leq t}^{ij}, $$ the effect of an exogenous covariate ($X$).
-We include several exogenous covariates based on this standard statistic formulation.
-These covariates are discussed below.\\
-\indent The structure of this model is very similar to a conventional (T)ERGM setup.
-The main difference is that  $\mathcal{C}_t$ excludes adjacency matrices in which there are loops that include edges sent at different time points.
-Loops can exist in the Supreme Court citation network, but only among cases decided in the same term.
-Furthermore, a case $i$ can only cite a case $j$ if $j$ entered the network before or at the same time as $i$ did.
-\\
-
-\section{Network Approaches to Studying Citations}
-
-The dependence processes discussed in the previous section represent new theoretical claims regarding the factors that account for citation formation among US Supreme Court opinions.
-Before we can test these claims, we consider methods for analyzing citation data as a network.
-Researchers from several fields have used network analysis to analyze citations---legal, patent, and scientific.
-Before describing our approach to modeling Supreme Court citations, we review the methods researchers have previously used to study citations.
-A citation is a directed link between two documents that indicates the citing document attributes the cited to be relevant to the evidentiary basis of the citing document.
-Collective patterns of citations within a certain domain make up citation networks.
-When raw citation data is formally represented as a network, the nodes are usually the documents themselves, and each directed arc is the existence of one or more cites from the sending to the receiving document.
-Given the permanent nature of documents in citations networks, these networks are acyclic \citep{leicht2007large,karrer2009random}.
-Relatedly, citation networks grow as new documents and citations enter the network, but established arcs persist in all but exceptional cases.
-
-Networks researchers have developed approaches to understanding the mechanisms that drive citations between documents.
-Work in this area generally proceeds by specifying a set of citation behaviors, then proposing a model to capture the combination of these behavioral rules \citep{simkin2007mathematical}.
-Assessment of the model is based on how well it fits citation distributions.
-Researchers working in this area tend to focus on modeling the growth of the citation network as governed by a degree-based mechanism such as preferential attachment \citep{barabasi1999emergence, vazquez2001statistics}, and the age of the paper \citep{jeong2003measuring,eom2011characterizing,wang2013quantifying}.
-Regarding age, the previous standard approach was to treat the probability of citation as function of degree (i.e., the number of citations previously accrued by the document) and age \citep{hajra2005aging,hajra2006modelling}, then examine the distribution of citations.
-More recently, work has been done in relaxing the assumption that the effect of degree is static, instead allowing it to vary with time \citep{wang2008measuring}.
-
-Another approach to understanding the generative process of citation networks is to examine the existence of network motifs, or subnetwork structures, that can be interpreted as measuring different generative mechanisms.
-For example, a triangle in which case A cites both cases B and C, and case B cites case C, is a type of motif---a triadic motif.
-If researchers find that a triangle is a highly prevalent motif in a network, that finding would suggest that triad closure is a prevalent process by which the network is generated.
-The analysis of network motifs is done by comparing the prevalence of the motifs in the observed network to the prevalence of the motifs in networks drawn from a relevant null distribution.
-This null model must also be characterized by features of citation networks discussed earlier, meaning that it must be a directed acyclic graph with unweighted arcs \citep{carstens2016topology}.
-\citet{karrer2009random} proposes the use of null models based on fixed-degree sequences.
-In these models, the nodes in the networks (i.e., opinions) have the same, or close to the same, numbers of ties.
-Conditional on the number of ties accrued by each node, ties are randomly re-wired such that there is no systematic pattern regarding the nodes to which any node connects.
-
-From this brief review we see two broad components of past methods for analyzing citation networks.
-First, node features, age in particular, are used to model the rate at which a document is cited.
-Second, theoretical models are used to account for prevalent graph structures.
-In the next section we describe a methodology that can be used to integrate case attributes and graph structures/motifs into a comprehensive model of US Supreme Court citations.
-
-%\textit{Comment from Benjamin Kassow: In section 5, I think the discussion is mostly good, but could be organized a bit differently. 
-%I think the discussion of the random intercepts is a bit confusing early on (it is explained more effectively later). }
-%%%% DONT DELETE THESE COMMENTS, INCLUDE IDEAS FOR ADDITIONAL COVARIATES %%%%%
-
-%\citet{cross2010determinants} find that the number of citations to Supreme Court opinions over a ten year period of the Rehnquist Court.
-(case level).
-Found case issue area, opinion ideology, opinion author, vote margin was not significant.
-Negative citations decreased future positive citations.
-
-%\item \citet{black2013citation} Study the number of citations to a case over time, in order to understand the lifecycle of citation rates.
-Show that the rate of citations to a case goes down very quickly.
-THE EFFECT OF AN INDEPENDENT VARIABLE ON A PRECEDENT IS DEPENDENT UPON THE AGE OF THE CASE.
-
-% \item \citet{spriggs2001explaining} Find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case, when the legal issues under consideration are complex, whn the cases ha been negatively treated previously, when there were many concurring opinions, and less likely to be overruled if it was a unanimous coalition.
-
-%\item \citet{lupu2013strategic} The hub score, which is a measure of the degree to which a case cites other important cases increases with salience [GET THE AMICUS BRIEFS, NYT COUNTS].
-Lower if stautory cases.
-Increases with ideological variance of the majority coalition.
-Increases with the number of concurring opinions.

### 2020-05-05 RCode/Plot Figures/Plot_SCC.R

@@ -285,7 +285,7 @@ p6
 data<- data.frame(Year, coM[7,], P[7,])
 colnames(data)<- c('Year', 'Theta', 'pi')
 eb <- aes(ymax = coM[7,]+1.96*steM[7,], ymin = coM[7,]-1.96*steM[7,])
-p7<- ggplot(data = data, aes(x = Year, y = Theta, shape=factor(pi), group=1)) +ggtitle('Martin-Quinn Score')+
+p7<- ggplot(data = data, aes(x = Year, y = Theta, shape=factor(pi), group=1)) +ggtitle('Ideological Distance')+
   geom_line() + 
   geom_ribbon(eb, alpha = 0.3)
 p7<-p7+scale_colour_manual(values=c('21'='springgreen4', '24'='orange1', '22'='red', '23'='black'))
@@ -326,12 +326,12 @@ p9
 data<- data.frame(Year, coM[10,], P[10,])
 colnames(data)<- c('Year', 'Theta', 'pi')
 eb <- aes(ymax = coM[10,]+1.96*steM[10,], ymin = coM[10,]-1.96*steM[10,])
-p10<- ggplot(data = data, aes(x = Year, y = Theta, shape=factor(pi), group=1)) +ggtitle('Absolute Difference in Martin Quinn Scores')+
+p10<- ggplot(data = data, aes(x = Year, y = Theta, shape=factor(pi), group=1)) +ggtitle('Ideological Breadth')+
   geom_line() + 
   geom_ribbon(eb, alpha = 0.3)
 p10<-p10+scale_colour_manual(values=c('21'='springgreen4', '24'='orange1', '22'='red', '23'='black'))
 p10<-p10+theme(legend.position='none', axis.text=element_text(size=12),axis.title=element_text(size=14),
-               plot.title=element_text(size=12, face='bold'))+ geom_point(aes(colour = factor(pi), size = 3), size=2.5) #+
+               plot.title=element_text(size=16, face='bold'))+ geom_point(aes(colour = factor(pi), size = 3), size=2.5) #+
 #geom_point(aes(colour = factor(pi)), size=1.5) 
 p10
 

### 2020-05-05 Tex/paper.tex
Updated Table 1

@@
 We found considerable agreement---over 95\% of the citations recorded in the CourtListener dataset were also in the Fowler dataset, and over 96\% of the citations recorded in the Fowler data were also in the CourtListener dataset.}.
-The Supreme Court citation network from $1937 - 2015$ consists of $9,945$ cases.
+The Supreme Court citation network from $1937 - 2015$ consists of $10,020$ cases.
 The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
-The network has a total of $111,986$ citation ties.\footnote{In order to focus on citation actions that were not intended to totally invalidate an opinion, we exclude $315$ citations that caused the cited case to be overruled.
+The network has a total of $112,939$ citation ties.\footnote{In order to focus on citation actions that were not intended to totally invalidate an opinion, we exclude $315$ citations that caused the cited case to be overruled.
 The data on overruling citation came from \citet{senate2016constitution}.
@@
 W Rehnquist            & 1987 - 2005                                          & 2022                                                              & 106.42                                                      \\ \hline
-J Roberts **           & 2006 - 2015                                          & 796                                                              & 79.6                                           \\ \hline
+J Roberts **           & 2006 - 2015                                          & 871                                                              & 87.1                                           \\ \hline
 \end{tabular}
 \caption[caption]{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
-\\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 4/2018).}
+\\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 5/2020).}
 \label{tab:chiefs}

### 2020-05-06 Tex/appendix.tex
Minor changes in Appendix
Updated Titlepage

@@
 \usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
+%\usepackage[T1]{fontenc}
 
@@
    \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
-  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
+%  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
   }
@@
 OTP stands for \textit{outgoing two-paths} and refers to the number of cases $r$ that are cited by case $i$ and that cite case $j$, while $j$ is also directly cited by $i$.
-The second ESP statistic is the OSP specification that has been introduced in section \ref{dependence_terms}.
+The second ESP statistic is the OSP specification that has been introduced in section 4.1.2 in the paper.
 Figure \ref{GOF} visualizes the goodness-of-fit results for the citation network for the 1950 (top) and 2015 (bottom) term.

### 2020-05-06 Tex/paper.tex
Updated Titlepage

@@
 The dependence effects are as substantively and statistically significant as the effects of the exogenous covariates we include in the model, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.\footnote{
-Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. 
 This work was supported in part by NIH, 

### 2020-05-06 Tex/appendix.tex
working to trim

@@
 \maketitle
+
+\section{c-ERGM Estimation}
+
+The normalizing constant in Equation 1 is intractable.
+For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure \ref{fig:ctergm}---there are 16,777,216 unique configurations of $C_t$ that could be observed.
+The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases.
+This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the c-ERGM.
+
+The common alternative relies on Monte Carlo methods to approximate the normalizing constant by simulating a large set of networks \citep{hunter2006inference,hummel2012improving}.
+The resulting estimator, the Monte Carlo MLE (MCMLE), is approximately consistent, meaning that it converges to the MLE as the sample size, i.e., the number of simulated networks, increases.
+However, one drawback is that with the number of nodes in the Supreme Court citation network being in the order of $10,000$, obtaining the MCMLE is computationally expensive \citep{schmid2017exponential} and the success of the algorithm heavily relies on the starting parameter vector $\theta_0$, which is ideally chosen in the proximity of the unknown MLE \citep{hummel2012improving}.
+The prevailing choice for $\theta_0$ is the maximum pseudo-likelihood estimation (MPLE) \citep{strauss1990pseudolikelihood}, a fast estimation method that is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
+The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated above, assume a logit form.
+The MPLE is simple to obtain, but does not guarantee a starting value close to the MLE \citep{SchmidHunter2020}.
+
+We estimate the MCMLE for every term of the c-ERGM between 1950 and 2015.
+While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows.
+The growth in the network also serves as our reason for estimating a separate model for each term.
+First, the network is large, with each term introducing the potential for hundreds-of-thousands of new citations.
+This means we have the statistical power necessary to estimate a separate set of parameter values for each term.
+Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network.
+This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term.
+
+The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant.
+However, the estimation of most networks in the 90s with the MPLE as starting values was not feasible in an reasonable time frame anymore.
+Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$.
+But even this approach started to fail for networks around the turn of the millennium.
+Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}.
+The MCMLE for these large citation networks was obtained by setting the starting value according to an approach introduced by \citet{SchmidHunter2020}.
+This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$.
+However, the MPLE of networks with the same sufficient statistics is not necessarily the same.
+Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a network $C_t^*$ using simulated annealing \citep{Kirkpatrick83}, with $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$ and a weak dependence structure among unfixed ties.
+The MPLE of such a network is similar to the MLE, making it an effective starting value for the MCMLE algorithm.
 

### 2020-05-06 Tex/paper.tex
working to trim

@@
 
+\usepackage[moderate]{savetrees}
+
 \usepackage[table,xcdraw]{xcolor}
@@
 
-\usepackage{geometry}
  \usepackage{float}
@@
 \usepackage{url}
+
+\usepackage{geometry}
 
@@
 
-\begin{figure}[htp]
+\begin{figure}[ht]
 \centering
@@
 
-\subsection{c-ERGM definition}
-
 We develop the citation ERGM (c-ERGM), to account for the structural constraints that apply to the network of Supreme Court citations.
@@
 
-\begin{figure}[htp]
+\begin{figure}[H]
 \centering
@@
 The logit form conditional probability is well known for the ERGM family \citep{goodreau2009birds}.
+Estimation of the c-ERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix.
 
-We should note that there is nothing about the c-ERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another.
+%We should note that there is nothing about the c-ERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another.
 For example, if we find a positive popularity effect, the term for that dynamic would indicate that there is an increase in the log odds that case $i$ cites case $j$ with each other case that has cited case $j$.
@@
 However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-ERGM can be specified to account for the dependence in the data.
-
-\subsection{c-ERGM Estimation}
-
-The normalizing constant in Equation 1 is intractable.
-For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure \ref{fig:ctergm}---there are 16,777,216 unique configurations of $C_t$ that could be observed.
-The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases.
-This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the c-ERGM.
-
-The common alternative relies on Monte Carlo methods to approximate the normalizing constant by simulating a large set of networks \citep{hunter2006inference,hummel2012improving}.
-The resulting estimator, the Monte Carlo MLE (MCMLE), is approximately consistent, meaning that it converges to the MLE as the sample size, i.e., the number of simulated networks, increases.
-However, one drawback is that with the number of nodes in the Supreme Court citation network being in the order of $10,000$, obtaining the MCMLE is computationally expensive \citep{schmid2017exponential} and the success of the algorithm heavily relies on the starting parameter vector $\theta_0$, which is ideally chosen in the proximity of the unknown MLE \citep{hummel2012improving}.
-The prevailing choice for $\theta_0$ is the maximum pseudo-likelihood estimation (MPLE) \citep{strauss1990pseudolikelihood}, a fast estimation method that is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
-The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated above, assume a logit form.
-The MPLE is simple to obtain, but does not guarantee a starting value close to the MLE \citep{SchmidHunter2020}.
-
-We estimate the MCMLE for every term of the c-ERGM between 1950 and 2015.
-While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows.
-The growth in the network also serves as our reason for estimating a separate model for each term.
-First, the network is large, with each term introducing the potential for hundreds-of-thousands of new citations.
-This means we have the statistical power necessary to estimate a separate set of parameter values for each term.
-Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network.
-This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term.
-
-The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant.
-However, the estimation of most networks in the 90s with the MPLE as starting values was not feasible in an reasonable time frame anymore.
-Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$.
-But even this approach started to fail for networks around the turn of the millennium.
-Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}.
-The MCMLE for these large citation networks was obtained by setting the starting value according to an approach introduced by \citet{SchmidHunter2020}.
-This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$.
-However, the MPLE of networks with the same sufficient statistics is not necessarily the same.
-Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a network $C_t^*$ using simulated annealing \citep{Kirkpatrick83}, with $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$ and a weak dependence structure among unfixed ties.
-The MPLE of such a network is similar to the MLE, making it an effective starting value for the MCMLE algorithm.
+Estimation of the c-ERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix.
 
@@
 
-\begin{figure}[htp]
+\begin{figure}[H]
 \includegraphics[scale=0.5]{degree_distribution}
-\caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2001. There are cases with an indegree >50, but they are not captured in this figure.}
+\caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2001. There are cases with an indegree  $>$50, but they are not captured in this figure.}
  \label{degree_dist}
@@
 
-\begin{figure}[htp]
+\begin{figure}[H]
 \includegraphics[scale=0.35,clip=true,trim=.5cm 0cm 0cm 2cm]{images/citations1}
@@
 
-  \begin{figure}[t]
+  \begin{figure}[H]
 \includegraphics[width=15cm]{SCC_AIC_BIC}
@@
 
-We recognize two important limitations of the current analysis, which should be addressed in future iterations of this work or in future studies.
+%We recognize two important limitations of the current analysis, which should be addressed in future iterations of this work or in future studies.
 First, our theoretical claims and analyses treat all citations---positive and negative---as equivalent.
@@
 
-\newpage
-
 \bibliography{bib} 

### 2020-05-06 Tex/paper.tex
trimmed some references

@@
 United States Supreme Court opinions exercise authority and influence, in part, through their roles as precedents affecting future jurisprudence in the US.
-The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,gillman2001s,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained,hitt2016measuring}.
+The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained,hitt2016measuring}.
 Despite a considerable body of research that focuses on the way in which relevant precedents shape decision making on the Court, relatively little work has focused on understanding which past opinions are cited by an opinion.
@@
 This is a special case of a very common process on networks referred to as ``triad closure''.
-Complex dependence is theoretically interesting on its own merits, but the effects of covariates cannot be reliably identified---either in terms of coefficient values or standard errors---without accounting for the interdependence inherent in networks \citep{desmaraisstatistical}.
+Complex dependence is theoretically interesting on its own merits, but the effects of covariates cannot be reliably identified---either in terms of coefficient values or standard errors---without accounting for the interdependence inherent in networks \citep{cranmer2016critique}.
 
@@
 We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
-To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{welles2015individual,duque2018recognizing, graif2017neighborhood,osei2018elite}.
+To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{duque2018recognizing,osei2018elite}.
 %The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
@@
 The full model is the one we present below.
-We use both AIC and BIC to compare the fit of the two models, with lower values indicating better fit \citep{ripplinger2008does}.
+We use both AIC and BIC to compare the fit of the two models, with lower values indicating better fit.
 

### 2020-08-05 Tex/PA_rrmemo/PA_RnR_memo.tex
adding memo

%\title{Overleaf Memo Template}
% Using the texMemo package by Rob Oakes
\documentclass[a4paper,11pt]{texMemo}
\usepackage[english]{babel}
\usepackage{graphicx, lipsum}

%% Edit the header section here. To include your
%% own logo, upload a file via the files menu.
\memosubject{Manuscript Revisions for \emph{Political Analysis}}
\memoto{PA-2020-075}
\memofrom{Generative Dynamics of Supreme Court Citations: Analysis with a New Statistical Network Model}
\memodate{\today}

\begin{document}
\maketitle

\noindent We are grateful for the opportunity to revise and resubmit our manuscript. In this memo
we have separated the reviewers' comments into separate criticisms and
suggestions. Under each comment, we describe how we have revised the manuscript in
response to the feedback provided. We generally agree with the criticisms offered, and think
that our manuscript has improved substantially as a result of incorporating this feedback.

\section*{Reviewer: 1}

\noindent \textbf{R1.1:} \emph{ In terms of estimation, it reads as if the smaller networks used MCMLE and the later the simulated annealing approach. Is that right? If so, does the simulated approach arrive at the same values for the smaller year networks as the MCMLE? Showing that might help validate the approach for the bigger/later year networks. At the very least it might mitigate my next point on the overtime changes.   }\\

\noindent \textbf{Addressed:}   \\

\noindent  \textbf{R1.2:} \emph{ For at least a few of the variables --- reciprocity, gwidegree, and Justices in majority --- we see a substantial change in recent periods. Does this have anything to do with the change in estimation strategy? How convinced can we be that these fairly big changes are substantively driven? }\\

\noindent \textbf{Addressed:}   \\

\noindent \textbf{R1.3:} \emph{ Relatedly, why isn’t overtime change in the covariates discussed? The authors instead choose to make broad statements: “in/consistently significant” over the period of study.  This seems to be a wasted opportunity and contrary to their motivation (a la Shalizi and Rinaldo 2013). If the authors want to make a statement about whether the covariate generally holds, then why not run a single model across all the years? Otherwise, wouldn’t it make sense to discuss these over time changes, at least when they appear to be quite meaningful?  }\\

\noindent \textbf{Addressed:}   \\

\noindent \textbf{R1.4:} \emph{ Not to belabor the point, but the motivation for the exogenous and endogenous terms are nicely situated in the literature, but the discussion of the results are quite anticlimactic. It merely notes that some results are generally in line with expectations while others are not (the ideology finding being terribly interesting and surprising!). I’d like to see a more substantive engagement with the results, particularly any trends / disruptions in the series of coefficients. This has the potential to be a much bigger contribution to the judicial literature than the authors acknowledge.  }\\

\noindent \textbf{Addressed:}   \\

\noindent \textbf{R1.5:} \emph{ While this is a well-executed paper (and visually beautiful), it’s not clear where the contribution is meant to be. Is this paper an introduction of a new model or an application of the ERGM to citation data?  If this is a new network model that would merit its own name, c-ERGM, what other networks might this model apply to?  Does it really make sense to have a model that just applies to this one judicial network...? I suppose this model works for any citations --- e.g., academic papers -- so maybe discussing a bit of that literature and data would help. In short, the change to the ERGM is slight, even when combining that with the necessary simulated estimation approach, so telling us why these changes are so useful beyond this network could drive home the methodological contribution.  }\\

\noindent \textbf{Addressed:}   \\

\noindent \textbf{R1.6:} \emph{ Alternatively, one could think about this paper more in terms of its substantive contributions. The front end is deeply ingrained in the citations and judicial literatures, as are the data, of course, but little attention is given in the end to what we learn from this paper. It seems like the authors decided not to engage either frame fully, which leaves the contribution lacking on both. Having said that, I think one could go either way with this paper, and I liked a lot about it on both fronts. }\\

\noindent \textbf{Addressed:}   \\

\section*{Reviewer: 2}

\noindent \textbf{R2.1:} \emph{ }\\

\noindent \textbf{Addressed:}  \\

\end{document}

### 2020-08-06 Tex/PA_rrmemo/PA_RnR_memo.tex
add R2 to memo

@@
 
-\noindent \textbf{R2.1:} \emph{ }\\
+\noindent \textbf{R2.1:} \emph{ I would really like to see discussion near the end of section 4.1 potentially expanded and/or cleaned up considerably to make some things more clear.}\\
+
+\noindent \textbf{Addressed:}  \\
+
+\noindent \textbf{R2.2:} \emph{ Additionally, implementation details are almost wholly relegated to the appendix; while this is largely appropriate, there is not even much information in the appendix, so I would recommend that one or both of the estimation details in the main paper and the appendix are expanded.
+What software was used?
+Do the authors generate novel convenience software to handle the specifics of the cERGM model and/or the computational overhead they discuss in the appendix?}\\
+
+\noindent \textbf{Addressed:}  \\
+
+\noindent \textbf{R2.3:} \emph{ These issues that I mention are particularly important if the authors intend for this model to be used by the judicial politics community rather than only network modelers in other contexts where the cERGM structure makes sense; a bit more practical information would be necessary I think for this to travel within the judicial politics community.}\\
+
+\noindent \textbf{Addressed:}  \\
+
+\noindent \textbf{R2.4:} \emph{ In many places the authors do a very good job of explaining the technical side of what they are doing precisely, as well as providing easy to digest intuitive explanations.
+However, one place where we seem to somewhat lack both clarity as well as thoroughness and precision is unfornately the section where it was arguably needed the most: the section on dependence terms.
+For example, the second paragraph of this section warns that "Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy," as far as I can tell, only for degeneracy as a result of the in-2-star statistic is a plan to guard against degeneracy addressed at all (surrounding equation (2)).
+I think more information about whether and how degeneracy resulting from the other statistics can be guarded against would be a good addition.
+Additionally, the authors fix the decay parameters for gwidegree and gwesp to 1 and 0.25 respectively with no discussion of the justification for those specific values; perhaps setting lambda for gwidegree to unity makes intuitive sense, but what drives the decision for gwesp?
+Considerably more guidance for applied researchers is needed in this section if the authors intend this model to be used by judicial politics researchers, which seems to be the case considering their discussion in the introduction.
+This would be a very good advance to the state of the art in citation studies in judicial politics, but the end of section 4.1 is not clear and thorough enough in my opinion to be useful to most applied judicial scholars whose work could be improved by the authors' methodological advance.}\\
+
+\noindent \textbf{Addressed:}  \\
+
+\noindent \textbf{R2.5:} \emph{ As a small comment, in relation to the degeneracy discussion, I think footnote 4 would be best served by not being a footnote.
+After your discussion of degeneracy I think the typical reader will be wondering, "Did this end up being a problem for their application?" The answer is, "No, we checked that and it was fine," and I just think that as it's such a natural thing for a reader to want to know after you spent multiple full paragraphs discussing the issue, this sentence deserves its place in the main text.}\\
+
+\noindent \textbf{Addressed:}  \\
+
+\noindent \textbf{R2.6:} \emph{ I think the manuscript would be greatly improved as well by pointing practitioners toward implementation software, even if this is a very short, simple statement.
+For example, if the answer for that is, "We used software X which is often used for ERGMs, and our custom calls to that software to handle the cERGM and large network we analyze can be found in the replication archive," great.
+If the authors have written convenience wrappers for existing software to handle these issues, or wholly novel software, even better; advertise it both for your own sake and for the sake of the substantive scholars who want to use your method because they care about modeling their data in the most appropriate way.}\\
+
+\noindent \textbf{Addressed:}  \\
+
+\noindent \textbf{R2.7:} \emph{ I think the manuscript would be greatly improved as well by pointing practitioners toward implementation softwarA point which may be useful to the authors in future work (though I'm not sure whether they find it has great utility for this methods piece): The authors find that there are not reliable effects for the ideological variables, and thus wonder, "whether, and if so, how, citations are shaped by ideological factors." I think one natural response could be that the ideological effects may actually be found within the dependence effects they investigate in some way, and I wonder if there are creative ways within the context of their model to test for that.}\\
 

### 2020-08-09 Tex/PA_rr/main.tex
added pa rr version

--- 2020-05-06 Tex/paper.tex
@@
-\documentclass[headsepline=true, abstracton]{scrartcl}
-
+\documentclass{cup-pan}
 \usepackage[utf8]{inputenc}
-%\usepackage[T1]{fontenc}
-
+\usepackage{blindtext}
+\usepackage{bm}
 \usepackage{amssymb}
@@
 \usepackage{amsthm}
-\usepackage{bm}
+\usepackage{geometry}
 
-\usepackage{natbib}
+\title{Generative Dynamics of Supreme Court Citations: Analysis with a New Statistical Network Model}
 
-\usepackage[moderate]{savetrees}
+\author[1]{Christian S. Schmid}
+\author[2]{Ted Hsuan Yun Chen}
+\author[3]{Bruce A. Desmarais}
 
-\usepackage[table,xcdraw]{xcolor}
+\affil[1]{Department of Statistics, The Pennsylvania State University, University Park, PA. Email: \url{schmid@psu.edu}}
+\affil[2]{Faculty of Social Sciences, University of Helsinki, Finland. Email: \url{ted.hsuanyun.chen@gmail.com}}
+\affil[3]{Department of Political Science, The Pennsylvania State University, University Park, PA. Email: \url{bdesmarais@psu.edu}}
 
-\usepackage{graphicx}
+%% Corresponding author
+\corrauthor{Christian S. Schmid}
 
- \usepackage{float}
+%% Abbreviated author list for the running footer
+\runningauthor{Schmid et al.}
 
-\usepackage{setspace}
-
-\usepackage{url}
-
-\usepackage{geometry}
+\addbibresource{bib.bib}
 
@@
 
-\renewcommand{\refname}{Bibliography}
+\maketitle
 
-\onehalfspacing
-\setlength{\headsep}{15mm}
-
-\thispagestyle{plain}
-
-\title{\Large Generative Dynamics of Supreme Court Citations: \\ Analysis with a New Statistical Model}
-%\author{}
- % Toggle % below to blind and unblind the manuscript
-%  \author[1]{Christian Schmid \thanks{cxs5700@psu.edu}}
-%   \author[2]{Ted Hsuan Yun Chen\thanks{thc126@psu.edu}}
-% \author[2]{Bruce A. Desmarais\thanks{bdesmarais@psu.edu}}
-% \author[1]{David R. Hunter \thanks{dhunter@stat.psu.edu}}
-% \affil[2]{Department of Political Science, Pennsylvania State University}
- %\affil[1]{Department of Statistics, Pennsylvania State University}
-
-\author{%
-  Christian S. Schmid\footnote{Department of Statistics, The Pennsylvania State University, schmid@psu.edu}%
-  \and Ted Hsuan Yun Chen \footnote{Department of Political Science, The Pennsylvania State University, thc126@psu.edu}%
-   \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
-%  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
-  }
-
-\maketitle
 \begin{abstract}
@@
 The dependence effects are as substantively and statistically significant as the effects of the exogenous covariates we include in the model, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.\footnote{
-This work was supported in part by NIH, 
-R01 AI36664-01. NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219. We would like to thank Rachael Hinkle and Benjamin Kassow for detailed comments on an earlier draft of this paper.}
+This work was supported in part by NSF grants SES-1558661, SES-1637089, SES-1619644, and CISE-1320219. }
+
+\keywords{US Supreme Court, majority opinions, citations, network, ERGM.}
 \end{abstract}
 
-\doublespacing
 \section{Introduction}
+\label{sec:overview}
 
 United States Supreme Court opinions exercise authority and influence, in part, through their roles as precedents affecting future jurisprudence in the US.
-The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained,hitt2016measuring}.
-Despite a considerable body of research that focuses on the way in which relevant precedents shape decision making on the Court, relatively little work has focused on understanding which past opinions are cited by an opinion.
+The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,gillman2001s,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained,pang2012endogenous}.
+Despite a considerable body of research that focuses on how precedents shape decision making on the Court, relatively little work has focused on understanding which past opinions are cited by an opinion.
 Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited in an opinion.
@@
 
-\begin{figure}[htp]
+\begin{figure}[bt]
 \centering
-\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{images/citations_trans.pdf}
-\caption{Illustration of transitive triangle connecting US Supreme Court opinions through citations.}
+\includegraphics[width = 0.49\textwidth,trim= 2cm 2cm 3cm 2cm,clip=true ]{citations_trans.pdf}\hspace{0.2cm}
+\includegraphics[width = 0.49\textwidth,trim= 2cm 2cm 3cm 2cm,clip=true ]{citations_recip.pdf}
+\caption{Illustrations of transitive triangle connecting US Supreme Court opinions through citations (left) and a reciprocal tie between two US Supreme Court opinions (right).}
 \label{fig:transitivity}
@@
 
-The second network property we consider in the context of Supreme Court citations is reciprocity.
-Reciprocity (also referred to as mutuality) is the tendency for node B to send a tie to node A in response to or coordination with A sending a tie to B \citep{garlaschelli2004patterns}.
-It is typically not possible for reciprocated ties to form in legal opinions.
-Most citations reference past opinions that were issued before the citing opinion's case was even argued before the Court.
-However, opinions written within the same Supreme Court term are often drafted in tandem, and can cite each other reciprocally.
-Opinion A citing opinion B within the same term represents a signal that opinion B is relevant to the legal reasoning underpinning opinion A.
-Unlike the citations themselves, the applicability of legal rules or lines of reasoning across cases is not directed---if A is relevant to B, B is highly likely to be relevant to A.
-We expect opinions written within the same term to exhibit a high degree of reciprocity.
-Below we provide two example passages from opinions that illustrate the phenomenon of within-term reciprocity.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{images/citations_recip.pdf}
-\caption{Illustration of a reciprocal tie between two US Supreme Court opinions.}
-\label{fig:reciprocity}
-\end{figure}
+% \begin{figure}[bt]
+% \centering
+% \includegraphics[width = 0.85\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{citations_recip.pdf}
+% \caption{Illustration of a reciprocal tie between two US Supreme Court opinions.}
+% \label{fig:reciprocity}
+% \end{figure}
 
@@
 353, 1985) is another case considering whether mandatory retirement violates the Age Discrimination in Employment Act.
-The mutual edge connecting these two cases is visualized in Figure \ref{fig:reciprocity}.
+The mutual edge connecting these two cases is visualized in Figure \ref{fig:transitivity}.
 These cases addressed very similar legal questions, which increased the likelihood that they would inform each other, and the opinions were written within the same term, which made it possible for them to cite each other.
@@
 
-\begin{figure}[ht]
+\begin{figure}[bt]
 \centering
-\includegraphics[width = 0.95\textwidth,trim= 0cm 3cm 6cm 8cm,clip=true ]{images/citations_pop.pdf}
+\includegraphics[width = 0.8\textwidth,trim= 1cm 4cm 1cm 3cm,clip=true ]{citations_pop.pdf}
 \caption{Illustration of ties sent to a landmark Supreme Court opinion via citations.}
@@
 We also seek to model the effects of covariates (i.e., case features) on citation formation.
-In the next section, we introduce a modeling framework that can incorporate all of these effects into a single model, and adapt to the peculiar structure of citation matrices.
+Next we introduce a modeling framework that can incorporate all of these effects into a single model, and adapt to the structural constraints of citation networks.
 
@@
 We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
-To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{duque2018recognizing,osei2018elite}.
+To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data, and has seen extensive application in recent political networks research---the exponential random graph model (ERGM) \citep[e.g.,][]{box2014evolution,duque2018recognizing,osei2018elite}.
 %The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
@@
 
-\begin{figure}[H]
+\begin{figure}[bt]
 \centering
-\includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 0cm,clip=true  ]{images/daggish.pdf}
+\includegraphics[width = 0.85\textwidth ]{daggish.pdf}
 \caption{Illustration of temporal structure of the Supreme Court Citation Network.
@@
 
-%We should note that there is nothing about the c-ERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another.
-For example, if we find a positive popularity effect, the term for that dynamic would indicate that there is an increase in the log odds that case $i$ cites case $j$ with each other case that has cited case $j$.
-That finding could be driven by one of two (or some combination of both) dynamics.
-First, it could be the case that a large volume of citations to a case in past opinions draws the attention of the justices to the ``popular'' case---as in the Baker v.
-Carr passage cited above.
-Second, it could be that we have omitted a case-level variable that is strongly associated with the number of citations accumulated by a case, and the number of citations in past cases is operating as a proxy for this unmeasured case characteristic.
-We cannot statistically differentiate between these two dynamics.
-However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-ERGM can be specified to account for the dependence in the data.
-Estimation of the c-ERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix.
-
-%Methods of Monte Carlo approximation are well established in the ERGM literature, \citep{hunter2006inference,van2009framework,hummel2012improving}, but with a number of nodes on the order of $10,000$, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}.
-However, because the network is large, we can rely on a fast estimation method---maximum pseudo-likelihood estimation (MPLE), which is consistent for discrete exponential family models \citep{nguyen2017near}, and has long been used with ERGM \citep{strauss1990pseudolikelihood}.
-The MPLE is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
-The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated above, assume a logit form.
-
-%One challenge that arises with the use of MPLE is that the usual estimate of the asymptotic covariance matrix (i.e., inverting the negated Hessian of the objective function at the maximum), results in standard errors that are biased downward.
-This leads to a high Type-1 error rate in hypothesis testing with MPLE if conventional standard errors are used \citep{van2009framework}.
-A form of non-parametric cluster bootstrap results in confidence intervals with correct coverage probability with MPLE \citep{desmarais2012statistical,desmarais2010consistent}.
-In the case of c-TERGM, the clusters are defined by $t$---the time points in which the cases are decided, and $T$ time-specific citation matrices $C_{t}$ are re-sampled with replacement to form a non-parametric cluster bootstrap sample.
-Confidence intervals are defined by the quantiles in the empirical distributions of estimates that result from re-estimating the MPLE on non-parametric cluster bootstrap samples.
-
  \section{Empirical Analysis}
@@
 
-\begin{table}[htp]
-\centering
-\begin{tabular}{|
->{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
-\hline
-{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Terms} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Total Number Cases} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Cases/Term} \\ \hline
-CE Hughes*              & 1937 - 1941                                          & 628                                                               & 125.6                                                     \\ \hline
-HF Stone                & 1942 - 1946                                          & 756                                                               & 151.2                                                     \\ \hline
-FM Vinson               & 1946 - 1953                                          & 789                                                               & 98.63                                                      \\ \hline
-E Warren                & 1954 - 1969                                          & 2149                                                              & 126.41                                                     \\ \hline
-WE Burger               & 1970 - 1986                                          & 2805                                                             & 155.83                                                     \\ \hline
-W Rehnquist            & 1987 - 2005                                          & 2022                                                              & 106.42                                                      \\ \hline
-J Roberts **           & 2006 - 2015                                          & 871                                                              & 87.1                                           \\ \hline
-\end{tabular}
-\caption[caption]{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
+\begin{table}[bt]
+\caption{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
 \\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 5/2020).}
 \label{tab:chiefs}
+\centering
+\begin{tabular}{l l l l r}
+\headrow \thead{Chief Justice} & \thead{Terms} & \thead{Total Number Cases} & \thead{Cases/Term}  \\
+CE Hughes*              & 1937 - 1941                                          & 628                                                               & 125.6                                                     \\
+HF Stone                & 1942 - 1946                                          & 756                                                               & 151.2                                                     \\
+FM Vinson               & 1946 - 1953                                          & 789                                                               & 98.63                                                      \\
+E Warren                & 1954 - 1969                                          & 2149                                                              & 126.41                                                     \\
+WE Burger               & 1970 - 1986                                          & 2805                                                             & 155.83                                                     \\
+W Rehnquist            & 1987 - 2005                                          & 2022                                                              & 106.42                                                      \\
+J Roberts **           & 2006 - 2015                                          & 871                                                              & 87.1                                           \\ 
+\end{tabular}
+
 \end{table}
@@
 These long-tailed (i.e., high kurtosis) distributions provide preliminary evidence of substantial heterogeneity in the features that drive citations to and from cases \citep{strogatz2001exploring}.
-Figure \ref{fig:networkviz} displays the citation network network from 1937--2001.
+Figure \ref{degree_dist} displays the citation network network from 1937--2015.
 We see here that the densest rates of tie formation tend to be between consecutive courts (e.g., the Stone Court is much more tightly tied to the Hughes Court than the Rehnquist Court is to the Hughes Court).
-This pattern lends preliminary/descriptive support to the hypothesis that the rate of citations to a case decrease over time.
+This pattern lends preliminary support to the hypothesis that the rate of citations to a case decrease over time.
 
-\begin{figure}[H]
-\includegraphics[scale=0.5]{degree_distribution}
-\caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2001. There are cases with an indegree  $>$50, but they are not captured in this figure.}
+\begin{figure}[bt]
+\centering
+\begin{minipage}{.37\linewidth}
+\centering
+\includegraphics[width = 0.99\textwidth]{degree_distribution.pdf}
+\includegraphics[width = 0.8\textwidth]{data_mat.png}%
+\end{minipage}%
+\begin{minipage}{.63\linewidth}
+\includegraphics[width = \textwidth]{courts_layer.png}
+% \includegraphics[width = \textwidth,clip=true,trim=0.5cm 3.75cm 0.75cm 0.5cm ]{courts_layer_updated.pdf}
+\end{minipage}
+\caption{Supreme Court Citation Network, 1937-2015.
+Network visualization on the right.
+Nodes are Supreme Court cases, color-coded based on the chief justice presiding over the court.
+On the top left is the in- and outdegree distribution of the network.
+There are cases with an in- or outdegree  $>$50, but they are not captured in this figure.
+The bottom left shows the citation data in adjacency matrix format following \autoref{fig:ctergm}.}
  \label{degree_dist}
-\vspace{-.25cm}
 \end{figure}
 
-\begin{figure}[H]
-\includegraphics[scale=0.35,clip=true,trim=.5cm 0cm 0cm 2cm]{images/citations1}
-\caption{ Supreme Court Citation Network, 1937-2001.
-Nodes are Supreme Court cases, with size based on incoming citations.
-Salient cases ({\em Oxford Guide}) are in yellow.
+% \begin{figure}[bt]
+% \centering
+% \includegraphics[scale = 0.45 ]{degree_distribution.pdf}
+% \caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2015. There are cases with an in- or outdegree  $>$50, but they are not captured in this figure.}
+%  \label{degree_dist}
+% \end{figure}
+
+% \begin{figure}[h]
+% \centering
+% \includegraphics[scale=0.20,clip=true,trim=.5cm 2cm 0cm 0cm ]{courts_layer_updated.pdf}
+% \caption{ Supreme Court Citation Network, 1937-2015.
+Nodes are Supreme Court cases, color-coded based on the chief justice presiding over the court.
 Each network layer contains cases decided under a different chief justice.
-Because our data is temporally bound, cases for the Hughes and Rehnquist courts are not complete.
+Because our data is temporally bound, cases for the Hughes and Roberts courts are not complete.
 }
- \label{fig:networkviz}
-\vspace{-.25cm}
-\end{figure}
+%  \label{fig:networkviz}\end{figure}
 
@@
 We include one covariate that accounts for the rate at which cases that fall under the same issue area cite each other.
-The variable $X^{ij}$ indicates whether cases $i$ and $j$ share the same issue area, e.g., $X^{ij}=1$ if cases $i$ and $j$ are classified into the same issue area, and $0$ otherwise.
+The variable $X_{ij}$ indicates whether cases $i$ and $j$ share the same issue area, e.g., $X_{ij}=1$ if cases $i$ and $j$ are classified into the same issue area, and $0$ otherwise.
 The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
@@
 
-\begin{table}[]
+\begin{table}[bt]
+\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
+\label{issue_area_coding}
 \centering
@@
 \end{tabular}
-\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
-\label{issue_area_coding}
 \end{table}
@@
 We model the way in which citations to a case depend upon the age of a case.
-For this we use a second-order polynomial in which one covariate $X^{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X^{ij}$ is the squared age of case $j$ at the time that case $i$ is decided.
+For this we use a second-order polynomial in which one covariate $X_{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X_{ij}$ is the squared age of case $j$ at the time that case $i$ is decided.
 We include these covariates because \citet{black2013citation} find that the number of citations to a Supreme Court case over time depends significantly on the age of the case, characterized by a sharp drop off and leveling out with age.
@@
 In our data we do not differentiate between negative and positive citations, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis.
-We include one covariate in which $X^{ij}$ is the number of justices in the majority coalition for case $j$.
-We also include another covariate in which $X^{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$.
-We expect both of these covariates to have positive effects.
+We include one covariate in which $X_{ij}$ is the number of justices in the majority coalition for case $j$.
+We also include another covariate in which $X_{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$.
+We expect these covariates to have positive effects.
 
@@
 
-%The log odds interpretation of the triangle statistic is that  $\theta$ gives the change in the log odds of case $i$ citing case $j$ associated  with each additional triangle that would be created by case $i$ citing case $j$ (e.g., if opinions in cases $i$ and $j$ cite one common case $k$, the log odds that case $i$ cites case $j$ change by theta relative to the configuration in which cases $i$ and $j$ do not cite any of the same cases).
-
- % \subsubsection{Dynamics in effects}
-
-% For the current model this section is not relevant anymore
+\begin{figure}[bt]
+\centering
+\includegraphics[width=10cm ]{SCC_AIC_BIC.pdf}
+\caption{AIC and BIC for the full and the independent model for the time frame 1950-2015 }
+ \label{AIC_BIC}
+\end{figure}
 
@@
 
-  \begin{figure}[H]
-\includegraphics[width=15cm]{SCC_AIC_BIC}
-\caption{AIC and BIC for the full and the independent model for the time frame 1950-2015 }
- \label{AIC_BIC}
-\vspace{-.25cm}
-\end{figure}
-
 Figure \ref{AIC_BIC} depicts the AIC and BIC comparison for the two different models for each term from 1950 to 2015.
@@
 We find that the full model fits the degree and edgewise shared partner distributions well, and is not degenerate.}
-
-%In our prediction experiment we randomly split the directed dyads into an 80\% training set and a 20\% test set.
-The parameters of the model are estimated using the directed dyads in the training set, and the parameters are used to form the conditional probability of a tie for all of the directed dyads in the test set.
-Directed dyads for which the conditional probability of a tie exceeds 0.5 are predicted to be citations.
-The experiment is run with the full model, and with a model that excludes all of the dependence terms (i.e., all terms involving reciprocity, in and out stars, and triangles)---the independent dyads model.
-We run this experiment for 10 iterations.
-Predictive performance is evaluated with three common and related measures---precision (i.e., the proportion of predicted citations that are actually citations), recall (the proportion of actual citations that are predicted to be citations), and the F1 score (the harmonic mean of precision and recall) \citep[see, e.g.,][for discussion of these measures]{makhoul1999performance}.
-All three measures are bounded between 0 and 1, with higher scores indicating better performance.
-
-%\begin{table}[H]
-%\centering
-
-%\begin{tabular}{rllll}
-%\hline \hline
-%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
-% & mean & range & mean & range \\ 
-%  \hline
-%precision & 0.6007 & (0.5934, 0.6074) & 0.8666 & (0.8647, 0.8701) \\ 
-%  recall & 0.1522 & (0.1495, 0.1549) & 0.5969 & (0.5945, 0.6019) \\ 
-%  F1 score & 0.2429 & (0.2389, 0.2462) & 0.7069 & (0.7048, 0.7107) \\ 
-%   \hline \hline
-%\end{tabular}
-%\caption{The predictive performance of the directed dyadic models with linear trends, over ten 80/20 train/test splits.}
-%\label{tab:prediction_linear}
-
-%\end{table}
-
-%We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms.
-The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set.
-This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.
-
-%\begin{table}[H]
-%\centering
-%\begin{tabular}{rllll}
-%\hline \hline
-%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
-% & mean & range & mean & range \\ 
-%  \hline
-%precision & 0.5938 & (0.5844, 0.5996) & 0.8668 & (0.8653, 0.8693) \\ 
-%  recall & 0.1551 & (0.1515, 0.1575) & 0.5969 & (0.5937, 0.6005) \\ 
-%  F1 score & 0.246 & (0.2406, 0.249) & 0.707 & (0.7045, 0.7095) \\  
-%   \hline \hline
-%\end{tabular}
-%\caption{The predictive performance of the directed dyadic models with no trends, over ten 80/20 train/test splits.}
-%\label{tab:prediction_notrend}
-%\vspace{0.4cm}
-%\begin{tabular}{rllll}
-%\hline \hline
-%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
-% & mean & range & mean & range \\ 
-%  \hline
-%precision & 0.6209 & (0.611, 0.6351) & 0.8669 & (0.8612, 0.8711) \\ 
-%  recall & 0.1511 & (0.1451, 0.1558) & 0.5962 & (0.5884, 0.6011) \\ 
-%  F1 score & 0.243 & (0.2347, 0.2486) & 0.7065 & (0.7011, 0.7103) \\ 
-%   \hline \hline
-%\end{tabular}
-%\caption{The predictive performance of the directed dyadic models with quadratic trends, over ten 80/20 train/test splits.}
-%\label{tab:prediction_quadratic}
-%\end{table}
-%We also conduct a predictive experiment for the model that does not consider any trends and for the model that does consider quadratic instead of linear trends.
-The predictive performance results are given in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic}.
-By comparing the full model results in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic} with the results of the full linear trend model in table \ref{tab:prediction_linear} we can conclude all three models perform similarly well.
-\\[0,5cm]
-
-%\textit{Comment from Benjamin Kassow: In section 6.2 (predictive performance), I like what is here, but was hoping to see a bit more discussion about the predictive 
-%performance of the model, with some explanation of what the predictive performance means. 
-%I understand it (I think), but depending on where this gets sent to, I am not sure if all reviewers will. 
-%I also think tying this into the overall model explanation or discussion of model performance overall will be helpful for better 
-%integrating this section into the rest of the manuscript.}
 
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=14cm ]{SCC_results_1.pdf}
+\caption{ERGM results for the dependence terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1. }
+ \label{SCC_results_1}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=14cm ]{SCC_results_2.pdf}
+\caption{ERGM results for the covariate terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1.}
+ \label{SCC_results_2}
+\end{figure}
+
 %The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes.
@@
 In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.
-
-%\begin{figure}[H]
-% \begin{tabular}{cc}
-
-%   (a) Reciprocity & (b) Transitivity \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\
-
-%    (c) Activity & (d) Popularity \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\
-
-%\end{tabular}
-%\caption{Trending effects for dependence effect estimates from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
-% \label{fig:coeftrends_endo}
-%\vspace{-.25cm}
-%\end{figure}
-
-%  \begin{figure}[H]
-%  \begin{tabular}{cc}
-% (e) MQ Score Difference & (f) Ideological Breadth \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\
-
-%  (g) Same Issue & (h) Majority Size \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\
-
-% (i) Difference in Years & (j) Difference in Years$^2$ \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\
-
-% (k) Overruled Cases & (l) Justice Homophily \\
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} & 
-%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/justicehomophily_coef_trend_linear.pdf} \\
-
-% \end{tabular}
-%\caption{Trending effects for exogenous variables from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
-% \label{fig:coeftrends_exo}
-%\vspace{-.25cm}
-%\end{figure}
-
-%\begin{itemize}
-%\item visualize Sender/receiver Issue Area Results
-%\item heatplot for issue area nodemix
-%\item visualize Sender/receiver for majority opinion writer
-%\item add Network effect plots for model without network statistics (add to existing plots?)
-%\end{itemize}
-
-%\newpage
-%   \thispagestyle{empty}
-
-  \begin{figure}[H]
-\includegraphics[width=16cm]{SCC_Results_1}
-\caption{ERGM results for the dependence terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1. }
- \label{SCC_results_1}
-
-\vspace{-.25cm}
-\end{figure}
-
-% \thispagestyle{empty}
- \begin{figure}[H]
-\includegraphics[width=16cm]{SCC_Results_2}
-\caption{ERGM results for the covariate terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1.}
- \label{SCC_results_2}
-\vspace{-.25cm}
-\end{figure}
 
@@
 
-%We recognize two important limitations of the current analysis, which should be addressed in future iterations of this work or in future studies.
-First, our theoretical claims and analyses treat all citations---positive and negative---as equivalent.
-Both data and methodological limitations will need to be overcome to consider these two citation types.
-Second, there is nothing in the c-ERGM that governs the generation of new cases (i.e., cases are assumed to arise exogenously, unrelated to the history of the citation network).
-As new cases are often taken up by the Supreme Court in order to test, clarify, and/or challenge existing precedents, it is likely inappropriate to assume that new cases arise exogenous to the existing citation network.
-Overcoming this limitation would require methodological innovation in the c-ERGM as well as theoretical development regarding exactly how citation structure might predict case emergence.
-
-\bibliography{bib} 
-\bibliographystyle{apsr}
-
+\printbibliography
 \end{document}

### 2020-08-14 Tex/PA_rrmemo/PA_RnR_memo.tex
wrote response to R1.1 and R1.2

@@
 \usepackage{graphicx, lipsum}
+\usepackage{bm}
+\usepackage[backend=bibtex]{biblatex}    
+\usepackage[english]{babel}
 
+\addbibresource{bib.bib}
 %% Edit the header section here. To include your
@@
 
-\noindent \textbf{Addressed:}   \\
+\noindent \textbf{Addressed:} All networks were estimated using MCMLE, however, for smaller networks we used the network's MPLE as the algorithms's starting value, while for larger networks the starting value was obtained using simulated annealing, since the MPLE as starting value did not lead to converged MCMLEs anymore.
+While in theory the algorithm should converge to the MLE regardless of the starting value, in practice this is not the case   \cite{hummel2012improving}.
+\\
+In order to be consistent with the used estimation approach, we estimate the MCMLE now for all networks using simulated annealing starting value. The results have not changed noticeably.  \\
 
@@
 
-\noindent \textbf{Addressed:}   \\
+\noindent \textbf{Addressed:} We have re-estimated all networks using the simulated annealing approach and the results look very similar to the initially submitted results.
+This is not too surprising, since the MCMLE algorithm only reports a succesful estimation if it has converged to the unique MLE.
+However, with a poorly chosen starting value the algorithm might fail to converge towards the MLE.
+This is what happened for some large networks using the MPLE.
+\\
 
@@
 
+\printbibliography
+
 \end{document}

### 2020-08-14 Tex/PA_appendix_rr/appendix.tex
Added and updated appendix

--- 2020-05-06 Tex/appendix.tex
@@
 \usepackage{amsthm}
-\usepackage{bm}
-
-\usepackage{natbib}
-
-\usepackage[table,xcdraw]{xcolor}
-
-\usepackage{graphicx}
-
+\usepackage[moderate]{savetrees}
 \usepackage{geometry}
@@
 
-\usepackage{setspace}
+\title{Appendix: Generative Dynamics of Supreme Court Citations: Analysis with a New Statistical Network Model}
 
-\usepackage{url}
+\author[1]{Christian S. Schmid}
+\author[2]{Ted Hsuan Yun Chen}
+\author[3]{Bruce A. Desmarais}
+
+\affil[1]{Department of Statistics, The Pennsylvania State University, University Park, PA. Email: \url{schmid@psu.edu}}
+\affil[2]{Faculty of Social Sciences, University of Helsinki, Finland. Email: \url{ted.hsuanyun.chen@gmail.com}}
+\affil[3]{Department of Political Science, The Pennsylvania State University, University Park, PA. Email: \url{bdesmarais@psu.edu}}
+
+%% Corresponding author
+\corrauthor{Christian S. Schmid}
+
+%% Abbreviated author list for the running footer
+\runningauthor{Schmid et al.}
+
+\addbibresource{bib.bib}
 
 \begin{document}
-
-\renewcommand{\refname}{Bibliography}
-
-\onehalfspacing
-\setlength{\headsep}{15mm}
-
-\thispagestyle{plain}
-
-\title{\Large \textbf{Appendix:} \\ Generative Dynamics of Supreme Court Citations: \\ Analysis with a New Statistical Model}
-%\author{}
- % Toggle % below to blind and unblind the manuscript
-%  \author[1]{Christian Schmid \thanks{cxs5700@psu.edu}}
-%   \author[2]{Ted Hsuan Yun Chen\thanks{thc126@psu.edu}}
-% \author[2]{Bruce A. Desmarais\thanks{bdesmarais@psu.edu}}
-% \author[1]{David R. Hunter \thanks{dhunter@stat.psu.edu}}
-% \affil[2]{Department of Political Science, Pennsylvania State University}
- %\affil[1]{Department of Statistics, Pennsylvania State University}
-
-\author{%
-  Christian S. Schmid\footnote{Department of Statistics, The Pennsylvania State University, schmid@psu.edu}%
-  \and Ted Hsuan Yun Chen \footnote{Department of Political Science, The Pennsylvania State University, thc126@psu.edu}%
-   \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
-%  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
-  }
 
@@
 
-The normalizing constant in Equation 1 is intractable.
-For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure \ref{fig:ctergm}---there are 16,777,216 unique configurations of $C_t$ that could be observed.
+The normalizing constant in Equation 1 of the main text is intractable.
+For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure 4 of the main text---there are 16,777,216 unique configurations of $C_t$ that could be observed.
 The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases.
@@
 The MPLE of such a network is similar to the MLE, making it an effective starting value for the MCMLE algorithm.
+Since for some networks the MCMLE was only obtainable using simulated annealing starting value, the final results in the paper have all been estimated using simulated annealing.
 
@@
 
-We see that our models do a good job capturing the out and indegree distribution of the citation network, since the black line falls almost exclusively within the ranges spanned by the boxplots.
-For the ESP distributions we can observe that the number of ties with $r=0$ shared partners is captured well for both the OTP as well as for the OSP statistic.
-However, the model overestimates the number of $r=1$ shared partners and then, especially in the 2015 term network, underestimates the number of ties with more than $r=1$ shared partners.
-
- \begin{figure}[H]
+\begin{figure}[bt]
  \begin{center}
-\includegraphics[width=13.8cm]{GOF_1950}\\
+\includegraphics[width=0.9\textwidth]{GOF_1950}\\
 %\caption{Goodness-of-fit diagnostic for the 1950 network.}
@@
 \vspace{1cm}
-\includegraphics[width=13.8cm]{GOF_2015}
+\includegraphics[width=0.9\textwidth]{GOF_2015}
 \vspace{0.1cm}
@@
 \end{figure}
+
+We see that our models do a good job capturing the out and indegree distribution of the citation network, since the black line falls almost exclusively within the ranges spanned by the boxplots.
+For the ESP distributions we can observe that the number of ties with $r=0$ shared partners is captured well for both the OTP as well as for the OSP statistic.
+However, the model overestimates the number of $r=1$ shared partners and then, especially in the 2015 term network, underestimates the number of ties with more than $r=1$ shared partners.
 
@@
 
-\bibliography{bib} 
-\bibliographystyle{apsr}
-
-\begin{figure}[H]
+\begin{figure}[bt]
  \begin{center}
-\includegraphics[width=14.5cm]{Deg_1950}
+\includegraphics[width=1\textwidth]{Deg_1950}
 \caption{Density and trace plots for the dependency terms of the 1950 term citation network.}
@@
 
- \begin{figure}[H]
+ \begin{figure}[bt]
  \begin{center}
-\includegraphics[width=14.5cm]{Deg_2015}
+\includegraphics[width=1\textwidth]{Deg_2015}
 \caption{Density and trace plots for the dependency terms of the 2015 term citation network.}
@@
 
+\newpage
+
+\printbibliography
 \end{document}

### 2020-08-16 RCode/Plot Figures/GOF.R
added updated SA results and code

@@ -16,7 +16,7 @@ year.total <- scc1[,11]-1936 #-1936 => the first year 1937 is 1, 1938 is 2 aso, 
 set.seed(12123)
 #sampled.years <- c(14, 24, 34, 44, 54, 64, 74)
 
-t<-  79 # 79 stands for 2015, 14 for 1950
+t<-  14 # 79 stands for 2015, 14 for 1950
 print(t+1936)
 
 
@@ -100,8 +100,8 @@ unfixed.esp.osp<- esp
 
 
 # load simulated models
-#load("GOF_Deg_1950.RData")
-load("2015_GOF_Deg_just_deg.RData")
+load("GOF_Deg_1950.RData")
+#load("2015_GOF_Deg_just_deg.RData")
 
 odeg.matrix <- li[[1]]
 ideg.matrix <- li[[2]]
@@ -189,5 +189,5 @@ plot(sim[,6], main="GWESP", col="grey45", xlab="Iterations", ylab="GWESP", cex.m
 abline(h=erg.est[6],lwd=2)
 
 
-title('Degeneracy Check 2015', outer=TRUE, cex.main=3)
+title('Degeneracy Check 1950', outer=TRUE, cex.main=3)
 

### 2020-08-16 RCode/Plot Figures/Plot_SCC.R

@@ -5,69 +5,69 @@
 #### Plot SCC Results
 ############################################
 
-load(file="1950_2.RData")
-load(file="1951_2.RData")
-load(file="1952_2.RData")
-load(file="1953_2.RData")
-load(file="1954_2.RData")
-load(file="1955_2.RData")
-load(file="1956_2.RData")
-load(file="1957_2.RData")
-load(file="1958_2.RData")
-load(file="1959_2.RData")
-load(file="1960_2.RData")
-load(file="1961_2.RData")
-load(file="1962_2.RData")
-load(file="1963_2.RData")
-load(file="1964_2.RData")
-load(file="1965_2.RData")
-load(file="1966_2.RData")
-load(file="1967_2.RData")
-load(file="1968_2.RData")
-load(file="1969_2.RData")
-load(file="1970_2.RData")
-load(file="1971_2.RData")
-load(file="1972_2.RData")
-load(file="1973_2.RData")
-load(file="1974_2.RData")
-load(file="1975_2.RData")
-load(file="1976_2.RData")
-load(file="1977_2.RData")
-load(file="1978_2.RData")
-load(file="1979_2.RData")
-load(file="1980_2.RData")
-load(file="1981_2.RData")
-load(file="1982_2.RData")
-load(file="1983_2.RData")
-load(file="1984_2.RData")
-load(file="1985_2.RData")
-load(file="1986_2.RData")
-load(file="1987_2.RData")
-load(file="1988_2.RData")
-load(file="1989_2.RData")
-load(file="1990_2.RData")
-load(file="1991_2.RData")
-load(file="1992_2.RData")
-load(file="1993_2.RData")
-load(file="1994_2.RData")
-load(file="1995_2.RData")
-load(file="1996_2.RData")
-load(file="1997_2.RData")
-load(file="1998_2.RData")
-load(file="1999_2.RData")
-load(file="2000_2.RData")
+load(file="1950_san.RData")
+load(file="1951_san.RData")
+load(file="1952_san.RData")
+load(file="1953_san.RData")
+load(file="1954_san.RData")
+load(file="1955_san.RData")
+load(file="1956_san.RData")
+load(file="1957_san.RData")
+load(file="1958_san.RData")
+load(file="1959_san.RData")
+load(file="1960_san.RData")
+load(file="1961_san.RData")
+load(file="1962_san.RData")
+load(file="1963_san.RData")
+load(file="1964_san.RData")
+load(file="1965_san.RData")
[…]

### 2020-08-17 RCode/Plot Figures/Plot_SCC.R
updated results figures
added background shading by justice; added horizontal line at 0; changed color to colorblind friendlier palette

@@ -148,7 +148,28 @@ grid.arrange(p1, p2, ncol=2, nrow=1)
 ## Plot Coefficients, Std.error and P-values
 ##################################################
 ##################################################
+# Defining Colour
+orng.c<-"#fcae91FF"
+purp.c<-"#7828a0FF"
+gren.c<-"#46aa96FF"
+lgry.c<-"#dcdcdcFF"
+mgry.c<-"#8c8c8cFF"
+dgry.c<-"#333333FF"
 
+mkrzywinski = c('#006E82FF', 
+                '#8214A0FF',
+                '#005AC8FF',
+                '#00A0FAFF', 
+                '#FA78FAFF',
+                '#14D2DCFF',
+                '#AA0A3CFF',
+                '#FA7850FF',
+                '#0AB45AFF',
+                '#F0F032FF',
+                '#A0FA82FF',
+                '#FAE6BEFF')
+
+mkrzywinski[c(1, 8, 7)] <- c('#0098b3', '#f85320', '#910833')
 
 ## loop to create a matrix with all coefficients, matrix of std.errors and matrix of p-values
 
@@ -193,77 +214,114 @@ P[is.na(P)]<- 23
 
 steM[is.na(steM)] <- 0
 
+# Old plotting colors toggle
+# mkrzywinski[c(1, 8, 7)] <- c('springgreen4', 'orange1', 'red')
 
 data<- data.frame(Year, coM[1,], P[1,])
 colnames(data)<- c('Year', 'Theta', 'pi')
 eb <- aes(ymax = coM[1,]+1.96*steM[1,], ymin = coM[1,]-1.96*steM[1,])
-p1<- ggplot(data = data, aes(x = Year, y = Theta, shape=factor(pi), group=1)) +ggtitle('Edges')+
+p1<- ggplot(data = data, aes(x = Year, y = Theta,  group=1)) +ggtitle('Edges')+
+  annotate("rect", fill = grey(0.5), alpha = 0.15, 
+           xmin = c(1953, 1986) + 0.5, xmax = c(1969, 2005) + 0.5,
+           ymin = -Inf, ymax = Inf)+
+  geom_ribbon(eb, fill = grey(0.7))+
   geom_line() + 
-  geom_ribbon(eb, alpha = 0.3)
-p1<-p1+scale_colour_manual(values=c('21'='springgreen4', '24'='orange1', '22'='red', '23'='black'))
[…]

### 2020-08-17 NetworkVisualizations/visualizations_PA.R
updated network visualization script

@@ -1,0 +1,241 @@
+#----Code to produce citation network plot----#
+# The 'network' package is required for the force-directed algorithm on line 57
+
+#----Preparation----#
+# Set Directory
+rm(list=ls())
+setwd("")
+
+# Loading Data
+#load("SCOTUS.RData")
+#rm(mq.matrix,Overruled.matrix,same.issue.area,SCDB_justice,year.diff.matrix, cluster.mat, MQ.scores, overruled, overruled_cases_supremecourt, same.opinion.writer, scc)
+
+# Defining Colour
+orng.c<-"#fcae91FF"
+purp.c<-"#7828a0FF"
+gren.c<-"#46aa96FF"
+lgry.c<-"#dcdcdcFF"
+mgry.c<-"#8c8c8cFF"
+dgry.c<-"#333333FF"
+
+mkrzywinski = c('#006E82FF', 
+                '#8214A0FF',
+                '#005AC8FF',
+                '#00A0FAFF', 
+                '#FA78FAFF',
+                '#14D2DCFF',
+                '#AA0A3CFF',
+                '#FA7850FF',
+                '#0AB45AFF',
+                '#F0F032FF',
+                '#A0FA82FF',
+                '#FAE6BEFF')
+
+# Preparing Network
+# Remove error citations
+adjacency.matrix[8822, 9409] <- 0
+adjacency.matrix[8822, 9540] <- 0
+
+# Remove Isolates (there aren't many (352), and it helps with the rest of the visualization)
+isolates <- which((rowSums(adjacency.matrix) + colSums(adjacency.matrix)) == 0)
+mat <- adjacency.matrix[-isolates, -isolates]
+rm(adjacency.matrix)
+
+# Defining layers
+justices <- unique(scc1$chief[-isolates])
+n.layer <- length(justices)
+justice.id <- as.numeric(factor(scc1$chief[-isolates], levels = justices))
+rm(scc1)
+
+# Partitioning matrix into multilayer blocks
+mat.block <- mat
+for(i in 1:n.layer){
+  #id the intralayer edges
+  mat.block[which(justice.id == i), which(justice.id == i)] <- mat[which(justice.id == i), which(justice.id == i)] * i
+  
+  if(i == n.layer){break} else {
+    for(j in (i + 1):n.layer){
[…]

### 2020-08-17 RCode/Plot Figures/Plot_SCC.R
fixed results plot horizontal spacing

@@ -233,7 +233,7 @@ p1<- ggplot(data = data, aes(x = Year, y = Theta,  group=1)) +ggtitle('Edges')+
   scale_color_manual(values=c('21'=mkrzywinski[1], '24'=mkrzywinski[8], '22'=mkrzywinski[7], '23'='black'))
 
 p1<-p1+theme_minimal() + theme(legend.position='none', axis.text=element_text(size=10),axis.title=element_text(size=14),
-                               plot.title=element_text(size=16, face='bold'), plot.margin = unit(c(0.1,0,0.1,0), "cm"))
+                               plot.title=element_text(size=16, face='bold'), plot.margin = unit(c(0.1,0,0.1,0.1), "cm"))
 p1
 
 
@@ -260,7 +260,7 @@ p2<- ggplot(data = data, aes(x = Year, y = Theta,  group=1)) +ggtitle('Reciproci
   scale_color_manual(values=c('21'=mkrzywinski[1], '24'=mkrzywinski[8], '22'=mkrzywinski[7], '23'='black'))
 
 p2<-p2+theme_minimal() + theme(legend.position='none', axis.text=element_text(size=10),axis.title=element_text(size=14),
-                               plot.title=element_text(size=16, face='bold'), plot.margin = unit(c(0.1,0,0.1,0), "cm"))
+                               plot.title=element_text(size=16, face='bold'), plot.margin = unit(c(0.1,0,0.1,0.1), "cm"))
 p2
 
 
@@ -282,7 +282,7 @@ p3<- ggplot(data = data, aes(x = Year, y = Theta,  group=1)) +ggtitle('Receiver 
   scale_color_manual(values=c('21'=mkrzywinski[1], '24'=mkrzywinski[8], '22'=mkrzywinski[7], '23'='black'))
 
 p3<-p3+theme_minimal() + theme(legend.position='none', axis.text=element_text(size=10),axis.title=element_text(size=14),
-                               plot.title=element_text(size=16, face='bold'), plot.margin = unit(c(0.1,0,0.1,0), "cm"))
+                               plot.title=element_text(size=16, face='bold'), plot.margin = unit(c(0.1,0,0.1,0.1), "cm"))
 p3
 
 data<- data.frame(Year, coM[4,], P[4,])
@@ -301,7 +301,7 @@ p4<- ggplot(data = data, aes(x = Year, y = Theta,  group=1)) +ggtitle('Different
[…]

### 2020-08-17 Tex/PA_rr/main.tex
fixed clashes with PA class
Original main.tex had settings that clashed with the PA class. These were ignored by overleaf but miktex (20.6) was not happy.
New Plots using only simulated annealing
updated description of results figures

@@
 \usepackage{bm}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{geometry}
+%\usepackage{amssymb}
+%\usepackage{amsmath}
+%\usepackage{amsthm}
+%\usepackage{geometry}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
 
@@
 The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in the respective variable.
+In our figures, a circle indicates that the estimate was significant at an $\alpha=0.05$ level, while an square translates to a p-value between $0.05$ and $0.1$.
+A triangle signals that the estimate for a given year was not significant at an $\alpha=0.1$ level.
+In each panel, the background shading indicates different chief justice courts, with the two darker shades representing, respectively, the Warren (1954-1969) and Rehnquist (1987-2005) courts.
+Finally, note that the panels showing results of the reciprocity statistic and the \textit{Overruled Cases} variable do not provide estimates for every term.
+In these missing terms there were no two cases that cited each other or no overruled cases being cited, making the estimate of these terms equivalent to $-\infty$.
+%For those years, the reciprocity term is included in the model, but the estimate of $-\infty$.
+
 We first discuss the dependence effect estimates in Figure \ref{SCC_results_1}.
-A circle indicates that the estimate was significant at an $\alpha=0.05$ level, while and square translates to a p-value between $0.05$ and $0.1$.
-A triangle signals that the estimate for a given year was not significant at an $\alpha=0.1$ level.
-The panel showing the results of the reciprocity statistic does not provide estimates for every term.
-In these missing terms there were no two cases that cited each other, making the estimate of the reciprocity statistic term equivalent to $-\infty$.
-%For those years, the reciprocity term is included in the model, but the estimate of $-\infty$.
-A similar scenario is visible for the \textit{Overruled Cases} variable. In the time period from 1953-1956 no cases that have been overruled have been cited which results in a non-estimable effect.
-
-We first discuss our findings regarding the dependence terms.
 In most of the years that it is estimable, the reciprocity effect is statistically significantly positive, and substantial in magnitude.
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
 \includegraphics[width=14cm ]{SCC_results_1.pdf}
-\caption{ERGM results for the dependence terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1. }
+\caption{ERGM results for the dependence terms.
+Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1.
+Different chief justice terms are indicated by shading in the background; the two grey areas indicate the Warren and Rehnquist courts.}
  \label{SCC_results_1}
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering
 \includegraphics[width=14cm ]{SCC_results_2.pdf}
-\caption{ERGM results for the covariate terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1.}
+\caption{ERGM results for the covariate terms.
+Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1.
+Different chief justice terms are indicated by shading in the background; the two grey areas indicate the Warren and Rehnquist courts.}
  \label{SCC_results_2}

### 2020-08-17 Tex/PA_rr/main.tex
changed all log-odds to log odds

@@
 where $C_{ij,t} = 1$ indicates that case $i$ cites case $j$, $C_{ij,t} = 0$ indicates that case $i$ does not cite case $j$, $C_{-ij,t}$ is the observed elements of $C_{t}$ except $C_{ij,t}$, and $\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)$ is the change in $\bm{h}(C_{t},C_{<t})$ that results from toggling $C_{ij,t} = 0$ to $C_{ij,t} = 1$.
-This re-arrangement illustrates that the parameters can be interpreted in terms of the change in the log odds of a citation from $i$ to $j$ given a one-unit increase in the corresponding element of $\bm{h}$, conditional on the other citations observed in the network.
-For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log odds of observing $C_{ij,t} = 1$ increases by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$).
+This re-arrangement illustrates that the parameters can be interpreted in terms of the change in the log-odds of a citation from $i$ to $j$ given a one-unit increase in the corresponding element of $\bm{h}$, conditional on the other citations observed in the network.
+For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log-odds of observing $C_{ij,t} = 1$ increases by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$).
 The logit form conditional probability is well known for the ERGM family \citep{goodreau2009birds}.
@@
 $$h_{covariate}(C_t,C_{<t},X) =  \sum_{ij} C_{ij,t}X_{ij}.$$ Since $C_{ij,t}$ is a binary indicator of whether case $i$ cites case $j$, $h_{covariate}$ amounts to the sum of covariate values among directed dyads for which we observed a citation.
-The dyadic interpretation of the coefficient attributed to this term is the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in $X_{ij}$ (i.e., exactly the interpretation of the effect of a covariate in logistic regression).
+The dyadic interpretation of the coefficient attributed to this term is the change in the log-odds of a tie from $i$ to $j$ given a one-unit increase in $X_{ij}$ (i.e., exactly the interpretation of the effect of a covariate in logistic regression).
 We include several exogenous covariates based on this standard formulation.
@@
 In practice, this is the number of mutual dyads in $C_{\leq t}$, since mutual dyads must emerge among opinions that were written in the same term.
-The log odds interpretation of the reciprocity statistic is that if the opinion in case $i$ cites case $j$, the log odds that case $j$ also cites $i$ change by the estimated coefficient relative to the configuration in which case $j$ does not cite case $i$.
+The log-odds interpretation of the reciprocity statistic is that if the opinion in case $i$ cites case $j$, the log-odds that case $j$ also cites $i$ change by the estimated coefficient relative to the configuration in which case $j$ does not cite case $i$.
 We expect the reciprocity effect to be positive.
@@
  \ell \in \{0, \dots , N_t -1 \} \setminus \{r, r-1\} $ are
-made. This changes the log odds in the following way
+made. This changes the log-odds in the following way
 $$\text{logit} P(C_{ij,t}=1 | C_{-ij,t}, C_{<t}) = \exp \bigl(~\theta \cdot( 1- \frac{1}{\lambda})^r \bigr).$$
@@
 In these figures we depict coefficient estimates and 95\% confidence intervals for all of the effects for all the variables we discussed in sections \ref{covariate_terms} and \ref{dependence_terms}.
-The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in the respective variable.
+The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log-odds of a tie from $i$ to $j$ given a one-unit increase in the respective variable.
 In our figures, a circle indicates that the estimate was significant at an $\alpha=0.05$ level, while an square translates to a p-value between $0.05$ and $0.1$.
@@
 In a typical year, the presence of a citation from case $i$ to case $j$ increases the log-odds of a citation from case $j$ to case $i$ by approximately 0.50.
-Both transitivity effects---different term transitivity, and gwesp---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect.
+Both transitivity effects---different term transitivity and gwesp---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect.
 The log-odds of a citation from case $i$ to case $j$ increases by more than 2 if there is at least one third case, $k$ that is cited by $i$, and cites to $j$.

### 2020-08-18 Tex/PA_rr/main.tex
expanded on tempral component of hub effect

--- 2020-08-17 Tex/PA_rr/main.tex
@@
 In most of the years that it is estimable, the reciprocity effect is statistically significantly positive, and substantial in magnitude.
-In a typical year, the presence of a citation from case $i$ to case $j$ increases the log-odds of a citation from case $j$ to case $i$ by approximately 0.50.
-Both transitivity effects---different term transitivity, and gwesp---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect.
-The log-odds of a citation from case $i$ to case $j$ increases by more than 2 if there is at least one third case, $k$ that is cited by $i$, and cites to $j$.
-Results regarding the popularity effect (i.e., gwidegree) are generally supportive of the hypothesis of a popularity effect in Supreme Court citations, as the gwidegree estimate is negative in every term.
-There are some terms for which the result is not statistically significant at conventional levels.
-The popularity effect looks to have grown substantially stronger after the 2000 term.
+In a typical year, the presence of a citation from case $i$ to case $j$ increases the log odds of a citation from case $j$ to case $i$ by approximately 0.50.
+Both transitivity effects---different term transitivity and gwesp---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect.
+The log odds of a citation from case $i$ to case $j$ increases by more than 2 if there is at least one third case, $k$ that is cited by $i$, and cites to $j$.
+Results regarding the popularity effect (i.e., gwidegree) are generally supportive of the hypothesis of a popularity effect in Supreme Court citations, as the gwidegree estimate is negative and statistically significant at conventional levels in nearly every term.
+This effect looks to have grown substantially after the 2000 term, which also coincides closely with the beginning of the Roberts court in 2006.
+
 Contrary to expectations, we find that the hub effect is negative.
 That is, each additional citation sent by a case reduces the likelihood that the case is cited.
-For every citation sent, the log-odds of any citation to a case decreases by approximately 0.025---a result that is statistically significant in each term.
+For every citation sent, the log odds of any citation to a case decreases by an average of approximately 0.025---a result that is statistically significant in each term.
+This negative effect of receiver outdegree, notably, appears to be decreasing with time.
+Specifically, term to term variation aside, there is an upward trend toward zero in the parameter estimates.
+We believe this finding is particularly illustrative of why our approach of fitting separate models for each term is appropriate.
+As time passes, the number of nodes in the network (i.e., cases) grows.
+With the increasing number of citable cases and a relatively stable baseline rate of citation (i.e., the edges effect), there is an increase in outdegree mean and variation.
+This increasing difference between outdegree count by case makes it unlikely for the per unit change effect in outdegree to remain at a constant level.
+We have then an instance of network growth leading to changes in the model parameters even if the underlying network process remains unchanged.
 

### 2020-08-20 Tex/PA_rr/main.tex
expanded on covariate term write up

@@
 
+\subsubsection{Dependence Terms}
 We first discuss the dependence effect estimates in Figure \ref{SCC_results_1}.
@@
 
-The effects of the exogenous covariates included in the c-ERGM are presented in Figure \ref{SCC_results_2}.
-Most of the exogenous covariate effects align with expectations.
-We find that cases are more likely to cite those that (1) have been decided most recently, (2) were decided by large majorities, (3) have not been overruled, (4) were authored by the same justice, and (5) are in the same issue area.
-Surprisingly, we do not find consistent effects of either of the covariates that involve justice ideology.
-The ideological distance between two cases, as measured by the absolute difference in the median of the majority coalitions; and the ideological breadth of the cited cases, as measured by the absolute difference between the most liberal and most conservative justices on the potential receiving case, both exhibit signs and significance levels that are highly volatile.
-Our results leave open the question of whether, and if so, how, citations are shaped by ideological factors.
+\subsubsection{Covariate Terms}
+Effects of the exogenous covariates included in the c-ERGM are presented in Figure \ref{SCC_results_2}.
+Most of the exogenous covariate effects align with expectations and have relatively stable parameter estimates over time.
+We find that cases are more likely to cite those that (1) have been decided most recently, (2) were authored by the same justice, (3) are in the same issue area, and (4) have not been overruled.
+Surprisingly, we do not find consistent effects for majority coalition size or for the two covariates that involve justice ideology.
+The effect of majority coalition size was generally not statistically significant until the later portion of the period examined; ideological distance between two cases had signs and significance levels that are highly inconsistent; and the ideological breadth of the potentially cited case was effectively zero for the entire period.
+
+These results leave open the question of whether and how citations are shaped by ideological factors.
+While we did not formally hypothesize about temporal dynamics in the effects, a closer look at the observed trends allows us to identify some patterns that suggest directions for future research in this area.
+In this discussion, we focus on the shift in the effect of ideological distance beginning with the Rehnquist court and extending into the Roberts court.
+
+The existing body of evidence indicates that chief justices impact the courts they preside over beyond casting one of the nine votes \citep{cross2005decisional,danelski2016chief}.
+It is therefore within expectations that there are discernible differences in citation patterns across different courts.
+With regard to the effect of ideological distance, we see that the Vinson (1950-1953) and Warren (1954-1969) courts were highly volatile, the Burger court (1970-1986) had a generally higher likelihood of citing opinions that were ideologically closer to itself, and the more recent Rehnquist (1987-2005) and Roberts (2006-2015) courts showed some tendency toward citing cases that are ideologically distant.
+The latter finding might be initially puzzling, but can be explained in light of observations from legal scholars that the Rehnquist court was characterized by a `split-the-difference' jurisprudence toward its later half \citep{wilkinson2005rehnquist,basiak2006roberts}.
+This means that the court strove to take moderate positions that compromised on both sides of the debated issues, which in large differed from the jurisprudential approach of prior courts.
+Less has been written about the Roberts court on this topic, but Roberts shares many similarities with Rehnquist especially in comparison to previous chief justices, and has himself noted the desirability of widely accepted rulings even at the expense of their scope \citep{pomerance2018center,sunstein2008trimming}.
+
+One way to achieve the kind of balance required for support is to cite broadly from both sides of the argument \citep{wilkinson2005rehnquist}, which increases the likelihood of citing ideologically distant cases.
+Consider for example Hamdi v.
+Rumsfeld (542 U.S.
+1, 2004) which ruled on the question of whether a U.S.
+citizens can be detained as an ``enemy combatant''.
+The court, in a four-justice plurality opinion supported by two additional concurring justices, decided at the same time that ``a state of war is not a blank check for the President'' and the due process assessment must ``pay keen attention to the particular burdens faced by the Executive in the context of military action''.
+In total, the plurality opinion in Hamdi v.
+Rumsfeld cited 47 opinions, with cites as distant as 2.42 and 2.13 on either side of its median position.
+
+In the above discussion, we identified an explanation for the observed temporal dynamics in the effect of ideological distance.
+Specifically, ideologically distant cases become more important as courts adopt the relatively new `split-the-difference' jurisprudential approach which requires broad citations.
+We recognize that there are alternative explanations---perhaps one that is related to the coterminous change in the effect of majority coalition size---but a deeper look into the relevant literature is beyond the scope of this paper.
+Our discussion serves to illustrate that term-by-term c-ERGMs can uncover important temporal dynamics that courts researchers can use to study a much greater range of citation-related phenomenon beyond what we present here.
+To the extent that the US Supreme Court is a political entity that interacts with the broader society, the body of court opinions and citations between them is corpus that continues to grow based on both its history and its contemporary context.
+The methods we present here afford researchers the tools to better understand these complexities.
 

### 2020-08-21 RCode/EDA.R
added code to get ideological distance info (paper sec 4.3)

@@ -103,3 +103,13 @@ roberts<- sum(scc1$Roberts==1) # 871
 roberts/10 #87.1
 
 
+# Some additional exploration
+cid <- which(scc1$usCite == '542 U.S. 507') # Getting row/col for specific case (this one is Hamdi v Rumsfeld)
+mqs <- mq.matrix[which(adjacency.matrix[cid,] == 1), cid] # Getting all the ideological distances to the cited opinions
+range(mqs)
+
+# Change first argument to subset data (currently rehnquist court); change function to get statistic of interest (currently range)
+mqdist <- sapply(which(scc1$Rehnquist == 1), function(x){range(mq.matrix[which(adjacency.matrix[x,] == 1), x])})
+plot(density(mqdist, na.rm = T)) # Plotting
+abline(v = range(mqs))
+

### 2020-08-25 Tex/PA_rr/main.tex
extended negative hub effect explanation

@@
 Results regarding the popularity effect (i.e., gwidegree) are generally supportive of the hypothesis of a popularity effect in Supreme Court citations, as the gwidegree estimate is negative and statistically significant at conventional levels in nearly every term.
-This effect looks to have grown substantially after the 2000 term, which also coincides closely with the beginning of the Roberts court in 2006.
 
@@
 For every citation sent, the log odds of any citation to a case decreases by an average of approximately 0.025---a result that is statistically significant in each term.
-This negative effect of receiver outdegree, notably, appears to be decreasing with time.
+We initially expected receiver outdegree to have a positive effect because opinions with broad scopes can be useful citation hubs.
+Instead, our finding indicate that high outdegree more importantly reflects case characteristics that negatively impact incoming citations.
+One likely characteristic is case complexity.
+Prior studies show that justices tend to take a more balanced approach when faced with complex cases \citep{collins2008amici, lindquist2007splitting}, which can yield opinions that cite widely \citep{wilkinson2005rehnquist}.
+At the same time, it has been shown that complex cases are more likely to be overruled \citep{spriggs2001explaining}.
+To the extent, then, that outdegree count captures case complexity, our finding suggests that justices cite more direct alternatives in anticipation of complex opinions being overruled.
+
+The negative effect of receiver outdegree, which seems to result from strategic anticipation by the justices, ostensibly decreases with time.
 Specifically, term to term variation aside, there is an upward trend toward zero in the parameter estimates.
@@
 We have then an instance of network growth leading to changes in the model parameters even if the underlying network process remains unchanged.
+Approaches that cannot encompass these complexities are likely to produce poor fitting models.
 

### 2020-08-25 Tex/PA_appendix_rr/appendix.tex
included software information in main and appendix and explained SA algorithm in more detail

@@
 Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}.
-The MCMLE for these large citation networks was obtained by setting the starting value according to an approach introduced by \citet{SchmidHunter2020}.
+The MCMLE for these large citation networks was obtained by setting the starting value according to an novel approach introduced by \citet{SchmidHunter2020}.
 This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$.
 However, the MPLE of networks with the same sufficient statistics is not necessarily the same.
-Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a network $C_t^*$ using simulated annealing \citep{Kirkpatrick83}, with $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$ and a weak dependence structure among unfixed ties.
-The MPLE of such a network is similar to the MLE, making it an effective starting value for the MCMLE algorithm.
+Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a new network $C_t^*$ on the same nodes as the observed network that satisfies $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$, and has a weak dependence structure among unfixed ties.
+Such a network can be found using simulated annealing algorithms \citep{Kirkpatrick83}.
+For networks with a weak dependence structure among unfixed ties, the MPLE is similar to the MLE.
+In addition, the same sufficient statistic between $C_t^*$ and the observed network guarantees the same MLE between the two networks.
+This makes the new network's MPLE an effective starting value for the MCMLE algorithm.
 Since for some networks the MCMLE was only obtainable using simulated annealing starting value, the final results in the paper have all been estimated using simulated annealing.
+The simulated annealing algorithm for finding an improved starting value for c-ERGMs was implemented in the \textbf{cERGM}-package for \textbf{R} \citep{RCore} and can be found at \url{http://github.com/schmid86/cERGM}.
 

### 2020-08-25 Tex/PA_rrmemo/PA_RnR_memo.tex
Added and updated appendix
updated memo to reflect expanded results in paper

@@
 In order to be consistent with the used estimation approach, we estimate the MCMLE now for all networks using simulated annealing starting value. The results have not changed noticeably.  \\
+The text in the appendix has been updated accordingly.\\
 
@@
 Otherwise, wouldn’t it make sense to discuss these over time changes, at least when they appear to be quite meaningful?
-}\\
-
-\noindent \textbf{Addressed:}   \\
-
-\noindent \textbf{R1.4:} \emph{ Not to belabor the point, but the motivation for the exogenous and endogenous terms are nicely situated in the literature, but the discussion of the results are quite anticlimactic.
+Not to belabor the point, but the motivation for the exogenous and endogenous terms are nicely situated in the literature, but the discussion of the results are quite anticlimactic.
 It merely notes that some results are generally in line with expectations while others are not (the ideology finding being terribly interesting and surprising!).
@@
 
-\noindent \textbf{Addressed:}   \\
+\noindent \textbf{Addressed:} We thank the reviewer for pushing us to expand on our results; we believe the extended discussion greatly improves the paper.
+We changed the results section in two main ways (pp.
+13-14).
 
-\noindent \textbf{R1.5:} \emph{ While this is a well-executed paper (and visually beautiful), it’s not clear where the contribution is meant to be.
+First, we situated our more unexpected findings (i.e.
+negative outdegree count and ideological distance) in the courts and legal studies literatures.
+For these findings, we offer explanations that comport with the field's understanding of justice behavior.
+The hub effect is negative because outdegree count reflects case complexity, and justices avoid citing these opinions in anticipation of them being overruled.
+Judges are more likely to cite ideologically distant cases because they increasingly attempt to provide more balanced opinions.
+
+Second, we extend this discussion to account for temporal dynamics.
+Using these same effects, we illustrate how our approach of fitting term-by-term models allows us to better capture the complexities that arise from both the temporal context (e.g.
+chief justice characteristics) and from the network growth characteristic of citation networks.\\
+
+\noindent \textbf{R1.4:} \emph{ While this is a well-executed paper (and visually beautiful), it’s not clear where the contribution is meant to be.
 Is this paper an introduction of a new model or an application of the ERGM to citation data?
@@
 
-\noindent \textbf{R1.6:} \emph{ Alternatively, one could think about this paper more in terms of its substantive contributions.
+\noindent \textbf{R1.5:} \emph{ Alternatively, one could think about this paper more in terms of its substantive contributions.
 The front end is deeply ingrained in the citations and judicial literatures, as are the data, of course, but little attention is given in the end to what we learn from this paper.

### 2020-08-25 Tex/PA_rr/main.tex
fixed problem with missing parenthesis and sum signs
added missing tex-lines

@@
 \usepackage{bm}
-%\usepackage{amssymb}
-%\usepackage{amsmath}
+%\usepackage{lmodern}
 %\usepackage{amsthm}
@@
 \usepackage[T1]{fontenc}
-\usepackage{lmodern}
+\usepackage{amssymb}
+\usepackage{amsmath}
+
+% original loaded packages
+%\documentclass{cup-pan}
+%\usepackage[utf8]{inputenc}
+%\usepackage{blindtext}
+%\usepackage{bm}
+%\usepackage[T1]{fontenc}
+%\usepackage{lmodern}
 
@@
 Estimation of the c-ERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix.
+The basis of our computation is the \textbf{ergm}-package \citep{HunterDavidR..2008} in \textbf{R} \citep{RCore}.
+In addition, a wrapper \textbf{R}-package called \textbf{cERGM} was developed to conveniently fit citation network data using the c-ERGM\footnote{The \textbf{cERGM}-package can be found on \url{http://github.com/schmid86}.}.
 

### 2020-08-25 Tex/PA_appendix_rr/appendix.tex
addressed R2.2 and R2.6

@@
 Such a network can be found using simulated annealing algorithms \citep{Kirkpatrick83}.
-For networks with a weak dependence structure among unfixed ties, the MPLE is similar to the MLE.
-In addition, the same sufficient statistic between $C_t^*$ and the observed network guarantees the same MLE between the two networks.
-This makes the new network's MPLE an effective starting value for the MCMLE algorithm.
-Since for some networks the MCMLE was only obtainable using simulated annealing starting value, the final results in the paper have all been estimated using simulated annealing.
+For networks with a weak dependence structure among unfixed ties, the MPLE is similar to the MLE, in addition, the same sufficient statistic between $C_t^*$ and the observed network $C_t$ guarantees the same MLE between these two networks.
+This makes the MPLE of $C_t^*$ an effective starting value for the MCMLE algorithm.
+Since for some networks the MCMLE was only obtainable using simulated annealing method to find a starting value, the final results in the paper have all been estimated using simulated annealing.
 The simulated annealing algorithm for finding an improved starting value for c-ERGMs was implemented in the \textbf{cERGM}-package for \textbf{R} \citep{RCore} and can be found at \url{http://github.com/schmid86/cERGM}.

### 2020-08-25 Tex/PA_rrmemo/PA_RnR_memo.tex
updated courts literature to judicial politics literature
addressed R2.2 and R2.6

@@
 \noindent \textbf{Addressed:} We have re-estimated all networks using the simulated annealing approach and the results look very similar to the initially submitted results.
-This is not too surprising, since the MCMLE algorithm only reports a succesful estimation if it has converged to the unique MLE.
+This is not surprising, since the MCMLE algorithm only reports a succesful estimation if it has converged to the unique MLE.
 However, with a poorly chosen starting value the algorithm might fail to converge towards the MLE.
-This is what happened for some large networks using the MPLE.
+This is what happened for some large networks using the MPLE as starting value.
 \\
@@
 First, we situated our more unexpected findings (i.e.
-negative outdegree count and ideological distance) in the courts and legal studies literatures.
+negative outdegree count and ideological distance) in the judicial politics and legal studies literatures.
 For these findings, we offer explanations that comport with the field's understanding of justice behavior.
@@
 \noindent \textbf{Addressed:}  \\
+We discuss implementation details, especially the simulated annealing approach in greater detail in the appendix now.
+In addition, we now provide information in the main paper about the used statistical software.
+A new R-package called cERGM was developed to conveniently fit cERGMs for citation network data.
+The existence of this novel R-package and the corresponding hyperlink to the R-package are being mentioned in the main paper as well as in the appendix.
+\\
 
@@
 
-\noindent \textbf{Addressed:}  \\
+\noindent \textbf{Addressed:} We now mention in the paper that we us \textbf{R} for the implementation of our model and that the code is based on the \textbf{ergm}-package.
+Furthermore, we note that we have developed a wrapper R-package of \textbf{ergm}, called \textbf{cERGM}, to conveniently fit citation networks using c-ERGMs.
+\\
 

### 2020-08-28 Tex/PA_appendix_rr/appendix.tex
updated estimation details in main text
Specifically, term by term from 1950 to 2015. Justified because temporal context matters and network growth.

@@
 \usepackage{bm}
+%\usepackage{lmodern}
+%\usepackage{amsthm}
+%\usepackage{geometry} 
+\usepackage[T1]{fontenc}
 \usepackage{amssymb}
 \usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage[moderate]{savetrees}
-\usepackage{geometry}
-\usepackage{float}
 
@@
 The prevailing choice for $\theta_0$ is the maximum pseudo-likelihood estimation (MPLE) \citep{strauss1990pseudolikelihood}, a fast estimation method that is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
-The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated above, assume a logit form.
+The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated in the main text, assume a logit form.
 The MPLE is simple to obtain, but does not guarantee a starting value close to the MLE \citep{SchmidHunter2020}.
 
-We estimate the MCMLE for every term of the c-ERGM between 1950 and 2015.
+%We estimate the MCMLE for every term of the c-ERGM between 1950 and 2015.
 While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows.

### 2020-08-28 Tex/PA_rr/main.tex
updated estimation details in main text
Specifically, term by term from 1950 to 2015. Justified because temporal context matters and network growth.

@@
 The logit form conditional probability is well known for the ERGM family \citep{goodreau2009birds}.
-Estimation of the c-ERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix.
-The basis of our computation is the \textbf{ergm}-package \citep{HunterDavidR..2008} in \textbf{R} \citep{RCore}.
-In addition, a wrapper \textbf{R}-package called \textbf{cERGM} was developed to conveniently fit citation network data using the c-ERGM\footnote{The \textbf{cERGM}-package can be found on \url{http://github.com/schmid86}.}.
 
@@
 In the next section we explain the variables we construct using these data sources.
-We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015)\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the CourtListener data.
+We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015).\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the CourtListener data.
 We decided to exclude these 145 cases from our analysis.
 Additionally, since the most commonly used data on Supreme Court citations in political science come from \citet{fowler2007network}, we compared the Fowler data to the CourtListener data in the time interval that they overlap (1937--2001).
-We found considerable agreement---over 95\% of the citations recorded in the CourtListener dataset were also in the Fowler dataset, and over 96\% of the citations recorded in the Fowler data were also in the CourtListener dataset.}.
-The Supreme Court citation network from $1937 - 2015$ consists of $10,020$ cases.
+We found considerable agreement---over 95\% of the citations recorded in the CourtListener dataset were also in the Fowler dataset, and over 96\% of the citations recorded in the Fowler data were also in the CourtListener dataset.} The Supreme Court citation network from $1937 - 2015$ consists of $10,020$ cases.
 The breakdown of the data by the Court's Chief Justice is presented in Table \ref{tab:chiefs}.
@@
 We see here that the densest rates of tie formation tend to be between consecutive courts (e.g., the Stone Court is much more tightly tied to the Hughes Court than the Rehnquist Court is to the Hughes Court).
-This pattern lends preliminary support to the hypothesis that the rate of citations to a case decrease over time.
+This pattern lends preliminary support to the hypothesis that the rate of citations to a case decreases over time.
+
+We fit a c-ERGM for every Supreme Court term between 1950 and 2015, meaning that we have 66 models where the outcome variables are the citations sent during the given term.
+Our approach of fitting separate term-by-term models is motivated by the temporally-dynamic data generating process.
+First, as our results will show, the temporal context and justice composition of a court will lead to differences in how opinions are written and therefore how citations are used.
+Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network.
+This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term.
+Practically, our dataset is large enough to support this approach, with each term introducing the potential for hundreds-of-thousands of new citations.
+This gives us the statistical power necessary to estimate a separate set of parameter values for each term.
+Further, while the entire set of citations can in theory be fit in a single network with varying parameters, it is technically difficult.
+While the network for the 1950 term consists of a manageable $1,962$ nodes, the size of the network monotonically increases to $10,020$ nodes in the 2015 term, making estimation more challenging as the network grows.
+
+Estimation of the c-ERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix.
+The basis of our computation is the \textbf{ergm}-package \citep{HunterDavidR..2008} in \textbf{R} \citep{RCore}.
+In addition, a wrapper \textbf{R}-package called \textbf{cERGM} was developed to conveniently fit citation network data using the c-ERGM\footnote{The \textbf{cERGM}-package can be found on \url{http://github.com/schmid86/cERGM}.}.
 

### 2020-08-28 Tex/PA_rrmemo/PA_RnR_memo.tex
updated estimation details in main text
Specifically, term by term from 1950 to 2015. Justified because temporal context matters and network growth.

@@
 
-\noindent \textbf{R2.7:} \emph{ I think the manuscript would be greatly improved as well by pointing practitioners toward implementation softwarA point which may be useful to the authors in future work (though I'm not sure whether they find it has great utility for this methods piece): The authors find that there are not reliable effects for the ideological variables, and thus wonder, "whether, and if so, how, citations are shaped by ideological factors." I think one natural response could be that the ideological effects may actually be found within the dependence effects they investigate in some way, and I wonder if there are creative ways within the context of their model to test for that.}\\
+\noindent \textbf{R2.7:} \emph{ A point which may be useful to the authors in future work (though I'm not sure whether they find it has great utility for this methods piece): The authors find that there are not reliable effects for the ideological variables, and thus wonder, "whether, and if so, how, citations are shaped by ideological factors." I think one natural response could be that the ideological effects may actually be found within the dependence effects they investigate in some way, and I wonder if there are creative ways within the context of their model to test for that.}\\
 

### 2020-08-28 Tex/PA_rr/main.tex
addressing contribution points

@@
 We find robust support for the inherent complexity underpinning the formation of citation ties, and show that incorporating complex dependence into the model of citation formation significantly improves the model's fit.
+
+We offer three contributions .
+First, we advance our understanding of the factors that drive citations between U.S.
+Supreme Court opinions.
+Second, for those who study judicial citations in general (e.g., U.S.
+state supreme courts \citep{hinkle2016transmission}, international courts \citep{lupu2012precedent},  German lower courts \citep{berlemann2020disposition}), we illustrate network-theoretic considerations that are likely to apply beyond the context of Supreme Court opinions.
+Third, we offer a novel extension of a statistical model for networks, and disseminate this model as a package for the R statistical software \citep{cergm}, which can be used for any form of citations (e.g., patent citations are used to measure both causes and consequences of innovation in the field of political economy \citep{akcigit2018growth,dincer2019does},  citations among academic journal articles have been used to study gender bias in political methodology \citep{dion2018gendered} and international relations \citep{maliniak2013gender},  and citations in documents produced in the policymaking process have been used to study links between public policy and scientific expertise \citep{costa2016science,koontz2018use}).
+Our contributions are both substantive and methodological.
+We focus most heavily on the first---the study of the U.S.
+Supreme Court citation network, since we see it as the most substantial innovation with respect to the existing literature that we offer in the current paper.
 
@@
 The basis of our computation is the \textbf{ergm}-package \citep{HunterDavidR..2008} in \textbf{R} \citep{RCore}.
-In addition, a wrapper \textbf{R}-package called \textbf{cERGM} was developed to conveniently fit citation network data using the c-ERGM\footnote{The \textbf{cERGM}-package can be found on \url{http://github.com/schmid86/cERGM}.}.
+In addition, we developed a wrapper \textbf{R}-package called \textbf{cERGM} \citep{cergm} to conveniently fit citation network data using the c-ERGM.
 
@@
 
+Though we see the advancements in modeling the Supreme Court citations as our central contribution, we make two contributions to broader literatures on judicial citations and citation analysis more generally.
+First, our arguments regarding the dependencies that shape Supreme Court citations are likely to apply to citation networks formed among other court opinions, and we have provided a road map and tools for modeling such dependencies.
+Second, we have provided an extended version of ERGM that is appropriate for all forms of citation network analysis, and is available in a convenient statistical software package.
+
 \printbibliography

### 2020-08-28 Tex/PA_rr/main.tex
revising footnote and memo

@@
 The maximum difference is reached in 1983 with a difference of 10,990 points for the AIC and 10,930 points for the BIC.
-This fit comparison provides robust evidence that the dependence terms contribute significantly to the explanatory power of the model.\footnote{In the appendix we present conventional goodness-of-fit plots that are used to assess the structural fit for ERGM \citep{hunter2008goodness}, as well as degeneracy diagnostics \citep{mukherjee2020degeneracy}.
-We find that the full model fits the degree and edgewise shared partner distributions well, and is not degenerate.}
+This fit comparison provides robust evidence that the dependence terms contribute significantly to the explanatory power of the model.
+In the appendix we present conventional goodness-of-fit plots that are used to assess the structural fit for ERGM \citep{hunter2008goodness}, as well as degeneracy diagnostics \citep{mukherjee2020degeneracy}.
+We find that the full model fits the degree and edgewise shared partner distributions well, and is not degenerate.
 

### 2020-08-28 Tex/PA_rrmemo/PA_RnR_memo.tex
revising footnote and memo

@@
 In short, the change to the ERGM is slight, even when combining that with the necessary simulated estimation approach, so telling us why these changes are so useful beyond this network could drive home the methodological contribution.
-}\\
-
-\noindent \textbf{Addressed:}   \\
-
-\noindent \textbf{R1.5:} \emph{ Alternatively, one could think about this paper more in terms of its substantive contributions.
+Alternatively, one could think about this paper more in terms of its substantive contributions.
 The front end is deeply ingrained in the citations and judicial literatures, as are the data, of course, but little attention is given in the end to what we learn from this paper.
@@
 
-\noindent \textbf{Addressed:}   \\
+\noindent \textbf{Addressed:}   Both reviewers' comments on the lack of clarity regarding contributions were very helpful.
+In our revisions we have both strengthened and clearly stated our substantive and methodological contributions.
+We offer three contributions---advancing our understanding of Supreme Court citations, illustrating theoretical considerations and methodological choices that should be useful to researchers who study other forms of judicial citations, and a new extension of ERGM that makes it applicable to any sort of citation network.
+We now state these precisely in both the introduction and conclusion of the paper.
+To strengthen our substantive contributions, we have expanded the discussion of our results, including integrating that discussing with the literature on Supreme Court citations.
+To strengthen our methodological contributions, we have expanded our discussion of the c-ERGM itself, and developed a publicly-available R package in which c-ERGM is implemented.
 
@@
 
-\noindent \textbf{Addressed:}  \\
+\noindent \textbf{Addressed:}  BRUCE, TO DO.\\
 
@@
 What software was used?
-Do the authors generate novel convenience software to handle the specifics of the cERGM model and/or the computational overhead they discuss in the appendix?}\\
+Do the authors generate novel convenience software to handle the specifics of the cERGM model and/or the computational overhead they discuss in the appendix?
+These issues that I mention are particularly important if the authors intend for this model to be used by the judicial politics community rather than only network modelers in other contexts where the cERGM structure makes sense; a bit more practical information would be necessary I think for this to travel within the judicial politics community.}\\
 
-\noindent \textbf{Addressed:}  \\
-We discuss implementation details, especially the simulated annealing approach in greater detail in the appendix now.
-In addition, we now provide information in the main paper about the used statistical software.
-A new R-package called cERGM was developed to conveniently fit cERGMs for citation network data.
-The existence of this novel R-package and the corresponding hyperlink to the R-package are being mentioned in the main paper as well as in the appendix.
+\noindent \textbf{Addressed:}  We discuss implementation details, especially the simulated annealing approach in greater detail in the appendix now.
+In addition, we now provide information in the main paper about the statistical software we used.
+We have now developed a new R package called cERGM to conveniently fit c-ERGMs for citation network data.
+The existence of this novel R-package and the corresponding citation to the R-package are now mentioned in the main paper as well as in the appendix.
 \\
 
-\noindent \textbf{R2.3:} \emph{ These issues that I mention are particularly important if the authors intend for this model to be used by the judicial politics community rather than only network modelers in other contexts where the cERGM structure makes sense; a bit more practical information would be necessary I think for this to travel within the judicial politics community.}\\
-
-\noindent \textbf{Addressed:}  \\
-
-\noindent \textbf{R2.4:} \emph{ In many places the authors do a very good job of explaining the technical side of what they are doing precisely, as well as providing easy to digest intuitive explanations.
+\noindent \textbf{R2.3:} \emph{ In many places the authors do a very good job of explaining the technical side of what they are doing precisely, as well as providing easy to digest intuitive explanations.
 However, one place where we seem to somewhat lack both clarity as well as thoroughness and precision is unfornately the section where it was arguably needed the most: the section on dependence terms.
@@
 
-\noindent \textbf{R2.5:} \emph{ As a small comment, in relation to the degeneracy discussion, I think footnote 4 would be best served by not being a footnote.
+\noindent \textbf{R2.4:} \emph{ As a small comment, in relation to the degeneracy discussion, I think footnote 4 would be best served by not being a footnote.
 After your discussion of degeneracy I think the typical reader will be wondering, "Did this end up being a problem for their application?" The answer is, "No, we checked that and it was fine," and I just think that as it's such a natural thing for a reader to want to know after you spent multiple full paragraphs discussing the issue, this sentence deserves its place in the main text.}\\
 
-\noindent \textbf{Addressed:}  \\
+\noindent \textbf{Addressed:}  We have moved this text from the footnote to the body of the paper.\\
 
-\noindent \textbf{R2.6:} \emph{ I think the manuscript would be greatly improved as well by pointing practitioners toward implementation software, even if this is a very short, simple statement.
+\noindent \textbf{R2.5:} \emph{ I think the manuscript would be greatly improved as well by pointing practitioners toward implementation software, even if this is a very short, simple statement.
 For example, if the answer for that is, "We used software X which is often used for ERGMs, and our custom calls to that software to handle the cERGM and large network we analyze can be found in the replication archive," great.
@@
 
-\noindent \textbf{R2.7:} \emph{ A point which may be useful to the authors in future work (though I'm not sure whether they find it has great utility for this methods piece): The authors find that there are not reliable effects for the ideological variables, and thus wonder, "whether, and if so, how, citations are shaped by ideological factors." I think one natural response could be that the ideological effects may actually be found within the dependence effects they investigate in some way, and I wonder if there are creative ways within the context of their model to test for that.}\\
+\noindent \textbf{R2.6:} \emph{ A point which may be useful to the authors in future work (though I'm not sure whether they find it has great utility for this methods piece): The authors find that there are not reliable effects for the ideological variables, and thus wonder, "whether, and if so, how, citations are shaped by ideological factors." I think one natural response could be that the ideological effects may actually be found within the dependence effects they investigate in some way, and I wonder if there are creative ways within the context of their model to test for that.}\\
 

### 2020-08-30 Tex/PA_rr/main.tex
revised dependence effects

@@
 We include dependence terms to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, and popularity.
-%We expect that the coefficients associated with each of these dependence effects will be positive.
-
-The common dependence effects for these three dynamics are the number of triangles (case $i$ cites cases $j$ and $k$, and case $j$ cites case $k$), the number of mutual dyads (case $i$ cites $j$ and vice versa), the number of out-2-stars (case $i$ cites $j$ and $k$), and the number of in-2-stars (case $i$ is being cited by cases $j$ and $k$) in $C_{\leq t}$.
-Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy \citep{Handcock.2003}.
+The common dependence effects for these three dynamics are the number of triangles, the number of mutual dyads, and the number of in-2-stars.
+A triangle is a configuration in which there are citations connecting all three cases in a triad.
+From the perspective of a single tie closing a triad, a positive triangle coefficient indicates that a tie is more likely to form between nodes $i$ and $j$ if nodes $i$ and $j$ have citation ties with one or more cases $k$.
+A mutual dyad is a configuration in which two nodes within a dyad exchange directed ties.
+From the perspective of a node closing a mutual dyad with a tie, a positive coefficient indicates that case $i$ is more likely to cite case $j$ if case $j$ has cited case $i$ (mutual dyads are only possible within the same term).
+The in-2-star configuration is one in which two nodes in a triad send ties to the same third node.
+From the perspective of a case sending a tie to close an in-2-star, a positive coefficient indicates that case $i$ is more likely to send a tie to case $j$ if one or more other cases also cite case $j$.
+Unfortunately, adding the triangle or in-2-star statistic causes model degeneracy for the Supreme Court Citation Network---a common issue with these terms \citep{Handcock.2003}.
 Model degeneracy is a common obstacle when fitting ERGMs, and occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network.
@@
 In an attempt to stabilize the model against model degeneracy as well as to still capture the popularity effect, \citet{SnijdersTomA.B..2006} introduces the \textit{alternating-in-k-star} statistic, which was shown to be equivalent to the \textit{geometrically weighted indegree distribution} (gwidegree) statistic introduced by \citet{hunter2006inference}.
+The indegree distribution is the distribution of the number of ties sent to nodes, across all nodes in the network.
 Define $D_r(C_t, C_{<t})$ as the number of nodes with indegree $r, r \in \{0, \dots , N_t -1 \}$, where $N_t$ is the number of cases in the network at time $t$, then gwidegree represents a network's indegree distribution in a single statistic by geometrically weighting the degree distribution
@@
 \end{equation}
-The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on $D_r(C_t, C_{<t})$ as $r$ increases, which means that the higher $\lambda$ the more the statistic responds to changes for cases with a high indegree.
+The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on each node in the network based on their indegree ($D_r(C_t, C_{<t})$) as the indegree value ($r$ ) increases, which means that the higher $\lambda$ the more the statistic weighs cases with a high indegree.
 We fixed $\lambda=1$.
-Adding one single edge to the network changes the indegree distribution of the network so that a case with an indegree of $s$ turns into a case with an indegree of $s + 1$.
-With our notation this means that $D_r$ and $D_{r+1}$ get replaced by
-$D_r -1$ and $D_{r+1} +1$, while no changes for $D_\ell,
- \ell \in \{0, \dots , N_t -1 \} \setminus \{r, r-1\} $ are
-made. This changes the log odds in the following way
-$$\text{logit} P(C_{ij,t}=1 | C_{-ij,t}, C_{<t}) = \exp \bigl(~\theta \cdot( 1- \frac{1}{\lambda})^r \bigr).$$
+We chose this value for two reason.
+First, it results in a fairly straightforward interpretation of the coefficient for gwidegree, as Equation \ref{gwidegree} reduces to a function that counts the number of nodes that receive at least one tie.
+Our second reason for selecting $\lambda=1$ is that it results in an accurate fit in the indegree distribution of the observed network (as demonstrated in the appendix).
+At $\lambda=1$, the weight attributed to each case is equivalent regardless of the indegree value, as long as the indegree is greater than zero.
+This means that the statistic does not grow with the addition of high indegree cases to the network.
+It only grows through the addition of cases with at least one in-citation.
+With  $\lambda=1$, the coefficient associated with the gwidegree statistic governs the degree to which the edges are sent to a small number of popular nodes, or distributed fairly evenly across the nodes.
+A positive coefficient encourages edges to be distributed evenly across a large number of nodes---assuring that as many nodes as possible receive at least one tie.
+A negative gwidegree coefficient discourages network configurations in which many nodes receive ties, placing higher probability instead on configurations in which many of the ties are sent to a relatively small number of popular nodes.
 As stated by \citet{Levy2016}, this means for the interpretation that a negative parameter value indicates the centralization of edges (i.e., popularity) while a positive parameter indicates the dispersion of edges (i.e., new ties going to less popular nodes).
@@
 We will refer to this statistic as the \textit{different term transitivity} statistic.
-As illustrated in Section 2, we theorize Supreme Court citations to pattern a high degree of transitivity.
+This statistic captures the central form of transitivity that we hypothesize above---when a justice writes the opinion for case $i$, cites a past opinion ($k$) as precedent, and then cites one or more of the opinions ($j$) that were cited in opinion $k$.
+The two-path from opinion $i$ to $j$ is created when opinion $i$ cites an opinion ($k$) that cites an earlier opinion ($j$).
+The closed transitive triangle is created when opinion $i$ cites directly to $j$, one of the precedents used in opinion $k$.
+The edge $i \to j$ counts as a ``transitive tie'' since it closes at least one transitive triangle.
+The \textit{different term transitivity}  statistic is the count of the number of transitive ties in the network that connect opinions written in different terms.
+As illustrated in Section 2, we theorize Supreme Court citations to exhibit a high degree of transitivity.
 As a result, we expect positive parameter estimates for the \textit{different term transitivity} statistic.
 
-The second transitivity statistic, intended to capture clustering that includes more than one tie formed in the same term, is the \textit{geometrically weighted edgewise shared partners} (gwesp) statistic introduced by \citet{hunter2006inference}.
-For directed networks the edgewise shared $r$-partners statistic ($ESP_r$) can take multiple forms, but we integrate the definition where cases $i$ and $j$ that are connected by a citation, also cite $r$ common other cases.
-We follow the notation of \citet{Butts.2008} and call this ESP specification \textit{outgoing shared partners} (OSP).
-Similar to the gwidegree-statistic the gwesp-statistic incorporates a network's edgewise shared partners distribution by replacing $D_r$ by $ESP_r$.
-We fix the decay parameter at $\lambda=0.25$.
-Triangles that are incident to the outgoing shared partners configuration arise when there is a cite between two cases decided at the same time, and those two cases cite one or more of the same third cases (decided either previously or also during the same term).
-This would happen if citations were clustering around legal issues that are not completely captured by our same-issue-area effects.
-As we do not expect that the issue area variable will account for all of the clustering around legal issues, we expect the gwesp statistic to also carry a positive coefficient value.
+Whereas the first transitivity statistic is focused on modeling triangles that form through the citation of past cases, the second transitivity statistic we include captures clustering that includes ties formed between cases decided in the same term.
+We both expect and observe a higher level of transitivity among same-term cases since justices have the opportunity to confer about cases and coordinate citations among related opinions.
+For modeling within-term transitivity, we use the \textit{geometrically weighted edgewise shared partners} (gwesp) statistic introduced by \citet{hunter2006inference}.
+We use the form of the gwesp statistic that captures configurations in which cases $i$ and $j$ are connected by a citation ($i$ cites $j$, $j$ cites $i$, or both), and also cite $r$ common other cases (i.e., shared partners).
+\citet{Butts.2008} call this conceptualization of $r$ the number of \textit{outgoing shared partners}.
+This form of gwesp captures the degree to which two related cases in the same term are likely to be based on similar bodies of precedent.
+If the gwesp coefficient is positive, we interpret the effect as saying that the more shared partners $r$ to which cases $i$ and $j$ send ties, the greater the likelihood that there is a citation connecting cases $i$ and $j$.
+As with the gwdegree term, there is a decay parameter.
+The gwesp statistic exhibits decreasing marginal returns with respect to the number of shared partners (e.g., the fourth shared partner between $i$ and $j$ does not contribute as much to gwesp as does the first shared partner between $i$ and $j$).
+The decay parameter governs the degree of marginal return decrease, with a value of 0 indicating that additional shared partners do not contribute anything, and a value of $\infty$ indicating a linear (non-decreasing) function.
+We fix the decay parameter at $\lambda=0.25$, as this accurately captures the edgewise shared partner distribution, and avoids degeneracy (which would be a problem at higher values of $\lambda$).
+In general, researchers can adjust the gwesp parameter down to avoid degeneracy, and increase it to more accurately capture the shared partner distribution in the network.
+As with cross-term transitivity, we expect the gwesp statistic to also carry a positive coefficient value.
 
@@
 We expect that new cases will be more likely to cite cases that themselves cited many cases, because cases that themselves cite a large number of cases address a broader array of legal issues.
-We include the outdegree of every node that entered the network prior time $t$ as a receiver effect.
-We expect the coefficient on this term to be positive.
+We include the outdegree (i.e., the number of citations sent) of every node that entered the network prior time $t$ as a receiver effect.
+We expect the coefficient on this term to be positive, as a positive coefficient would mean that cases that send more citations are likely to also receive more citations.
 This is not technically a dependence term, as it is formulated as a covariate, but we discuss it in the current section because it is purely a network structure effect.

### 2020-08-30 Tex/PA_rrmemo/PA_RnR_memo.tex
revising memo
revised dependence effects

@@
 
-\noindent \textbf{Addressed:}   Both reviewers' comments on the lack of clarity regarding contributions were very helpful.
+\noindent \textbf{Addressed:}  The reviewer's comments on the lack of clarity regarding contributions were very helpful.
 In our revisions we have both strengthened and clearly stated our substantive and methodological contributions.
@@
 
-\noindent \textbf{Addressed:}  \\
+\noindent \textbf{Addressed:}  We agree that the original write-up of the dependence effects fell short in terms of both explaining how to interpret the effects, and providing guidance regarding how we made a number of specification choices.
+For each dependence effect we revised the coverage in three directions.
+First, we revised and expanded our explanations of both the intuitive and substantive interpretations of the effect.
+Second, we provided justifications for functional form and hyperparameter specifications (e.g., the decay parameters in gwidegree and gwesp).
+Third, we removed technical jargon and mathematical expressions that were not necessary for explaining and justifying what we did.
 

### 2020-08-31 Tex/PA_rr/main.tex
completed first round of revisions

@@
 Surprisingly, we do not find consistent effects for majority coalition size or for the two covariates that involve justice ideology.
-The effect of majority coalition size was generally not statistically significant until the later portion of the period examined; ideological distance between two cases had signs and significance levels that are highly inconsistent; and the ideological breadth of the potentially cited case was effectively zero for the entire period.
+The effect of majority coalition size was generally not statistically significant until the later portion of the period examined; ideological distance between two cases had signs and significance levels that are highly inconsistent; and the effect of the ideological breadth of the potentially cited case was effectively zero for the entire period.
+A purely measurement-based explanation of our findings is that we need a more precise measure of the ideological position of an opinion, as far as citation behavior goes, than the median ideal point of the justice in the majority coalition.
+Clustering that results from ideological homophily that is not effectively accounted for by our measure would be picked up, in part, by our measures of transitivity, which are statistically and substantively significant throughout the period studied.
+However, we do observe connections between our findings and the literature on decision-making on the Supreme Court.
 
-These results leave open the question of whether and how citations are shaped by ideological factors.
+Our results leave open the question of whether and how citations are shaped by ideological factors.
 While we did not formally hypothesize about temporal dynamics in the effects, a closer look at the observed trends allows us to identify some patterns that suggest directions for future research in this area.

### 2020-08-31 Tex/PA_rrmemo/PA_RnR_memo.tex
completed first round of revisions

@@
 
-\noindent \textbf{Addressed:}  BRUCE, TO DO.\\
+\noindent \textbf{Addressed:}  Please see our response to R2.3. .\\
 
@@
 Third, we removed technical jargon and mathematical expressions that were not necessary for explaining and justifying what we did.
+\\
 
@@
 
-\noindent \textbf{Addressed:} We now mention in the paper that we us \textbf{R} for the implementation of our model and that the code is based on the \textbf{ergm}-package.
-Furthermore, we note that we have developed a wrapper R-package of \textbf{ergm}, called \textbf{cERGM}, to conveniently fit citation networks using c-ERGMs.
+\noindent \textbf{Addressed:} We now mention in the paper that we use R for the implementation of our model and that the code is based on extensions of the ergm-package.
+Furthermore, we note that we have developed a wrapper R-package of ergm, called cERGM, to conveniently fit citation networks using c-ERGMs.
 \\
@@
 
-\noindent \textbf{Addressed:}  \\
+\noindent \textbf{Addressed:}  This comment raises an important point that we now discuss in the results section.
+If the median justice in the majority coalition is not a precise measure of the ideological position of an opinion, some of the clustering that may result from ideological homophily in citations would be picked up in the transitivity terms.
+Unfortunately, there is no way (at least that we can think of) to test for this without improving upon the measure of opinion ideology.
 

### 2020-08-31 Tex/PA_rr/main.tex
meshed ideological explanations

@@
 
-We offer three contributions .
+We offer three contributions.
 First, we advance our understanding of the factors that drive citations between U.S.
@@
 state supreme courts \citep{hinkle2016transmission}, international courts \citep{lupu2012precedent},  German lower courts \citep{berlemann2020disposition}), we illustrate network-theoretic considerations that are likely to apply beyond the context of Supreme Court opinions.
-Third, we offer a novel extension of a statistical model for networks, and disseminate this model as a package for the R statistical software \citep{cergm}, which can be used for any form of citations (e.g., patent citations are used to measure both causes and consequences of innovation in the field of political economy \citep{akcigit2018growth,dincer2019does},  citations among academic journal articles have been used to study gender bias in political methodology \citep{dion2018gendered} and international relations \citep{maliniak2013gender},  and citations in documents produced in the policymaking process have been used to study links between public policy and scientific expertise \citep{costa2016science,koontz2018use}).
+Third, we offer a novel extension of a statistical model for networks, and disseminate this model as a package for the R statistical software \citep{cergm}, which can be used for any form of citations.
+For example, patent citations are used to measure both causes and consequences of innovation in the field of political economy \citep{akcigit2018growth,dincer2019does}, citations among academic journal articles have been used to study gender bias in political methodology \citep{dion2018gendered} and international relations \citep{maliniak2013gender}, and citations in documents produced in the policymaking process have been used to study links between public policy and scientific expertise \citep{costa2016science,koontz2018use}.
 Our contributions are both substantive and methodological.
@@
 
-These three properties---transitivity, reciprocity, and popularity form the core of our network theory of U.S.
+These three properties---transitivity, reciprocity, and popularity---form the core of our network theory of U.S.
 Supreme Court citations.
@@
 The effect of majority coalition size was generally not statistically significant until the later portion of the period examined; ideological distance between two cases had signs and significance levels that are highly inconsistent; and the effect of the ideological breadth of the potentially cited case was effectively zero for the entire period.
+
+Our results leave open the question of whether and how citations are shaped by ideological factors.
 A purely measurement-based explanation of our findings is that we need a more precise measure of the ideological position of an opinion, as far as citation behavior goes, than the median ideal point of the justice in the majority coalition.
 Clustering that results from ideological homophily that is not effectively accounted for by our measure would be picked up, in part, by our measures of transitivity, which are statistically and substantively significant throughout the period studied.
-However, we do observe connections between our findings and the literature on decision-making on the Supreme Court.
 
-Our results leave open the question of whether and how citations are shaped by ideological factors.
-While we did not formally hypothesize about temporal dynamics in the effects, a closer look at the observed trends allows us to identify some patterns that suggest directions for future research in this area.
+Beyond potential measurement issues, we do observe connections between our findings and the literature on decision-making on the Supreme Court.
+These connections are evident from a closer look at temporal dynamics in the effects.
 In this discussion, we focus on the shift in the effect of ideological distance beginning with the Rehnquist court and extending into the Roberts court.

### 2020-09-02 Tex/PA_appendix_rr/appendix.tex
final revisions

@@
 
-\section{c-ERGM Estimation}
+\section{cERGM Estimation}
 
@@
 The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases.
-This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the c-ERGM.
+This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the cERGM.
 
@@
 
-%We estimate the MCMLE for every term of the c-ERGM between 1950 and 2015.
+%We estimate the MCMLE for every term of the cERGM between 1950 and 2015.
 While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows.
@@
 Since for some networks the MCMLE was only obtainable using simulated annealing method to find a starting value, the final results in the paper have all been estimated using simulated annealing.
-The simulated annealing algorithm for finding an improved starting value for c-ERGMs was implemented in the \textbf{cERGM}-package for \textbf{R} \citep{RCore} and can be found at \url{http://github.com/schmid86/cERGM}.
+The simulated annealing algorithm for finding an improved starting value for cERGMs was implemented in the \textbf{cERGM}-package for \textbf{R} \citep{RCore} and can be found at \url{http://github.com/schmid86/cERGM}.
 

### 2020-09-02 Tex/PA_rr/main.tex
final revisions

@@
 The first network property that we theorize in the context of Supreme Court citations is transitivity.
-In a network of directed relations (e.g., A cites B, but B doesn't cite A) transitivity refers to the tendency for A to send a tie to C if A sends a tie to B and B sends a tie to C \citep{holland1971transitivity}.
+In a network of directed relations (e.g., $i$ cites $j$, but $j$ doesn't cite $i$) transitivity refers to the tendency for $i$ to send a tie to $k$ if $i$ sends a tie to $j$ and $j$ sends a tie to $k$ \citep{holland1971transitivity}.
 In undirected networks, transitivity is simply the process by which friends of friends become friends (i.e., a friend of a friend is a friend).
-The term, ``transitive closure'' refers to a tie forming from A to C in response to extant ties from A to B and B to C.
+The term, ``transitive closure'' refers to a tie forming from $i$ to $k$ in response to extant ties from $i$ to $j$ and $j$ to $k$.
 When writing opinions, Supreme Court justices present the legal bases for their rulings, which often involves discussing the most primary/relevant precedents underpinning these legal bases, but also the precedents and legal rules on which the primary precedents were based.
-This process of presenting several layers/levels of precedent in an opinion follows the structure of transitive closure exactly---opinion A cites opinion B as a primary precedent, and then cites opinion C because opinion B cites opinion C.
+This process of presenting several layers/levels of precedent in an opinion follows the structure of transitive closure exactly---opinion $i$ cites opinion $j$ as a primary precedent, and then cites opinion $k$ because opinion $j$ cites opinion $k$.
 The two examples presented below illustrate this process.
@@
 
-We develop the citation ERGM (c-ERGM), to account for the structural constraints that apply to the network of Supreme Court citations.
+We develop the citation ERGM (cERGM), to account for the structural constraints that apply to the network of Supreme Court citations.
 These structural constraints amount to three departures from the structure of networks for which ERGMs are currently designed.
@@
 The structure of the citation network is depicted in Figure \ref{fig:ctergm}---a hypothetical citation network established over three time periods, with three cases decided in each time period.
-We denote $C_t$ to be the set of citation and non-citations added to the network at time $t$ (i.e., via the addition of three cases), $C_{ <t}$ to be the citations among cases decided before time $t$, and $C_{ \leq t}$ to denote the entire set of citations and non-citations on which $C_t$ can depend through the c-ERGM specification.
+We denote $C_t$ to be the set of citation and non-citations added to the network at time $t$ (i.e., via the addition of three cases), $C_{ <t}$ to be the citations among cases decided before time $t$, and $C_{ \leq t}$ to denote the entire set of citations and non-citations on which $C_t$ can depend through the cERGM specification.
 
@@
 
-The likelihood function of the c-ERGM is given by
+The likelihood function of the cERGM is given by
 \begin{equation}
@@
 
-Though the likelihood function of the c-ERGM may appear quite different from that of statistical models conventionally used in political science, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients.
+Though the likelihood function of the cERGM may appear quite different from that of statistical models conventionally used in political science, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients.
 $$ P(C_{ij,t} = 1 | C_{-ij,t}, C_{ < t}) = \frac{\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right]}{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right] + \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 0) } \right]}, $$ $$ \text{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~} = \frac{1}{ 1 + \exp \left[ - {\bm{\theta}'\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)} \right]}, $$
@@
 
-We fit a c-ERGM for every Supreme Court term between 1950 and 2015, meaning that we have 66 models where the outcome variables are the citations sent during the given term.
+We fit a cERGM for every Supreme Court term between 1950 and 2015, meaning that we have 66 models where the outcome variables are the citations sent during the given term.
 Our approach of fitting separate term-by-term models is motivated by the temporally-dynamic data generating process.
@@
 
-Estimation of the c-ERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix.
+Estimation of the cERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix.
 The basis of our computation is the \textbf{ergm}-package \citep{HunterDavidR..2008} in \textbf{R} \citep{RCore}.
-In addition, we developed a wrapper \textbf{R}-package called \textbf{cERGM} \citep{cergm} to conveniently fit citation network data using the c-ERGM.
+In addition, we developed a wrapper \textbf{R}-package called \textbf{cERGM} \citep{cergm} to conveniently fit citation network data using the cERGM.
 
@@
 
-\subsection{c-ERGM specification}
+\subsection{cERGM specification}
 
-The c-ERGM we estimate includes two classes of terms---one that captures the effects of covariates on tie formation, and another that captures the complex dependence processes that we expect to observe in the Supreme Court Citation Network.
+The cERGM we estimate includes two classes of terms---one that captures the effects of covariates on tie formation, and another that captures the complex dependence processes that we expect to observe in the Supreme Court Citation Network.
 We describe our model specification by defining the terms within these two classes.
@@
 
-Covariate effects are accounted for in the c-ERGM via the term that is used to specify the effect of covariates in other ERGM family models, as
+Covariate effects are accounted for in the cERGM via the term that is used to specify the effect of covariates in other ERGM family models, as
 $$h_{covariate}(C_t,C_{<t},X) =  \sum_{ij} C_{ij,t}X_{ij}.$$ Since $C_{ij,t}$ is a binary indicator of whether case $i$ cites case $j$, $h_{covariate}$ amounts to the sum of covariate values among directed dyads for which we observed a citation.
@@
 
-\subsection{c-ERGM Coefficient Estimates}
-The estimates from the c-ERGM are presented in Figures \ref{SCC_results_1} and \ref{SCC_results_2}.
-In these figures we depict coefficient estimates and 95\% confidence intervals for all of the effects for all the variables we discussed in sections \ref{covariate_terms} and \ref{dependence_terms}.
+\subsection{cERGM Coefficient Estimates}
+The estimates from the cERGM are presented in Figures \ref{SCC_results_1} and \ref{SCC_results_2}.
+In these figures we depict coefficient estimates and 95\% confidence intervals for all of the effects for all the variables we discussed in section \ref{covariate_terms}.
 The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in the respective variable.
@@
 \subsubsection{Covariate Terms}
-Effects of the exogenous covariates included in the c-ERGM are presented in Figure \ref{SCC_results_2}.
+Effects of the exogenous covariates included in the cERGM are presented in Figure \ref{SCC_results_2}.
 Most of the exogenous covariate effects align with expectations and have relatively stable parameter estimates over time.
@@
 We recognize that there are alternative explanations---perhaps one that is related to the coterminous change in the effect of majority coalition size---but a deeper look into the relevant literature is beyond the scope of this paper.
-Our discussion serves to illustrate that term-by-term c-ERGMs can uncover important temporal dynamics that courts researchers can use to study a much greater range of citation-related phenomenon beyond what we present here.
+Our discussion serves to illustrate that term-by-term cERGMs can uncover important temporal dynamics that courts researchers can use to study a much greater range of citation-related phenomenon beyond what we present here.
 To the extent that the US Supreme Court is a political entity that interacts with the broader society, the body of court opinions and citations between them is corpus that continues to grow based on both its history and its contemporary context.
@@
 We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network.
-This methodology---the citation-ERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation.
+This methodology---the citation ERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation.
 We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015.

### 2020-09-03 Tex/PA_rr/main.tex
final revisions

@@
 Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited in an opinion.
-We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{harris1982structural,caldeira1988legal,fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics,ethayarajh2018rose}.
+We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{caldeira1988legal,fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics,ethayarajh2018rose}.
 
@@
 
-There exist two broad benefits of conducting empirical analyses of citations at the directed dyad level.
+There are two benefits of analyzing citations at the directed dyad level.
 The first is that directed dyadic analyses can test both dyadic and case-level hypotheses.
 For example, case-level analyses can model whether opinions supported by a liberal majority coalition are more likely than those supported by a conservative majority coalition to be cited heavily in the future, but they cannot precisely model whether liberal cases will be cited more by liberal cases than by conservative cases.
-Thus, the first reason for analyzing citations at the dyadic level is to expand the set of hypotheses that can be represented in the model.
+Thus, the first reason for analyzing citations at the dyadic level is to expand the set of hypotheses that can be tested.
 The second reason for studying citations at the directed dyadic level is that, as articulated in the growing literature on legal citation networks, citations form complex networks in which a citation at one point in time may influence future citations.
@@
 Third, we offer a novel extension of a statistical model for networks, and disseminate this model as a package for the R statistical software \citep{cergm}, which can be used for any form of citations.
-For example, patent citations are used to measure both causes and consequences of innovation in the field of political economy \citep{akcigit2018growth,dincer2019does}, citations among academic journal articles have been used to study gender bias in political methodology \citep{dion2018gendered} and international relations \citep{maliniak2013gender}, and citations in documents produced in the policymaking process have been used to study links between public policy and scientific expertise \citep{costa2016science,koontz2018use}.
+For example, patent citations are used to measure both causes and consequences of innovation in the field of political economy \citep{akcigit2018growth,dincer2019does}, citations in academic journal articles and syllabi have been used to study gender bias in political science  \citep{dion2018gendered,maliniak2013gender,hardt2019gender,atchison2017negating}, and citations in documents produced in the policymaking process have been used to study links between public policy and scientific expertise \citep{costa2016science,koontz2018use,pattyn2020knowledge}.
 Our contributions are both substantive and methodological.
@@
 The first network property that we theorize in the context of Supreme Court citations is transitivity.
-In a network of directed relations (e.g., $i$ cites $j$, but $j$ doesn't cite $i$) transitivity refers to the tendency for $i$ to send a tie to $k$ if $i$ sends a tie to $j$ and $j$ sends a tie to $k$ \citep{holland1971transitivity}.
+In a network of directed relations (e.g., $i$ cites $j$, but $j$ doesn't cite $i$) transitivity refers to the tendency for $i$ to send a tie to $k$ if $i$ sends a tie to $j$ and $j$ sends a tie to $k$ \citep{holland1971transitivity,hallinan1990sex}.
 In undirected networks, transitivity is simply the process by which friends of friends become friends (i.e., a friend of a friend is a friend).
@@
 
+The second network property that we hypothesize is reciprocity \citep{erikson2013formalist} among cases are decided in the same term.
+We expect that opinions that are written within the same term, and cover highly similar cases will cite each other.
 The first case in our example reciprocal dyad is a passage from Western Air Lines v.
@@
 The third, and final, network property we consider in the context of Supreme Court citations is popularity.
-Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent \citep{barabasi1999emergence}.
+Popularity, also termed ``preferential attachment'' is the tendency for ties to be sent to nodes to which many ties have already been sent \citep{barabasi1999emergence,chayes2013mathematics}.
 Citations to an opinion signal both the Court's awareness of the legal reasoning of the case and the Court's evaluation that the opinion is an authoritative precedent.
@@
 We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above.
-To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data, and has seen extensive application in recent political networks research---the exponential random graph model (ERGM) \citep[e.g.,][]{box2014evolution,duque2018recognizing,osei2018elite}.
+To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data, and has seen extensive application in recent political networks research---the exponential random graph model (ERGM) \citep[e.g.,][]{bratton2011networks, box2014evolution,duque2018recognizing,osei2018elite}.
 %The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
@@
 \centering
-\includegraphics[width = 0.85\textwidth ]{daggish.pdf}
+\includegraphics[width = 0.775\textwidth ]{daggish.pdf}
 \caption{Illustration of temporal structure of the Supreme Court Citation Network.
@@
 The shaded small squares are hypothetical observed citations, and the white small squares are citations that could have been observed but were not.
-The regions of the citation matrix that are represented by large white rectangles are citations that could not have been observed since the citing case would have been decided in a term that preceded the term of the cited case.
+The regions of the matrix that are represented by large white rectangles are citations that could not have been observed since the citing case would have been decided in a term that preceded the term of the cited case.
 The citing case ID is given in the row and the prospective cited case is given in the column.
@@
 
-Though the likelihood function of the cERGM may appear quite different from that of statistical models conventionally used in political science, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients.
+Though the likelihood function of the cERGM is quite different from those of conventional regression models, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients.
 $$ P(C_{ij,t} = 1 | C_{-ij,t}, C_{ < t}) = \frac{\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right]}{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right] + \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 0) } \right]}, $$ $$ \text{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~} = \frac{1}{ 1 + \exp \left[ - {\bm{\theta}'\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)} \right]}, $$
@@
 
-We fit a cERGM for every Supreme Court term between 1950 and 2015, meaning that we have 66 models where the outcome variables are the citations sent during the given term.
+We fit the cERGM separately for every Supreme Court term between 1950 and 2015, meaning that we have 66 models where the outcome variables are the citations sent during the given term.
 Our approach of fitting separate term-by-term models is motivated by the temporally-dynamic data generating process.
@@
 Further, while the entire set of citations can in theory be fit in a single network with varying parameters, it is technically difficult.
-While the network for the 1950 term consists of a manageable $1,962$ nodes, the size of the network monotonically increases to $10,020$ nodes in the 2015 term, making estimation more challenging as the network grows.
+While the network for the 1950 term consists of a manageable $1,962$ nodes, the size of the network increases to $10,020$ nodes in the 2015 term, making estimation more challenging.
 
@@
 The basis of our computation is the \textbf{ergm}-package \citep{HunterDavidR..2008} in \textbf{R} \citep{RCore}.
-In addition, we developed a wrapper \textbf{R}-package called \textbf{cERGM} \citep{cergm} to conveniently fit citation network data using the cERGM.
+In addition, we developed a wrapper \textbf{R}-package called \textbf{cERGM} \citep{cergm} to conveniently fit the cERGM.
 
@@
 We account for this effect following \citet{spriggs2001explaining}, who find that cases are more likely to be overruled when the Court is ideologically distant from the median justice in the majority coalition that decided the case.
-\citet{clark2010locating} estimates a latent coordinate model of Supreme Court opinions based on the network of case-to-case citations.
-They find that the majority opinion falls at the ideal point of the median member of the majority coalition in the case.
+\citet{clark2010locating} find that the majority opinion falls at the ideal point of the median member of the majority coalition in the case.
 We include a covariate term in which $X_{ij}$ is the absolute difference between the Martin-Quinn scores \citep{martin2002dynamic} of the median justices in the majority coalitions for cases $i$ and $j$.
-We expect this variable to have a negative effect, which would correspond to cases citing those to which they are ideologically similar.
+We expect this variable to have a negative effect, which would mean that cases cite those that are ideologically similar.
 %WE COULD USE THIS DATASET AS A SECOND APPLICATION

### 2020-09-03 Tex/PA_rrmemo/PA_RnR_memo.tex
final revisions

@@
 \usepackage{bm}
-\usepackage[backend=bibtex]{biblatex}    
+%\usepackage[backend=bibtex]{biblatex}    
+\usepackage[authordate,backend=bibtex]{biblatex-chicago} 
 \usepackage[english]{babel}
@@
 \noindent We are grateful for the opportunity to revise and resubmit our manuscript. In this memo
-we have separated the reviewers' comments into separate criticisms and
+we have organized the reviewers' comments into separate criticisms and
 suggestions. Under each comment, we describe how we have revised the manuscript in
@@
 
-\noindent \textbf{Addressed:} All networks were estimated using MCMLE, however, for smaller networks we used the network's MPLE as the algorithms's starting value, while for larger networks the starting value was obtained using simulated annealing, since the MPLE as starting value did not lead to converged MCMLEs anymore.
-While in theory the algorithm should converge to the MLE regardless of the starting value, in practice this is not the case   \cite{hummel2012improving}.
+\noindent \textbf{Addressed:} All networks were estimated using MCMLE, however, for smaller networks we used the network's MPLE as the algorithms's starting value, while for larger networks the starting value was obtained using simulated annealing, since the MPLE as starting value did not lead to converged MCMLE.
+While in theory the algorithm should converge to the MLE regardless of the starting value, in practice this is not the case   (\cite{hummel2012improving}).
+In order to be consistent with the used estimation approach, we estimate the MCMLE now for all networks using simulated annealing starting value. The results have not changed noticeably. 
+The text in the appendix has been updated accordingly.
+In the appendix we also now cite a recently-released paper on arXiv (\cite{SchmidHunter2020}), in which the simulated annealing approach to finding start values for ERGM is described in detail.
 \\
-In order to be consistent with the used estimation approach, we estimate the MCMLE now for all networks using simulated annealing starting value. The results have not changed noticeably.  \\
-The text in the appendix has been updated accordingly.\\
 
@@
 \noindent \textbf{Addressed:} We have re-estimated all networks using the simulated annealing approach and the results look very similar to the initially submitted results.
-This is not surprising, since the MCMLE algorithm only reports a succesful estimation if it has converged to the unique MLE.
+This is not surprising, since the MCMLE algorithm only reports a succesful estimation if it has converged to the unique approximate MLE.
 However, with a poorly chosen starting value the algorithm might fail to converge towards the MLE.
@@
 
-First, we situated our more unexpected findings (i.e.
+First, we situated our more surprising findings (i.e.
 negative outdegree count and ideological distance) in the judicial politics and legal studies literatures.
-For these findings, we offer explanations that comport with the field's understanding of justice behavior.
+For these findings, we offer explanations that comport with the field's understanding of judicial behavior.
 The hub effect is negative because outdegree count reflects case complexity, and justices avoid citing these opinions in anticipation of them being overruled.
@@
 To strengthen our substantive contributions, we have expanded the discussion of our results, including integrating that discussing with the literature on Supreme Court citations.
-To strengthen our methodological contributions, we have expanded our discussion of the c-ERGM itself, and developed a publicly-available R package in which c-ERGM is implemented.
+To strengthen our methodological contributions, we have expanded our discussion of the cERGM itself, and developed a publicly-available R package in which cERGM is implemented.
 
+\newpage
 \section*{Reviewer: 2}
@@
 In addition, we now provide information in the main paper about the statistical software we used.
-We have now developed a new R package called cERGM to conveniently fit c-ERGMs for citation network data.
-The existence of this novel R-package and the corresponding citation to the R-package are now mentioned in the main paper as well as in the appendix.
+We have now developed a new R package called cERGM to conveniently fit cERGMs for citation network data.
+The existence of this new R-package and the corresponding citation to the package are now mentioned in the main paper as well as in the appendix.
 \\
@@
 
-\noindent \textbf{Addressed:} We now mention in the paper that we use R for the implementation of our model and that the code is based on extensions of the ergm-package.
-Furthermore, we note that we have developed a wrapper R-package of ergm, called cERGM, to conveniently fit citation networks using c-ERGMs.
+\noindent \textbf{Addressed:} We now mention in the paper that we use R for the implementation of our model and that the code is based on extensions of the ergm package.
+Furthermore, we note that we have developed a wrapper R-package of ergm, called cERGM, to conveniently analyze citation networks using cERGMs.
 \\

### 2020-09-03 Tex/PA_rr/main.tex
finished revisions

@@
 We present a methodology for studying citations between Supreme Court opinions at the dyadic level, as a network, that overcomes these limitations.
-This methodology---the citation exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation.
+This methodology---the citation exponential random graph model, for which we provide user-friendly software---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation.
 We then analyze a network that includes all Supreme Court cases decided between 1950 and 2015.
 We find evidence for dependence processes, including reciprocity, transitivity, and popularity.
-The dependence effects are as substantively and statistically significant as the effects of the exogenous covariates we include in the model, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.\footnote{
+The dependence effects are as substantively and statistically significant as the effects of exogenous covariates, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.
+\footnote{
 This work was supported in part by NSF grants SES-1558661, SES-1637089, SES-1619644, and CISE-1320219. }
@@
 
-The second network property that we hypothesize is reciprocity \citep{erikson2013formalist} among cases are decided in the same term.
+The second network property that we hypothesize is reciprocity \citep{erikson2013formalist} among cases that are decided in the same term.
 We expect that opinions that are written within the same term, and cover highly similar cases will cite each other.

### 2021-01-03 NetworkVisualizations/scc_intro_plots_PA.R
added subgraph plotting in intro

@@ -1,0 +1,87 @@
+# Prep
+library(statnet)
+load("supreme.RData"); rm(model20.mle, scnetwork, edge_col, edge_row)
+par(mar=c(0,0,0,0))
+
+orng.c<-"#fcae91FF"
+purp.c<-"#7828a0FF"
+gren.c<-"#46aa96FF"
+lgry.c<-"#dcdcdcFF"
+mgry.c<-"#8c8c8cFF"
+dgry.c<-"#333333FF"
+
+# Plotting function
+plot.local<-function(cases,seed,mode="both",label.pos=1){
+  cid<-node_attributes$caseid[which(node_attributes$usid%in%uscases)]
+  if(mode=="in"){
+    tel<-edgelist[which(edgelist[,2]%in%cid),]
+  }
+  if(mode=="out"){
+    tel<-edgelist[which(edgelist[,1]%in%cid),]
+  }
+  if(mode=="both"){
+    tel<-edgelist[which(edgelist[,1]%in%cid|edgelist[,2]%in%cid),]
+  }
+  temp.net<-network(tel)
+  
+  e.col<-rep(0,nrow(tel))
+  e.col[which(tel[,1]%in%cid&tel[,2]%in%cid)]<-1
+  
+  set.seed(seed)
+  par(mar=c(0,0,0,0))
+  plot(temp.net,
+       vertex.col=ifelse((temp.net%v%"vertex.names")%in%cid,orng.c,adjustcolor(mgry.c,alpha.f=0.4)),
+       vertex.border=ifelse((temp.net%v%"vertex.names")%in%cid,rgb(0,0,0),adjustcolor(mgry.c,alpha.f=0.4)),
+       vertex.cex=ifelse((temp.net%v%"vertex.names")%in%cid,2,1),
+       edge.col=ifelse(e.col,adjustcolor(purp.c,alpha.f=0.9),rgb(0,0,0,0.1)),
+       label=sub("US"," U.S. ",node_attributes$names[match(temp.net%v%"vertex.names",node_attributes$caseid)]),
+       label.border=rgb(0,0,0,0),label.bg=ifelse((temp.net%v%"vertex.names")%in%cid,rgb(1,1,1,0.5),rgb(0,0,0,0)),boxed.labels=T,label.pad=1,label.cex=1.2,
[…]

### 2021-01-07 Tex/PA_rr/main.tex
Added citation of replication data
updated back matter

@@
 The dependence effects are as substantively and statistically significant as the effects of exogenous covariates, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.
-\footnote{
-This work was supported in part by NSF grants SES-1558661, SES-1637089, SES-1619644, and CISE-1320219. }
 
@@
 Our three data sources for this study include the Supreme Court Database (SCDB) \citep{spaeth2014supreme}, Martin-Quinn scores \citep{martin2002dynamic}, and Supreme Court citation data provided by the CourtListener Free Law Project \citep{CourtListener}.
-In the next section we explain the variables we construct using these data sources.
+In the next section we explain the variables we construct using these data sources\footnote{Data and replication code are provided by \citet{replicationdata}}.
 We limit the Supreme Court terms included in our analysis to those that are covered by all three of these data source (1937--2015).\footnote{There were 145 cases that were listed in the SCDB but could not be matched to a case in the CourtListener data.
@@
 
-One way to achieve the kind of balance required for support is to cite broadly from both sides of the argument \citep{wilkinson2005rehnquist}, which increases the likelihood of citing ideologically distant cases.
-Consider for example Hamdi v.
-Rumsfeld (542 U.S.
-1, 2004) which ruled on the question of whether a U.S.
-citizens can be detained as an ``enemy combatant''.
-The court, in a four-justice plurality opinion supported by two additional concurring justices, decided at the same time that ``a state of war is not a blank check for the President'' and the due process assessment must ``pay keen attention to the particular burdens faced by the Executive in the context of military action''.
-In total, the plurality opinion in Hamdi v.
-Rumsfeld cited 47 opinions, with cites as distant as 2.42 and 2.13 on either side of its median position.
-
-In the above discussion, we identified an explanation for the observed temporal dynamics in the effect of ideological distance.
-Specifically, ideologically distant cases become more important as courts adopt the relatively new `split-the-difference' jurisprudential approach which requires broad citations.
-We recognize that there are alternative explanations---perhaps one that is related to the coterminous change in the effect of majority coalition size---but a deeper look into the relevant literature is beyond the scope of this paper.
-Our discussion serves to illustrate that term-by-term cERGMs can uncover important temporal dynamics that courts researchers can use to study a much greater range of citation-related phenomenon beyond what we present here.
-To the extent that the US Supreme Court is a political entity that interacts with the broader society, the body of court opinions and citations between them is corpus that continues to grow based on both its history and its contemporary context.
-The methods we present here afford researchers the tools to better understand these complexities.
-
 \begin{figure}
@@
 \end{figure}
+
+One way to achieve the kind of balance required for support is to cite broadly from both sides of the argument \citep{wilkinson2005rehnquist}, which increases the likelihood of citing ideologically distant cases.
+Consider for example Hamdi v.
+Rumsfeld (542 U.S.
+1, 2004) which ruled on the question of whether a U.S.
+citizens can be detained as an ``enemy combatant''.
+The court, in a four-justice plurality opinion supported by two additional concurring justices, decided at the same time that ``a state of war is not a blank check for the President'' and the due process assessment must ``pay keen attention to the particular burdens faced by the Executive in the context of military action''.
+In total, the plurality opinion in Hamdi v.
+Rumsfeld cited 47 opinions, with cites as distant as 2.42 and 2.13 on either side of its median position.
+
+In the above discussion, we identified an explanation for the observed temporal dynamics in the effect of ideological distance.
+Specifically, ideologically distant cases become more important as courts adopt the relatively new `split-the-difference' jurisprudential approach which requires broad citations.
+We recognize that there are alternative explanations---perhaps one that is related to the coterminous change in the effect of majority coalition size---but a deeper look into the relevant literature is beyond the scope of this paper.
+Our discussion serves to illustrate that term-by-term cERGMs can uncover important temporal dynamics that courts researchers can use to study a much greater range of citation-related phenomenon beyond what we present here.
+To the extent that the US Supreme Court is a political entity that interacts with the broader society, the body of court opinions and citations between them is corpus that continues to grow based on both its history and its contemporary context.
+The methods we present here afford researchers the tools to better understand these complexities.
 
@@
 
+\section*{Funding}
+This work was supported in part by NSF grants SES-1558661, SES-1637089, SES-1619644, and CISE-1320219.
+
+\section*{Acknowledgements}
+We thank participants at the 2018 SPSA and 2018 APSA meetings for helpful feedback.
+
+\section*{Data Availability Statement}
+The replication code and data for this paper can be found at \citet{replicationdata}.
+
 \printbibliography

### 2021-01-14 Tex/accepted_preprint/accepted.tex
updated accepted preprint files

--- 2020-09-03 Tex/PA_rr/main.tex
@@
-\documentclass{cup-pan}
+\documentclass[headsepline=true, abstracton]{scrartcl}
+
 \usepackage[utf8]{inputenc}
-\usepackage{blindtext}
-\usepackage{bm}
-%\usepackage{lmodern}
-%\usepackage{amsthm}
-%\usepackage{geometry} 
-\usepackage[T1]{fontenc}
+%\usepackage[T1]{fontenc}
+
 \usepackage{amssymb}
 \usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{bm}
 
-% original loaded packages
-%\documentclass{cup-pan}
-%\usepackage[utf8]{inputenc}
-%\usepackage{blindtext}
-%\usepackage{bm}
-%\usepackage[T1]{fontenc}
-%\usepackage{lmodern}
+\usepackage{natbib}
+\usepackage{hyperref}
+\hypersetup{colorlinks=true, citecolor=darkgray, linkcolor=darkgray, urlcolor=darkgray}
 
-\title{Generative Dynamics of Supreme Court Citations: Analysis with a New Statistical Network Model}
+\usepackage[moderate]{savetrees}
 
-\author[1]{Christian S. Schmid}
-\author[2]{Ted Hsuan Yun Chen}
-\author[3]{Bruce A. Desmarais}
+\usepackage[table,xcdraw]{xcolor}
 
-\affil[1]{Department of Statistics, The Pennsylvania State University, University Park, PA. Email: \url{schmid@psu.edu}}
-\affil[2]{Faculty of Social Sciences, University of Helsinki, Finland. Email: \url{ted.hsuanyun.chen@gmail.com}}
-\affil[3]{Department of Political Science, The Pennsylvania State University, University Park, PA. Email: \url{bdesmarais@psu.edu}}
+\usepackage{graphicx}
 
-%% Corresponding author
-\corrauthor{Christian S. Schmid}
+ \usepackage{float}
 
-%% Abbreviated author list for the running footer
-\runningauthor{Schmid et al.}
+\usepackage{setspace}
 
-\addbibresource{bib.bib}
+\usepackage{url}
+
+\usepackage{geometry}
 
@@
 
+\renewcommand{\refname}{Bibliography}
+
+\onehalfspacing
+\setlength{\headsep}{15mm}
+
+\thispagestyle{plain}
+
+\title{\Large Generative Dynamics of Supreme Court Citations: \\ Analysis with a New Statistical Network Model\thanks{
+		This work was supported in part by NSF grants SES-1558661, SES-1637089, SES-1619644, and CISE-1320219.
+We thank Rachel Hinkle, Benjamin Kassow, Lauren Santoro, and Michael Nelson for helpful feedback on this project.
+The replication code and data for this paper can be found at \citet{replicationdata}.}}
+%\author{}
+ % Toggle % below to blind and unblind the manuscript
+%  \author[1]{Christian Schmid \thanks{cxs5700@psu.edu}}
+%   \author[2]{Ted Hsuan Yun Chen\thanks{thc126@psu.edu}}
+% \author[2]{Bruce A. Desmarais\thanks{bdesmarais@psu.edu}}
+% \author[1]{David R. Hunter \thanks{dhunter@stat.psu.edu}}
+% \affil[2]{Department of Political Science, Pennsylvania State University}
+ %\affil[1]{Department of Statistics, Pennsylvania State University}
+
+\author{%
+  Christian S. Schmid\footnote{Department of Statistics, The Pennsylvania State University, schmid@psu.edu}%
+  \and Ted Hsuan Yun Chen\footnote{Faculty of Social Sciences, University of Helsinki, ted.hsuanyun.chen@gmail.com}%
+   \and Bruce A. Desmarais\footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
+%  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
+  }
+
+\date{Conditionally Accepted at \textit{Political Analysis}\\ December 29, 2020}
+
 \maketitle
-
 \begin{abstract}
@@
 The dependence effects are as substantively and statistically significant as the effects of exogenous covariates, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.
-\footnote{
-This work was supported in part by NSF grants SES-1558661, SES-1637089, SES-1619644, and CISE-1320219. }
-
-\keywords{US Supreme Court, majority opinions, citations, network, ERGM.}
 \end{abstract}
 
+\onehalfspacing
 \section{Introduction}
@@
 	\centering
-\includegraphics[width = 0.49\textwidth,trim= 2cm 2cm 3cm 2cm,clip=true ]{citations_trans.pdf}\hspace{0.2cm}
+	\includegraphics[width = 0.49\textwidth,trim= 2cm 2cm 3cm 2cm,clip=true ]{citations_trans.pdf}\hspace{0.2cm}%
 	\includegraphics[width = 0.49\textwidth,trim= 2cm 2cm 3cm 2cm,clip=true ]{citations_recip.pdf}
@@
 The citations to Baker v.
-Carr are visualized in Figrue \ref{fig:popularity}.
+Carr are visualized in Figure \ref{fig:popularity}.
 \begin{quotation}
@@
 \begin{table}[bt]
-\caption{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
-\\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 5/2020).}
 	\label{tab:chiefs}
 	\centering
-\begin{tabular}{l l l l r}
-\headrow \thead{Chief Justice} & \thead{Terms} & \thead{Total Number Cases} & \thead{Cases/Term}  \\
+	\begin{tabular}{l l c r}
+		 {Chief Justice} & {Terms} & {Total Number Cases} & {Cases/Term} \\ \hline
 		CE Hughes*              & 1937 - 1941                                          & 628                                                               & 125.6                                                     \\
@@
 	\end{tabular}
-
+		\caption{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941.
+\\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 5/2020).}
 \end{table}
@@
 \begin{table}[bt]
-\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
+
 	\label{issue_area_coding}
@@
 	\end{tabular}
+	\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
 \end{table}
@@
 
-\printbibliography
+\singlespacing
+\bibliography{bib} 
+\bibliographystyle{apsr}
+
+\onehalfspacing
+\appendix
+\section{cERGM Estimation}
+
+The normalizing constant in Equation 1 of the main text is intractable.
+For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure 4 of the main text---there are 16,777,216 unique configurations of $C_t$ that could be observed.
+The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases.
+This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the cERGM.
+
+The common alternative relies on Monte Carlo methods to approximate the normalizing constant by simulating a large set of networks \citep{hunter2006inference,hummel2012improving}.
+The resulting estimator, the Monte Carlo MLE (MCMLE), is approximately consistent, meaning that it converges to the MLE as the sample size, i.e., the number of simulated networks, increases.
+However, one drawback is that with the number of nodes in the Supreme Court citation network being in the order of $10,000$, obtaining the MCMLE is computationally expensive \citep{schmid2017exponential} and the success of the algorithm heavily relies on the starting parameter vector $\theta_0$, which is ideally chosen in the proximity of the unknown MLE \citep{hummel2012improving}.
+The prevailing choice for $\theta_0$ is the maximum pseudo-likelihood estimation (MPLE) \citep{strauss1990pseudolikelihood}, a fast estimation method that is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network.
+The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated in the main text, assume a logit form.
+The MPLE is simple to obtain, but does not guarantee a starting value close to the MLE \citep{SchmidHunter2020}.
+
+%We estimate the MCMLE for every term of the cERGM between 1950 and 2015.
+While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows.
+The growth in the network also serves as our reason for estimating a separate model for each term.
+First, the network is large, with each term introducing the potential for hundreds-of-thousands of new citations.
+This means we have the statistical power necessary to estimate a separate set of parameter values for each term.
+Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network.
+This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term.
+
+The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant.
+However, the estimation of most networks in the 90s with the MPLE as starting values was not feasible in an reasonable time frame anymore.
+Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$.
+But even this approach started to fail for networks around the turn of the millennium.
+Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}.
+The MCMLE for these large citation networks was obtained by setting the starting value according to an novel approach introduced by \citet{SchmidHunter2020}.
+This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$.
+However, the MPLE of networks with the same sufficient statistics is not necessarily the same.
+Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a new network $C_t^*$ on the same nodes as the observed network that satisfies $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$, and has a weak dependence structure among unfixed ties.
+Such a network can be found using simulated annealing algorithms \citep{Kirkpatrick83}.
+For networks with a weak dependence structure among unfixed ties, the MPLE is similar to the MLE, in addition, the same sufficient statistic between $C_t^*$ and the observed network $C_t$ guarantees the same MLE between these two networks.
+This makes the MPLE of $C_t^*$ an effective starting value for the MCMLE algorithm.
+Since for some networks the MCMLE was only obtainable using simulated annealing method to find a starting value, the final results in the paper have all been estimated using simulated annealing.
+The simulated annealing algorithm for finding an improved starting value for cERGMs was implemented in the \textbf{cERGM}-package for \textbf{R} \citep{RCore} and can be found at \url{http://github.com/schmid86/cERGM}.
+
+\section{Goodness-of-Fit}
+We evaluate the goodness-of-fit of the model following \citet{hunter2008goodness} by examining the distribution of four hyper statistics, e.g., the out- and indegree distribution and the distribution for two different edgewise shared partners statistics.
+OTP stands for \textit{outgoing two-paths} and refers to the number of cases $r$ that are cited by case $i$ and that cite case $j$, while $j$ is also directly cited by $i$.
+The second ESP statistic is the OSP specification that has been introduced in section 4.1.2 in the paper.
+Figure \ref{GOF} visualizes the goodness-of-fit results for the citation network for the 1950 (top) and 2015 (bottom) term.
+The solid black line indicates the statistic's distribution in the Supreme Court citation network of that given term and the boxplots depict the statistic's distribution of $1000$ networks that have been simulated from the ERGM defined by the MCMLE.
+This means that in the ideal case the solid black line passes through ever single boxplot.
+
+\begin{figure}[bt]
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{GOF_1950}\\
+		%\caption{Goodness-of-fit diagnostic for the 1950 network.}
+		% \label{GOF_1950}
+		\vspace{1cm}
+		\includegraphics[width=0.9\textwidth]{GOF_2015}
+		\vspace{0.1cm}
+		\caption{Goodness-of-fit diagnostic for the 1950 network (top) and the 2015 network (bottom).}
+		\label{GOF}
+		\vspace{-.25cm}
+	\end{center}
+\end{figure}
+
+We see that our models do a good job capturing the out and indegree distribution of the citation network, since the black line falls almost exclusively within the ranges spanned by the boxplots.
+For the ESP distributions we can observe that the number of ties with $r=0$ shared partners is captured well for both the OTP as well as for the OSP statistic.
+However, the model overestimates the number of $r=1$ shared partners and then, especially in the 2015 term network, underestimates the number of ties with more than $r=1$ shared partners.
+
+\section{Checking for Model Degeneracy}
+
+A common challenge when fitting ERGMs is model degeneracy.
+Model degeneracy occurs when the probability distribution defined by the parameter vector does not predominantly yield networks with similar statistics as the observed network.
+Generally, model degeneracy results in simulated networks with no ties or all possible ties.
+In a non-degenerate model the statistics of the networks that were simulated from the probability distribution defined by the MCMLE fall in the proximity of the observed network's statistics.
+Figures \ref{mcmcdiagnostics_1950} and \ref{mcmcdiagnostics_2015} depict trace and density plots for the dependence terms in the 1950 and 2015 term citation network.
+The histograms on the left visualize a statistic's density from $1000$ simulated networks, while the right side shows the statistic's trace plot of the same $1000$ networks.
+The solid black line indicates the statistic's value in the actual citation network.
+Both figures indicate that this model is non-degenerate and that the simulated network's statistics fall almost evenly around the observed statistic.
+The density and trace plots for the ERGM of the terms not depicted provide similar results.
+
+\begin{figure}[bt]
+	\begin{center}
+		\includegraphics[width=1\textwidth]{Deg_1950}
+		\caption{Density and trace plots for the dependency terms of the 1950 term citation network.}
+		\label{mcmcdiagnostics_1950}
+		\vspace{-.25cm}
+	\end{center}
+\end{figure}
+
+\begin{figure}[bt]
+	\begin{center}
+		\includegraphics[width=1\textwidth]{Deg_2015}
+		\caption{Density and trace plots for the dependency terms of the 2015 term citation network.}
+		\label{mcmcdiagnostics_2015}
+		\vspace{-.25cm}
+	\end{center}
+\end{figure}
+
 \end{document}

### 2021-01-15 Tex/accepted_preprint/accepted.tex
fixed typo in acknowledgements
removed comments

@@
 		This work was supported in part by NSF grants SES-1558661, SES-1637089, SES-1619644, and CISE-1320219.
-We thank Rachel Hinkle, Benjamin Kassow, Lauren Santoro, and Michael Nelson for helpful feedback on this project.
+We thank Rachael Hinkle, Benjamin Kassow, Lauren Santoro, and Michael Nelson for helpful feedback on this project.
 The replication code and data for this paper can be found at \citet{replicationdata}.}}
-%\author{}
- % Toggle % below to blind and unblind the manuscript
-%  \author[1]{Christian Schmid \thanks{cxs5700@psu.edu}}
-%   \author[2]{Ted Hsuan Yun Chen\thanks{thc126@psu.edu}}
-% \author[2]{Bruce A. Desmarais\thanks{bdesmarais@psu.edu}}
-% \author[1]{David R. Hunter \thanks{dhunter@stat.psu.edu}}
-% \affil[2]{Department of Political Science, Pennsylvania State University}
- %\affil[1]{Department of Statistics, Pennsylvania State University}
 
@@
    \and Bruce A. Desmarais\footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
-%  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
   }
@@
 
-% \begin{figure}[bt]
-% \centering
-% \includegraphics[width = 0.85\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{citations_recip.pdf}
-% \caption{Illustration of a reciprocal tie between two US Supreme Court opinions.}
-% \label{fig:reciprocity}
-% \end{figure}
-
 The second network property that we hypothesize is reciprocity \citep{erikson2013formalist} among cases that are decided in the same term.
@@
 To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data, and has seen extensive application in recent political networks research---the exponential random graph model (ERGM) \citep[e.g.,][]{bratton2011networks, box2014evolution,duque2018recognizing,osei2018elite}.
-%The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).
 The network structures for which ERGMs are currently designed are insufficient to account for the structure of citation networks.
@@
 The denominator of Equation 1 represents a normalizing constant, in which the positive weight is summed over all possible configurations of $C_{t}$, from the network in which the cases added to the network at time $t$ send no citations at all, to the network in which the cases added at time $t$ cite every possible case, and everything in between.
-%The probability of observing the citation network up to time $T$---the final time point in the dataset---is given by the product over the probability of adding the citations at each time point $t$, conditional on the citations that were added to the network previously.
 
@@
 		\includegraphics[width = \textwidth]{courts_layer.png}
-		% \includegraphics[width = \textwidth,clip=true,trim=0.5cm 3.75cm 0.75cm 0.5cm ]{courts_layer_updated.pdf}
 	\end{minipage}
@@
 \end{figure}
-
-% \begin{figure}[bt]
-% \centering
-% \includegraphics[scale = 0.45 ]{degree_distribution.pdf}
-% \caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2015. There are cases with an in- or outdegree  $>$50, but they are not captured in this figure.}
-%  \label{degree_dist}
-% \end{figure}
-
-% \begin{figure}[h]
-% \centering
-% \includegraphics[scale=0.20,clip=true,trim=.5cm 2cm 0cm 0cm ]{courts_layer_updated.pdf}
-% \caption{ Supreme Court Citation Network, 1937-2015.
-Nodes are Supreme Court cases, color-coded based on the chief justice presiding over the court.
-Each network layer contains cases decided under a different chief justice.
-Because our data is temporally bound, cases for the Hughes and Roberts courts are not complete.
-}
-%  \label{fig:networkviz}\end{figure}
 
@@
 We expect this variable to have a negative effect, which would mean that cases cite those that are ideologically similar.
-%WE COULD USE THIS DATASET AS A SECOND APPLICATION
-
-%We include three sets of dummy variables that account for the issue areas of cases.
-In the first set of dummy variables---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$) (e.g., for the Criminal Procedure indicator, $X_{ij} = 1$ if case $i$ is a criminal procedure case, and $0$ otherwise).
-The second set of dummy variables represent receiver issue area intercepts in which $X^{ij}$ indicates the issue of the receiver case.
-The third set of dummy variables account for the rate of citation within each issue area and between each combination of issue areas.
-There is a separate dummy variable for each directed pair of issue areas.
-For example the A$\rightarrow$B dummy variable $X_{ij} = 1$ if case $i$ is issue area A and case $j$ is issue area B, and 0 otherwise.
-Note that, given the high dimension of these dummy variables, we omit one each of the sender and recipient indicators, and 28 of the 196 mixing rate indicators.
-Unfortunately, this leaves these indicators largely uninterpretable, but these variables are nonetheless important to absorb variation in citations that is attributable to the legal domains of two cases in a dyad.
-The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}.
-We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case.
-Table \ref{issue_area_coding} reports the issue areas covered in our data.
 
@@
 Table \ref{issue_area_coding} reports the issue areas covered in our data.
-
-%We also include a set of covariates that account for justice-specific variation in citations.
-We include two types of terms at the justice level.
-First, we include an indicator variable that models the rate at which justices cite themselves.
-This effect is modeled with a single indicator variable in which $X_{ij} = 1$ if the majority opinions in cases $i$ and $j$ were written by the same justice.
-Second, we use two sets of dummy variables to account for differences across justices both in terms of their tendencies to include citations in the opinions they write and to be cited by other opinions.
-The justice sender dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $i$.
-The justice recipient dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $j$.
-This set of variables is motivated by the use of citations by \citet{kosma1998measuring} to measure the influence of individual Supreme Court justices.
 
@@
 This is not technically a dependence term, as it is formulated as a covariate, but we discuss it in the current section because it is purely a network structure effect.
-
-%In order to account for the popularity effect, we add the \textit{geometrically weighted indegree distribution} (gwidegree) statistic to the model.
-The geometrically weighted distribution statistics were introduced by \citet{Hunter.2006} to counteract model degeneracy.
 
@@
 
-%The results from the c-TERGM provide evidence that citations between US Supreme Court cases are driven both by exogenous case attributes and complex dependence processes.
-The scales of covariate and dependence effects are comparable.
-%The simple linear trends represent one limitation of the current analysis.
-In future work, more complex polynomials dynamical functions should be considered to determine whether effects truly flip signs.
-
 \section{Conclusion}
@@
 
-%We estimate the MCMLE for every term of the cERGM between 1950 and 2015.
-While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows.
-The growth in the network also serves as our reason for estimating a separate model for each term.
-First, the network is large, with each term introducing the potential for hundreds-of-thousands of new citations.
-This means we have the statistical power necessary to estimate a separate set of parameter values for each term.
-Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network.
-This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term.
-
 The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant.
@@
 		\includegraphics[width=0.9\textwidth]{GOF_1950}\\
-		%\caption{Goodness-of-fit diagnostic for the 1950 network.}
-		% \label{GOF_1950}
 		\vspace{1cm}
